### 2022-08-17 code/L0/L0_1_Ecuador_plant_species_lists.Rmd
set up file and folder structure

---
title: "Ecuador plant species list"
authors: "Hazel Anderson"
collaborators: "None"
data input: "None"
data output: "Ecuador_plant_species_list.csv"
project: "Frugivoria"
date: "2022-08-17"
output: html_document
---
# Set Working directory & file path
```{r}
setwd()
data_path <- 'G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0'
```

# Load required packages
```{r}
library(BIEN)
library(kewr)
library(gbifdb)
```

# Pull data from various sources

## BIEN
```{r}
Ecuador <- BIEN_occurrence_country("Ecuador")
```

## Kew
```{r}

```

## GBIF
```{r}

```