### 2022-09-21 archive/BIENDataVisualization.Rmd
add archive files -- old code from Hazel undergrad

---
title: "BIEN data visualization"
author: "Hazel Anderson"
date: "3/1/2021"
output: html_document
---

```{r}
library(BIEN)
library(sp)
library(sf)
library(raster)
library(ggplot2)
library(dplyr)
```

```{r}
#load files
traits <- read.csv("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/L0/Ecuador_species_withTraits.csv")
allplant_rich <- raster("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/analysis/MODIS_data_2019/plant_ric.tif")
fruiting_rich <- raster("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/analysis/MODIS_data_2019/fruitingplant_ric.tif")
forestfruiting_rich <- raster("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/analysis/MODIS_data_2019/fruitingplant_rich_great_50.tif")
```

```{r}
allplant_rich@crs
fruiting_rich@crs
forestfruiting_rich@crs

plot(allplant_rich)
plot(fruiting_rich)
plot(forestfruiting_rich)
```
```{r}
allplant_rich@extent
fruiting_rich@extent
forestfruiting_rich@extent
```
```{r}
allplant_rich.pts <- rasterToPoints(allplant_rich, spatial=TRUE)
fruiting_rich.pts <- rasterToPoints(fruiting_rich, spatial=TRUE)
forestfruiting_rich.pts <- rasterToPoints(forestfruiting_rich, spatial=TRUE)
```

```{r}
# extract values from maps
all_rich.data <- extract(allplant_rich, allplant_rich.pts)
fruiting_rich.data <- extract(fruiting_rich, fruiting_rich.pts)
forestfruiting_rich.data <- extract(forestfruiting_rich, forestfruiting_rich.pts)
```

```{r}
elevation <-raster("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/analysis/MODIS_data_2019/srtm_1km.tif")
elevation@crs
#THE LINE BELOW DOESN"T WORK
ecuador_elevation <-crop(elevation, c(xmin=-91.3884, xmax=-70.00002, ymin=--10.0014, ymax=10))
#suggestion--intersect elevation with richness points; check to make sure they are the same resolution
```
```{r}
#Plotting fruiting species richness vs elevation
#Combine .data and .pts into a dataframe
fruiting_rich.pts$richness <- fruiting_rich.data

#NEED to get richness data and elevation data to be the same length
plot(fruiting_rich.pts, ecuador_elevation)
```

```{r}

subset_traits <- table(traits$scrubbed_species_binomial, traits$trait_name)
subset_traits

#WANT TO PLOT Trait_name by number of species with record for that trait
#ALSO want to map one trait in Space; NEED to subset to all records with values for selected trait
```