### 2022-10-10 code/L0/L0_1_Ecuador_plants_BIEN.Rmd
update top header information

---
title: "Ecuador plant species BIEN data"
authors: "Hazel J. Anderson"
collaborators: "None"
data input: "none"
data output: "Ecuador_BIEN_species_list.csv, Ecuador_BIEN_occurrence.csv, Ecuador_BIEN_traits.csv"
project: "Frugivoria"
date: "2022-08-17"
output: html_document
---
# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
for (package in c("BIEN", "tidyr", "mde")) {
  if (!require(package, character.only=T, quietly=T)) {
    install.packages(package)
    library(package, character.only=T)
  }
}
```

# Get data for Ecuador from BIEN database

```{r}
# occurrence data
Ecuador_BIEN_occ <- BIEN_occurrence_country("Ecuador")
# species list
Ecuador_BIEN_species_list <- BIEN_list_country(country = "Ecuador")
# trait data
Ecuador_BIEN_traits <- BIEN_trait_country("Ecuador")
```

## Data summary
```{r}
summary(Ecuador_BIEN_occ)
na_summary(Ecuador_BIEN_occ, exclude_cols = c("date_collected"))
```

```{r}
summary(Ecuador_BIEN_species_list)
na_summary(Ecuador_BIEN_species_list)
```

```{r}
summary(Ecuador_BIEN_traits)
na_summary(Ecuador_BIEN_traits)
```

# Write data to csv
```{r}
Ecuador_BIEN_occurrence_path <- file.path(output_path,"Ecuador_BIEN_occurrence.csv") 
write.csv(Ecuador_BIEN_occ, file = Ecuador_BIEN_occurrence_path)

Ecuador_BIEN_species_list_path <- file.path(output_path, "Ecuador_BIEN_species_list.csv")
write.csv(Ecuador_BIEN_species_list, file = Ecuador_BIEN_species_list_path)

Ecuador_BIEN_traits_path <- file.path(output_path, "Ecuador_BIEN_traits.csv")
write.csv(Ecuador_BIEN_traits, file = Ecuador_BIEN_traits_path)
```