### 2022-10-10 code/L0/L0_Ecuador_elevation.Rmd
Initalizing

---
title: "Ecuador Elevation"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "None"
data output: "TBD"
project: "Frugivoria"
date: "2022-10-10"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Install and load required packages
```{r}
library(elevatr)
library(sf)
library(raster)
```
# Take location coordinates and convert to sf object
```{r}
#x= long, y = lat
ecuador <- data.frame(x = -79.3666965, y = -1.3397668) %>%
  st_as_sf(coords = c("x", "y"), crs = 4326)
```

```{r}
elevation <- get_elev_raster(ecuador, z = 9)
```