### 2022-10-10 code/L0/L0_Ecuador_plants_GBIF.Rmd
added summary information and write data to csv

---
title: "Ecuador plant species GBIF"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "None"
data output: "TBD"
project: "Frugivoria"
date: "2022-10-10"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Install and load required packages
```{r}
library(rgbif)
library(mde)
```

# Set GBIF User information
```{r}
#only run once
#install.packages("usethis")
#usethis::edit_r_environ()
```
# Get Taxon Key for Plants
```{r}
name_backbone("Plantae")
```

```{r}
occ_download(
pred("taxonKey", 6), 
pred("hasGeospatialIssue", FALSE),
pred("hasCoordinate", TRUE),
pred("country","EC"),
pred("occurrenceStatus","PRESENT"), 
pred_not(pred_in("basisOfRecord",c("FOSSIL_SPECIMEN","LIVING_SPECIMEN"))),
format = "SIMPLE_CSV"
)
```
# Retrive GBIF download, save file, and load into r
```{r}
d <- occ_download_get("0065583-220831081235567", path = output_path) %>%
  occ_download_import()
```

# Summary 
```{r}
summary(d)
na_summary(d, exclude_cols = c("gbifID"))
```