### 2022-10-10 code/L1/L1_Ecuador_plant_species_BIEN.Rmd
Inital commit; create L1 folder

---
title: "Ecuador plant species BIEN Cleaning"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "Ecuador_BIEN_species_list.csv, Ecuador_BIEN_occurrence.csv, Ecuador_BIEN_traits.csv"
data output: "TBD"
project: "Frugivoria"
date: "2022-10-10"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(taxize)
library(taxList)
```

# Read in BIEN data
```{r}
Ecuador_BIEN_occurrence <- read.csv(file.path(data_path,"Ecuador_BIEN_occurrence.csv"))

Ecuador_BIEN_species_list <- read.csv(file.path(data_path, "Ecuador_BIEN_species_list.csv"))

Ecuador_BIEN_traits_path <- read.csv(file.path(data_path, "Ecuador_BIEN_traits.csv"))
```

```{r}

```