### 2022-10-12 code/L0/L0_Ecuador_elevation.Rmd
change package used to geodata Still getting an error and can't download data (connection error)

---
title: "Ecuador Elevation"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "None"
data output: "TBD"
project: "Frugivoria"
date: "2022-10-10"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Install and load required packages
```{r}
library(geodata)
```

```{r}
ecuador_outline <- gadm(country = "ECU", level=1, path = data_path, version="latest", resolution=1)

```

# Download elevation data for Ecuador
```{r}
ecuador <- elevation_30s(country = "ECU", path = data_path)
ecuador
```