### 2022-10-12 code/L0/L0_Ecuador_plants_WFO.Rmd
Initial commit download WFO backbone

---
title: "Ecuador plant species WorldFlora"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "TBD"
data output: "TBD"
project: "Frugivoria"
date: "2022-10-11"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(WorldFlora)
```

```{r}
if(! file.exists(file.path(data_path,'WFO.zip'))){
 download.file("http://104.198.143.165/files/WFO_Backbone/_WFOCompleteBackbone/WFO_Backbone.zip", dest=file.path(data_path,"WFO.zip"), mode="wb") 
}
```

```{r}
unzip (file.path(data_path,"WFO.zip"), exdir = data_path)
```