### 2022-10-27 code/L0/L0_Ecuador_plants_TRY.Rmd
inital commit with header information

---
title: "Ecuador plant species TRY data"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "None"
data output: "TBD"
project: "Frugivoria"
date: "2022-10-27"
output: html_document
---

Public data path "G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/TRY_Ecuador_PublicData_10192020.txt"