### 2022-10-31 code/L0/L0_Ecuador_plants_TRY.Rmd
add file paths, load required packages, read in data, explore TRY data using rtry package

---
title: "Ecuador plant species TRY data"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "None"
data output: "TBD"
project: "Frugivoria"
date: "2022-10-27"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(rtry)
```

# Read in TRY data
Data was obtained from https://www.try-db.org/TryWeb/Prop0.php 
```{r}
data <- rtry_import("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/TRY_Ecuador_PublicData_10192020.txt")
```
# Explore TRY data
```{r}
# explore by trait id and trait name
TRYdata_explore_trait <- rtry_explore(data, TraitID, TraitName)
TRYdata_explore_trait

# explore by species id, species name, trait id, and trait name
TRYdata_explore_species <- rtry_explore(data, AccSpeciesID, AccSpeciesName, TraitID, TraitName)
TRYdata_explore_species

# explore by data id, data name, trait id, and trait name sorted by trait id
TRYdata_explore_anc <- rtry_explore(data, DataID, DataName, TraitID, TraitName, sortBy = TraitID)
TRYdata_explore_anc
```