### 2022-11-09 code/L1/L1_Ecuador_plants_BIEN.Rmd
change packages, add pre-filtering code for BIEN_occurrences

---
title: "Ecuador plant species BIEN Cleaning"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "Ecuador_BIEN_species_list.csv, Ecuador_BIEN_occurrence.csv, Ecuador_BIEN_traits.csv"
data output: "TBD"
project: "Frugivoria"
date: "2022-10-10"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(bdc)
```

# Read in BIEN data
```{r}
Ecuador_BIEN_occurrence <- read.csv(file.path(data_path,"Ecuador_BIEN_occurrence.csv"))
Ecuador_BIEN_species_list <- read.csv(file.path(data_path, "Ecuador_BIEN_species_list.csv"))
Ecuador_BIEN_traits_path <- read.csv(file.path(data_path, "Ecuador_BIEN_traits.csv"))
```

# Pre-filter
```{r}
# Flag records missing species names
check_pf <-
  bdc_scientificName_empty(
  data = Ecuador_BIEN_occ,
  sci_name = "scrubbed_species_binomial")
```

```{r}
# Flag records missing partial or complete information on geographic coordinates
check_pf <- bdc_coordinates_empty(
  data = check_pf,
  lat = "latitude",
  lon = "longitude")
```

```{r}
# Flag records with out-of-range coordinates: latitude > 90 or -90; longitude >180 or -180
check_pf <- bdc_coordinates_outOfRange(
  data = check_pf,
  lat = "latitude",
  lon = "longitude")

```
```{r}
# Deriving country names for records missing country names
check_pf <- bdc_country_from_coordinates(
  data = check_pf,
  lat = "latitude",
  lon = "longitude",
  country = "country")
```

```{r}
# Country names are standardized against a list of country names in several languages retrieved from Wikipedia
check_pf <- bdc_country_standardized(
  data = check_pf,
  country = "country"
)
```

```{r}
# Correcting latitude and longitude transposed
check_pf <-
  bdc_coordinates_transposed(
    data = check_pf,
    id = "database_id",
    sci_names = "scrubbed_species_binomial",
    lat = "decimalLatitude",
    lon = "decimalLongitude",
    country = "country_suggested",
    countryCode = "countryCode",
    border_buffer = 0.2, # in decimal degrees (~22 km at the equator)
    save_outputs = FALSE
  )
```

```{r}
# Records outside one or multiple reference countries
check_pf <-
  bdc_coordinates_country_inconsistent(
    data = check_pf,
    country_name = "Ecuador",
    country = "country_suggested",
    lon = "longitude",
    lat = "latitude",
    dist = 0.1 # in decimal degrees (~11 km at the equator)
  )
```

```{r}
# Report
check_pf <- bdc_summary_col(data = check_pf)
report <-
  bdc_create_report(data = check_pf,
                    database_id = "database_id",
                    workflow_step = "prefilter",
                    save_report = FALSE)
report
```

```{r}
# Figures
figures <-
bdc_create_figures(data = check_pf,
                   database_id = "database_id",
                   workflow_step = "prefilter",
                   save_figures = FALSE)

# Check figures using
figures$.coordinates_country_inconsistent

```

```{r}
# Filter out flags
output <-
  check_pf %>%
  dplyr::filter(.summary == TRUE) %>%
  bdc_filter_out_flags(data = ., col_to_remove = "all")
```