### 2022-11-14 code/L1/L1_Ecuador_plants_TRY.Rmd
inital commit set file paths, load packages, read in try data

---
title: "Ecuador plant trait TRY Cleaning"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "Ecuador_TRY_plant_traits.csv"
data output: "TBD"
project: "Frugivoria"
date: "2022-11-14"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(rtry)
```

# Read in IUCN occurrence data
```{r}
try_data <- read.csv(file.path(data_path,"Ecuador_TRY_plant_traits.csv"))
```

# Explore data
```{r}
TRYdata_explore_trait <- rtry_explore(try_data, TraitID, TraitName)
View(TRYdata_explore_trait)
```