### 2022-11-16 code/L1/L1_Ecuador_plants_TRY.Rmd
Add code to transform dataframe to wide table

---
title: "Ecuador plant trait TRY Cleaning"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "Ecuador_TRY_plant_traits.csv"
data output: "TBD"
project: "Frugivoria"
date: "2022-11-14"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(rtry)
```

# Read in IUCN occurrence data
```{r}
try_data <- read.csv(file.path(data_path,"Ecuador_TRY_plant_traits.csv"))
```

# Transform to wide table
```{r}
#-------------------------------------------------
# Exclude
# 1. All entries with "" in TraitID
# 2. Potential categorical traits that don't have a StdValue
# 3. Traits that have not yet been standardized in TRY
# Then select the relevant columns for transformation
# Note: The complete.cases() is used to ensure the cases are complete,
#       i.e. have no missing values
#-------------------------------------------------
num_traits <- rtry_select_row(try_data, complete.cases(TraitID) & complete.cases(StdValue))
num_traits <- rtry_select_col(num_traits, ObservationID, AccSpeciesID, AccSpeciesName, TraitID, TraitName, StdValue, UnitName)
# Extract the unique value of latitude (DataID 59) and the corresponding ObservationID
workdata_lat <- rtry_select_anc(try_data, 59)

# Extract the unique value of longitude (DataID 60) and the corresponding ObservationID
workdata_lon <- rtry_select_anc(try_data, 60)

# To merge the extracted ancillary data with the numerical traits
# Merge the relevant data frames based on the ObservationID using rtry_join_left (left join)
num_traits_georef <- rtry_join_left(num_traits, workdata_lat, baseOn = ObservationID)
num_traits_georef <- rtry_join_left(num_traits_georef, workdata_lon, baseOn = ObservationID)

# Perform wide table transformation on TraitID, TraitName and UnitName
# With cell values to be the mean values calculated for StdValue
num_traits_georef_wider <- rtry_trans_wider(num_traits_georef, names_from = c(TraitID, TraitName, UnitName), values_from = c(StdValue), values_fn = list(StdValue = mean))
```