### 2023-01-12 code/L0/L0_Ecuador_plants_GIFT.Rmd
Inital commit with file paths, load packages and header information

---
title: "Ecuador plant species GIFT data"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "None"
data output: "TBD"
project: "Frugivoria"
date: "2023-01-12"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
remotes::install_github("https://github.com/BioGeoMacro/GIFT",
dependencies = TRUE)
library(GIFT)
library(mde)
library(DataExplorer)
library(summarytools)
```

# Get data for Ecuador from GIFT database