### 2023-01-22 code/L0/L0_Ecuador_plants_TRY.Rmd
add bdvis package & corresponding code convert TRY data to wide table format

---
title: "Ecuador plant species TRY data"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "None"
data output: "Ecuador_TRY_plant_traits.csv"
project: "Frugivoria"
date: "2022-10-27"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(rtry)
library(mde)
library(DataExplorer)
library(summarytools)
library(bdvis)
```

# Read in TRY data
Data was obtained from https://www.try-db.org/TryWeb/Prop0.php 
```{r}
data <- rtry_import("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/TRY_Ecuador_PublicData_10192020.txt", encoding = "Latin-1")
data <- as.data.frame(data)
```

# Explore TRY data
```{r}
# explore by trait id and trait name
TRYdata_explore_trait <- rtry_explore(data, TraitID, TraitName)
TRYdata_explore_trait

# explore by species id, species name, trait id, and trait name
TRYdata_explore_species <- rtry_explore(data, AccSpeciesID, AccSpeciesName, TraitID, TraitName)
TRYdata_explore_species

# explore by data id, data name, trait id, and trait name sorted by trait id
TRYdata_explore_anc <- rtry_explore(data, DataID, DataName, TraitID, TraitName, sortBy = TraitID)
TRYdata_explore_anc
```

# Convert to wide table
From https://cran.r-project.org/web/packages/rtry/vignettes/rtry-workflow-general.html
```{r}
num_traits <- rtry_select_row(data, complete.cases(TraitID) & complete.cases(StdValue), getAncillary = TRUE, rmDuplicates = TRUE)
num_traits <- rtry_select_col(num_traits, ObservationID, AccSpeciesID, AccSpeciesName, TraitID, TraitName, StdValue, UnitName)

# Extract the unique value of latitude (DataID 59) and the corresponding ObservationID
workdata_lat <- rtry_select_anc(data, 59)

# Extract the unique value of longitude (DataID 60) and the corresponding ObservationID
workdata_lon <- rtry_select_anc(data, 60)

# To merge the extracted ancillary data with the numerical traits
# Merge the relevant data frames based on the ObservationID using rtry_join_left (left join)
num_traits_georef <- rtry_join_left(num_traits, workdata_lat, baseOn = ObservationID)
num_traits_georef <- rtry_join_left(num_traits_georef, workdata_lon, baseOn = ObservationID)

TRY_data <- rtry_trans_wider(num_traits_georef, names_from = c(TraitID, TraitName, UnitName), values_from = c(StdValue), values_fn = list(StdValue = mean))
```

# Summary of data
```{r}
summary(TRY_data)
```

### using DataExplorer package
```{r}
plot_str(TRY_data)
introduce(TRY_data)
plot_intro(TRY_data)
plot_missing(TRY_data)
plot_bar(TRY_data)
```

### using summarytools package
```{r}
ecuador_TRY_plant_traits_summary <- dfSummary(TRY_data, 
            na.col       = FALSE,
            style        = "multiline",
            plain.ascii  = TRUE,
            graph.magnif = .8)
print(ecuador_TRY_plant_traits_summary, method = "render")
view(ecuador_TRY_plant_traits_summary, file = file.path(output_path,"Ecuador_TRY_summary.html"))
# gives the following error: Error in withCallingHandlers(expr, warning = function(w) if (inherits(w, : 
#   invalid multibyte string at '<b0>45''
```

```{r}
#using bdvis package
d <- format_bdvis(d, , Latitude = "Latitude", Longitude = "Longitude", Date_collected = "Date_collected", Scientific_name = "AccSpeciesName")
bdsummary(d)
d$Date_collected <- as.Date(d$Date_collected)
bdcalendarheat(d)
getcellid(d, gridscale = 10)
distrigraph(d, ptype = "species")
distrigraph(d, ptype = "efforts")
distrigraph(d, ptype = "cell")
distrigraph(d, ptype = "effortspecies")
bdcomplete(d)
mapgrid(d, ptype = "species", region = "Ecuador")
```

# Write data to csv
```{r}
ecuador_TRY_plant_traits_path <- file.path(output_path,"Ecuador_TRY_plant_traits.csv")
write.csv(data, file = ecuador_TRY_plant_traits_path)

```