### 2023-01-23 code/L0/L0_Ecuador_plants_NPP.Rmd
read in data downloaded from Neotropical plant portal add summary information and export to csv

---
title: "Ecuador plant species Neotropical Plant Portal data"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "None"
data output: "Ecuador_NPP_plant_occ.csv"
project: "Frugivoria"
date: "2023-1-23"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(DataExplorer)
library(summarytools)
library(bdvis)
```

# Read in Neotropical Plant Portal data
Data was obtained from https://serv.biokic.asu.edu/neotrop/plantae/collections/map/index.php?clid=1&cltype=vouchers&taxonfilter=&db=all&type=1&reset=1
```{r}
npp_data <- read.csv(file.path(data_path,"NeotropicalPlantPortal_SymbOutput_2022-11-28_084151_DwC-A/occurrences.csv"))
```

# Summary of data
```{r}
summary(npp_data)
```

### using DataExplorer package
```{r}
plot_str(npp_data)
introduce(npp_data)
plot_intro(npp_data)
plot_missing(npp_data)
```

### using summarytools package
```{r}
ecuador_NPP_plant_occ_summary <- dfSummary(npp_data, 
            na.col       = FALSE,
            style        = "multiline",
            plain.ascii  = TRUE,
            graph.magnif = .8)
print(ecuador_NPP_plant_occ_summary, method = "render")
view(ecuador_NPP_plant_occ_summary, file = file.path(output_path,"Ecuador_NPP_summary.html"))
```

```{r}
#using bdvis package
d <- format_bdvis(npp_data, , Latitude = "decimalLatitude", Longitude = "decimalLongitude", Date_collected = "eventDate", Scientific_name = "scientificName")
bdsummary(d)

bdcalendarheat(d)
getcellid(d, gridscale = 10)
distrigraph(d, ptype = "species")
distrigraph(d, ptype = "efforts")
distrigraph(d, ptype = "cell")
distrigraph(d, ptype = "effortspecies")
bdcomplete(d)
mapgrid(d, ptype = "species", region = "Ecuador")
```

# Write data to csv
```{r}
ecuador_npp_path <- file.path(output_path,"Ecuador_NPP_plant_occ.csv")
write.csv(npp_data, file = ecuador_npp_path)
```