### 2023-03-23 archive/L0_Ecuador_plants_WFO.Rmd
move unused code to archive

---
title: "Ecuador plant species WorldFlora"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "None"
data output: "WFO_Backbone"
project: "Frugivoria"
date: "2022-10-11"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(WorldFlora)
```

```{r}
dir.create(file.path(data_path, "WFO_Backbone"))
WFO_path <- file.path(data_path, "WFO_Backbone")
setwd(WFO_path)
WFO.download()
```