### 2023-03-23 code/L1/L1_NorthAndes_plants_IUCN.Rmd
rename from Ecuador to Northern Andes

---
title: "Ecuador plant species IUCN Cleaning"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "Ecuador_IUCN_plant_occ.csv, Ecuador_IUCN_plant_species.csv"
data output: "TBD"
project: "Frugivoria"
date: "2022-11-14"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(WorldFlora)
library(taxize)
```

# Read in IUCN occurrence data
```{r}
iucn_occ <- read.csv(file.path(data_path,"Ecuador_IUCN_plant_occ.csv"))
iucn_names <- read.csv(file.path(data_path, "Ecuador_IUCN_plant_species.csv"))
iucn_data <- merge(iucn_names, iucn_occ, by = "taxonid")
iucn_data$X.x <- NULL
iucn_data$X.y <- NULL
```

# Read in World Flora Backbone
```{r}
WFO.backbone <- read.delim(file.path(data_path,"WFO_Backbone/classification.txt"))
```

# Prepare IUCN data for analysis with WFO.match
```{r}
iucn_data.prepare <- WFO.prepare(spec.data = iucn_data, spec.full = "scientific_name", verbose = TRUE)
```

# Standarize plant names according to the World Flora Online taxonimic backbone
```{r}
iucn_data.standard <- WFO.match(spec.data = iucn_data.prepare, WFO.data = WFO.backbone, spec.name = "scientific_name", counter = TRUE)
```

# Resolve IUCN data taxonomic names with taxize
```{r}
iucn_data.resolved <- gnr_resolve(iucn_data$scientific_name)
```

# Merge resolved and standardized data
```{r}

```