### 2023-03-23 archive/L0_Ecuador_plants_POWO.Rmd

---
title: "Northern Andes plant species POWO"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "None"
data output: "NorthAndes_POWO_species_list.csv"
project: "Frugivoria"
date: "2023-02-27"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
devtools::install_github("barnabywalker/kewr")
library(kewr)
library(dplyr)
library(DataExplorer)
library(summarytools)
library(visdat)
```

# Search Plants of the World Online [https://powo.science.kew.org/] for species in Northern Andes
Best to break up request into smaller chunks so R handle it without timing out
## Ecuador
```{r}
Ecuador_POWO_page1 <- search_powo(list(distribution="Ecuador"), limit = 4000)
Ecuador_POWO_page2 <- request_next(Ecuador_POWO_page1)
Ecuador_POWO_page3 <- request_next(Ecuador_POWO_page2)
Ecuador_POWO_page4 <- request_next(Ecuador_POWO_page3)
Ecuador_POWO_page5 <- request_next(Ecuador_POWO_page4)
Ecuador_POWO_page6 <- request_next(Ecuador_POWO_page5)
```

```{r}
# Combine pages
Ecuador_POWO <- bind_rows(
  tidy(Ecuador_POWO_page1),
  tidy(Ecuador_POWO_page2),
  tidy(Ecuador_POWO_page3),
  tidy(Ecuador_POWO_page4),
  tidy(Ecuador_POWO_page5),
  tidy(Ecuador_POWO_page6)
)
```

## Colombia
```{r}
Colombia_POWO_page1 <- search_powo(list(distribution="Colombia"), limit = 4000)
Colombia_POWO_page2 <- request_next(Colombia_POWO_page1)
Colombia_POWO_page3 <- request_next(Colombia_POWO_page2)
Colombia_POWO_page4 <- request_next(Colombia_POWO_page3)
Colombia_POWO_page5 <- request_next(Colombia_POWO_page4)
Colombia_POWO_page6 <- request_next(Colombia_POWO_page5)
Colombia_POWO_page7 <- request_next(Colombia_POWO_page6)
Colombia_POWO_page8 <- request_next(Colombia_POWO_page7)
```

```{r}
# Combine pages
Colombia_POWO <- bind_rows(
  tidy(Colombia_POWO_page1),
  tidy(Colombia_POWO_page2),
  tidy(Colombia_POWO_page3),
  tidy(Colombia_POWO_page4),
  tidy(Colombia_POWO_page5),
  tidy(Colombia_POWO_page6)
)
```

## Venezuela
```{r}
Venezuela_POWO_page1 <- search_powo(list(distribution="Venezuela"), limit = 4000)
Venezuela_POWO_page2 <- request_next(Venezuela_POWO_page1)
Venezuela_POWO_page3 <- request_next(Venezuela_POWO_page2)
Venezuela_POWO_page4 <- request_next(Venezuela_POWO_page3)
Venezuela_POWO_page5 <- request_next(Venezuela_POWO_page4)
Venezuela_POWO_page6 <- request_next(Venezuela_POWO_page5)
```

```{r}
# Combine pages
Venezuela_POWO <- bind_rows(
  tidy(Venezuela_POWO_page1),
  tidy(Venezuela_POWO_page2),
  tidy(Venezuela_POWO_page3),
  tidy(Venezuela_POWO_page4),
  tidy(Venezuela_POWO_page5),
  tidy(Venezuela_POWO_page6)
)
```

## Combine Ecuador, Colombia, and Venezuela
```{r}
NorthAndes_POWO <- rbind(Ecuador_POWO,Colombia_POWO, Venezuela_POWO)
```

# Summary
```{r}
summary(NorthAndes_POWO)
```

```{r}
# using DataExplorer package
plot_str(NorthAndes_POWO)
introduce(NorthAndes_POWO)
plot_intro(NorthAndes_POWO)
plot_missing(NorthAndes_POWO)
plot_bar(NorthAndes_POWO)
```

```{r}
# using summarytools package
NorthAndes_POWO_summary <- dfSummary(NorthAndes_POWO, 
            na.col       = FALSE,
            style        = "multiline",
            plain.ascii  = TRUE,
            graph.magnif = .8)
print(NorthAndes_POWO_summary, method = "render")
view(ENorthAndes_POWO_summary, file = file.path(output_path,"Ecuador_POWO_summary.html"))
```

```{r}
#using visdat package
vis_dat(NorthAndes_POWO, warn_large_data = FALSE)
vis_miss(NorthAndes_POWO, warn_large_data = FALSE)
NorthAndes_POWO %>%
  select_if(is.numeric)%>%
  vis_cor(warn_large_data = FALSE)
```

# Write data to csv
```{r}
#convert NorthAndes_POWOO from tbl_df to dataframe
NorthAndes_POWO <- as.data.frame(NorthAndes_POWO)
# remove image column from dataframe
NorthAndes_POWO <- NorthAndes_POWO %>% select(-images)

NorthAndes_POWO_path <- file.path(output_path, "NorthAndes_POWO.csv")
write.csv(NorthAndes_POWO, file = NorthAndes_POWO)
```