### 2023-03-23 code/L0/L0_NorthAndes_plants_BIEN.Rmd

---
title: "Northern Andes plant species BIEN data"
authors: "Hazel J. Anderson"
collaborators: "None"
data input: "none"
data output: "NorthAndes_BIEN_species_list.csv, NorthAndes_BIEN_occurrence.csv, NorthAndes_BIEN_traits.csv"
project: "Frugivoria"
date: "2023-02-13"
output: html_document
---
# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(BIEN)
library(tidyr)
library(dplyr)
library(DataExplorer)
library(summarytools)
library(bdvis)
library(visdat)
```

# Get data for Ecuador from BIEN database
```{r}
# occurrence data
Ecuador_BIEN_occ <- BIEN_occurrence_country("Ecuador", all.taxonomy = TRUE, observation.type = TRUE, collection.info = TRUE, political.boundaries = TRUE, native.status = TRUE)
# species list
Ecuador_BIEN_species_list <- BIEN_list_country(country = "Ecuador")
# trait data
Ecuador_BIEN_traits <- BIEN_trait_country("Ecuador", all.taxonomy = TRUE, political.boundaries = TRUE, source.citation = TRUE)
```

```{r}
#remove duplicate data_collected column from Ecuador_BIEN_occ
Ecuador_BIEN_occ <- subset (Ecuador_BIEN_occ, select = -date_collected)
```

# Get data for Venezuela from BIEN database
```{r}
# occurrence data
Venezuela_BIEN_occ <- BIEN_occurrence_country("Venezuela", all.taxonomy = TRUE, observation.type = TRUE, collection.info = TRUE, political.boundaries = TRUE, native.status = TRUE)
# species list
Venezuela_BIEN_species_list <- BIEN_list_country(country = "Venezuela")
# trait data
Venezuela_BIEN_traits <- BIEN_trait_country("Venezuela", all.taxonomy = TRUE, political.boundaries = TRUE, source.citation = TRUE)
```

```{r}
#remove duplicate data_collected column from Ecuador_BIEN_occ
Venezuela_BIEN_occ <- subset (Venezuela_BIEN_occ, select = -date_collected)
```

# Get data for Colombia from BIEN database
```{r}
# occurrence data
Colombia_BIEN_occ <- BIEN_occurrence_country("Colombia", all.taxonomy = TRUE, observation.type = TRUE, collection.info = TRUE, political.boundaries = TRUE, native.status = TRUE)
# species list
Colombia_BIEN_species_list <- BIEN_list_country(country = "Colombia")
# trait data
Colombia_BIEN_traits <- BIEN_trait_country("Colombia", all.taxonomy = TRUE, political.boundaries = TRUE, source.citation = TRUE)
```

```{r}
#remove duplicate data_collected column from Ecuador_BIEN_occ
Colombia_BIEN_occ <- subset (Colombia_BIEN_occ, select = -date_collected)
```

# Combine BIEN data for Ecuador, Colombia, and Venezuela
```{r}
NorthAndes_BIEN_occ <- rbind(Ecuador_BIEN_occ, Venezuela_BIEN_occ, Colombia_BIEN_occ)
NorthAndes_BIEN_species_list <- rbind(Ecuador_BIEN_species_list, Venezuela_BIEN_species_list, Colombia_BIEN_species_list)
NorthAndes_BIEN_traits <- rbind(Ecuador_BIEN_traits, Venezuela_BIEN_traits, Colombia_BIEN_traits)
```

# Data summary
```{r}
glimpse(NorthAndes_BIEN_occ)
```

```{r}
glimpse(NorthAndes_BIEN_species_list)
```

```{r}
glimpse(NorthAndes_BIEN_traits)
```

### using DataExplorer package
```{r}
plot_str(NorthAndes_BIEN_occ)
introduce(NorthAndes_BIEN_occ)
plot_intro(NorthAndes_BIEN_occ)
plot_missing(NorthAndes_BIEN_occ)
plot_bar(NorthAndes_BIEN_occ)
```

```{r}
plot_str(NorthAndes_BIEN_species_list)
introduce(NorthAndes_BIEN_species_list)
plot_intro(NorthAndes_BIEN_species_list)
```

```{r}
plot_str(NorthAndes_BIEN_traits)
introduce(NorthAndes_BIEN_traits)
plot_intro(NorthAndes_BIEN_traits)
plot_missing(NorthAndes_BIEN_traits)
plot_bar(NorthAndes_BIEN_traits)
```

```{r}
# using summarytools package
NorthAndes_BIEN_occ_summary <- dfSummary(NorthAndes_BIEN_occ, 
            na.col       = FALSE,
            style        = "multiline",
            plain.ascii  = TRUE,
            graph.magnif = .8)
print(NorthAndes_BIEN_occ_summary, method = "render")
view(NorthAndes_BIEN_occ_summary, file = file.path(output_path,"NorthAndes_BIEN_occ_summary.html"))
```

```{r}
NorthAndes_BIEN_species_list_summary <- dfSummary(NorthAndes_BIEN_species_list, 
            na.col       = FALSE,
            style        = "multiline",
            plain.ascii  = TRUE,
            graph.magnif = .8)
print(NorthAndes_BIEN_species_list_summary, method = "render")
view(NorthAndes_BIEN_species_list_summary, file = file.path(output_path,"NorthAndes_BIEN_species_list_summary.html"))
```

```{r}
NorthAndes_BIEN_traits_summary <- dfSummary(NorthAndes_BIEN_traits, 
            na.col       = FALSE,
            style        = "multiline",
            plain.ascii  = TRUE,
            graph.magnif = .8)
print(NorthAndes_BIEN_traits_summary, method = "render")
view(NorthAndes_BIEN_traits_summary, file = file.path(output_path,"NorthAndes_BIEN_traits_summary.html"))
```
```{r}
#using bdvis package
NorthAndes_BIEN_occ_vis <- format_bdvis(NorthAndes_BIEN_occ, Latitude = "latitude", Longitude = "longitude", Date_collected = "date_collected", Scientific_name = "verbatim_scientific_name")
bdsummary(NorthAndes_BIEN_occ_vis)
```
```{r}
## ERROR - Error in !dbPreExists : invalid argument type
## ERROR - Error in postgresqlNewConnection(drv, ...) :
# RPosgreSQL error: could not connect postgres@localhost:5432 on dbname "test": could not connect to  server: Connection refused (0x0000274D/10061)

#bdcomplete(NorthAndes_BIEN_occ_vis)
#distrigraph(NorthAndes_BIEN_occ_vis, ptype = "species")
#distrigraph(NorthAndes_BIEN_occ_vis, ptype = "efforts")
#distrigraph(NorthAndes_BIEN_occ_vis, ptype = "cell")
#distrigraph(NorthAndes_BIEN_occ_vis, ptype = "effortspecies")
```

```{r}
#using visdat package
vis_dat(NorthAndes_BIEN_occ, warn_large_data = FALSE)
vis_miss(NorthAndes_BIEN_occ, warn_large_data = FALSE)
NorthAndes_BIEN_occ %>%
  select_if(is.numeric)%>%
  vis_cor(warn_large_data = FALSE)
```

```{r}
#using visdat package
vis_dat(NorthAndes_BIEN_traits, warn_large_data = FALSE)
vis_miss(NorthAndes_BIEN_traits, warn_large_data = FALSE)
NorthAndes_BIEN_traits %>%
  select_if(is.numeric)%>%
  vis_cor(warn_large_data = FALSE)
```

# Write data to csv
```{r}
NorthAndes_BIEN_occurrence_path <- file.path(output_path,"NorthAndes_BIEN_occurrence.csv") 
write.csv(NorthAndes_BIEN_occ, file = NorthAndes_BIEN_occurrence_path)

NorthAndes_BIEN_species_list_path <- file.path(output_path, "NorthAndes_BIEN_species_list.csv")
write.csv(NorthAndes_BIEN_species_list, file = NorthAndes_BIEN_species_list_path)

NorthAndes_BIEN_traits_path <- file.path(output_path, "NorthAndes_BIEN_traits.csv")
write.csv(NorthAndes_BIEN_traits, file = NorthAndes_BIEN_traits_path)
```