### 2023-03-23 archive/L0_Ecuador_plants_GIFT.Rmd

---
title: "Ecuador plant species GIFT data"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "None"
data output: "TBD"
project: "Frugivoria"
date: "2023-01-12"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
remotes::install_github("https://github.com/BioGeoMacro/GIFT",dependencies = TRUE)
library(GIFT)
library(mde)
library(DataExplorer)
library(summarytools)
```

# Get data for Ecuador from GIFT database
```{r}
# retrive entity_ID for Ecuador
regions <- GIFT_regions()
str(regions)
regions[regions$geo_entity %in% "Ecuador",]

```

```{r}
Ecuador_shape <- GIFT_shape(entity_ID = 10348)
```

```{r}
GIFT_data <- GIFT_checklist(floristic_group = 'all', geo_type = 'All', shp = Ecuador_shape, overlap = 'extent_intersect', taxonomic_group = TRUE)
```

```{r}
richness <- GIFT_richness(what = "species_richness")
trait_coverage <- GIFT_richness(what = "trait_coverage")
```

```{r}
richness[richness$geo_entity %in% 10348,]
trait_coverage[trait_coverage$geo_entity %in% 10348,]
```

```{r}

```