### 2023-03-23 code/L0/L0_NorthAndes_frugivores_GBIF.Rmd
inital commit

---
title: "Northern Andes frugivore species GBIF"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "None"
data output: "NorthAndes_GBIF_frugivore_occ.csv"
project: "Frugivoria"
date: "2023-03-23"
output: html_document
---
# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Install and load required packages
```{r}
library(rgbif)
library(DataExplorer)
library(bdvis)
library(visdat)
library(dplyr)
```

# Get Taxon Key for Mammals & Birds
```{r}
name_backbone("Mammalia")
name_backbone("Aves")
```

```{r}
occ_download(
pred("taxonKey", c("359", "212")), 
pred("hasGeospatialIssue", FALSE),
pred("hasCoordinate", TRUE),
pred_in("country",c("EC","CO","VE")), 
pred_not(pred_in("basisOfRecord",c("FOSSIL_SPECIMEN","LIVING_SPECIMEN"))),
format = "SIMPLE_CSV"
)
```

# Retrive GBIF download, save file, and load into r
```{r}
d <- occ_download_get("0282867-220831081235567", path = output_path) %>%
  occ_download_import()
```

# Summary 
```{r}
glimpse(d)
```

```{r}
# using DataExplorer package
plot_str(d)
introduce(d)
plot_intro(d)
plot_missing(d)
plot_bar(d)
```

```{r}
#using bdvis package
d_vis <- format_bdvis(d, Latitude = "decimalLatitude", Longitude = "decimalLongitude", Date_collected = "eventDate", Scientific_name = "scientificName", Family = "family", Genus = "genus")
bdsummary(d_vis)
```

```{r}
#using visdat package
vis_dat(d, warn_large_data = FALSE)
vis_miss(d, warn_large_data = FALSE)
d %>%
  select_if(is.numeric)%>%
  vis_cor(warn_large_data = FALSE)
```

# Write data to csv
```{r}
NorthAndes_GBIF_path <- file.path(output_path,"NorthAndes_GBIF_frugivore_occ.csv") 
write.csv(d, file = NorthAndes_GBIF_path)
```