### 2023-07-11 code/L1/CleanGBIFData.qmd
inital commit of cleaning plant gbif data

---
title: "Clean GBIF Data"
format: html
editor: visual
---

```{r}
library(countrycode)
library(CoordinateCleaner)
library(dplyr)
library(ggplot2)
library(rgbif)
library(sp)
```

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Read in Data
```{r}
TropicalAndes_GBIF_plant_occ <- read.csv(file.path(data_path,"TropAndes_GBIF_plant_occ.csv"))
TropicalAndes_GBIF_frugivore_occ <- read.csv(file.path(data_path,"TropAndes_GBIF_frugivore_occ.csv"))
```

```{r}
names(TropicalAndes_GBIF_plant_occ)
```

```{r}
dim(TropicalAndes_GBIF_plant_occ)
dim(TropicalAndes_GBIF_frugivore_occ)
```

```{r}
# remove records without coordinates
TropicalAndes_GBIF_plant_occ_cleaned <- TropicalAndes_GBIF_plant_occ%>%
  filter(!is.na(decimalLongitude))%>%
  filter(!is.na(decimalLatitude))
TropicalAndes_GBIF_frugivore_occ_cleaned <- TropicalAndes_GBIF_frugivore_occ%>%
  filter(!is.na(decimalLongitude))%>%
  filter(!is.na(decimalLatitude))
```

```{r}
dim(TropicalAndes_GBIF_plant_occ_cleaned)
dim(TropicalAndes_GBIF_frugivore_occ_cleaned)
```

```{r}
#convert country code from ISO2c to ISO3c
TropicalAndes_GBIF_plant_occ_cleaned$countryCode <-  countrycode(TropicalAndes_GBIF_plant_occ_cleaned$countryCode, origin =  'iso2c', destination = 'iso3c')
TropicalAndes_GBIF_frugivore_occ_cleaned$countryCode <-  countrycode(TropicalAndes_GBIF_frugivore_occ_cleaned$countryCode, origin =  'iso2c', destination = 'iso3c')
```

```{r}
#flag problems
TropicalAndes_GBIF_plant_occ_cleaned <- data.frame(TropicalAndes_GBIF_plant_occ_cleaned)
plant_flags <- clean_coordinates(x = TropicalAndes_GBIF_plant_occ_cleaned, 
                           lon = "decimalLongitude", 
                           lat = "decimalLatitude",
                           countries = "countryCode",
                           species = "species",
                          tests = c("capitals", "centroids", "equal","gbif", "institutions",
                                    "duplicates","zeros", "countries"))
TropicalAndes_GBIF_frugivore_occ_cleaned<- data.frame(TropicalAndes_GBIF_frugivore_occ_cleaned)
frugivore_flags <- clean_coordinates(x = TropicalAndes_GBIF_frugivore_occ_cleaned, 
                           lon = "decimalLongitude", 
                           lat = "decimalLatitude",
                           countries = "countryCode",
                           species = "species",
                          tests = c("capitals", "centroids", "equal","gbif", "institutions",
                                    "duplicates","zeros", "countries"))
```

```{r}
summary(plant_flags)
plot(plant_flags, lon = "decimalLongitude", lat = "decimalLatitude")
summary(frugivore_flags)
plot(frugivore_flags, lon = "decimalLongitude", lat = "decimalLatitude")
```

```{r}
#Exclude problematic records
TropicalAndes_GBIF_plant_occ_cleaned <- TropicalAndes_GBIF_plant_occ_cleaned[plant_flags$.summary,]
TropicalAndes_GBIF_frugivore_occ_cleaned <- TropicalAndes_GBIF_frugivore_occ_cleaned[frugivore_flags$.summary,]
```

```{r}
dim(TropicalAndes_GBIF_plant_occ_cleaned)
dim(TropicalAndes_GBIF_frugivore_occ_cleaned)
```

# Sample Bias
```{r}
library(sampbias)
library(dplyr)
```

```{r}
#remove columns
TropicalAndes_GBIF_plant_occ_cleaned <- subset(TropicalAndes_GBIF_plant_occ_cleaned, select = -c(X, decimallongitude, decimallatitude))

TropicalAndes_GBIF_frugivore_occ_cleaned <- subset(TropicalAndes_GBIF_frugivore_occ_cleaned, select = -c(X, decimallongitude, decimallatitude))
```

```{r}
bias.out.plants <- calculate_bias(x = TropicalAndes_GBIF_plant_occ_cleaned, res = 1)
```

```{r}
bias.out.frugivores <- calculate_bias(x = TropicalAndes_GBIF_frugivore_occ_cleaned, res = 1)
```