### 2023-07-11 code/L2/GBIF_PlantData_GapAnalysis.Rmd
inital commit

GBIF Plant Gap Analyis

# Load required packages
```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(smoothr)
library(purrr)
library(raster)
library(scico)
library(ggspatial)
```

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Read in Data
```{r}
TropicalAndes_GBIF_plant_occ <- read.csv(file.path(data_path,"TropAndes_GBIF_plant_occ.csv"))
```

```{r}
summary(TropicalAndes_GBIF_plant_occ)
```

```{r}
TropicalAndes_GBIF_plant_occ %>% count(family, sort = TRUE, name = "Occurrences records")
```

```{r}
occ_by_country <- TropicalAndes_GBIF_plant_occ %>% count(countryCode, sort = TRUE, name = "occurrencesRecords") 
occ_by_country
occ_by_country$occurrencesRecords <- as.numeric(occ_by_country$occurrencesRecords)
occ_by_country
tbl <- with(occ_by_country, table(occ_by_country$countryCode, occ_by_country$occurrencesRecords))
barplot(tbl, beside = TRUE, legend = TRUE)
```

```{r}
TropicalAndes_GBIF_plant_occ %>% count(species, sort = TRUE, name = "Occurrences records")
```

```{r}
spc <- tapply(TropicalAndes_GBIF_plant_occ$species, TropicalAndes_GBIF_plant_occ$countryCode, function(x)length(unique(x)) )
as.data.frame(spc)
```

# Sample Bias
```{r}
library(sampbias)
```

```{r}
bias.out.plants <- calculate_bias(x = TropicalAndes_GBIF_plant_occ, res = 1)
```

```{r}
bias.out.frugivores <- calculate_bias(x = TropicalAndes_GBIF_frugivore_occ, res = 1)
```

```{r}
summary(bias.out.plants)
```

```{r}
summary(bias.out.frugivores)
```

```{r}
plot(bias.out.plants)
proj.plants <- project_bias(bias.out.plants)
map_bias(proj.plants, type = "log_sampling_rate")
map_bias(proj.plants, type = "sampling_rate")
map_bias(proj.plants, type = "diff_to_max")
```

```{r}
plot(bias.out.frugivores)
proj.frugivore <- project_bias(bias.out.frugivores)
map_bias(proj.frugivore, type = "log_sampling_rate")
map_bias(proj.frugivore, type = "sampling_rate")
map_bias(proj.frugivore, type = "diff_to_max")
```

# Number of Occurences per grid cell
Based on https://azizka.github.io/Biodiversity_Data_from_Field_to_Yield/04_species_richness_maps_tutorial.html

```{r}
library(readr)
library(ggplot2)
library(magrittr)
library(raster)
library(rgdal)
devtools::install_github("azizka/speciesgeocodeR")
library(speciesgeocodeR)
options(scipen=10000)
```

```{r}
# Visualize
world.inp <- map_data("world")

# The number of occurrences per grid cell
gocc <- RichnessGrid(TropicalAndes_GBIF_plant_occ, type = "abu")

plot(gocc)

# Or more sofisticated plotting Prepare for plotting
plo_gocc <- data.frame(rasterToPoints(gocc))

# Visualize
ggplot() + 
  geom_map(data = world.inp, map = world.inp,
           aes(x = long, y = lat, map_id = region), fill = "grey80") +
  xlim(min(TropicalAndes_GBIF_plant_occ$decimalLongitude, na.rm = T),
    max(TropicalAndes_GBIF_plant_occ$decimalLongitude, na.rm = T)) +
  ylim(min(TropicalAndes_GBIF_plant_occ$decimalLatitude, na.rm = T), 
    max(TropicalAndes_GBIF_plant_occ$decimalLatitude, na.rm = T)) + 
  geom_raster(data = plo_gocc, aes(x = x, y = y, fill = layer)) +
  scale_fill_viridis_c(name="Count") +
  coord_fixed() +
  theme_bw() +
  ggtitle("Number of GBIF Plant occurrence records") +
  ylab("Latitude") +
  xlab("Longitude")

# The number of species per grid cell
gri <- RichnessGrid(TropicalAndes_GBIF_plant_occ)

plot(gri)
```

```{r}
# Visualize
world.inp <- map_data("world")

# The number of occurrences per grid cell
gocc <- RichnessGrid(TropicalAndes_GBIF_frugivore_occ, type = "abu")

plot(gocc)

# Or more sofisticated plotting Prepare for plotting
plo_gocc <- data.frame(rasterToPoints(gocc))

# The number of species per grid cell
gri <- RichnessGrid(TropicalAndes_GBIF_frugivore_occ)

plot(gri)
```

```{r}
ggplot() +
  geom_map(data = world.inp, map = world.inp,
           aes(x = long, y = lat, map_id = region), fill = "grey80") +
  xlim(min(TropicalAndes_GBIF_frugivore_occ$decimalLongitude, na.rm = T), 
    max(TropicalAndes_GBIF_frugivore_occ$decimalLongitude, na.rm = T)) +
  ylim(min(TropicalAndes_GBIF_frugivore_occ$decimalLatitude, na.rm = T), 
    max(TropicalAndes_GBIF_frugivore_occ$decimalLatitude, na.rm = T)) +
  geom_raster(data = plo_gocc, aes(x = x, y = y, fill = layer)) +
  scale_fill_viridis_c(name="Count") +
  coord_fixed() +
  theme_bw() +
  ggtitle("Number of GBIF Bird & Mammal occurrence records") +
  ylab("Latitude") +
  xlab("Longitude")
```