### 2023-07-11 code/L2/EvaluatingIUCNAssessedTropicalAndesPlantSpecies.Rmd
inital commit

Evaluating Assessed Tropical Andes plant species from the IUCN and POWO

Modified from https://barnabywalker.github.io/kewr/articles/mapping-assessed-species.html

# Load required packages
```{r}
library(kewr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
```

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# 1. Getting assessment information

```{r}
# Load in data
data <- read.csv(file.path(data_path,"IUCN/Tropical Andes Plants - Search Results/simple_summary.csv"))
```

```{r}
# Subset to just species names
TropicalAndes_plants <- data[ , c('internalTaxonId', 'scientificName', 'authority', 'redlistCategory')]
```

# 2. Match names to an IPNI ID
For large data sets use http://namematch.science.kew.org/csv
```{r}
TropicalAndes_plants <-
  TropicalAndes_plants %>%
  unite("full_name", scientificName, authority, sep=" ", remove=FALSE)
```

```{r}
full_matches <- match_knms(TropicalAndes_plants$full_name)
full_matches
```

```{r}
full_matches <- tidy(full_matches)

unmatched <- filter(full_matches, !matched)

to_match <- 
  TropicalAndes_plants %>%
  filter(full_name %in% unmatched$submitted)

part_matches <- match_knms(to_match$scientificName)
part_matches
```

```{r}
part_matches <- tidy(part_matches)

full_matches <- 
  full_matches %>%
  filter(matched) %>%
  left_join(
    TropicalAndes_plants %>% dplyr::select(internalTaxonId, full_name, redlistCategory),
    by=c("submitted"="full_name")
  )

part_matches <- 
  part_matches %>%
  filter(matched) %>%
  left_join(
    TropicalAndes_plants %>% dplyr::select(internalTaxonId, scientificName, redlistCategory),
    by=c("submitted"="scientificName")
  )

matched_names <- bind_rows(full_matches, part_matches)

head(matched_names)
```

# 3. Resolve assessments to accepted names
```{r}
# wrap up the lookup_wcvp function to make sure it comes back as a list
f <- function(internalTaxonId) {
  list(lookup_wcvp(internalTaxonId))
}

resolved_names <- 
  matched_names %>%
  dplyr::nest_by(internalTaxonId, ipni_id) %>%
  dplyr::mutate(matched_record=f(ipni_id))

resolved_names <-
  resolved_names %>%
  dplyr::mutate(status=matched_record$status)
```

```{r}
resolved_names <-
  resolved_names %>%
  filter(status %in% c("accepted", "homotypic synonym"))
```

```{r}
resolved_names %>%
  ungroup() %>%
  add_count(internalTaxonid) %>%
  summarise(multiple_matches=sum(n > 1))
```

```{r}
resolved_names <-
  resolved_names %>%
  mutate(accepted_id=ifelse(status != "accepted",
                              wcvp_record$accepted$id,
                              wcvp_record$id),
         accepted_name=ifelse(status != "accepted",
                              wcvp_record$accepted$name,
                              wcvp_record$name),
         accepted_author=ifelse(status != "accepted",
                              wcvp_record$accepted$author,
                              wcvp_record$authors),
         accepted_rank=ifelse(status != "accepted",
                              wcvp_record$accepted$rank,
                              wcvp_record$rank))

resolved_names <- 
  resolved_names %>%
  select(-wcvp_record) %>%
  unnest(cols=c(data)) %>%
  rename(match_id=ipni_id)

accepted_species <-
  resolved_names %>%
  filter(accepted_rank == "Species") %>%
  ungroup()

nrow(accepted_species)
```

# 4. Get a list of all species in the country
```{r}
Ecuador_checklist <- search_powo(list(distribution= "Ecuador"),
                         filters=c("accepted", "species"),
                        limit=5000)
Ecuador_checklist
Venezuela_checklist <- search_powo(list(distribution= "Venezuela"),
                         filters=c("accepted", "species"),
                        limit=5000)
Venezuela_checklist

Colombia_checklist <- search_powo(list(distribution= "Colombia"),
                         filters=c("accepted", "species"),
                        limit=5000)
Colombia_checklist

Peru_checklist <- search_powo(list(distribution= "Peru"),
                         filters=c("accepted", "species"),
                        limit=5000)
Peru_checklist

Bolivia_checklist <- search_powo(list(distribution= "Bolivia"),
                         filters=c("accepted", "species"),
                        limit=5000)
Bolivia_checklist
```
```{r}
Bolivia_checklist <- tidy(Bolivia_checklist)
Colombia_checklist <- tidy(Colombia_checklist)
Ecuador_checklist <- tidy(Ecuador_checklist)
Peru_checklist <- tidy(Peru_checklist)
Venezuela_checklist <- tidy(Venezuela_checklist)
```

```{r}
checklist <- rbind(Bolivia_checklist, Colombia_checklist, Ecuador_checklist, Peru_checklist, Venezuela_checklist)
```

```{r}
checklist <-
  checklist %>%
  mutate(ipni_id=str_extract(fqId, "[0-9\\-]+")) %>%
  select(ipni_id, family, name, author) %>%
  distinct(ipni_id, .keep_all=TRUE) %>%
  left_join(
    accepted_species %>% select(accepted_id, category),
    by=c("ipni_id"="accepted_id")
  )
```

# 5. Calculating the proportion of assessed species
```{r}
checklist %>%
  summarise(p_assessed=mean(! is.na(category)))
```

```{r}
iucn_colours <- c("NE"="#ffffff", 
                  "DD"="#d1d1d6", 
                  "LC"="#60c659", 
                  "NT"="#cce226", 
                  "VU"="#f9e814", 
                  "EN"="#fc7f3f", 
                  "CR"="d81e05", 
                  "EW"="#542344", 
                  "EX"="#000000")
checklist %>%
  replace_na(list(category="NE")) %>%
  mutate(category=factor(category, levels=names(iucn_colours),
                         ordered=TRUE)) %>%
  ggplot(mapping=aes(y=category, fill=category,
                     colour=category == "NE")) +
  geom_bar() +
  scale_fill_manual(values=iucn_colours, drop=FALSE) +
  scale_colour_manual(values=c(`TRUE`="black", `FALSE`=NA)) +
  scale_y_discrete(drop=FALSE) +
  guides(fill=FALSE, colour=FALSE) +
  labs(x="Number of species", y="IUCN Red List category")
```