### 2023-07-11 archive/L2_NorthAndes_plant_occ_option2.Rmd
move old code to archive

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(bdDwC)
```

# Load data
```{r}
gbif_data <- read.csv(file.path(data_path,"Ecuador_GBIF_occurences.csv"))
iucn_data <- read.csv(file.path(data_path, "Ecuador_IUCN_plant_occ.csv"))
bien_data <- read.csv(file.path(data_path, "Ecuador_BIEN_occurrence.csv"))
```

```{r}
# get newest version of Darwin Cloud Data
downloadCloudData()
```

```{r}
# darwinize GBIF dataframe
gbif_result <- darwinizeNames(dataUser = gbif_data,
                            dataDWC   = bdDwC:::dataDarwinCloud$data)
new_gbif_data <- renameUserData(gbif_data, gbif_result)
```

```{r}
# darwinize IUCN dataframe
iucn_result <- darwinizeNames(dataUser = iucn_data,
                            dataDWC   = bdDwC:::dataDarwinCloud$data)
new_iucn_data <- renameUserData(iucn_data, iucn_result)
```

```{r}
# darwinize BIEN dataframe
bien_result <- darwinizeNames(dataUser = bien_data,
                            dataDWC   = bdDwC:::dataDarwinCloud$data)
new_bien_data <- renameUserData(bien_data, bien_result)
```

# Write data to csv
```{r}
# renamed GBIF data
ecuador_GBIF_path <- file.path(output_path,"Ecuador_GBIF_plant_occ.csv")
write.csv(new_gbif_data, file = ecuador_GBIF_path)
# GBIF configuration table
ecuador_GBIF_table_path <- file.path(output_path,"Ecuador_plant_GBIF_ConfigurationTable.csv")
write.csv(gbif_result, file = ecuador_GBIF_table_path)
```

```{r}
# renamed IUCN data
ecuador_IUCN_path <- file.path(output_path,"Ecuador_IUCN_plant_occ.csv")
write.csv(new_iucn_data, file = ecuador_IUCN_path)
# IUCN configuration table
ecuador_IUCN_table_path <- file.path(output_path,"Ecuador_plant_IUCN_ConfigurationTable.csv")
write.csv(iucn_result, file = ecuador_IUCN_table_path)
```

```{r}
# renamed BIEN data
ecuador_BIEN_path <- file.path(output_path,"Ecuador_BIEN_plant_occ.csv")
write.csv(new_bien_data, file = ecuador_BIEN_path)

```