### 2023-07-13 archive/L0_NorthAndes_plants_IUCN.Rmd

---
title: "Northern Andes plant species IUCN data"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "None"
data output: "NorthAndes_IUCN_plant_species.csv, NorthAndes_IUCN_plant_occ.csv"
project: "Frugivoria"
date: "2023-02-27"
output: html_document
---
# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Install and load required packages
```{r}
library(dplyr)
library(mde)
library(DataExplorer)
library(summarytools)
```

# Read in IUCN data
Data was obtained from https://www.iucnredlist.org/search?dl=true&permalink=a44f4942-2b56-47b9-8122-9896ac4af647
```{r}
assessments <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/assessments.csv"))
common_names <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/common_names.csv"))
conservation_needed <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/conservation_needed.csv"))
countries <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/countries.csv"))
credits <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/credits.csv"))
dois <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/dois.csv"))
fao <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/fao.csv"))
habitats <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/habitats.csv"))
plant_specific <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/plant_specific.csv"))
references <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/references.csv"))
research_needed <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/research_needed.csv"))
simple_summary <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/simple_summary.csv"))
synonyms <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/synonyms.csv"))
taxonomy <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/taxonomy.csv"))
threats <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/threats.csv"))
usetrade <- read.csv(file.path(data_path,"IUCN/Northern Andes 1 - Search Results/usetrade.csv"))
points_data <- read.csv(file.path(data_path,"IUCN/Northern Andes 3 - Range data Points/points_data.csv"))
```

the different tables joined using the "internalTaxonId" when viewing taxon information, or the "assessmentID" field to see assessment information, especially in cases where downloads include multiple assesments for a taxon (e.g. regional and global assessments).
```{r}

```

# Summary of data
```{r}
summary(northAndes_plants)
na_summary(northAndes_plants, exclude_cols = c("subpopulation"))
```

```{r}
summary(northAndes_IUCN_plant_occ)
na_summary(northAndes_IUCN_plant_occ)
```

```{r}
summary(northAndes_IUCN_plant_habitat)
na_summary(northAndes_IUCN_plant_habitat)
```

### using DataExplorer package
```{r}
plot_str(northAndes_plants)
introduce(northAndes_plants)
plot_intro(northAndes_plants)
plot_missing(northAndes_plants)
plot_bar(northAndes_plants)
```

```{r}
plot_str(northAndes_IUCN_plant_occ)
introduce(northAndes_IUCN_plant_occ)
plot_intro(northAndes_IUCN_plant_occ)
plot_missing(northAndes_IUCN_plant_occ)
plot_bar(northAndes_IUCN_plant_occ)
```

```{r}
plot_str(northAndes_IUCN_plant_habitat)
introduce(northAndes_IUCN_plant_habitat)
plot_intro(northAndes_IUCN_plant_habitat)
plot_missing(northAndes_IUCN_plant_habitat)
plot_bar(northAndes_IUCN_plant_habitat)
```

### using summarytools package
```{r}
northAndes_plants_summary <- dfSummary(northAndes_plants, 
            na.col       = FALSE,
            style        = "multiline",
            plain.ascii  = TRUE,
            graph.magnif = .8)
print(northAndes_plants_summary, method = "render")
view(northAndes_plants_summary, file = file.path(output_path,"northAndes_plants_summary.html"))
```

```{r}
northAndes_IUCN_plant_occ_summary <- dfSummary(northAndes_IUCN_plant_occ, 
            na.col       = FALSE,
            style        = "multiline",
            plain.ascii  = TRUE,
            graph.magnif = .8)
print(northAndes_IUCN_plant_occ_summary, method = "render")
view(northAndes_IUCN_plant_occ_summary, file = file.path(output_path,"northAndes_IUCN_plant_occ_summary.html"))
```

```{r}
northAndes_IUCN_plant_habitat_summary <- dfSummary(northAndes_IUCN_plant_habitat, 
            na.col       = FALSE,
            style        = "multiline",
            plain.ascii  = TRUE,
            graph.magnif = .8)
print(northAndes_IUCN_plant_habitat_summary, method = "render")
view(northAndes_IUCN_plant_habitat_summary, file = file.path(output_path,"northAndes_IUCN_plant_habitat_summary.html"))
```

# Write data to csv
```{r}
northAndes_ICUN_species_list_path <- file.path(output_path,"northAndes_IUCN_plant_species.csv")
write.csv(northAndes_plants, file = northAndes_ICUN_species_list_path)

northAndes_IUCN_plant_occ_path <- file.path(output_path, "northAndes_IUCN_plant_occ.csv")
write.csv(northAndes_IUCN_plant_occ, file = northAndes_IUCN_plant_occ_path)

northAndes_IUCN_plant_habitat_path <- file.path(output_path, "northAndes_IUCN_plant_habitat.csv")
write.csv(northAndes_IUCN_plant_habitat, file = northAndes_IUCN_plant_habitat_path)
```