### 2023-07-18 code/L0/L0_TropicalAndes_plants_BIEN_traits.Rmd
add summary, fill in overview, add title, save output to csv

---
title: "Tropical Andes plant trait data BIEN"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script retrieves plant trait data from the BIEN database for plant species in the countries that make up the Tropical Andes."
data input: "none"
data output: "TropicalAndes_BIEN_traits.csv"
date: "2023-07-18"
output: html_document
---

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(BIEN)
library(tidyr)
library(dplyr)
```

# List of countries of interest
```{r}
country_vector <- c("Ecuador", "Venezuela", "Peru", "Colombia", "Bolivia")
```

```{r}
TropicalAndes_BIEN_traits <- BIEN_trait_country(country_vector, all.taxonomy = TRUE, political.boundaries = TRUE, source.citation = TRUE)
```

# Data summary
```{r}
glimpse(TropicalAndes_BIEN_traits)
```

```{r}
print("The number of records is")
nrow(TropicalAndes_BIEN_traits)
print("The number of species is")
length(unique(TropicalAndes_BIEN_traits$scrubbed_species_binomial))
print("The number of families is")
length(unique(TropicalAndes_BIEN_traits$family_matched))
```

# Write data to csv
```{r}
TropicalAndes_BIEN_traits_path <- file.path(output_path,"TropicalAndes_BIEN_plant_traits.csv") 
write.csv(TropicalAndes_BIEN_traits, file = TropicalAndes_BIEN_traits_path)
```