### 2023-07-18 code/L0/L0_TropicalAndes_frugivores_frugivoria_traits.Rmd
inital committ

---
title: "Tropical Andes frugivore trait data Frugivoria"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: ""
data input: ""
data output: ""
date: "2023-07-18"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}

```

```{r}

```

# Summary 
```{r}
glimpse(d)
```

```{r}
print("The number of occurrences is")
nrow(d)
print("The number of species is" )
length(unique(d$species))
print("The number of families is" )
length(unique(d$family))
```

# Write data to csv
```{r}
TropicalAndes_Frugivoria_path <- file.path(output_path,"TropicalAndes_Frugivoria_frugivore_traits.csv") 
write.csv(d2, file = TropicalAndes_Frugivoria_path)
```