### 2023-07-18 code/L0/L0_TropicalAndes_plants_TRY_traits.Rmd
inital commit

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}

```

# Load TRY data
Data was obtained from https://www.try-db.org/TryWeb/Prop0.php 
```{r}
try_data <- rtry_import("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/23650_24112022083128/23650.txt", encoding = "Latin-1")
try_data <- as.data.frame(try_data)
```

```{r}

```