### 2023-07-24 code/L0/L0_TropicalAndes_plants_TRY_traits.Rmd
subset by tropical andes georefferenced records species list code save as csv

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(rtry)
```

# Load TRY data
Data was obtained from https://www.try-db.org/TryWeb/Prop0.php 
```{r}
try_data <- rtry_importujkgh("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY/25496_17032023010024/25496.txt")
```

#Subset to Tropical Andes
```{r}
head(try_data)
```

# Summary 
```{r}
glimpse(try_data)
```

```{r}
print("The number of records is")
nrow(try_data)
print("The number of species is" )
length(unique(try_data$AccSpeciesName))
```

# Add Lat & Long as columns in database

```{r}
# Extract the unique value of latitude (DataID 59) and the corresponding ObservationID
workdata_lat <- rtry_select_anc(try_data, 59)

# Extract the unique value of longitude (DataID 60) and the corresponding ObservationID
workdata_lon <- rtry_select_anc(try_data, 60)
```

```{r}
# To merge the extracted ancillary data with the numerical traits
# Merge the relevant data frames based on the ObservationID using rtry_join_left (left join)
try_data_georef <- rtry_join_left(try_data, workdata_lat, baseOn = ObservationID)
try_data_georef <- rtry_join_left(try_data_georef, workdata_lon, baseOn = ObservationID)
glimpse(try_data_georef)
```
```{r}
# remove rows with na in latitude
try_data_georef_noNA <- try_data_georef[!is.na(try_data_georef$Latitude),]
glimpse(try_data_georef_noNA)
```

```{r}
#subset by bounding box of Tropical Andes
try_data_georef_lat <- subset(try_data_georef_noNA, Latitude< 12.5 & Latitude> -22.8)
glimpse(try_data_georef_lat)
try_data_georef_lon <- subset(try_data_georef_lat, Longitude> -96.6 & Longitude< -57.5)
glimpse(try_data_georef_lon)
```

```{r}
#remove unnecessary column V29
try_data_TropicalAndes <- subset(try_data_georef_lon, select =-V29)
```

# Extract species list
```{r}
TRY_SpeciesList <- unique(try_data_TropicalAndes$AccSpeciesName)
```

# Summary

```{r}
glimpse(try_data_TropicalAndes)
```

```{r}
print("The number of records is")
nrow(try_data_TropicalAndes)
print("The number of species is" )
length(unique(try_data_TropicalAndes$AccSpeciesName))
```

# Write data to csv
```{r}
TropicalAndes_TRY_path <- file.path(output_path,"TropicalAndes_TRY_plant_traits.csv") 
write.csv(try_data_TropicalAndes, file = TropicalAndes_TRY_path)
TropicalAndes_TRY_species_path <- file.path(output_path,"TropicalAndes_TRY_plant_species.csv") 
write.csv(TRY_SpeciesList, file = TropicalAndes_TRY_species_path)
```