### 2023-07-25 archive/L0_TropicalAndes_plants_BIEN_ranges.Rmd
move to archive

---
title: "Tropical Andes plant species BIEN Ranges"
authors: "Hazel J. Anderson"
collaborators: "None"
data input: "none"
data output: "TBD"
project: "Frugivoria"
date: "2023-02-13"
output: html_document
---
# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(BIEN)
library(tidyr)
library(dplyr)
library(rnaturalearth)
```

# Polygons of countries
```{r}
worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")
#polygon of Tropical Andes
TApoly <- worldMap %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
```

# Get data for Tropical Andes from BIEN database
```{r}
# range map data
TropicalAndes_BIEN_ranges <- BIEN_ranges_box(min.lat = -24, max.lat = 14, min.long = -94, max.long = -54, directory = output_path, species.names.only = TRUE)

#BIEN_ranges_sf(TApoly, directory = output_path, return.species.list = TRUE, crop.ranges = TRUE)

```

# Data summary
```{r}
glimpse(TropicalAndes_BIEN_occ)
```