### 2023-07-25 archive/MappingTopicalAndes.Rmd
move old code to archive

Map of Tropical Andes

```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(ggspatial)
```

```{r}
worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")
# country subset. 
ECpoly <- worldMap %>% filter(sovereignt == "Ecuador")
VEpoly <- worldMap %>% filter(sovereignt == "Venezuela")
COpoly <- worldMap %>% filter(sovereignt == "Colombia")
PEpoly <- worldMap %>% filter(sovereignt == "Peru")
BOpoly <- worldMap %>% filter(sovereignt == "Bolivia")
#polygon of Tropical Andes
TApoly <- worldMap %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
# set color for Tropical Andes
Tropical_Andes <- "lightgreen"
```

```{r}
MapPlot <-
ggplot() +
geom_sf(data = worldMap, fill = "White") +
geom_sf(data = ECpoly, fill = Tropical_Andes) +
geom_sf(data = VEpoly, fill = Tropical_Andes) +
geom_sf(data = COpoly, fill = Tropical_Andes) +
geom_sf(data = PEpoly, fill = Tropical_Andes) +
geom_sf(data = BOpoly, fill = Tropical_Andes) +
labs(title = "Tropical Andes", x = "Latitude", y = "Longitude") +
coord_sf(xlim = c(-94, -54), ylim = c(-24, 14), expand = FALSE) +
annotation_scale(location = "bl", width_hint = 0.5) +
annotation_north_arrow(location = "bl", which_north = "true", 
  pad_x = unit(0.75, "in"), pad_y = unit(0.5, "in"), style = north_arrow_fancy_orienteering) +
theme(panel.background = element_rect(fill = "lightblue"))

MapPlot
```

```{r}
ggsave(filename = "Tropical Andes Map.png", dpi = 300, device="png")
```