### 2023-07-27 code/L1/L1_TropicalAndes_plant_lookupTable.Rmd
inital commit

---
title: "Tropical Andes plant Lookup Table"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: ""
data input: ""
data output: ""
date: "2023-07-25"
output: html_document
---

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}

```

# Read in species lists
```{r}
TRY_species <- read.csv(file.path(data_path,"TropicalAndes_TRY_plant_species.csv"))
BIEN_species <- read.csv(file.path(data_path,"TropicalAndes_BIEN_plant_trait_species.csv"))
```

```{r}
#convert species list into vectors
BIEN_species <- BIEN_species[,2]
TRY_species <- TRY_species[,2]
#what is in both
species <- intersect(TRY_species, BIEN_species)
# what doesn't match
all.equal(TRY_species, BIEN_species)

```