### 2023-07-27 code/L1/L1_TropicalAndes_plants_BIEN_occ.Rmd
inital commit

---
title: "Cleaning Tropical Andes plant BIEN occurrence records"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: ""
data input: "TropicalAndes_GBIF_plant_occ.csv"
data output: "TropicalAndes_GBIF_plant_occ_cleaned.csv"
date: "2023-07-25"
output: html_document
---

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(countrycode)
library(CoordinateCleaner)
library(dplyr)
```

# Read in data
```{r}
TropicalAndes_BIEN_plant_occ <- read.csv(file.path(data_path,"TropicalAndes_BIEN_plant_occ.csv"))
```

# Clean data
Adapted code from https://ropensci.github.io/CoordinateCleaner/articles/Cleaning_GBIF_data_with_CoordinateCleaner.html

## Select columns of interest
```{r}
# Get column names
names(TropicalAndes_BIEN_plant_occ)
```

```{r}
# Subset to columns of interest
TropicalAndes_BIEN_plant_occ <- TropicalAndes_BIEN_plant_occ %>%
  dplyr::select()
# remove records without coordinates
TropicalAndes_BIEN_plant_occ <- TropicalAndes_BIEN_plant_occ %>%
  filter(!is.na(decimalLongitude)) %>%
  filter(!is.na(decimalLatitude))
```

```{r}
 #to avoid specifying it in each function
names(dat)[2:3] <- c("decimallongitude", "decimallatitude")

TropicalAndes_BIEN_plant_occ_clean <- TropicalAndes_BIEN_plant_occ%>%
  cc_val()%>%
  cc_equ()%>%
  cc_cap()%>%
  cc_cen()%>%
  cc_coun(iso3 = "countryCode")%>%
  cc_gbif()%>%
  cc_inst()%>%
  cc_sea()%>%
  cc_zero()%>%
  cc_outl()%>%
  cc_dupl()
```