### 2023-08-02 code/L0/L0_TropicalAndes_plants_BIEN_occ.Rmd
load forest shape unsuccessfully get occurrence data from BIEN

---
title: "Tropical Andes plant species BIEN data"
authors: "Hazel J. Anderson"
collaborators: "None"
data input: "none"
data output: "TropicalAndes_BIEN_species_list.csv, TropicalAndes_BIEN_occurrence.csv"
project: "Frugivoria"
date: "2023-02-13"
output: html_document
---
# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(BIEN)
library(tidyr)
library(dplyr)
library(sf)
```

# Load Tropical Andes moist lowland and montane forest shape
```{r}
TropicalAndes_IUCNHabitat_Forest <- st_read(file.path(data_path, "TropicalAndes_IUCNHabitat_Forest.shp"))
```

# Get occurrence data for Tropical Andes from BIEN database
```{r}
TropicalAndes_BIEN_occ <- BIEN_occurrence_sf(TropicalAndes_IUCNHabitat_Forest, all.taxonomy = TRUE, observation.type = TRUE, political.boundaries = TRUE, collection.info = TRUE)
```

# Data summary
```{r}
glimpse(TropicalAndes_BIEN_occ)
```

```{r}
print("The number of records is")
nrow(TropicalAndes_BIEN_occ)
print("The number of species is")
length(unique(TropicalAndes_BIEN_occ$scrubbed_species_binomial))
print("The number of families is")
length(unique(TropicalAndes_BIEN_occ$family_matched))
```

# Extract species list
```{r}
BIEN_occ_SpeciesList <- unique(TropicalAndes_BIEN_occ$scrubbed_species_binomial)
```

# Write data to csv
```{r}
TropicalAndes_BIEN_occurrence_path <- file.path(output_path,"TropicalAndes_BIEN_plant_occ.csv") 
write.csv(TropicalAndes_BIEN_occ, file = TropicalAndes_BIEN_occurrence_path)

TropicalAndes_BIEN_species_list_path <- file.path(output_path, "TropicalAndes_BIEN_plant_occ_species.csv")
write.csv(TropicalAndes_BIEN_species_list, file = TropicalAndes_BIEN_species_list_path)
```