### 2023-08-03 code/L1/L1_TropicalAndes_subset_occ_forest.Rmd
fix code converting to SpatialPointsDataframe change plant_occ name fix issue with removing NAs

---
title: "Subset occurrence data to IUCN Tropical Andes Lowland and Montane Forest"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script retrives subsets plant and frugivore occurrence data to IUCN Tropical Andes Lowland and Montane Forest."
data input: "TropicalAndes_IUCNHabitat_Forest.tif, TropicalAndes_GBIF_plant_occ_cleaned.csv, TropicalAndes_GBIF_frugivore_occ_cleaned.csv"
data output: "TropicalAndes_plant_occ_forest.csv, TropicalAndes_frugivore_occ_forest.csv, TropicalAndes_plant_Forest_SpeciesList.csv, TropicalAndes_frugivore_Forest_SpeciesList.csv"
date: "2023-07-18"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
data_path2<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(dplyr)
library(raster)
library(sf)
```

# Read in data
```{r}
TropicalAndes_IUCNHabitat_Forest <- raster(file.path(data_path2, "TropicalAndes_IUCNHabitat_Forest.tif"))
#TropicalAndes_plant_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_plant_occ_cleaned.csv"))
TropicalAndes_frugivore_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_frugivore_occ_cleaned.csv"))
```

```{r}
TropicalAndes_IUCNHabitat_Forest
```

```{r}
TropicalAndes_IUCNHabitat_Forest0.3 <- aggregate(TropicalAndes_IUCNHabitat_Forest, 30, fun=mean, na.rm=TRUE)
TropicalAndes_IUCNHabitat_Forest0.3
```

# Rasterize occurrence data
```{r}
# convert to SpatialPointsDataFrame
coordinates(TropicalAndes_plant_occ) <- ~ decimalLongitude + decimalLatitude
coordinates(TropicalAndes_frugivore_occ) <- ~ decimalLongitude + decimalLatitude
proj4string(TropicalAndes_plant_occ) <- CRS("+proj=longlat +datum=WGS84 +no_defs")
proj4string(TropicalAndes_frugivore_occ) <- CRS("+proj=longlat +datum=WGS84 +no_defs")
TropicalAndes_plant_occ
TropicalAndes_frugivore_occ
```

# Crop occurrence data to forest extent
```{r}
TropicalAndes_plant_occ$ref<- extract(TropicalAndes_IUCNHabitat_Forest, TropicalAndes_plant_occ)
TropicalAndes_frugivore_occ$ref<- extract(TropicalAndes_IUCNHabitat_Forest, TropicalAndes_frugivore_occ)
```

```{r}
TropicalAndes_plant_occ$ref <- as.character(TropicalAndes_plant_occ$ref)
TropicalAndes_frugivore_occ$ref <- as.character(TropicalAndes_frugivore_occ$ref)
```

```{r}
plant_occ_forest <- TropicalAndes_plant_occ@data
frugivore_occ_forest <- TropicalAndes_frugivore_occ@data
```

```{r}
#remove row if ref = NA
plant_occ_forest <- plant_occ_forest[!is.na(plant_occ_forest$ref), ]
frugivore_occ_forest <- frugivore_occ_forest[!is.na(frugivore_occ_forest$ref), ]
```

# Summary
```{r}
glimpse(plant_occ_forest)
glimpse(frugivore_occ_forest)
```

```{r}
print("The number of plant records is")
nrow(plant_occ_forest)
print("The number of plant species is")
length(unique(plant_occ_forest$species))
print("The number of plant families is")
length(unique(plant_occ_forest$family))
```

```{r}
print("The number of frugivore records is")
nrow(frugivore_occ_forest)
print("The number of frugivore species is")
length(unique(frugivore_occ_forest$species))
print("The number of frugivore families is")
length(unique(frugivore_occ_forest$family))
```

# Extract species list for plants and frugivore
```{r}
TropicalAndes_plant_Forest_SpeciesList <- unique(plant_occ_forest$species)
TropicalAndes_frugivore_Forest_SpeciesList <- unique(frugivore_occ_forest$species)
```

# Write data to csv
```{r}
# subset plant occ, subset frugivore occ, plant species list, frugivore species list
write.csv(plant_occ_forest, file = file.path(output_path,"TropicalAndes_plant_occ_forest.csv"))
write.csv(frugivore_occ_forest, file = file.path(output_path,"TropicalAndes_frugivore_occ_forest.csv"))
write.csv(TropicalAndes_plant_SpeciesList, file = file.path(output_path,"TropicalAndes_plant_Forest_SpeciesList.csv"))
write.csv(TropicalAndes_frugivore_Forest_SpeciesList, file = file.path(output_path,"TropicalAndes_frugivore_Forest_SpeciesList"))
```