### 2023-08-03 code/L2/L2_1_TropicalAndes_plant_traits.Rmd
inital commit

---
title: "Tropical Andes plant traits"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "Combinging TRY and BIEN plant trait data"
data input: ""
data output: ""
date: "2023-08-03"
output: html_document
---

# Load required packages
```{r}
library(dplyr)
library(tidyr)
```

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Read in Data
```{r}
BIEN_traits <- read.csv(file.path(data_path,"TropicalAndes_BIEN_plant_traits_forest.csv"))
TRY_traits <- read.csv(file.path(data_path,"TropicalAndes_TRY_plant_traits_forestSpecies.csv"))
```

# Subset to columns of interest
```{r}
# remove TRY columns without value in TraitID
TRY_traits_2 <- TRY_traits[!is.na(TRY_traits$TraitID), ]
```

```{r}
BIEN_traits_sub <- BIEN_traits[ , c("scrubbed_species_binomial", "trait_name", "trait_value", "unit")]
TRY_traits_sub <- TRY_traits_2[ , c("AccSpeciesName", "TraitName", "OrigValueStr", "OrigUnitStr")]
```

# Standardize column names
```{r}
BIEN_traits_stan <- BIEN_traits_sub %>% rename(species = scrubbed_species_binomial)
TRY_traits_stan <- TRY_traits_sub %>% rename(species = AccSpeciesName, trait_name = TraitName, trait_value = OrigValueStr, unit = OrigUnitStr)
```

```{r}
plant_traits <- dplyr::bind_rows(BIEN_traits_stan, TRY_traits_stan, .id = 'source')
```

```{r}
#replace source 1 or 2 with BIEN or TRY
plant_traits['source'][plant_traits['source'] == 1] <- "BIEN"
plant_traits['source'][plant_traits['source'] == 2] <- "TRY"
```

# Summary
```{r}
glimpse(plant_traits)
```

```{r}
plant_traits %>%
  count(trait_name)
```

# Dispersal Traits
```{r}
# subset plant traits to just dispersal traits
plant_traits_dispersal <- plant_traits[plant_traits$trait_name == "Dispersal syndrome" | plant_traits$trait_name == "whole plant dispersal syndrome",]
# look at possible trait values
plant_traits_dispersal_values <- unique(plant_traits_dispersal$trait_value)
# save list to manually classify biotic, abiotic, none, and NA
#write.csv(plant_traits_dispersal_values, file.path(data_path,"plant_trait_dispersal_values.csv"))
# read in classified table
dispersal_classified <- read.csv(file.path(data_path,"plant_trait_dispersal_values_classified.csv"))
# reclassify dispersal trait value to table
plant_traits_dispersal$trait_value <- 
dispersal_classified$Classification[match(plant_traits_dispersal$trait_value,dispersal_classified$OrigValue )]
# check to make sure it worked
unique(plant_traits_dispersal$trait_value)
```

```{r}
#need to remove none & NA
plant_traits_dispersal <- plant_traits_dispersal[plant_traits_dispersal$trait_value == "biotic" | plant_traits_dispersal$trait_value == "abiotic",]
#need one value per species

plant_traits_dispersal_wide <- pivot_wider(data = plant_traits_dispersal, id_cols = species, names_from = trait_name, values_from = trait_value, values_fn = list)
```

# Plant fruit mass traits
```{r}
plant_traits_mass <- plant_traits[plant_traits$trait_name == "Fruit dry mass",]
plant_traits_mass <- plant_traits_mass[plant_traits_mass$unit == "mg",]
plant_traits_mass$trait_value <- as.numeric(plant_traits_mass$trait_value)
plant_traits_mass_wide <- pivot_wider(data = plant_traits_mass, id_cols = species, names_from = trait_name, values_from = trait_value, values_fn = mean)
# ready to combine
```

# Plant height traits
```{r}
plant_traits_height <- plant_traits[plant_traits$trait_name == "Plant height vegetative" |plant_traits$trait_name == "whole plant height" ,]
plant_traits_height$trait_value <- as.numeric(plant_traits_height$trait_value)
plant_traits_height_wide <- pivot_wider(data = plant_traits_height, id_cols = species, names_from = trait_name, values_from = trait_value, values_fn = mean)
plant_traits_height_wide$plant_height <- rowMeans(plant_traits_height_wide[,2:3], na.rm = TRUE)
# ready to combine
```

# Plant longevity traits
```{r}
plant_traits_longevity <- plant_traits[plant_traits$trait_name == "Plant lifespan (longevity)" | plant_traits$trait_name == "maximum whole plant longevity",]
#trait values are a mix of numeric and characters
```

# Combine all traits
```{r}
# combine all trait wide tables
plant_traits_wide <- bind_cols(plant_traits_height_wide, plant_traits_mass_wide)
# remove species with NAs for one or more traits

```

# Write data to csv
```{r}
write.csv(plant_traits, file.path(output_path,"TropicalAndes_plant_traits_forest.csv"))
```