### 2023-08-04 code/L0/L0_TropicalAndes_plants_GIFT_traits.Rmd
inital commit

---
title: "Tropical Andes plant trait data GIFT"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script retrieves plant trait data from the GIFT database for plant species list."
data input: ""
data output: ""
date: "2023-08-03"
output: html_document
---

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(GIFT)
library(dplyr)
```

# Load in species list
```{r}
species_list <- read.csv("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1/TropicalAndes_plant_Forest_SpeciesList.csv")
species_list <- species_list$x
```

# Retrive GIFT data
```{r}
trait_meta <- GIFT_traits_meta()
```
Traits of interest: plant height (ID = 1.6.3), lifespan (ID = 2.2.1), dispersal syndrome (ID = 3.3)

```{r}
# 
height <- GIFT_traits(trait_IDs = c("1.6.2"))
lifespan <- GIFT_traits(trait_IDs = c("2.2.1"))
dispersal_syndrome <- GIFT_traits(trait_IDs = c("3.3.1", "3.3.2"))
```

```{r}
# filter by species list
height_sub <- height %>%
  filter(work_species %in% species_list)
dispersal_syndrome_sub <- dispersal_syndrome %>%
  filter(work_species %in% species_list)
lifespan_sub <- lifespan %>%
  filter(work_species %in% species_list)
```

```{r}
#rename columns to trait name
height_sub <- height_sub %>% rename(trait_value = trait_value_1.6.2, species = work_species)
dispersal_syndrome_sub <- dispersal_syndrome_sub %>% rename(trait_value = trait_value_3.3.1, species = work_species)
lifespan_sub <- lifespan_sub %>% rename(trait_value = trait_value_2.2.1, species = work_species)
```

```{r}
#keep just species and trait columns
height_sub <- height_sub %>% select(species, trait_value)
dispersal_syndrome_sub <- dispersal_syndrome_sub %>% select(species, trait_value)
lifespan_sub <- lifespan_sub %>% select(species, trait_value)
```

```{r}
# add trait name column
height_sub$trait_name <- "plant_height"
dispersal_syndrome_sub$trait_name <- "dispersal_syndrome"
lifespan_sub$trait_name <- "plant_lifespan"
```

```{r}
#convert numeric trait values to character
height_sub$trait_value <- as.character(height_sub$trait_value)
lifespan_sub$trait_value <- as.character(lifespan_sub$trait_value)
```

```{r}
#combine all trait dataframes into one long table
GIFT_traits <- dplyr::bind_rows(height_sub, dispersal_syndrome_sub, lifespan_sub)
```

# Summary
```{r}
glimpse(GIFT_traits)
```

```{r}
print("The number of records is")
nrow(GIFT_traits)
print("The number of species is")
length(unique(GIFT_traits$species))
```

# Write data to csv
```{r}
write.csv(GIFT_traits, file = file.path(output_path,"TropicalAndes_GIFT_plant_traits_forest.csv"))
```