### 2023-08-04 code/L2/L2_2_TropicalAndes_funcdiv.Rmd
inital commit

---
title: "Tropical Andes Functional diversity for plants and Frugivores"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: ""
data input: ""
data output: ""
date: "2023-08-03"
output: html_document
---

# Load required packages
```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(smoothr)
library(purrr)
library(raster)
library(scico)
library(ggspatial)
library(letsR)
library(mFD)
```

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
data_path2<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
```

# Read in Data
```{r}
TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_plant_occ_forest.csv"))
TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_frugivore_occ_forest.csv"))
TropicalAndes_IUCNHabitat_Forest <- raster(file.path(data_path2, "TropicalAndes_IUCNHabitat_Forest.tif"))
frugivore_traits <- read.csv(file.path(data_path,"TropicalAndes_Frugivoria_traits_Forest.csv"))
plant_traits <- 
```

```{r}
lets.presab.points
```