### 2023-08-04 code/L2/L2_2_TropicalAndes_funcdiv.Rmd
add PAM code

---
title: "Tropical Andes Functional diversity for plants and Frugivores"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: ""
data input: ""
data output: ""
date: "2023-08-03"
output: html_document
---

# Load required packages
```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(smoothr)
library(purrr)
library(raster)
library(scico)
library(ggspatial)
library(letsR)
library(mFD)
```

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
data_path2<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
```

# Read in Data
```{r}
TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_plant_occ_forest.csv"))
TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_frugivore_occ_forest.csv"))
TropicalAndes_IUCNHabitat_Forest <- raster(file.path(data_path2, "TropicalAndes_IUCNHabitat_Forest.tif"))
frugivore_traits <- read.csv(file.path(data_path,"TropicalAndes_Frugivoria_traits_Forest.csv"))
plant_traits <- file.path(data_path,"TropicalAndes_plant_traits_forest.csv")
```

# Remove species with NAs for one or more traits
```{r}
frugivore_traits <- na.omit(frugivore_traits)
```

# Forest shape conversion
```{r}
TropicalAndes_IUCNHabitat_Forest <- aggregate(TropicalAndes_IUCNHabitat_Forest, 50, fun=mean, na.rm=TRUE)
TropicalAndes_IUCNHabitat_Forest
```

```{r}
Forest_poly <- rasterToPolygons(TropicalAndes_IUCNHabitat_Forest, na.rm = TRUE, dissolve = TRUE)
Forest_poly
plot(Forest_poly)
```

```{r}
Forest_sf <- st_as_sf(Forest_poly)
Forest_sf
plot(Forest_sf)
st_write(Forest_sf, file.path(data_path, "Forest_sf.shp"))
```

# Create Presence-Absense matrix
```{r}
x <- TropicalAndes_plant_occ_forest$decimalLongitude
y <- TropicalAndes_plant_occ_forest$decimalLatitude
plant_xy <- cbind(x, y)
plant_species <- TropicalAndes_plant_occ_forest$species
plants_PAM <- lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -54, ymn = -24, ymx = 14)
summary(plants_PAM)
plot(plants_PAM, xlab = "Longitude", ylab = "Latitude", main = "Plant richness map")
```

```{r}
x <- TropicalAndes_frugivore_occ_forest$decimalLongitude
y <- TropicalAndes_frugivore_occ_forest$decimalLatitude
frugivore_xy <- cbind(x, y)
frugivore_species <- TropicalAndes_frugivore_occ_forest$species
frugivore_PAM <- lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -54, ymn = -24, ymx = 14)
summary(frugivore_PAM)
plot(frugivore_PAM, xlab = "Longitude", ylab = "Latitude", main = "Frugivore richness map")
```

# Create trait type table
```{r}
frug_trait_name <- c("diet_cat", "body_mass", "body_size", "generation_time")
frug_trait_type <- c("N", "Q", "Q", "Q")
frug_trait_cat <- as.data.frame(cbind(frug_trait_name, frug_trait_type))
plant_trait_name <- c("dispersal_syndrome", "plant_height", "plant_lifespan")
frug_trait_type <- c("N", "Q", "Q") 
```