### 2023-08-04 code/L1/L1_TropicalAndes_subset_occ_forest_taxResolution.Rmd
inital commit

---
title: "Tropical Andes plant species list"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script retrives updated names for GBIF species list."
data input: "none"
data output: ""
date: "2023-08-03"
output: html_document
---

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(TNRS)
devtools::install_github("ecoinfor/U.Taxonstand")
library(U.Taxonstand)
library(tidyr)
library(dplyr)
```

# Load species list
```{r}
#species list
TropicalAndes_plant_Forest_SpeciesList <- read.csv("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1/TropicalAndes_plant_Forest_SpeciesList.csv")
TropicalAndes_plant_Forest_SpeciesList$NAME <- TropicalAndes_plant_Forest_SpeciesList$x
species_list <- TropicalAndes_plant_Forest_SpeciesList$NAME
```

# Use TRNS package
```{r}
TNRS_results <- TNRS(TropicalAndes_plant_Forest_SpeciesList)
```

```{r}
write.csv(TNRS_results, file.path(output_path,"TropicalAndes_plant_Forest_SpeciesList_TNRS_results.csv"))
```

# Using U.Taxonstand package
```{r}
U.Taxonstand_results <- nameMatch(spList = species_list, spSource = TropicalAndes_plant_Forest_SpeciesList, author = FALSE, max.distance = 1)
```

```{r}
write.csv(U.Taxonstand_results, file.path(output_path,"TropicalAndes_plant_Forest_SpeciesList_U.Taxonstand_results.csv"))
```