### 2023-08-05 code/L2/L2_2_TropicalAndes_funcdiv.Rmd
add plotting code fix mismatched name issue

---
title: "Tropical Andes Functional diversity for plants and Frugivores"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: ""
data input: ""
data output: ""
date: "2023-08-03"
output: html_document
---

# Load required packages
```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(smoothr)
library(purrr)
library(raster)
library(scico)
library(ggspatial)
library(letsR)
library(mFD)
```

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
data_path2<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
```

# Read in Data
```{r}
TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_plant_occ_forest_sub.csv"))
TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path,"TropicalAndes_frugivore_occ_forest_sub.csv"))
TropicalAndes_IUCNHabitat_Forest <- raster(file.path(data_path2, "TropicalAndes_IUCNHabitat_Forest.tif"))
frugivore_traits <- read.csv(file.path(data_path,"TropicalAndes_Frugivoria_traits_Forest.csv"))
plant_traits <- read.csv(file.path(data_path,"TropicalAndes_plant_traits_forest.csv"))
```

# Forest shape conversion
```{r}
TropicalAndes_IUCNHabitat_Forest <- aggregate(TropicalAndes_IUCNHabitat_Forest, 50, fun=mean, na.rm=TRUE)
TropicalAndes_IUCNHabitat_Forest
```

```{r}
Forest_poly <- rasterToPolygons(TropicalAndes_IUCNHabitat_Forest, na.rm = TRUE, dissolve = TRUE)
Forest_poly
plot(Forest_poly)
```

```{r}
Forest_sf <- st_as_sf(Forest_poly)
Forest_sf
plot(Forest_sf)
st_write(Forest_sf, file.path(data_path, "Forest_sf.shp"))
```

# Create Presence-Absense matrix
```{r}
x <- TropicalAndes_plant_occ_forest$decimalLongitude
y <- TropicalAndes_plant_occ_forest$decimalLatitude
plant_xy <- cbind(x, y)
plant_species <- TropicalAndes_plant_occ_forest$species
plants_PAM <- lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -54, ymn = -24, ymx = 14)
summary(plants_PAM)
plot(plants_PAM, xlab = "Longitude", ylab = "Latitude", main = "Plant richness map")
plants_PAM_matrix <- lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -54, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE)
```

```{r}
x <- TropicalAndes_frugivore_occ_forest$decimalLongitude
y <- TropicalAndes_frugivore_occ_forest$decimalLatitude
frugivore_xy <- cbind(x, y)
frugivore_species <- TropicalAndes_frugivore_occ_forest$species
frugivore_PAM <- lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -54, ymn = -24, ymx = 14)
summary(frugivore_PAM)
plot(frugivore_PAM, xlab = "Longitude", ylab = "Latitude", main = "Frugivore richness map")
frugivore_PAM_matrix <- lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -54, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE)
```

# Prepping data for functional diversity calculations

```{r}
plant_traits <- na.omit(plant_traits)
frugivore_traits <- na.omit(frugivore_traits)
frugivore_traits$body_mass_e <- as.numeric(frugivore_traits$body_mass_e)
frugivore_traits$body_size_mm <- as.numeric(frugivore_traits$body_size_mm)
plant_traits <- plant_traits[, c("species", "plant_height", "dispersal_syndrome", "plant_lifespan")]
frugivore_traits <- frugivore_traits[, c("IUCN_species_name", "diet_cat", "body_mass_e", "body_size_mm", "generation_time")]
```

```{r}
coordinates_sf_plant <- st_as_sf(TropicalAndes_plant_occ_forest_sub, coords = c("decimalLongitude", "decimalLatitude"), crs = st_crs(Forest_sf))
coordinates_sf_frugivore <- st_as_sf(TropicalAndes_frugivore_occ_forest_sub, coords = c("decimalLongitude", "decimalLatitude"), crs = st_crs(Forest_sf))
```

```{r}
# Perform the spatial intersection
intersections_plant_TA <- st_intersection(coordinates_sf_plant, Forest_sf)
intersections_frugivore_TA <- st_intersection(coordinates_sf_frugivore, Forest_sf)
# Convert the intersected sf object back to a dataframe
coordinates_df_subset_plant <- as.data.frame(intersections_plant_TA)
coordinates_df_subset_frugivore <- as.data.frame(intersections_frugivore_TA)
```

```{r}
# Print the resulting subset of coordinates
print(coordinates_df_subset_plant)
print(coordinates_df_subset_frugivore)
# Convert the point geometry to a dataframe of longitude and latitude
subset_plant <- as.data.frame(st_coordinates(intersections_plant_TA))
subset_frugivore <- as.data.frame(st_coordinates(intersections_frugivore_TA))

# Match the coordinates to the original dataframe
#merge the subset and full dataframe together to get final TA dataset
plant_PAM_filter <- merge(plants_PAM_matrix, subset_plant, by.x = c("Longitude(x)", "Latitude(y)"), by.y = c("X", "Y"))
frugivore_PAM_filter <- merge(frugivore_PAM_matrix, subset_frugivore, by.x = c("Longitude(x)", "Latitude(y)"), by.y = c("X", "Y"))
```

```{r}
# Turn PAM into matrix
plant_PAM <- as.matrix(plant_PAM_filter)
frugivore_PAM <- as.matrix(frugivore_PAM_filter)

# Add rownames
# Get the number of rows in the matrix
num_rows_plant <- nrow(plant_PAM)
num_rows_frugivore <- nrow(frugivore_PAM)
# Generate unique row names
row_names_plant <- paste0("cell", 1:num_rows_plant)
row_names_frugivore <- paste0("cell", 1:num_rows_frugivore)
# Assign row names to the matrix
rownames(plant_PAM) <- row_names_plant
rownames(frugivore_PAM) <- row_names_frugivore
# Print the matrix with row names
print(plant_PAM)
column_names_plant <- colnames(plant_PAM)
clean_column_names_plant <- gsub("_", " ", column_names_plant)
print(frugivore_PAM)
column_names_frugivore <- colnames(frugivore_PAM)
clean_column_names_frugivore <- gsub("_", " ", column_names_frugivore)
#Insert clean column names
colnames(plant_PAM) <- clean_column_names_plant
colnames(frugivore_PAM) <- clean_column_names_frugivore
```

```{r}
#turn data into correct data types for inputs into the trait categories dataframe
frugivore_traits$diet_cat <- as.factor(frugivore_traits$diet_cat)
frugivore_traits$body_mass_e <- as.numeric(frugivore_traits$body_mass_e)
frugivore_traits$body_size_mm <- as.numeric(frugivore_traits$body_size_mm)
frugivore_traits$generation_time <- as.numeric(frugivore_traits$generation_time)

plant_traits$plant_height <- as.numeric(plant_traits$plant_height)
plant_traits$plant_lifespan<- as.numeric(plant_traits$plant_lifespan)
plant_traits$dispersal_syndrome <- as.factor(plant_traits$dispersal_syndrome)

# Remove the species from PAM that have no occurrences anymore after subsetting to TA
# Remove columns with sum equal to zero
PAM_plant_site_final <- plant_PAM[, colSums(plant_PAM) != 0]
PAM_frugivore_site_final <- frugivore_PAM[, colSums(frugivore_PAM) != 0]
#Save coordinates for later
site_loc_key_plant <- PAM_plant_site_final[,1:2]
site_loc_key_frugivore <- PAM_frugivore_site_final[,1:2]

columns_to_remove <- c(1,2)
PAM_plant_site_final <- PAM_plant_site_final[,-columns_to_remove]
PAM_frugivore_site_final <- PAM_frugivore_site_final[,-columns_to_remove]
colnames_plant <- colnames(PAM_plant_site_final)
colnames_frugivore <- colnames(PAM_frugivore_site_final)
```

```{r}
# Remove species names from trait matrix not in the PAM
plant_traits_df_subset <- plant_traits %>% filter(species %in% colnames_plant)
frugivore_traits_df_subset <- frugivore_traits %>% filter(IUCN_species_name %in% colnames_frugivore)

# Turn trait dataframe into a matrix
class(plant_traits)
class(frugivore_traits)
plant_traits_matrix <- as.matrix(plant_traits_df_subset)
frugivore_traits_matrix <- as.matrix(frugivore_traits_df_subset)

# Define row names as species names
row_names_plant <- plant_traits_matrix[,1]
row_names_frugivore <- frugivore_traits_matrix[,1]

# Assign row names to the matrix
rownames(plant_traits_matrix) <- row_names_plant
rownames(frugivore_traits_matrix) <- row_names_frugivore
# Turn back into dataframe
plant_traits_df_final <-as.data.frame(plant_traits_matrix)
frugivore_traits_df_final <-as.data.frame(frugivore_traits_matrix)
plant_traits_df$X <-NULL
frugivore_traits_df$X <-NULL
```

```{r}
#fix types
frugivore_traits_df_final$diet_cat <- as.factor(frugivore_traits_df_final$diet_cat)
frugivore_traits_df_final$body_mass_e <- as.numeric(frugivore_traits_df_final$body_mass_e)
frugivore_traits_df_final$body_size_mm<- as.numeric(frugivore_traits_df_final$body_size_mm)
frugivore_traits_df_final$generation_time <- as.numeric(frugivore_traits_df_final$generation_time)

plant_traits_df_final$plant_height <- as.numeric(plant_traits_df_final$plant_height)
plant_traits_df_final$plant_lifespan<- as.numeric(plant_traits_df_final$plant_lifespan)
plant_traits_df_final$dispersal_syndrome <- as.factor(plant_traits_df_final$dispersal_syndrome)

# remove duplicate species name column
plant_traits_df_final <- plant_traits_df_final[, c("plant_height", "dispersal_syndrome", "plant_lifespan")]
frugivore_traits_df_final <- frugivore_traits_df_final[, c("diet_cat", "body_mass_e", "body_size_mm", "generation_time")]
```

# Create trait type table
```{r}
trait_name <- c("diet_cat", "body_mass_e", "body_size_mm", "generation_time")
trait_type <- c("N", "Q", "Q", "Q")
frug_trait_cat <- as.data.frame(cbind(trait_name, trait_type))
trait_name <- c("dispersal_syndrome", "plant_height", "plant_lifespan")
trait_type <- c("N", "Q", "Q") 
plant_trait_cat <- as.data.frame(cbind(trait_name, trait_type))
```

```{r}
# Summary of the assemblages * species dataframe:
asb_sp_plant_summ <- mFD::asb.sp.summary(asb_sp_w = PAM_plant_site_final)
asb_sp_frugivore_summ <- mFD::asb.sp.summary(asb_sp_w = PAM_frugivore_site_final)
```

```{r}
# Species traits summary:
plant_traits_summ <- mFD::sp.tr.summary(
  tr_cat     = plant_trait_cat,   
  sp_tr      = plant_traits_df_final, 
  stop_if_NA = TRUE)
frugivore_traits_summ <- mFD::sp.tr.summary(
  tr_cat     = frug_trait_cat,   
  sp_tr      = frugivore_traits_df_final, 
  stop_if_NA = TRUE)
```

# Estimate functional trait-based distances between species
```{r}

sp_dist_plants <- mFD::funct.dist(
  sp_tr         = plant_traits_df_final,
  tr_cat        = plant_trait_cat,
  metric        = "gower",
  scale_euclid  = "scale_center",
  ordinal_var   = "classic",
  weight_type   = "equal",
  stop_if_NA    = TRUE)
```

```{r}
sp_dist_frugivore <- mFD::funct.dist(
  sp_tr         = frugivore_traits_df_final,
  tr_cat        = frug_trait_cat,
  metric        = "gower",
  scale_euclid  = "scale_center",
  ordinal_var   = "classic",
  weight_type   = "equal",
  stop_if_NA    = TRUE)
```

# Generate a multidimensional space
```{r}
fspaces_quality_plants <- mFD::quality.fspaces(
  sp_dist             = sp_dist_plants,
  maxdim_pcoa         = 10,
  deviation_weighting = "absolute",
  fdist_scaling       = FALSE,
  fdendro             = "average")
```

```{r}
fspaces_quality_frugivore <- mFD::quality.fspaces(
  sp_dist             = sp_dist_frugivore,
  maxdim_pcoa         = 10,
  deviation_weighting = "absolute",
  fdist_scaling       = FALSE,
  fdendro             = "average")
```

```{r}
# Look at the quality spaces only (MAD index looks at mean absolute deviation from the dissimilary matrix; want the deviation
# to be low, meaning that the true distances have been retained in the PCA)
round(fspaces_quality_plants$"quality_fspaces", 3)            
round(fspaces_quality_frugivore$"quality_fspaces", 3)     
# Plot the quality spaces (chose to look at 3D, 4D, and 5D since they had the lowest MAD). Will go with 4 dimensions.
mFD::quality.fspaces.plot(
  fspaces_quality            = fspaces_quality_plants,
  quality_metric             = "mad",
  fspaces_plot               = c("pcoa_3d", "pcoa_4d", 
                                 "pcoa_5d"),
  name_file                  = NULL,
  range_dist                 = NULL,
  range_dev                  = NULL,
  range_qdev                 = NULL,
  gradient_deviation         = c(neg = "darkblue", nul = "grey80", pos = "darkred"),
  gradient_deviation_quality = c(low = "yellow", high = "red"),
  x_lab                      = "Trait-based distance")
mFD::quality.fspaces.plot(
  fspaces_quality            = fspaces_quality_frugivore,
  quality_metric             = "mad",
  fspaces_plot               = c("pcoa_3d", "pcoa_4d", 
                                 "pcoa_5d"),
  name_file                  = NULL,
  range_dist                 = NULL,
  range_dev                  = NULL,
  range_qdev                 = NULL,
  gradient_deviation         = c(neg = "darkblue", nul = "grey80", pos = "darkred"),
  gradient_deviation_quality = c(low = "yellow", high = "red"),
  x_lab                      = "Trait-based distance")
```

```{r}
#testing correlation between functional axes and traits
sp_faxes_coord_plants <- fspaces_quality_plants$"details_fspaces"$"sp_pc_coord"
sp_faxes_coord_frugivore <- fspaces_quality_frugivore$"details_fspaces"$"sp_pc_coord"
# View the components of the PCA axes *Remember the firt few components explain the most variation in dissimilarity. Clusters into groups
# Computes linear model for continuous traits and Kruskall-Wallis tests for other types. 
plant_tr_faxes <- mFD::traits.faxes.cor(
  sp_tr          = plant_traits_df_final, 
  sp_faxes_coord = sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")], 
  plot           = TRUE)
frugivore_tr_faxes <- mFD::traits.faxes.cor(
  sp_tr          = frugivore_traits_df_final, 
  sp_faxes_coord = sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")], 
  plot           = TRUE)
# Print traits with significant effect:
plant_tr_faxes$"tr_faxes_stat"[which(plant_tr_faxes$"tr_faxes_stat"$"p.value" < 0.05), ]
frugivore_tr_faxes$"tr_faxes_stat"[which(frugivore_tr_faxes$"tr_faxes_stat"$"p.value" < 0.05), ]
# Return plots:
plant_tr_faxes$"tr_faxes_plot"
frugivore_tr_faxes$"tr_faxes_plot"
```

```{r}
#plotting functional space
sp_faxes_coord_plants <- fspaces_quality_plants$"details_fspaces"$"sp_pc_coord"
sp_faxes_coord_frugivore <- fspaces_quality_frugivore$"details_fspaces"$"sp_pc_coord"
big_plot_plants <- mFD::funct.space.plot(
  sp_faxes_coord  = sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")],
  faxes           = c("PC1", "PC2", "PC3", "PC4"),
  name_file       = NULL,
  faxes_nm        = NULL,
  range_faxes     = c(NA, NA),
  color_bg        = "grey95",
  color_pool      = "darkgreen",
  fill_pool       = "white",
  shape_pool      = 21,
  size_pool       = 1,
  plot_ch         = TRUE,
  color_ch        = "black",
  fill_ch         = "white",
  alpha_ch        = 0.5,
  plot_vertices   = TRUE,
  color_vert      = "blueviolet",
  fill_vert       = "blueviolet",
  shape_vert      = 6,
  size_vert       = 1,
  plot_sp_nm      = NULL,
  nm_size         = 3,
  nm_color        = "black",
  nm_fontface     = "plain",
  check_input     = TRUE)
big_plot_frugivore <- mFD::funct.space.plot(
  sp_faxes_coord  = sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")],
  faxes           = c("PC1", "PC2", "PC3", "PC4"),
  name_file       = NULL,
  faxes_nm        = NULL,
  range_faxes     = c(NA, NA),
  color_bg        = "grey95",
  color_pool      = "darkgreen",
  fill_pool       = "white",
  shape_pool      = 21,
  size_pool       = 1,
  plot_ch         = TRUE,
  color_ch        = "black",
  fill_ch         = "white",
  alpha_ch        = 0.5,
  plot_vertices   = TRUE,
  color_vert      = "blueviolet",
  fill_vert       = "blueviolet",
  shape_vert      = 6,
  size_vert       = 1,
  plot_sp_nm      = NULL,
  nm_size         = 3,
  nm_color        = "black",
  nm_fontface     = "plain",
  check_input     = TRUE)
#Need to remove parts of the PAM that have values less than or equal to the number of dimensions (4)
# Calculate row sums
row_sums_plant <- rowSums(PAM_plant_site_final)
subset_matrix_plant <- PAM_plant_site_final[row_sums_plant >= 4, ]
row_sums_frugivore <- rowSums(PAM_frugivore_site_final)
subset_matrix_frugivore <- PAM_frugivore_site_final[row_sums_frugivore >= 4, ]

```

```{r}
big_plot_plants
big_plot_frugivore
```

```{r}
# getting sp_faxes_coord_plants and subset_matrix_plants names to match
sp_faxes_coord_plants_sub <- sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")]
## check number of species names
nrow(sp_faxes_coord_plants_sub) # 54
ncol(subset_matrix_plant) #82

sp_faxes_coord_plants_sub_names <- row.names(sp_faxes_coord_plants_sub)
subset_matrix_plant_names <- colnames(subset_matrix_plant)

sp_faxes_coord_plants_sub_df <- as.data.frame(sp_faxes_coord_plants_sub)
subset_matrix_plant_df <- as.data.frame(subset_matrix_plant)

subset_matrix_plant <- subset_matrix_plant_df[, which((names(subset_matrix_plant_df) %in% sp_faxes_coord_plants_sub_names)==TRUE)]

ncol(subset_matrix_plant)
subset_matrix_plant <- as.matrix(subset_matrix_plant)
```

```{r}
# match frugivore names
sp_faxes_coord_frugivore_sub <- sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")]

## check number of species names
nrow(sp_faxes_coord_frugivore_sub) #718
ncol(subset_matrix_frugivore) #712

sp_faxes_coord_frugivore_sub_names <- row.names(sp_faxes_coord_frugivore_sub)
subset_matrix_frugivore_names <- colnames(subset_matrix_frugivore)

sp_faxes_coord_frugivore_sub_names <- chartr(".", " ", sp_faxes_coord_frugivore_sub_names)
sp_faxes_coord_frugivore_sub <- as.data.frame(sp_faxes_coord_frugivore_sub)
row.names(sp_faxes_coord_frugivore_sub) <- sp_faxes_coord_frugivore_sub_names

sp_faxes_coord_frugivore_sub <- sp_faxes_coord_frugivore_sub[ which((row.names(sp_faxes_coord_frugivore_sub) %in% subset_matrix_frugivore_names)==TRUE), ]

nrow(sp_faxes_coord_frugivore_sub)
sp_faxes_coord_frugivore_sub <- as.matrix(sp_faxes_coord_frugivore_sub)
```

```{r}
#computing FD
# The number of species per assemblage has to be higher or equal to the number of traits
alpha_fd_indices_plant <- mFD::alpha.fd.multidim(
  sp_faxes_coord   = sp_faxes_coord_plants_sub,
  asb_sp_w         = subset_matrix_plant,
  ind_vect         = c("fide", "fdis", "fmpd", "fnnd", "feve", "fori", "fspe"), #did not run fdiv or fric (slows it down)
  scaling          = TRUE,
  check_input      = TRUE,
  details_returned = TRUE)

alpha_fd_indices_frugivore <- mFD::alpha.fd.multidim(
  sp_faxes_coord   = sp_faxes_coord_frugivore_sub,
  asb_sp_w         = subset_matrix_frugivore,
  ind_vect         = c("fide", "fdis", "fmpd", "fnnd", "feve", "fori", "fspe"), #did not run fdiv or fric (slows it down)
  scaling          = TRUE,
  check_input      = TRUE,
  details_returned = TRUE)

details_list_plant <- alpha_fd_indices_plant$"details"
details_list_frugivore <- alpha_fd_indices_frugivore$"details"

#plot
plots_alpha_plant <- mFD::alpha.multidim.plot(
  output_alpha_fd_multidim = alpha_fd_indices_plant,
  plot_asb_nm              = c("cell4", "cell20"),
  ind_nm                   = c("fdis", "fide", "fnnd"
                        , "fori", "fspe"),
  faxes                    = NULL,
  faxes_nm                 = NULL,
  range_faxes              = c(NA, NA),
  color_bg                 = "grey95",
  shape_sp                 = c(pool = 3, asb1 = 21, asb2 = 21),
  size_sp                  = c(pool = 0.7, asb1 = 1, asb2 = 1),
  color_sp                 = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  color_vert               = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_sp                  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_vert                = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  color_ch          
  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_ch                  = c(pool = "white", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  alpha_ch                 = c(pool = 1, asb1 = 0.3, asb2 = 0.3),
  shape_centroid_fdis      = c(asb1 = 22,  asb2 = 24),
  shape_centroid_fdiv      = c(asb1 = 22,  asb2 = 24),
  shape_centroid_fspe      = 23,
  color_centroid_fspe      = "black",
  size_sp_nm               = 3, 
  color_sp_nm              = "black",
  plot_sp_nm               = NULL,
  fontface_sp_nm           = "plain",
  save_file                = FALSE,
  check_input              = TRUE)

plots_alpha_frugivore <- mFD::alpha.multidim.plot(
  output_alpha_fd_multidim = alpha_fd_indices_frugivore,
  plot_asb_nm              = c("cell4", "cell20"),
  ind_nm                   = c("fdis", "fide", "fnnd"
                        , "fori", "fspe"),
  faxes                    = NULL,
  faxes_nm                 = NULL,
  range_faxes              = c(NA, NA),
  color_bg                 = "grey95",
  shape_sp                 = c(pool = 3, asb1 = 21, asb2 = 21),
  size_sp                  = c(pool = 0.7, asb1 = 1, asb2 = 1),
  color_sp                 = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  color_vert               = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_sp                  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_vert                = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  color_ch          
  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_ch                  = c(pool = "white", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  alpha_ch                 = c(pool = 1, asb1 = 0.3, asb2 = 0.3),
  shape_centroid_fdis      = c(asb1 = 22,  asb2 = 24),
  shape_centroid_fdiv      = c(asb1 = 22,  asb2 = 24),
  shape_centroid_fspe      = 23,
  color_centroid_fspe      = "black",
  size_sp_nm               = 3, 
  color_sp_nm              = "black",
  plot_sp_nm               = NULL,
  fontface_sp_nm           = "plain",
  save_file                = FALSE,
  check_input              = TRUE)

plots_alpha_plant$"fric"$"patchwork"  #comparing convex hulls of different hulls
plots_alpha_frugivore$"fric"$"patchwork"
plots_alpha_plant$"fdiv"$"patchwork"
plots_alpha_frugivore$"fdiv"$"patchwork"
plots_alpha_plant$"fspe"$"patchwork"
plots_alpha_frugivore$"fspe"$"patchwork"
plots_alpha_plant$"fdis"$"patchwork"
plots_alpha_frugivore$"fdis"$"patchwork"
plots_alpha_plant$"fori"$"patchwork"
plots_alpha_frugivore$"fori"$"patchwork"
```

# Turning the FD values into a raster
```{r}

#Need to subset the coordinates made at the top to the same set of the assemblages (removed values less than 5 so these coords need to be removed to do this correctly)

#Have coordinates and associated values

library(raster)

# Create an empty raster with desired resolution and extent
resolution <- c(0.75, 0.75)  # Set the desired resolution
extent <- c(-85, -54, -24, 14)  # Set the desired extent
empty_raster <- raster(resolution = resolution, xmn = extent[1], xmx = extent[2], ymn = extent[3], ymx = extent[4])

# Generate coordinates
subset_coords_plant <- site_loc_key_plant[rowSums(PAM_plant_site_final) >= 5, ]
subset_coords_plant_sp <-subset_coords_plant[,1:2 ]
subset_coords_frugivore <- site_loc_key_frugivore[rowSums(PAM_frugivore_site_final) >= 5, ]
subset_coords_frugivore_sp <-subset_coords_frugivore[,1:2 ]

#Get functional dispersion
fdis_plant <- alpha_fd_indices_plant$functional_diversity_indices$fdis
fdis_frugivore <- alpha_fd_indices_frugivore$functional_diversity_indices$fdis

plant_fd_sp <- data.frame(subset_coords_plant_sp, fdis_plant)
frugivore_fd_sp <- data.frame(subset_coords_frugivore_sp, fdis_frugivore)

# Convert the dataframe to sf format
spatial_fdis_plant <- st_as_sf(plant_fd_sp, coords = c("Longitude.x.", "Latitude.y."))
spatial_fdis_frugivore <- st_as_sf(frugivore_fd_sp, coords = c("Longitude.x.", "Latitude.y."))

```

```{r}
# Rasterize the sf data to create the FD raster
fd_raster_plant <- rasterize(spatial_fdis_plant, empty_raster)
writeRaster(fd_raster_plant$fdis_plant, filename = file.path(output_path,"FD_plant_TA.tif"), format="GTiff",overwrite=T)
writeRaster(fd_raster_plant$ID, filename = file.path(output_path,"FD_plant_TA_ID.tif"), format="GTiff")

fd_raster_frugivore <- rasterize(spatial_fdis_frugivore, empty_raster)
writeRaster(fd_raster_frugivore$fdis_frugivore, filename = file.path(output_path,"FD_frugivore_TA.tif"), format="GTiff",overwrite=T)
writeRaster(fd_raster_frugivore$ID, filename = file.path(output_path,"FD_frugivore_TA_ID.tif"), format="GTiff")
```

# Functional diversity plots
```{r}
# set crs of sf objects
spatial_fdis_plant <- spatial_fdis_plant %>% st_set_crs(4326)
spatial_fdis_frugivore <- spatial_fdis_frugivore %>% st_set_crs(4326)
```

```{r}
# Raster to raster point
spec_FD_raster_plant_point <- rasterToPoints(fd_raster_plant, spatial = TRUE)
spec_FD_raster_frugivore_point <- rasterToPoints(fd_raster_frugivore, spatial = TRUE)
# Convert to a 'conventional' dataframe
fd_raster_plant_point <- data.frame(spec_FD_raster_plant_point)
fd_raster_frugivore_point <- data.frame(spec_FD_raster_frugivore_point)
```

```{r}
plantgridFDTA <-
  ggplot() +
  geom_sf(data = worldMap, fill = "white")+
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_tile(data= fd_raster_plant_point, aes(x=x, y=y, fill=fdis_plant)) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  ) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  labs(title = "Tropical Andes Plant Functional Dispersion", x = "Longitude", y = "Latitude", fill = "FD")
plantgridFDTA
```

```{r}
frugivoregridFDTA <-
  ggplot() +
  geom_sf(data = worldMap, fill = "white")+
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_tile(data= fd_raster_frugivore_point, aes(x=x, y=y, fill=fdis_frugivore)) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  ) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  labs(title = "Tropical Andes Frugivore Functional Dispersion", x = "Longitude", y = "Latitude", fill = "FD")
frugivoregridFDTA
```

# Extract cell values
```{r}
plantgridFDTA <- plantgridFDTA[["data"]]
frugivoregridFDTA <- frugivoregridFDTA[["data"]]
```

```{r}
plot(spatial_fdis_plant$fdis_plant, spatial_fdis_frugivore$fdis_frugivore,
     main = "Plant FD vs Frugivore FD",
     xlab = "Plant Functional Dispersion by cell", ylab = "Frugivore Functional Dispersion by cell")
abline(lm(spatial_fdis_frugivore$fdis_frugivore ~ spatial_fdis_plant$fdis_plant))
summary(lm(spatial_fdis_frugivore$fdis_frugivore ~ spatial_fdis_plant$fdis_plant))
```