### 2023-09-27 code/L2/L2_TropicalAndes_plant_taxonomiccoverage.Rmd
moved from L1 to L2

```{r}
#input trait datasets
#compare coverage of species, families, genera by trait coverage (plot)
#do the same for GBIF occurrence 
library(WorldFlora)
library(taxize)
library(BIEN)
library(GIFT)
library(dplyr)
```

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Read in Data
```{r}
BIEN_traits <- read.csv(file.path(data_path,"TropicalAndes_BIEN_plant_traits_forest.csv"))
TRY_traits <- read.csv(file.path(data_path,"TropicalAndes_TRY_plant_traits_forestSpecies.csv"))
GIFT_traits <- read.csv(file.path(data_path,"TropicalAndes_GIFT_plant_traits_forest.csv"))
GBIF_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_plant_occ.csv"))
```

```{r}
#species list
BIEN_species <- unique(BIEN_traits$scrubbed_species_binomial)
TRY_species <- unique(TRY_traits$SpeciesName)
GIFT_species <- unique(GIFT_traits$species)
GBIF_species <- unique(GBIF_occ$species)
```

# Get higher taxonomy for species lists
```{r}
BIEN_tax<- BIEN_taxonomy_species(BIEN_species)
```

```{r}
TRY_tax <- tax_name(TRY_species, get = c('family', 'genus'), db= "itis")
```

```{r}
gift_sp <- GIFT_species()
sapply(GIFT_species, function(x) grep(x, gift_sp$work_species))
gift_sp[sapply(GIFT_species, function(x) grep(x, gift_sp$work_species)), ]

#GIFT_tax <- GIFT_taxgroup(taxon_lvl = c("family", "genus")[1], species = GIFT_species)
```

# Plot coverage
```{r}
BIEN_family_coverage <- BIEN_tax %>% count(scrubbed_family)
BIEN_family_coverage 
BIEN_genus_coverage <- BIEN_tax %>% count(scrubbed_genus) 
BIEN_genus_coverage
barplot(height = BIEN_family_coverage$n, names = BIEN_family_coverage$scrubbed_family, horiz=T, las=1)
barplot(height = BIEN_genus_coverage$n, names = BIEN_genus_coverage$scrubbed_genus, horiz=T, las=1)
```