### 2023-10-04 code/L1/L1_F1_TropicalAndes_frugivores_GBIF.Rmd
clean repo: update names and move old code to archive

---
title: "Cleaning Tropical Andes frugivore GBIF occurrence records"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "Uses CoordinateCleaner to flag and remove problematic records from GBIF frugivore occurrence records."
data input: "TropicalAndes_GBIF_frugivore_occ.csv"
data output: "TropicalAndes_GBIF_frugivore_occ_cleaned.csv, TropicalAndes_GBIF_frugivore_occ_flagged.csv"
date: "2023-07-25"
output: html_document
---

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(countrycode)
library(CoordinateCleaner)
library(dplyr)
```

# Read in data
```{r}
TropicalAndes_GBIF_frugivore_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_frugivore_occ.csv"))
```

# Clean data
Adapted code from https://ropensci.github.io/CoordinateCleaner/articles/Cleaning_GBIF_data_with_CoordinateCleaner.html

```{r}
# remove records without coordinates
TropicalAndes_GBIF_frugivore_occ <- TropicalAndes_GBIF_frugivore_occ %>%
  filter(!is.na(decimalLongitude)) %>%
  filter(!is.na(decimalLatitude))
```

```{r}
flags <- clean_coordinates(x = TropicalAndes_GBIF_frugivore_occ,
                           lon = "decimalLongitude",
                           lat = "decimalLatitude",
                           countries = "countryCode",
                           species = "species",
                           tests = c("capitals", "centroids",
                                     "duplicates", "equal", "gbif",
                                     "institutions", "seas","zeros"))
summary(flags)
```

```{r}
#exclude problematic records
TropicalAndes_GBIF_frugivore_occ_cleaned <- TropicalAndes_GBIF_frugivore_occ[flags$.summary,]
TropicalAndes_GBIF_frugivore_occ_flagged <- TropicalAndes_GBIF_frugivore_occ[!flags$.summary,]
```

```{r}
table(TropicalAndes_GBIF_frugivore_occ_cleaned$year)
```

```{r}
# remove records from before second world war
TropicalAndes_GBIF_frugivore_occ_cleaned <- TropicalAndes_GBIF_frugivore_occ_cleaned %>%
  filter(year > 1945) 
```

# Summary
```{r}
glimpse(TropicalAndes_GBIF_frugivore_occ_cleaned)
```

```{r}
print("The number of records is")
nrow(TropicalAndes_GBIF_frugivore_occ_cleaned)
print("The number of species is")
length(unique(TropicalAndes_GBIF_frugivore_occ_cleaned$species))
print("The number of families is")
length(unique(TropicalAndes_GBIF_frugivore_occ_cleaned$family))
```

# Write data to csv
```{r}
write.csv(TropicalAndes_GBIF_frugivore_occ_cleaned, file = file.path(output_path,"TropicalAndes_GBIF_frugivore_occ_cleaned.csv"))

write.csv(TropicalAndes_GBIF_frugivore_occ_flagged, file = file.path(output_path,"TropicalAndes_GBIF_frugivore_occ_flagged.csv"))
```