### 2023-10-04 code/L1/L1_F2_TropicalAndes_frugivore_traits.Rmd

---
title: "Frugivoria frugivore trait subset"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "Subsetting Frugivoria to the species and traits of interest"
data input: ""
data output: ""
date: "2023-08-03"
output: html_document
---

# Load required packages
```{r}
library(dplyr)
```

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
data_path2<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Read in data
```{r}
Frugivoria <- read.csv(file.path(data_path, file = "TropicalAndes_Frugivoria_frugivore_traits.csv"))
frugivore_species <- read.csv(file.path(data_path2, file = "TropicalAndes_frugivore_Forest_SpeciesList.csv"))
```

```{r}
frugivore_species <- frugivore_species$x
```

```{r}
# subset Frugivoria by species list
Frugivoria_filtered <- Frugivoria %>%
  filter(IUCN_species_name %in% frugivore_species)
```

```{r}
# keep columns with traits of diet category, body size, body mass, generation time
Frugivoria_subset <- Frugivoria_filtered[ , c("IUCN_species_name", "family", "genus", "species", "diet_cat", "body_mass_e", "body_size_mm", "generation_time")]
```

# Summary
```{r}
glimpse(Frugivoria_subset)
```

# Write data to csv
```{r}
write.csv(Frugivoria_subset, file.path(output_path,"TropicalAndes_Frugivoria_traits_Forest.csv"))
```