### 2023-10-04 code/L1/L1_F3_TropicalAndes_frug_occ.Rmd
inital commit

---
title: "TITLE"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script subsets frugivore occurrence data to species with complete trait coverage."
data input: "ADD"
data output: "ADD"
date: "2023-10-04"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}

```