### 2023-10-18 code/L0/L0_TropicalAndes_plants_GIFT.Rmd
remove species list, add more traits

---
title: "Tropical Andes plant trait data GIFT"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script retrieves plant trait data from the GIFT database for plant species list."
data input: ""
data output: ""
date: "2023-08-03"
output: html_document
---

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(GIFT)
library(dplyr)
```

# Retrive GIFT data
```{r}
trait_meta <- GIFT_traits_meta()
```
Traits of interest: plant height (ID = 1.6.3), lifespan (ID = 2.2.1), dispersal syndrome (ID = 3.3), Plant_height_min (1.6.1), Plant_height_max (1.6.2), Fruit_type_1 (3.16.1), fruit_dryness_1 (3.18.1), Fruit_length_min (3.13.1), Fruit_length_max (3.13.2), Fruit_length_mean (3.13.3), Seed_mass_min (3.2.1), Seed_mass_max (3.2.2), Seed_mass_mean (3.2.3), Seed_length_min (3.10.1), Seed_length_max (3.10.2), Seed_length_mean (3.10.3), Seed_width_min (3.11.1), Seed_width_max (3.11.2), Seed_width_mean (3.11.3)

```{r}
# 
height <- GIFT_traits(trait_IDs = c("1.6.2", "1.6.1", "1.6.3"))
lifespan <- GIFT_traits(trait_IDs = c("2.2.1"))
dispersal_syndrome <- GIFT_traits(trait_IDs = c("3.3.1", "3.3.2"))
fruit_type <- GIFT_traits(trait_IDs = c("3.16.1", "3.18.1"))
fruit_length <- GIFT_traits(trait_IDs = c("3.13.1", "3.13.2", "3.13.3"))
seed_mass <- GIFT_traits(trait_IDs = c("3.2.1", "3.2.2", "3.2.3"))
seed_length <- GIFT_traits(trait_IDs = c("3.10.1", "3.10.2", "3.10.3"))
seed_width <- GIFT_traits(trait_IDs = c("3.11.1", "3.11.2", "3.11.3"))
```

```{r}
# filter by species list
height_sub <- height %>%
  filter(work_species %in% species_list)
dispersal_syndrome_sub <- dispersal_syndrome %>%
  filter(work_species %in% species_list)
lifespan_sub <- lifespan %>%
  filter(work_species %in% species_list)
```

```{r}
#rename columns to trait name
height_sub <- height_sub %>% rename(trait_value = trait_value_1.6.2, species = work_species)
dispersal_syndrome_sub <- dispersal_syndrome_sub %>% rename(trait_value = trait_value_3.3.1, species = work_species)
lifespan_sub <- lifespan_sub %>% rename(trait_value = trait_value_2.2.1, species = work_species)
```

```{r}
#keep just species and trait columns
height_sub <- height_sub %>% select(species, trait_value)
dispersal_syndrome_sub <- dispersal_syndrome_sub %>% select(species, trait_value)
lifespan_sub <- lifespan_sub %>% select(species, trait_value)
```

```{r}
# add trait name column
height_sub$trait_name <- "plant_height"
dispersal_syndrome_sub$trait_name <- "dispersal_syndrome"
lifespan_sub$trait_name <- "plant_lifespan"
```

```{r}
#convert numeric trait values to character
height_sub$trait_value <- as.character(height_sub$trait_value)
lifespan_sub$trait_value <- as.character(lifespan_sub$trait_value)
```

```{r}
#combine all trait dataframes into one long table
GIFT_traits <- dplyr::bind_rows(height_sub, dispersal_syndrome_sub, lifespan_sub)
```

# Summary
```{r}
glimpse(GIFT_traits)
```

```{r}
print("The number of records is")
nrow(GIFT_traits)
print("The number of species is")
length(unique(GIFT_traits$species))
```

```{r}
GIFT_traits %>% count(trait_name)
```

# Write data to csv
```{r}
write.csv(GIFT_traits, file = file.path(output_path,"TropicalAndes_GIFT_plant_traits_forest.csv"))
```