### 2023-10-18 code/L0/L0_TropicalAndes_plants_checklist.Rmd
inital commit, add read in shapefile, and checklist code

---
title: "Tropical Andes plant checklist using GIFT"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script retrieves plant checklist data from the GIFT database for Tropical Andes Forest region."
data input: "Forest_sf.sph"
data output: ""
date: "2023-10-17"
output: html_document
---

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(GIFT)
library(dplyr)
library(sf)
```

# Load Tropical Andes Forest shape
```{r}
TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path, "Forest_sf.shp"), layer = "Forest_sf")
```

```{r}
checklist <- GIFT_checklists(taxon_name = "Angiospermae", taxonomic_group = TRUE, floristic_group = "all", shp = TropicalAndes_IUCNHabitat_Forest, overlap = "extent_intersect")
```