### 2023-10-25 code/L0/L0_TropicalAndes_plants_GIFT.Rmd
update to include plant growth form

---
title: "Tropical Andes plant trait data GIFT"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script retrieves plant trait data from the GIFT database for plant species list."
data input: ""
data output: ""
date: "2023-08-03"
output: html_document
---

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(GIFT)
library(dplyr)
```

# Retrive GIFT data
```{r}
GIFT_traits_meta()
```
Traits of interest: plant height (ID = 1.6.3), lifespan (ID = 2.2.1), dispersal syndrome (ID = 3.3), Plant_height_min (1.6.1), Plant_height_max (1.6.2), Fruit_type_1 (3.16.1), fruit_dryness_1 (3.18.1), Fruit_length_min (3.13.1), Fruit_length_max (3.13.2), Fruit_length_mean (3.13.3), Seed_mass_min (3.2.1), Seed_mass_max (3.2.2), Seed_mass_mean (3.2.3), Seed_length_min (3.10.1), Seed_length_max (3.10.2), Seed_length_mean (3.10.3), Seed_width_min (3.11.1), Seed_width_max (3.11.2), Seed_width_mean (3.11.3), Growth_form (1.2.1)

```{r}
# 
height_mean <- GIFT_traits(trait_IDs = c("1.6.3"))
height_min <- GIFT_traits(trait_IDs = c("1.6.1"))
height_max <- GIFT_traits(trait_IDs = c("1.6.2"))

lifespan <- GIFT_traits(trait_IDs = c("2.2.1"))

dispersal_syndrome_1 <- GIFT_traits(trait_IDs = c("3.3.1"))
dispersal_syndrome_2 <- GIFT_traits(trait_IDs = c("3.3.2"))

fruit_type <- GIFT_traits(trait_IDs = c("3.16.1"))
fruit_dryness <- GIFT_traits(trait_IDs = c("3.18.1"))

fruit_length_mean <- GIFT_traits(trait_IDs = c("3.13.3"))
fruit_length_min <- GIFT_traits(trait_IDs = c("3.13.1"))
fruit_length_max <- GIFT_traits(trait_IDs = c("3.13.2"))

seed_mass_mean <- GIFT_traits(trait_IDs = c("3.2.3"))
seed_mass_min <- GIFT_traits(trait_IDs = c("3.2.1"))
seed_mass_max <- GIFT_traits(trait_IDs = c("3.2.2"))

seed_length_mean <- GIFT_traits(trait_IDs = c("3.10.3"))
seed_length_min <- GIFT_traits(trait_IDs = c("3.10.1"))
seed_length_max <- GIFT_traits(trait_IDs = c("3.10.2"))

seed_width_mean <- GIFT_traits(trait_IDs = c("3.11.3"))
seed_width_min <- GIFT_traits(trait_IDs = c("3.11.1"))
seed_width_max <- GIFT_traits(trait_IDs = c("3.11.2"))

growth_form <- GIFT_traits(trait_IDs = c("1.2.1"))
```

```{r}
#rename columns to trait name
height_mean <- height_mean %>% rename(trait_value = trait_value_1.6.3, species = work_species)
height_min <- height_min %>% rename(trait_value = trait_value_1.6.1, species = work_species)
height_max <- height_max %>% rename(trait_value = trait_value_1.6.2, species = work_species)

lifespan <- lifespan %>% rename(trait_value = trait_value_2.2.1, species = work_species)

dispersal_syndrome_1 <- dispersal_syndrome_1 %>% rename(trait_value = trait_value_3.3.1, species = work_species)
dispersal_syndrome_2 <- dispersal_syndrome_2 %>% rename(trait_value = trait_value_3.3.2, species = work_species)

fruit_type <- fruit_type %>% rename(trait_value = trait_value_3.16.1, species = work_species)
fruit_dryness <- fruit_dryness %>% rename(trait_value = trait_value_3.18.1, species = work_species)

fruit_length_mean <- fruit_length_mean %>% rename(trait_value = trait_value_3.13.3, species = work_species)
fruit_length_min <- fruit_length_min %>% rename(trait_value = trait_value_3.13.1, species = work_species)
fruit_length_max <- fruit_length_max %>% rename(trait_value = trait_value_3.13.2, species = work_species)

seed_mass_mean <- seed_mass_mean %>% rename(trait_value = trait_value_3.2.3, species = work_species)
seed_mass_min <- seed_mass_min %>% rename(trait_value = trait_value_3.2.1, species = work_species)
seed_mass_max <- seed_mass_max %>% rename(trait_value = trait_value_3.2.2, species = work_species)

seed_length_mean <- seed_length_mean %>% rename(trait_value = trait_value_3.10.3, species = work_species)
seed_length_min <- seed_length_min %>% rename(trait_value = trait_value_3.10.1, species = work_species)
seed_length_max <- seed_length_max %>% rename(trait_value = trait_value_3.10.2, species = work_species)

seed_width_mean <- seed_width_mean %>% rename(trait_value = trait_value_3.11.3, species = work_species)
seed_width_min <- seed_width_min %>% rename(trait_value = trait_value_3.11.1, species = work_species)
seed_width_max <- seed_width_max %>% rename(trait_value = trait_value_3.11.2, species = work_species)

growth_form <- growth_form %>% rename(trait_value = trait_value_1.2.1, species = work_species)
```

```{r}
#keep just species and trait columns
height_mean <- height_mean %>% select(species, trait_value)
height_min <- height_min %>% select(species, trait_value)
height_max <- height_max %>% select(species, trait_value)

lifespan <- lifespan %>% select(species, trait_value)

dispersal_syndrome_1 <- dispersal_syndrome_1 %>% select(species, trait_value)
dispersal_syndrome_2 <- dispersal_syndrome_2 %>% select(species, trait_value)

fruit_type <- fruit_type %>% select(species, trait_value)
fruit_dryness <- fruit_dryness %>% select(species, trait_value)

fruit_length_mean <- fruit_length_mean %>% select(species, trait_value)
fruit_length_min <- fruit_length_min %>% select(species, trait_value)
fruit_length_max <- fruit_length_max %>% select(species, trait_value)

seed_mass_mean <- seed_mass_mean %>% select(species, trait_value)
seed_mass_min <- seed_mass_min %>% select(species, trait_value)
seed_mass_max <- seed_mass_max %>% select(species, trait_value)

seed_length_mean <- seed_length_mean %>% select(species, trait_value)
seed_length_min <- seed_length_min %>% select(species, trait_value)
seed_length_max <- seed_length_max %>% select(species, trait_value)

seed_width_mean <- seed_width_mean %>% select(species, trait_value)
seed_width_min <- seed_width_min %>% select(species, trait_value)
seed_width_max <- seed_width_max %>% select(species, trait_value)

growth_form <- growth_form %>% select(species, trait_value)
```

```{r}
# add trait name column
height_mean$trait_name <- "plant_height"
height_min$trait_name <- "plant_height_min"
height_max$trait_name <- "plant_height_max"
lifespan$trait_name <- "plant_lifespan"
dispersal_syndrome_1$trait_name <- "dispersal_syndrome_1"
dispersal_syndrome_2$trait_name <- "dispersal_syndrome_2"
fruit_type$trait_name <- "fruit_type"
fruit_dryness$trait_name <- "fruit_dryness"
fruit_length_mean$trait_name <- "fruit_length_mean"
fruit_length_min$trait_name <- "fruit_length_min"
fruit_length_max$trait_name <- "fruit_length_max"
seed_mass_mean$trait_name <- "seed_mass_mean"
seed_mass_min$trait_name <- "seed_mass_min"
seed_mass_max$trait_name <- "seed_mass_max"
seed_length_mean$trait_name <- "seed_length_mean"
seed_length_min$trait_name <- "seed_length_min"
seed_length_max$trait_name <- "seed_length_max"
seed_width_mean$trait_name <- "seed_width_mean"
seed_width_min$trait_name <- "seed_width_min"
seed_width_max$trait_name <- "seed_width_max"
growth_form$trait_name <- "growth_form"
```

```{r}
# convert numberic traits into characters so we can combine into one dataframe
height_mean$trait_value <- as.character(height_mean$trait_value)
height_min$trait_value <- as.character(height_min$trait_value)
height_max$trait_value <- as.character(height_max$trait_value)
lifespan$trait_value <- as.character(lifespan$trait_value)
fruit_length_mean$trait_value <- as.character(fruit_length_mean$trait_value)
fruit_length_min$trait_value <- as.character(fruit_length_min$trait_value)
fruit_length_max$trait_value <- as.character(fruit_length_max$trait_value)
seed_mass_mean$trait_value <- as.character(seed_mass_mean$trait_value)
seed_mass_min$trait_value <- as.character(seed_mass_min$trait_value)
seed_mass_max$trait_value <- as.character(seed_mass_max$trait_value)
seed_length_mean$trait_value <- as.character(seed_length_mean$trait_value)
seed_length_min$trait_value <- as.character(seed_length_min$trait_value)
seed_length_max$trait_value <- as.character(seed_length_max$trait_value)
seed_width_mean$trait_value <- as.character(seed_width_mean$trait_value)
seed_width_min$trait_value <- as.character(seed_width_min$trait_value)
seed_width_max$trait_value <- as.character(seed_width_max$trait_value)
```

```{r}
#combine all trait dataframes into one long table
GIFT_traits <- dplyr::bind_rows(height_mean, height_min, height_max, lifespan, dispersal_syndrome_1, dispersal_syndrome_2, fruit_type, fruit_dryness, fruit_length_mean, fruit_length_min, fruit_length_max, seed_mass_mean, seed_mass_min, seed_mass_max, seed_length_mean, seed_length_min, seed_length_max, seed_width_mean, seed_width_min, seed_width_max, growth_form)
```

# Summary
```{r}
glimpse(GIFT_traits)
```

```{r}
print("The number of records is")
nrow(GIFT_traits)
print("The number of species is")
length(unique(GIFT_traits$species))
```

```{r}
GIFT_traits %>% count(trait_name)
```

# Write data to csv
```{r}
write.csv(GIFT_traits, file = file.path(output_path,"TropicalAndes_GIFT_plant_traits.csv"))
```