### 2023-10-31 code/L1/L1_P2_TropicalAndes_plant_species_synonyms.Rmd
add header information, update file path, remove unnecessary/old code, clean up unnecessary text

---
title: "Plant species synonym list"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "Get synonmys for the plant species list of GBIF occurrence data"
data input: "GBIF_species_list.csv"
data output: ""
date: "2023-08-03"
output: html_document
---

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

```{r}
library(taxadb)
#library(taxize)
library(dplyr)
```

```{r}
#create local itis database
td_create("itis",overwrite=FALSE)
```

# Read in species list
```{r}
GBIF_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_plant_occ_cleaned.csv"))
GBIF_species_list <- unique(GBIF_occ$species)
```

```{r}
GBIF_species_list <- as.data.frame(GBIF_species_list)
names(GBIF_species_list) <- c("Scientific.Name")
```

#Using taxize
```{r}
species_synonyms <- synonyms(GBIF_species_list$Scientific.Name, db = "pow", rows = 1)
```

# Using taxadb
```{r}
#get IDS for each scientific name
syn1<-GBIF_species_list %>%
  mutate(ID=get_ids(Scientific.Name,"itis"))

#Deal with NAs (one name corresponds to more than 1 ITIS code)
syn1_NA<-as.data.frame(syn1$Scientific.Name[is.na(syn1$ID)])
colnames(syn1_NA)<-c("name")

NA_IDS<-NULL
tmp <- NULL
tmp$name <- NA
for(i in unique(syn1_NA$name)){
  tmp <-as.data.frame(filter_name(i, 'itis')[5])
  tmp$name<-paste0(i)
  NA_IDS<-rbind(NA_IDS,tmp)
}

#join with original names
colnames(syn1)<-c("name","ID")
IDS<-left_join(syn1,NA_IDS,by="name")

#extract just the unique IDs
IDS<-data.frame(ID=c(IDS[,"ID"],IDS[,"acceptedNameUsageID"]))
IDS<-as.data.frame(unique(IDS$ID))
IDS<-as.data.frame(IDS[-is.na(IDS)])
colnames(IDS)<-"ID"
#extract all names with synonyms in ITIS that are at the species level
#set query
ITIS<-taxa_tbl("itis") %>%
  select(scientificName,taxonRank,acceptedNameUsageID,taxonomicStatus) %>%
  filter(taxonRank == "species")

#see query
ITIS %>% show_query()
#retrieve results
ITIS_names<-ITIS %>% collect()

#filter to only those that match ITIS codes for all my species
ITIS_names<-ITIS_names %>%
  filter(acceptedNameUsageID %in% IDS$ID)
```