### 2023-10-31 code/L2/L2_TropicalAndes_data_citizenscience.Rmd
inital commit, read in data, plot year vs number of records for occurrence data, subset cleaned data for citizen science

Citizen science assessment

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
```

# Load required packages
```{r}
library(dplyr)
library(ggplot2)
library(rgbif)
```

# Read in Data
```{r}
plant_GBIF_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_plant_occ_cleaned.csv"))
```

# Time (year) vs records plot
```{r}
records_time <- GBIF_occ %>% count(year)
names(records_time) <- c("year", "count")
```

```{r}
ggplot(records_time, aes(x=year, y=count)) +
  geom_line() +
  xlim(1950, 2023) +
  geom_area(fill= "darkgreen")+
  ggtitle("Plant occurrence records in GBIF by year since 1950")
```

# Subset citizen science data
```{r}
dataset_keys <- unique(GBIF_occ$datasetKey)
# 50c9509d-22c7-4a22-a47d-8c48425ef4a7 = iNaturalist
GBIF_occ_iNaturalist <- GBIF_occ %>% filter(datasetKey == "50c9509d-22c7-4a22-a47d-8c48425ef4a7")
```

```{r}
records_time_iNaturalist <- GBIF_occ_iNaturalist %>% count(year)
names(records_time_iNaturalist) <- c("year", "count")
```

```{r}
ggplot(records_time_iNaturalist, aes(x=year, y=count)) +
  geom_line() +
  xlim(1950, 2023) +
  geom_area(fill= "lightgreen") +
  ggtitle("iNaturalist Plant occurrence records in GBIF by year since 1950")
```
# iNaturalist proportion of total records
```{r}
records_time$group <- c("all")
records_time_iNaturalist$group <- c("iNaturalist")
records_time_both <- rbind(records_time, records_time_iNaturalist)
```

```{r}
ggplot(records_time_both, aes(x=year, y=count, color=group)) + 
    geom_line() +
  xlim(1970, 2023) +
  ggtitle("Plant occurrence records in GBIF by year since 1970")
```