### 2023-11-01 code/L0/L0_TropicalAndes_plants_speciesLink_occ.Rmd

---
title: "Tropical Andes plant speciesLink data"
authors: "Hazel J. Anderson"
collaborators: "None"
data input: "none"
data output: "TBD"
project: "Frugivoria"
date: "2023-10-18"
output: html_document
---
# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(Rocc)
library(tidyr)
library(dplyr)
library(sf)
```

## List of desired countries
```{r}
TA_countries <- c("Bolivia", "Ecuador", "Venezuela", "Colombia", "Peru")
```

```{r}
TropicalAndes_speciesLink_occ <- rspeciesLink(dir = output_path, filename = "TropicalAndes_speciesLink_occ", basisOfRecord = c('PreservedSpecimen', 'HumanObservation', 'MachineObservation'), country = TA_countries, Coordinates = c("Yes", "Original", "Automatic"), Scope = "plants")
```