### 2023-11-01 archive/L0_NorthAndes_plants_BIEN.Rmd

---
title: "Northern Andes plant species BIEN data"
authors: "Hazel J. Anderson"
collaborators: "None"
data input: "none"
data output: "NorthAndes_BIEN_species_list.csv, NorthAndes_BIEN_occurrence.csv, NorthAndes_BIEN_traits.csv"
project: "Frugivoria"
date: "2023-02-13"
output: html_document
---
# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(BIEN)
library(tidyr)
library(dplyr)
library(DataExplorer)
library(summarytools)
library(bdvis)
library(visdat)
```

# Read in IUCN data
Data was obtained from https://www.iucnredlist.org/search?permalink=9621ab77-2341-4bf8-9557-b7f19f7f2a35
```{r}
plants_habitats <- read.csv(file.path(data_path,"IUCN/Tropical Andes Plants - Search Results/habitats.csv"))
```

# Create species list from IUCN data
```{r}
plant_species <- unique(plants_habitats$scientificName)
```

# List of countries of interest
```{r}
country_vector <- c("Ecuador", "Venezuela", "Peru", "Colombia", "Bolivia")
```

# Get data for Tropical Andes from BIEN database
```{r}
# occurrence data
# old code -- TropicalAndes_BIEN_occ <- BIEN_occurrence_country(country_vector, all.taxonomy = TRUE, collection.info = TRUE, political.boundaries = TRUE, native.status = TRUE)
TropicalAndes_BIEN_occ <- BIEN_occurrence_box(min.lat = -94, max.lat = -54, min.long = -24, max.long = 14,species = plant_species, all.taxonomy = TRUE, political.boundaries = TRUE, source.citation = TRUE, observation.type = TRUE)

# species list
TropicalAndes_BIEN_species_list <- BIEN_list_country(country = country_vector)

# trait data
TropicalAndes_BIEN_traits <- BIEN_trait_country(country_vector, all.taxonomy = TRUE, political.boundaries = TRUE, source.citation = TRUE)
```

# Data summary
```{r}
glimpse(TropicalAndes_BIEN_occ)
```

```{r}
glimpse(TropicalAndes_BIEN_species_list)
```

```{r}
glimpse(TropicalAndes_BIEN_traits)
```

# Write data to csv
```{r}
TropicalAndes_BIEN_occurrence_path <- file.path(output_path,"TropicalAndes_BIEN_occurrence.csv") 
write.csv(TropicalAndes_BIEN_occ, file = TropicalAndes_BIEN_occurrence_path)

TropicalAndes_BIEN_species_list_path <- file.path(output_path, "TropicalAndes_BIEN_species_list.csv")
write.csv(TropicalAndes_BIEN_species_list, file = TropicalAndes_BIEN_species_list_path)

TropicalAndes_BIEN_traits_path <- file.path(output_path, "TropicalAndes_BIEN_traits.csv")
write.csv(TropicalAndes_BIEN_traits, file = TropicalAndes_BIEN_traits_path)
```