### 2023-11-01 code/L2/L2_TropicalAndes_plant_taxcoverage.Rmd

---
title: "Taxonomic coverage for plant data"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script summarize taxonomic coverage and gaps of plant data used."
data input: "ADD"
data output: "ADD"
date: "2023-10-04"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
```

# Load required packages
```{r}
library(dplyr)
```

# Read in Data
```{r}
GBIF_occ <- read.csv(file.path(data_path,"TropicalAndes_GBIF_plant_occ_cleaned.csv"))
GIFT_checklist <- read.csv(file.path(data_path,"TropicalAndes_plant_checklist.csv"))
```

# Plot coverage
GIFT checklist 
species 70376
genera  4381
families 273

GBIF occ
species 46,073
genera 4,885
families 581

```{r}
checklist <- c(70376, 4381, 273)
occ <- c(46073, 4885, 581)
compare_coverage <- data.frame(checklist, occ, row.names = c("species", "genera", "families"))
```

```{r}
barplot(species, df = compare_coverage)
```