### 2023-11-01 archive/L2_NorthAndes_plant_traits.Rmd

---
title: "Ecuador plant trait data merging"
author: "Hazel J. Anderson"
collaborators: "None"
data input: "TBD"
data output: "TBD"
project: "Frugivoria"
date: "2022-11-30"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
```

# Load required packages
```{r}
library(bdc)
```

# Read in the configuration table
```{r}
metadata <-
  readr::read_csv(system.file("ADD FILE PATH HERE",
                              package = "bdc"),
                  show_col_types = FALSE)
```

# Read in BIEN and TRY trait data
```{r}
# Path to the folder containing the example datasets. For instance:
path <- data_path

# Change in the Configuration table the path to the folder in your computer containing the example datasets
metadata$fileName <-
  gsub(pattern = "https://raw.githubusercontent.com/brunobrr/bdc/master/inst/extdata/input_files/",
       replacement = path,
       x = metadata$fileName)
```

```{r}
database <-
bdc_standardize_datasets(metadata = metadata,
                         format = "csv",
                         overwrite = TRUE,
                         save_database = TRUE)
```