### 2023-11-01 code/L1/L1_F3_TropicalAndes_frug_occ.Rmd

---
title: "Frugivore occurrence subset by species with complete trait records"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script subsets frugivore occurrence data to species with complete trait coverage."
data input: "ADD"
data output: "ADD"
date: "2023-10-04"
output: html_document
---

# Set file paths
```{r}
data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
data_path_L1<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(dplyr)
#run if not installed remotes::install_github("FRBCesab/funbiogeo")
library(funbiogeo)
```

# Read in data
```{r}
Frugivoria_subset <- read.csv(file.path(data_path_L1, file = "TropicalAndes_Frugivoria_traits_subset.csv"))
frugivore_occ <- read.csv(file.path(data_path_L1, file = "TropicalAndes_GBIF_frugivore_occ_cleaned.csv"))
```

# Check trait coverage
```{r}
summary(Frugivoria_subset)
```
```{r}
fb_plot_number_species_by_trait(Frugivoria_subset)
```
```{r}
Frugivoria_complete <- fb_filter_species_by_trait_coverage(Frugivoria_subset, threshold_traits_proportion =  1)
```

```{r}
summary(Frugivoria_complete)
```

```{r}
fb_plot_number_species_by_trait(Frugivoria_complete)
```

# Create list of species with complete trait records
```{r}
frugivoria_species <- unique(Frugivoria_complete$IUCN_species_name)
```

# Subset frugivore occurrence data by species list
```{r}
frugivore_occ_subset <- frugivore_occ %>%
  filter(species %in% frugivoria_species)
```

# Summary
```{r}
glimpse(frugivore_occ_subset)
```

```{r}
source("~/GitHub/neotropical_plants/code/L0/data_summary.R")
```

```{r}
data_summary(frugivore_occ_subset, frugivore_occ_subset$species, frugivore_occ_subset$genus, frugivore_occ_subset$family)
```

# Write data to csv
```{r}
write.csv(frugivor_occ_subset, file.path(output_path,"TropicalAndes_GBIF_frugivore_occ_cleaned_subset.csv"))
```