### 2023-11-02 code/L1/L1_P2_TropicalAndes_taxcleaning.Rmd
update file paths and input files, add new count code, update for updated files

---
title: "Tropical Andes plant Lookup Table"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: ""
data input: ""
data output: ""
date: "2023-07-25"
output: html_document
---

# Set file paths
```{r}
data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(knitr)
library(TNRS)
```

# Read in data
```{r}
GBIF_occ <- read.csv(file.path(data_path_L1, "TropicalAndes_GBIF_plant_occ_cleaned.csv"))
TRY_traits <- read.csv(file.path(data_path_L0,"TropicalAndes_TRY_plant_traits.csv"))
BIEN_traits <- read.csv(file.path(data_path_L0,"AllDesired_BIEN_plant_traits.csv"))
GIFT_traits <- read.csv(file.path(data_path_L0,"TropicalAndes_GIFT_plant_traits.csv"))
```

# Generate species lists
```{r}
GBIF_species <- unique(GBIF_occ$species)
TRY_species <- unique(TRY_traits$SpeciesName)
BIEN_species <- unique(BIEN_traits$scrubbed_species_binomial)
GIFT_species <- unique(GIFT_traits$species)
```

```{r}
length(GBIF_species)
length(TRY_species)
length(BIEN_species)
length(GIFT_species)
```

# Using TNRS to get harmonize taxonomy
## GBIF species
```{r}
#split up to speed up api call
GBIF_species1 <-GBIF_species[1:2245]
GBIF_species2 <-GBIF_species[2245:6745]
GBIF_species3 <-GBIF_species[6745:11245]
GBIF_species4 <-GBIF_species[11245:15745]
GBIF_species5 <-GBIF_species[15745:20245]
GBIF_species6 <-GBIF_species[16553:20245]
GBIF_species7 <-GBIF_species[20245:24745]
GBIF_species8 <-GBIF_species[24745:29245]
GBIF_species9 <-GBIF_species[29245:33745]
GBIF_species10 <-GBIF_species[33745:38245]
GBIF_species11 <-GBIF_species[38245:42745]
GBIF_species12 <-GBIF_species[42745:47245]
```

```{r}
GBIF_results1 <- TNRS(taxonomic_names = GBIF_species1)
GBIF_results2 <- TNRS(taxonomic_names = GBIF_species2)
GBIF_results3 <- TNRS(taxonomic_names = GBIF_species3)
GBIF_results4 <- TNRS(taxonomic_names = GBIF_species4)
GBIF_results5 <- TNRS(taxonomic_names = GBIF_species5)
GBIF_results6 <- TNRS(taxonomic_names = GBIF_species6)
GBIF_results7 <- TNRS(taxonomic_names = GBIF_species7)
GBIF_results8 <- TNRS(taxonomic_names = GBIF_species8)
GBIF_results9 <- TNRS(taxonomic_names = GBIF_species9)
GBIF_results10 <- TNRS(taxonomic_names = GBIF_species10)
GBIF_results11 <- TNRS(taxonomic_names = GBIF_species11)
GBIF_results12 <- TNRS(taxonomic_names = GBIF_species12)
```

```{r}
#combine results
GBIF_results <- rbind(GBIF_results1, GBIF_results2, GBIF_results3, GBIF_results4, GBIF_results5, GBIF_results6, GBIF_results7, GBIF_results8, GBIF_results9, GBIF_results10, GBIF_results11, GBIF_results12)
```

## TRY species
```{r}
#split up to speed up api call
TRY_species1 <-TRY_species[1:5000]
TRY_species2 <-TRY_species[5001:10000]
TRY_species3 <-TRY_species[10001:15000]
TRY_species4 <-TRY_species[15001:20000]
TRY_species5 <-TRY_species[20001:25000]
TRY_species6 <-TRY_species[25001:30000]
TRY_species7 <-TRY_species[30001:35000]
TRY_species8 <-TRY_species[35001:40000]
TRY_species9 <-TRY_species[40001:45000]
TRY_species10 <-TRY_species[45001:50000]
TRY_species11 <-TRY_species[50001:55000]
TRY_species12 <-TRY_species[55001:60000]
TRY_species13 <-TRY_species[60001:65000]
TRY_species14 <-TRY_species[65001:70000]
TRY_species15 <-TRY_species[70001:75000]
TRY_species16 <-TRY_species[75001:80000]
TRY_species17 <-TRY_species[80001:85000]
TRY_species18 <-TRY_species[85001:90000]
TRY_species19 <-TRY_species[90001:95000]
TRY_species20 <-TRY_species[95001:100000]
TRY_species21 <-TRY_species[100001:105000]
TRY_species22 <-TRY_species[105001:110000]
TRY_species23 <-TRY_species[110001:115000]
TRY_species24 <-TRY_species[115001:120000]
TRY_species25 <-TRY_species[120001:125000]
TRY_species26 <-TRY_species[125001:130000]
TRY_species27 <-TRY_species[130001:135000]
TRY_species28 <-TRY_species[135001:140000]
TRY_species29 <-TRY_species[140001:145000]
TRY_species30 <-TRY_species[145001:150000]
TRY_species31 <-TRY_species[150001:155000]
TRY_species32 <-TRY_species[155001:160000]
TRY_species33 <-TRY_species[160001:165000]
TRY_species34 <-TRY_species[165001:170000]
TRY_species35 <-TRY_species[170001:175000]
TRY_species36 <-TRY_species[175001:180000]
TRY_species37 <-TRY_species[180001:185000]
TRY_species38 <-TRY_species[185001:190000]
TRY_species39 <-TRY_species[190001:195000]
TRY_species40 <-TRY_species[195001:200000]
TRY_species41 <-TRY_species[200001:205000]
TRY_species42 <-TRY_species[205001:210000]
TRY_species43 <-TRY_species[210001:215000]
TRY_species44 <-TRY_species[215001:220000]
TRY_species45 <-TRY_species[220001:225000]
TRY_species46 <-TRY_species[225001:230000]
TRY_species47 <-TRY_species[230001:235000]
TRY_species48 <-TRY_species[235001:240000]
TRY_species49 <-TRY_species[240001:245000]
TRY_species50 <-TRY_species[245001:250000]
TRY_species51 <-TRY_species[250001:255000]
TRY_species52 <-TRY_species[255001:260000]
TRY_species53 <-TRY_species[260001:265000]
TRY_species54 <-TRY_species[265001:270000]
TRY_species55 <-TRY_species[270001:275000]
TRY_species56 <-TRY_species[275001:280000]
TRY_species57 <-TRY_species[280001:285000]
TRY_species58 <-TRY_species[285001:290000]
TRY_species59 <-TRY_species[290001:295000]
TRY_species60 <-TRY_species[295001:300000]
TRY_species61 <-TRY_species[300001:305000]
TRY_species62 <-TRY_species[305001:310000]
TRY_species63 <-TRY_species[310001:315000]
TRY_species64 <-TRY_species[315001:320000]
TRY_species65 <-TRY_species[320001:325000]
TRY_species66 <-TRY_species[325001:330000]
TRY_species67 <-TRY_species[330001:335000]
TRY_species68 <-TRY_species[335001:340000]
TRY_species69 <-TRY_species[340001:345000]
TRY_species70 <-TRY_species[345001:350000]
TRY_species71 <-TRY_species[350001:355000]
TRY_species72 <-TRY_species[355001:360000]
TRY_species73 <-TRY_species[360001:365000]
TRY_species74 <-TRY_species[365001:370000]
TRY_species75 <-TRY_species[370001:375000]
TRY_species76 <-TRY_species[375001:380000]
TRY_species77 <-TRY_species[380001:385000]
TRY_species78 <-TRY_species[385001:390000]
TRY_species79 <-TRY_species[390001:395000]
TRY_species80 <-TRY_species[395001:400000]
```

```{r}
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results2 <- TNRS(taxonomic_names = TRY_species2)
TRY_results3 <- TNRS(taxonomic_names = TRY_species3)
TRY_results4 <- TNRS(taxonomic_names = TRY_species4)
TRY_results5 <- TNRS(taxonomic_names = TRY_species5)
TRY_results6 <- TNRS(taxonomic_names = TRY_species6)
TRY_results7 <- TNRS(taxonomic_names = TRY_species7)
TRY_results8 <- TNRS(taxonomic_names = TRY_species8)
TRY_results9 <- TNRS(taxonomic_names = TRY_species9)
TRY_results10 <- TNRS(taxonomic_names = TRY_species10)
TRY_results11 <- TNRS(taxonomic_names = TRY_species11)
TRY_results12 <- TNRS(taxonomic_names = TRY_species12)
TRY_results13 <- TNRS(taxonomic_names = TRY_species13)
TRY_results14 <- TNRS(taxonomic_names = TRY_species14)
TRY_results15 <- TNRS(taxonomic_names = TRY_species15)
TRY_results16 <- TNRS(taxonomic_names = TRY_species16)
TRY_results17 <- TNRS(taxonomic_names = TRY_species17)
TRY_results18 <- TNRS(taxonomic_names = TRY_species18)
TRY_results19 <- TNRS(taxonomic_names = TRY_species19)
TRY_results20 <- TNRS(taxonomic_names = TRY_species20)
TRY_results21 <- TNRS(taxonomic_names = TRY_species21)
TRY_results22 <- TNRS(taxonomic_names = TRY_species22)
TRY_results23 <- TNRS(taxonomic_names = TRY_species23)
TRY_results24 <- TNRS(taxonomic_names = TRY_species24)
TRY_results25 <- TNRS(taxonomic_names = TRY_species25)
TRY_results26 <- TNRS(taxonomic_names = TRY_species26)
TRY_results27 <- TNRS(taxonomic_names = TRY_species27)
TRY_results28 <- TNRS(taxonomic_names = TRY_species28)
TRY_results29 <- TNRS(taxonomic_names = TRY_species1)
TRY_results30 <- TNRS(taxonomic_names = TRY_species1)
TRY_results31 <- TNRS(taxonomic_names = TRY_species1)
TRY_results32 <- TNRS(taxonomic_names = TRY_species1)
TRY_results33 <- TNRS(taxonomic_names = TRY_species1)
TRY_results34 <- TNRS(taxonomic_names = TRY_species1)
TRY_results35 <- TNRS(taxonomic_names = TRY_species1)
TRY_results36 <- TNRS(taxonomic_names = TRY_species1)
TRY_results37 <- TNRS(taxonomic_names = TRY_species1)
TRY_results38 <- TNRS(taxonomic_names = TRY_species1)
TRY_results39 <- TNRS(taxonomic_names = TRY_species1)
TRY_results40 <- TNRS(taxonomic_names = TRY_species1)
TRY_results41 <- TNRS(taxonomic_names = TRY_species1)
TRY_results42 <- TNRS(taxonomic_names = TRY_species1)
TRY_results43 <- TNRS(taxonomic_names = TRY_species1)
TRY_results44 <- TNRS(taxonomic_names = TRY_species1)
TRY_results45 <- TNRS(taxonomic_names = TRY_species1)
TRY_results46 <- TNRS(taxonomic_names = TRY_species1)
TRY_results47 <- TNRS(taxonomic_names = TRY_species1)
TRY_results48 <- TNRS(taxonomic_names = TRY_species1)
TRY_results49 <- TNRS(taxonomic_names = TRY_species1)
TRY_results50 <- TNRS(taxonomic_names = TRY_species1)
TRY_results51 <- TNRS(taxonomic_names = TRY_species1)
TRY_results52 <- TNRS(taxonomic_names = TRY_species1)
TRY_results53 <- TNRS(taxonomic_names = TRY_species1)
TRY_results54 <- TNRS(taxonomic_names = TRY_species1)
TRY_results55 <- TNRS(taxonomic_names = TRY_species1)
TRY_results56 <- TNRS(taxonomic_names = TRY_species1)
TRY_results57 <- TNRS(taxonomic_names = TRY_species1)
TRY_results58 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
TRY_results1 <- TNRS(taxonomic_names = TRY_species1)
```

```{r}
#combine results
GBIF_results <- rbind(GBIF_results1, GBIF_results2, GBIF_results3, GBIF_results4, GBIF_results5, GBIF_results6, GBIF_results7, GBIF_results8, GBIF_results9, GBIF_results10, GBIF_results11, GBIF_results12)
```

## BIEN species
```{r}
BIEN_results <- TNRS(taxonomic_names = BIEN_species)
```

## GIFT species
```{r}
GIFT_results <- TNRS(taxonomic_names = GIFT_species)
```

# Save full results
```{r}
write.csv(BIEN_results, file.path(output_path,"BIEN_plant_species_resolved_output_full.csv"))
write.csv(TRY_results, file.path(output_path,"TRY_plant_species_resolved_output_full.csv"))
write.csv(GBIF_results, file.path(output_path,"GBIF_plant_species_resolved_output_full.csv"))
write.csv(GIFT_results, file.path(output_path,"GIFT_plant_species_resolved_output_full.csv"))
```

# Subset results to species with matches, conflicts, and disparties
Name_submitted values not the same as Accepted_name
```{r}
GBIF_matches <- subset(GBIF_results, Taxonomic_status == "Accepted")
nrow(GBIF_matches)
GBIF_conflicts <- subset(GBIF_results, Taxonomic_status == "Synonym")
nrow(GBIF_conflicts)
GBIF_disparities <- subset(GBIF_results, Taxonomic_status %in% c("No opinion", "Illegitimate", "Invalid", ""))
nrow(GBIF_disparities)
```

```{r}
TRY_matches <- subset(TRY_results, Taxonomic_status == "Accepted")
nrow(TRY_matches)
TRY_conflicts <- subset(TRY_results, Taxonomic_status == "Synonym")
nrow(TRY_conflicts)
TRY_disparities <- subset(TRY_results, Taxonomic_status %in% c("No opinion", "Illegitimate", "Invalid", ""))
nrow(TRY_disparities)
```

```{r}
BIEN_matches <- subset(BIEN_results, Taxonomic_status == "Accepted")
nrow(BIEN_matches)
BIEN_conflicts <- subset(BIEN_results, Taxonomic_status == "Synonym")
nrow(BIEN_conflicts)
BIEN_disparities <- subset(BIEN_results, Taxonomic_status %in% c("No opinion", "Illegitimate", "Invalid", ""))
nrow(BIEN_disparities)
```

```{r}
GIFT_matches <- subset(GIFT_results, Taxonomic_status == "Accepted")
nrow(GIFT_matches)
GIFT_conflicts <- subset(GIFT_results, Taxonomic_status == "Synonym")
nrow(GIFT_conflicts)
GIFT_disparities <- subset(GIFT_results, Taxonomic_status %in% c("No opinion", "Illegitimate", "Invalid", ""))
nrow(GIFT_disparities)
```

# Checking 
```{r}

```