### 2023-11-08 code/L1/L1_P2_TropicalAndes_taxcleaning.Rmd
update packages, combine species list to simplify harmonization, remove NAs from species list, add code to check common scientific name issues, add bdc to harmonize species that TNRS had disparties with, create lookup table

---
title: "Tropical Andes plant Lookup Table"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: ""
data input: ""
data output: ""
date: "2023-07-25"
output: html_document
---

# Load required packages
```{r}
library(knitr)
library(TNRS)
library(bdc)
library(dplyr)
```

# Set file paths
```{r}
data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Read in data
```{r}
GBIF_occ <- read.csv(file.path(data_path_L1, "TropicalAndes_GBIF_plant_occ_cleaned.csv"))
TRY_traits <- read.csv(file.path(data_path_L0,"TropicalAndes_TRY_plant_traits.csv"))
BIEN_traits <- read.csv(file.path(data_path_L0,"AllDesired_BIEN_plant_traits.csv"))
GIFT_traits <- read.csv(file.path(data_path_L0,"TropicalAndes_GIFT_plant_traits.csv"))
```

# Generate species lists
```{r}
GBIF_species <- unique(GBIF_occ$species)
TRY_species <- unique(TRY_traits$SpeciesName)
BIEN_species <- unique(BIEN_traits$scrubbed_species_binomial)
GIFT_species <- unique(GIFT_traits$species)
```

```{r}
length(GBIF_species)
length(TRY_species)
length(BIEN_species)
length(GIFT_species)
```
## combine species list together
```{r}
all_species <- c(GBIF_species, TRY_species, BIEN_species, GIFT_species)
all_species <- unique(all_species)
length(all_species)
```
# Remove NAs
```{r}
all_species <- all_species[!is.na(all_species)]
length(all_species)
```

# Use Rocc to check scientific names 
Adapted from tutorial https://liibre.github.io/Rocc/articles/articles/check_taxa.html
```{r}
name_check_results <- bdc_clean_names(sci_names = all_species, save_outputs = FALSE)
```
```{r}
all_species_cleaned <- name_check_results$names_clean
```

# Using TNRS to get harmonize taxonomy

```{r}
all_species_hamonized_results <- TNRS(taxonomic_names = all_species_cleaned)
```

# Subset TNRS results to species with matches, conflicts, and disparties
```{r}
unique(all_species_hamonized_results$Taxonomic_status)
```

```{r}
all_species_matches <- subset(all_species_hamonized_results, Taxonomic_status == "Accepted")
nrow(all_species_matches)
all_species_conflicts <- subset(all_species_hamonized_results, Taxonomic_status == "Synonym")
nrow(all_species_conflicts)

# Disparties
all_species_no.opinion <- subset(all_species_hamonized_results, Taxonomic_status == "No opinion")
nrow(all_species_no.opinion)
all_species_illegitimate <- subset(all_species_hamonized_results, Taxonomic_status == "Illegitimate")
nrow(all_species_illegitimate)
all_species_invalid <- subset(all_species_hamonized_results, Taxonomic_status == "Invalid")
nrow(all_species_invalid)
all_species_other <- subset(all_species_hamonized_results, Taxonomic_status == "")
nrow(all_species_other)
```
# Create species list of names with disparities
```{r}
all_species_disparities <- all_species_hamonized_results %>%
                            filter(Taxonomic_status %in% c("No opinion", "Illegitimate", "Invalid", "")) %>%
                            select(Name_submitted)
```

# Use bdc to harmonize disparities
## Using ITIS database
```{r}
all_species_disparities_col_results <- bdc_query_names_taxadb(sci_name = all_species_disparities$Name_submitted, replace_synonyms = TRUE,suggest_names = TRUE, db = "col", rank_name = "Plantae", rank = "kingdom")
```

## Subset COL results to accepted, synonym, and NA

```{r}
unique(all_species_disparities_col_results$taxonomicStatus)
```

```{r}
all_species_col_accepted <- subset(all_species_disparities_col_results, taxonomicStatus == "accepted")
nrow(all_species_col_accepted)
all_species_col_synonym <- subset(all_species_disparities_col_results, taxonomicStatus == "synonym")
nrow(all_species_col_synonym)
all_species_col_NA <- subset(all_species_disparities_col_results, is.na(taxonomicStatus))
nrow(all_species_col_NA)
```

# Create lookup table
## From TNRS
```{r}
species_matches <- all_species_matches[ , c("Name_submitted", "Name_matched", "Source")]
species_corrected <- all_species_conflicts[ , c("Name_submitted", "Name_matched", "Source")]
```

## From bdc using COL
```{r}
species_col_accepted <- all_species_col_accepted[ , c("original_search", "suggested_name")]
species_col_accepted$source <- c("col")
species_col_synonmyn <- all_species_col_synonym[ , c("original_search", "suggested_name")]
species_col_synonmyn$source <- c("col")
```

```{r}
# combine to create lookup table
lookup_table <- rbind(species_matches, species_corrected)
```

# Save lookup table as csv
```{r}

```

# Update data files species names using lookup table
```{r}

```

# Write data to csv
```{r}

```