### 2023-11-15 code/L1/L1_P3_TropicalAndes_plant_traits.Rmd
add code that parrallels L1_F2_TropicalAndes_frugivore_traits

---
title: "TITLE"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script subsets plant trait data by synonyms species list and combine."
data input: "ADD"
data output: "ADD"
date: "2023-11-04"
output: html_document
---

# Load required packages
```{r}
library(dplyr)
```

# Set file paths
```{r}
data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Read in data
```{r}
plant_occ <- read.csv(file.path(data_path_L1, "TropicalAndes_GBIF_plant_occ_cleaned.csv"))
TRY_traits <- read.csv(file.path(data_path_L0,"TropicalAndes_TRY_plant_traits.csv"))
BIEN_traits <- read.csv(file.path(data_path_L0,"AllDesired_BIEN_plant_traits.csv"))
GIFT_traits <- read.csv(file.path(data_path_L0,"TropicalAndes_GIFT_plant_traits.csv"))
```

# Get species list from plant_occ data
```{r}
plant_species <- unique(plant_occ$species)
```

# Subset TRY, BIEN, and GIFT by species list
## TRY
```{r}
TRY_filtered <- TRY_traits %>%
  filter(SpeciesName %in% plant_species)
```

## BIEN
```{r}
BIEN_filtered <- BIEN_traits %>%
  filter(scrubbed_species_binomial %in% plant_species)
```

## GIFT
```{r}
GIFT_filtered <- GIFT_traits %>%
  filter(species %in% plant_species)
```