### 2023-12-06 code/L1/L1_P3_TropicalAndes_plant_traits.Rmd
update input filepaths, add row count code, add code to combine trait data, start reclassification code

---
title: "TITLE"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script subsets plant trait data by synonyms species list and combine."
data input: "ADD"
data output: "ADD"
date: "2023-11-04"
output: html_document
---

# Load required packages
```{r}
library(dplyr)
```

# Set file paths
```{r}
data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Read in data
```{r}
plant_occ <- read.csv(file.path(data_path_L1, "TropicalAndes_GBIF_plant_occ_harmonized.csv"))
TRY_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_TRY_plant_traits_harmonized.csv"))
BIEN_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_BIEN_plant_traits_harmonized.csv"))
GIFT_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_GIFT_plant_traits_harmonized.csv"))
```

# Get species list from plant_occ data
```{r}
plant_species <- unique(plant_occ$Accepted_species)
length(plant_species)
```

# Subset TRY, BIEN, and GIFT by species list
## TRY
```{r}
nrow(TRY_traits)
TRY_filtered <- TRY_traits %>%
  filter(Accepted_species %in% plant_species)
nrow(TRY_filtered)
```

## BIEN
```{r}
nrow(BIEN_traits)
BIEN_filtered <- BIEN_traits %>%
  filter(Accepted_species %in% plant_species)
nrow(BIEN_filtered)
```

## GIFT
```{r}
nrow(GIFT_traits)
GIFT_filtered <- GIFT_traits %>%
  filter(Accepted_species %in% plant_species)
nrow(GIFT_filtered)
```

# Combine trait information from TRY, BIEN, GIFT

```{r}
TRYTraitTypes <- TRY_filtered %>%
  group_by(TraitName) %>%
  distinct(DataName)
```

```{r}
colnames(TRY_filtered)
```
columns to keep : Accepted_species, TraitName, OrigValueStr

```{r}
TRY_filtered_subset <- TRY_filtered[ , c("Accepted_species", "TraitName", "OrigValueStr", "UnitName")]
colnames(TRY_filtered_subset) <- c("Accepted_species", "TraitName", "TraitValue", "Unit")
```

```{r}
colnames(BIEN_filtered)
```
columns to keep : Accepted_species, trait_name, trait_value, unit

```{r}
BIEN_filtered_subset <- BIEN_filtered[ , c("Accepted_species", "trait_name", "trait_value", "unit")]
colnames(BIEN_filtered_subset) <- c("Accepted_species", "TraitName", "TraitValue", "Unit")
```

```{r}
colnames(GIFT_filtered)
```
columns to keep : Accepted_species, trait_value, trait_name

```{r}
GIFT_filtered_subset <- GIFT_filtered[ , c("Accepted_species", "trait_name", "trait_value")]
colnames(GIFT_filtered_subset) <- c("Accepted_species", "TraitName", "TraitValue")
```

Add Unit column to GIFT dataframe
```{r}
GIFT_filtered_subset <- GIFT_filtered_subset %>%
  mutate(Unit = case_when(
  TraitName == "plant_height" ~ "m",
  TraitName == "plant_height_max" ~ "m",
  TraitName == "plant_height_min" ~ "m",
  TraitName == "seed_mass_mean" ~ "g",
  TraitName == "seed_mass_min" ~ "g",
  TraitName == "seed_mass_max" ~ "g",
  TraitName == "fruit_length_min" ~ "cm",
  TraitName == "fruit_length_max" ~ "cm",
  TraitName == "fruit_length_mean" ~ "cm",
  TraitName == "seed_length_max" ~ "mm",
  TraitName == "seed_length_min" ~ "mm",
  TraitName == "seed_length_mean" ~ "mm",
  TraitName == "seed_width_max" ~ "mm",
  TraitName == "seed_width_min" ~ "mm",
  TraitName == "seed_width_mean" ~ "mm",
  TraitName == "plant_lifespan" ~ "years",
  .default = ""
  ))
```

```{r}
traits <- rbind(TRY_filtered_subset, BIEN_filtered_subset, GIFT_filtered_subset)
```

```{r}
traits %>%
  count(TraitName)
```

# Standardize trait measurements
## Dispersal syndrome
```{r}
traits_dispersal <- traits %>%
  filter(grepl('Dispersal|dispersal', TraitName))
traits_dispersal$OrigTraitValue <- traits_dispersal$TraitValue
```

Reclassify to the following:
anemochorous - wind dispersal
anthropochorous - human dispersal
autochorous - non-aided dispersal
hydrochorous - water dispersal
myrmecochorous - ant dispersal
zoochorous - animal dispersal
unspecialized - multiple
```{r}
traits_dispersal <- traits_dispersal %>%
  filter() %>%
  mutate(TraitValue= case_when(
    grepl('Eaten|Animal|Bird|Mammal|Monkey', OrigTraitValue, ignore.case = TRUE) ~ "zoochorous",
    grepl('Ants', OrigTraitValue, ignore.case = TRUE) ~ "myrmecochorous",
    grepl('wind|Anemochory', OrigTraitValue, ignore.case = TRUE) ~ "anemochorous",
    grepl('human', OrigTraitValue, ignore.case = TRUE) ~ "anthropochorous",
    grepl('No', OrigTraitValue, ignore.case = TRUE) ~ "autochorous"
  ))
```

```{r}
print(unique(traits_dispersal$TraitValue))
```