### 2024-03-07 code/L1/L1_P3_TropicalAndes_plant_traits.Rmd
remove rows without accepted species names, create functions to average trait measurements (get one trait measure per species)

---
title: "TITLE"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script subsets plant trait data by synonyms species list and combine."
data input: "ADD"
data output: "ADD"
date: "2023-11-04"
output: html_document
---

# Load required packages
```{r}
library(dplyr)
```

# Set file paths
```{r}
data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Read in data
```{r}
plant_occ <- read.csv(file.path(data_path_L1, "TropicalAndes_GBIF_plant_occ_harmonized.csv"))
TRY_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_TRY_plant_traits_harmonized.csv"))
BIEN_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_BIEN_plant_traits_harmonized.csv"))
GIFT_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_GIFT_plant_traits_harmonized.csv"))
```

# Get species list from plant_occ data
```{r}
plant_species <- unique(plant_occ$Accepted_species)
length(plant_species)
```

# Subset TRY, BIEN, and GIFT by species list
## TRY
```{r}
nrow(TRY_traits)
TRY_filtered <- TRY_traits %>%
  filter(Accepted_species %in% plant_species)
nrow(TRY_filtered)
```

## BIEN
```{r}
nrow(BIEN_traits)
BIEN_filtered <- BIEN_traits %>%
  filter(Accepted_species %in% plant_species)
nrow(BIEN_filtered)
```

## GIFT
```{r}
nrow(GIFT_traits)
GIFT_filtered <- GIFT_traits %>%
  filter(Accepted_species %in% plant_species)
nrow(GIFT_filtered)
```

# Combine trait information from TRY, BIEN, GIFT

```{r}
TRYTraitTypes <- TRY_filtered %>%
  group_by(TraitName) %>%
  distinct(DataName)
```

```{r}
colnames(TRY_filtered)
```
columns to keep : Accepted_species, TraitName, OrigValueStr

```{r}
TRY_filtered_subset <- TRY_filtered[ , c("Accepted_species", "TraitName", "OrigValueStr", "UnitName")]
colnames(TRY_filtered_subset) <- c("Accepted_species", "TraitName", "TraitValue", "Unit")
```

```{r}
colnames(BIEN_filtered)
```
columns to keep : Accepted_species, trait_name, trait_value, unit

```{r}
BIEN_filtered_subset <- BIEN_filtered[ , c("Accepted_species", "trait_name", "trait_value", "unit")]
colnames(BIEN_filtered_subset) <- c("Accepted_species", "TraitName", "TraitValue", "Unit")
```

```{r}
colnames(GIFT_filtered)
```
columns to keep : Accepted_species, trait_value, trait_name

```{r}
GIFT_filtered_subset <- GIFT_filtered[ , c("Accepted_species", "trait_name", "trait_value")]
colnames(GIFT_filtered_subset) <- c("Accepted_species", "TraitName", "TraitValue")
```

Add Unit column to GIFT dataframe
```{r}
GIFT_filtered_subset <- GIFT_filtered_subset %>%
  mutate(Unit = case_when(
  TraitName == "plant_height" ~ "m",
  TraitName == "plant_height_max" ~ "m",
  TraitName == "plant_height_min" ~ "m",
  TraitName == "seed_mass_mean" ~ "g",
  TraitName == "seed_mass_min" ~ "g",
  TraitName == "seed_mass_max" ~ "g",
  TraitName == "fruit_length_min" ~ "cm",
  TraitName == "fruit_length_max" ~ "cm",
  TraitName == "fruit_length_mean" ~ "cm",
  TraitName == "seed_length_max" ~ "mm",
  TraitName == "seed_length_min" ~ "mm",
  TraitName == "seed_length_mean" ~ "mm",
  TraitName == "seed_width_max" ~ "mm",
  TraitName == "seed_width_min" ~ "mm",
  TraitName == "seed_width_mean" ~ "mm",
  TraitName == "plant_lifespan" ~ "years",
  .default = ""
  ))
```

```{r}
traits <- rbind(TRY_filtered_subset, BIEN_filtered_subset, GIFT_filtered_subset)
nrow(traits)
```

```{r}
# remove rows without accepted species names
traits_clean <- traits[!(is.na(traits$Accepted_species) | traits$Accepted_species==""), ]
nrow(traits_clean)
```

```{r}
traits %>%
  count(TraitName)
```

# Standardize trait measurements
## Create functions
```{r}
# Numeric data
numeric_data_average <- function(data, unit, traitname) {
# remove rows if TraitValue is non-numeric
data <- data %>%
  filter(!is.na(as.numeric(TraitValue))) %>%
  mutate(TraitValue = as.numeric(TraitValue))
# remove duplicate rows
data <- data %>%
  distinct(.keep_all = TRUE)
# calculate average value per trait per species
average_data <- aggregate(TraitValue ~ Accepted_species, data, FUN = mean, na.rm = TRUE)
# Remove rows with missing or empty species names
average_data <- average_data[!is.na(average_data$Accepted_species) & average_data$Accepted_species !="", ]
# Add unit & trait column
average_data$Unit <- unit
average_data$TraitName <- traitname

return(average_data)
}
```

```{r}
# categorical data
factor_data_merge <- function(data, traitname) {
# remove duplicate rows
data <- data %>%
  distinct(.keep_all = TRUE)
# remove rows with missing or empty species names
data <- data[!is.na(data$Accepted_species) & data$Accepted_species != "", ]
# Subset data to just species and TraitValue
data <- data[,c("Accepted_species", "TraitValue")]
# Find the most common trait value for each species
  most_common_values <- aggregate(. ~ Accepted_species, data, function(x) {
    # Find the mode (most common value) of the trait values
    mode_value <- names(sort(table(x), decreasing = TRUE))[1]
    return(mode_value)
  })
# Merge most common trait values back to the original data
merged_data <- merge(data, most_common_values, by = "Accepted_species", all.x = TRUE)

# Rename the merged column
colnames(merged_data)[ncol(merged_data)] <- "TraitValue"

# Remove column TraitValue.x
merged_data <- merged_data[, !colnames(merged_data) %in% "TraitValue.x", drop = FALSE]

# remove duplicate rows
merged_data <- merged_data %>%
  distinct(.keep_all = TRUE)

# Add unit & trait column
merged_data$TraitName <- traitname

  return(merged_data)
}
```

## Dispersal syndrome
```{r}
traits_dispersal <- traits %>%
  filter(grepl('Dispersal syndrome|whole plant dispersal syndrome|dispersal_syndrome_1|dispersal_syndrome_2', TraitName))
traits_dispersal$OrigTraitValue <- traits_dispersal$TraitValue
```

```{r}
# remove rows containing weight or mass
traits_dispersal <- traits_dispersal[!grepl("weight|mass|air-dry|oven|mean|SE",traits_dispersal$TraitValue, ignore.case = TRUE),]
nrow(traits_dispersal)
length(unique(traits_dispersal$Accepted_species))
```

Reclassify to the following:
anemochorous - wind dispersal
anthropochorous - human dispersal
autochorous - non-aided dispersal
hydrochorous - water dispersal
myrmecochorous - ant dispersal
zoochorous - animal dispersal
unspecialized - multiple
```{r}
traits_dispersal <- traits_dispersal %>%
  filter() %>%
  mutate(TraitValue= case_when(
    grepl('Eaten|Animal|Bird|Mammal|animals|endozoochor|dysochor|epizoochor|zoochor|Vertebrates|Reptiles+birds|Endo-zoochory|mammals|exozoochory|birds|Zoochory|vertebrate|Dys-zoochory|Ornithochory|Epi-zoochory|endozoochorous|endozoochory|ectozoochorous|zoochorous', OrigTraitValue, ignore.case = TRUE) ~ "zoochorous",
    grepl('Ants', OrigTraitValue, ignore.case = TRUE) ~ "myrmecochorous",
    grepl('wind|Anemochory|anemochor|meteorochor|chamaechor|Anemochory|anemochorous', OrigTraitValue, ignore.case = TRUE) ~ "anemochorous",
    grepl('human|agochor|man|hemerochor|anthropochorous', OrigTraitValue, ignore.case = TRUE) ~ "anthropochorous",
    grepl('No|Unassisted|autochor|blastochor|ombrochor|Autochory|Barochory|ballochor|herpochor|autochorous', OrigTraitValue, ignore.case = TRUE) ~ "autochorous",
    grepl('water|nautochor|floating|hydrochor|hydrochorous', OrigTraitValue, ignore.case = TRUE) ~ "hydrochorous",
    grepl('boleochor|combination|non specialized|unspecialized', OrigTraitValue, ignore.case = TRUE) ~ "unspecialized"
  )) %>%
  filter(!is.na(TraitValue))
nrow(traits_dispersal)
length(unique(traits_dispersal$Accepted_species))
```

```{r}
print(unique(traits_dispersal$TraitValue))
```

```{r}
traits_dispersal <- traits_dispersal %>%
  mutate(TraitName = "dispersal_syndrome")
```

```{r}
averagetraits_dispersal <- factor_data_merge(traits_dispersal, "Dispersal Syndrome")
```

## Fruit dry mass
```{r}
traits_fruitMass <- traits %>%
  filter(grepl('Fruit dry mass', TraitName))
traits_fruitMass$OrigTraitValue <- traits_fruitMass$TraitValue
```

```{r}
print(unique(traits_fruitMass$Unit))
```

```{r}
averagetraits_fruitMass <- numeric_data_average(traits_fruitMass, "mg", "Fruit Mass")
```

## Fruit length
```{r}
traits_fruitLength <- traits %>%
  filter(grepl('Fruit length|maximum fruit length|minimum fruit length|fruit_length_min|fruit_length_max|fruit_length_mean', TraitName))
traits_fruitLength$OrigTraitValue <- traits_fruitLength$TraitValue
```

```{r}
print(unique(traits_fruitLength$Unit))
```

```{r}
# remove rows if TraitValue is non-numeric
traits_fruitLength <- traits_fruitLength %>%
  filter(!is.na(as.numeric(TraitValue))) %>%
  mutate(TraitValue = as.numeric(TraitValue))
```

```{r}
# if units = cm, multiply trait value by 10 and change units to mm
traits_fruitLength <- traits_fruitLength %>%
  mutate(TraitValue = if_else(Unit == "cm", TraitValue * 10, TraitValue),
         Unit = if_else(Unit == "cm", "mm", Unit))
```

```{r}
averagetraits_fruitLength <- numeric_data_average(traits_fruitLength, "mm", "Fruit Length")
```

## Plant height
```{r}
traits_plantHeight <- traits %>%
  filter(grepl('Plant height generative|Plant height vegetative|maximum whole plant height|minimum whole plant height|plant_height|plant_height_max|plant_height_min|whole plant height', TraitName))
traits_plantHeight$OrigTraitValue <- traits_plantHeight$TraitValue
```

```{r}
print(unique(traits_plantHeight$Unit))
```

```{r}
averagetraits_plantHeight <- numeric_data_average(traits_plantHeight, "m", "Plant Height")
```

## Plant lifespan/longevity
```{r}
traits_plantLifespan <- traits %>%
  filter(grepl('Plant lifespan (longevity)|longest whole plant longevity|maximum whole plant longevity|plant_lifespan', TraitName))
traits_plantLifespan$OrigTraitValue <- traits_plantLifespan$TraitValue
```

```{r}
print(unique(traits_plantLifespan$Unit))
```

```{r}
averagetraits_plantLifespan <- numeric_data_average(traits_plantLifespan, "years", "Plant Lifespan")
```

## Fruit type
```{r}
traits_fruitType <- traits %>%
  filter(grepl('Fruit dehiscence type|Fruit surface type|Fruit texture inside (fleshy, leathrey, …)|Fruit type|fruit type|fruit_dryness|fruit_type', TraitName))
traits_fruitType$OrigTraitValue <- traits_fruitType$TraitValue
```

```{r}
print(unique(traits_fruitType$Unit))
```

```{r}
averagetraits_fruitType <- factor_data_merge(traits_fruitType, "Fruit Type")
```

## Seed mass
```{r}
traits_seedMass <- traits %>%
  filter(grepl('Seed dry mass|seed mass|seed_mass_max|seed_mass_mean|seed_mass_min', TraitName))
traits_seedMass$OrigTraitValue <- traits_seedMass$TraitValue
```

```{r}
print(unique(traits_seedMass$Unit))
```

```{r}
# remove rows if TraitValue is non-numeric
traits_seedMass <- traits_seedMass %>%
  filter(!is.na(as.numeric(TraitValue))) %>%
  mutate(TraitValue = as.numeric(TraitValue))
```

```{r}
# convert rows with unit = g to mg; multiply by 1000
traits_seedMass <- traits_seedMass %>%
  mutate(TraitValue = if_else(Unit == "g", TraitValue * 1000, TraitValue),
         Unit = if_else(Unit == "g", "mg", Unit))
```

```{r}
averagetraits_seedMass <- numeric_data_average(traits_seedMass, "mg", "Seed Mass")
```

## Seed length
```{r}
traits_seedLength <- traits %>%
  filter(grepl('Seed length|seed_length_max|seed_length_mean|seed_length_min', TraitName))
traits_seedLength$OrigTraitValue <- traits_seedLength$TraitValue
```

```{r}
print(unique(traits_seedLength$Unit))
```

```{r}
averagetraits_seedLength <- numeric_data_average(traits_seedLength, "mm", "Seed Length")
```

## Seed width
```{r}
traits_seedWidth <- traits %>%
  filter(grepl('Seed width|seed_width_max|seed_width_mean|seed_width_min'
, TraitName))
traits_seedWidth$OrigTraitValue <- traits_seedWidth$TraitValue
```

```{r}
print(unique(traits_seedWidth$Unit))
```

```{r}
averagetraits_seedWidth <- numeric_data_average(traits_seedWidth, "mm", "Seed Width")
```

## Growth form
```{r}
traits_growthForm <- traits %>%
  filter(grepl('Plant growth form|growth_form|whole plant growth form|whole plant growth form diversity'
, TraitName))
traits_growthForm$OrigTraitValue <- traits_growthForm$TraitValue
```

```{r}
averagetraits_growthForm <- factor_data_merge(traits_growthForm, "Growth Form")
```