### 2024-04-18 code/L1/L1_P4_TropicalAndes_plant_imputetraits.Rmd
add code for imputation

---
title: "TITLE"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script fill plant trait gaps with imputation."
data input: "ADD"
data output: "ADD"
date: "2023-10-04"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(tidyr)
library(mice)
library(dplyr)
```

# Read in data
```{r}
plant_traits <- read.csv(file.path(data_path,"TropicalAndes_all_plant_traits_harmonized.csv"))
```

# Set the seed for reproducibility
```{r}
set.seed(123)
```

```{r}
#replace " " with _ in column names
names(wide_traits) <- sub(" ", "_", names(wide_traits))
```

# show the missing data pattern
```{r}
md.pattern(wide_traits)
```

# Perform the imputation
```{r}
imp_model <- mice(wide_traits, method = "cart", maxit = 20)
```

```{r}
imputed_data <- complete(imp_model)
```