### 2024-04-25 code/L1/L1_P4_TropicalAndes_plant_imputetraits.Rmd
add code to use BIEN and GIFT to fill in gaps by genus or family

---
title: "Plant Trait Imputation"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script fill plant trait gaps with imputation."
data input: "ADD"
data output: "ADD"
date: "2023-10-04"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(tidyr)
library(BIEN)
library(GIFT)
library(mice)
library(dplyr)
```

# Read in data
```{r}
plant_traits <- read.csv(file.path(data_path,"TropicalAndes_all_plant_traits_harmonized.csv"))
```

# Use BIEN and GIFT to fill in gaps by genus or family.
Need a list of families and genus for each trait that needs requesting
2. add genus and family information to dataframe
3. list by trait of species and genus
## Create a dataframe of species and TraitNames with TraitValue = NA
```{r}
# remove X column
plant_traits <- plant_traits[, !colnames(plant_traits) %in% "X", drop = FALSE]
# Convert the wide dataframe to long format
long_plant_traits <- plant_traits %>%
  gather(TraitName, TraitValue, -species)

# Filter rows with NA values
na_traits <- long_plant_traits %>%
  filter(is.na(TraitValue))

# get a list of species
na_species_list <- unique(na_traits$species)
```

## Add genus and family information to dataframe
```{r}
taxonomic_info <- BIEN_taxonomy_species(na_species_list)
```

```{r}
# keep only columns with family, genus, species
subset_df <- taxonomic_info[, c("scrubbed_family", "scrubbed_genus", "scrubbed_species_binomial")]
# remove duplicates
na_species_taxonomy <- distinct(subset_df)
# rename column to family, genus, species
na_species_taxonomy <- na_species_taxonomy %>%
  rename(family = scrubbed_family,
         genus = scrubbed_genus,
         species = scrubbed_species_binomial)
```

```{r}
# add to species dataframe
na_species_df <- as.data.frame(na_species_list)
names(na_species_df) <- "species"
na_species_df <- merge(na_species_df, na_species_taxonomy, by = "species", all.x = TRUE)

# retrieve list of species without family & genus information
species_no_family <- na_species_df %>%
  filter(is.na(family) | family == "") %>%
  select(species) %>%
  pull()
species_no_genus <- na_species_df %>%
  filter(is.na(genus) | genus == "") %>%
  select(species) %>%
  pull()
# check if all the species in no_family & no_genus lists are the same
identical(species_no_family, species_no_genus)
```
## Use taxize for species with no family and genus info

```{r}
library(taxize)

# Function to retrieve taxonomic information for a chunk of species names
get_taxonomic_info_chunk <- function(chunk_species_names) {
  # Initialize an empty list to store taxonomic information for each chunk
  chunk_taxonomic_info <- list()

  # Loop through each species name in the chunk and retrieve taxonomic information
  for (species_name in chunk_species_names) {
    # Add a delay of 1 second between consecutive API requests
    Sys.sleep(1)

    # Make API request to retrieve taxonomic information
    taxon_info <- tax_name(species_name, get = c("genus", "family"), db = "both")

    # Store taxonomic information for the species in the list
    chunk_taxonomic_info[[species_name]] <- data.frame(Species = species_name, 
                                                       Genus = taxon_info$genus, 
                                                       Family = taxon_info$family)
  }

  # Return the list of taxonomic information for the chunk
  return(chunk_taxonomic_info)
}

# Define the chunk size
chunk_size <- 100  # Adjust the chunk size as needed

# Split the dataframe into chunks based on the chunk size
chunks <- split(species_no_family, ceiling(seq_along(species_no_family) / chunk_size))

# Initialize an empty list to store taxonomic information for all chunks
all_taxonomic_info <- list()

# Iterate over each chunk and retrieve taxonomic information
for (chunk_index in seq_along(chunks)) {
  # Get the species names in the current chunk
  chunk_species_names <- chunks[[chunk_index]]

  # Retrieve taxonomic information for the chunk
  chunk_taxonomic_info <- get_taxonomic_info_chunk(chunk_species_names)

  # Append taxonomic information for the chunk to the list
  all_taxonomic_info <- c(all_taxonomic_info, list(chunk_taxonomic_info))
}

# Combine taxonomic information from all chunks into a single list
all_taxonomic_info <- do.call(c, all_taxonomic_info)

# Combine the list of dataframes into a single dataframe
taxonomic_df <- do.call(rbind, all_taxonomic_info)
```

# Set the seed for reproducibility
```{r}
set.seed(123)
```

```{r}
#replace " " with _ in column names
names(plant_traits) <- sub(" ", "_", names(plant_traits))
```

# show the missing data pattern
```{r}
md.pattern(plant_traits)
```
# Trying to prevent error
Error in edit.setup(data, setup, ...) : 
  `mice` detected constant and/or collinear variables. No predictors were left after their removal.
```{r}
# Subset to numeric plant traits
num_plant_traits <- plant_traits %>%
  select_if(is.numeric)
# Check for constant variables
constant_vars <- sapply(num_plant_traits, function(x) length(unique(x))) == 1
constant_vars <- names(constant_vars[constant_vars])

# Remove constant variables
num_plant_traits <- num_plant_traits[, !names(num_plant_traits) %in% constant_vars]

# Check for collinearity
cor_matrix <- cor(num_plant_traits)

# Find highly correlated variables
highly_correlated_pairs <- which(abs(cor_matrix) > 0.8 & upper.tri(cor_matrix, diag = TRUE), arr.ind = TRUE)

# Print highly correlated pairs
print(highly_correlated_pairs)
```

# Perform the imputation
```{r}
imp_model <- mice(plant_traits, method = "cart", maxit = 20)
```

```{r}
imputed_data <- complete(imp_model)
```