### 2024-05-01 code/L1/L1_P4_TropicalAndes_plant_imputetraits.Rmd
add genus and family info to na_species dataframes, get species with no family info and retrive that info from powo and add it to the dataframe, combine family and genus info to dataframe with species and traits, create a list of unique genus and families for each trait to pull from BIEN and GIFT

---
title: "Plant Trait Imputation"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script fill plant trait gaps with imputation."
data input: "ADD"
data output: "ADD"
date: "2023-10-04"
output: html_document
---

# Set file paths
```{r}
data_path<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path<- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(tidyr)
library(BIEN)
library(GIFT)
library(mice)
library(dplyr)
```

# Read in data
```{r}
plant_traits <- read.csv(file.path(data_path,"TropicalAndes_all_plant_traits_harmonized.csv"))
```

# Use BIEN and GIFT to fill in gaps by genus or family.
Need a list of families and genus for each trait that needs requesting
2. add genus and family information to dataframe
3. list by trait of species and genus

## Create a dataframe of species and TraitNames with TraitValue = NA
```{r}
# remove X column
plant_traits <- plant_traits[, !colnames(plant_traits) %in% "X", drop = FALSE]
# Convert the wide dataframe to long format
long_plant_traits <- plant_traits %>%
  gather(TraitName, TraitValue, -species)

# Filter rows with NA values
na_traits <- long_plant_traits %>%
  filter(is.na(TraitValue))

# get a list of species
na_species_list <- unique(na_traits$species)
```

## Add genus and family information to dataframe
```{r}
taxonomic_info <- BIEN_taxonomy_species(na_species_list)
```

```{r}
# keep only columns with family, genus, species
subset_df <- taxonomic_info[, c("scrubbed_family", "scrubbed_genus", "scrubbed_species_binomial")]
# remove duplicates
na_species_taxonomy <- distinct(subset_df)
# rename column to family, genus, species
na_species_taxonomy <- na_species_taxonomy %>%
  rename(family = scrubbed_family,
         genus = scrubbed_genus,
         species = scrubbed_species_binomial)
```

```{r}
# add to species dataframe
na_species_df <- as.data.frame(na_species_list)
names(na_species_df) <- "species"
na_species_df <- merge(na_species_df, na_species_taxonomy, by = "species", all.x = TRUE)

# retrieve list of species without family & genus information
species_no_family <- na_species_df %>%
  filter(is.na(family) | family == "") %>%
  select(species) %>%
  pull()
species_no_genus <- na_species_df %>%
  filter(is.na(genus) | genus == "") %>%
  select(species) %>%
  pull()
# check if all the species in no_family & no_genus lists are the same
identical(species_no_family, species_no_genus)
```
Use taxize for species with no family and genus info
```{r}
library(taxize)

# Function to retrieve taxonomic information for a chunk of species names
get_taxonomic_info_chunk <- function(chunk_species_names) {
  # Initialize an empty list to store taxonomic information for each chunk
  chunk_taxonomic_info <- list()

  # Loop through each species name in the chunk and retrieve taxonomic information
  for (species_name in chunk_species_names) {
    tryCatch({
      # Add a delay between consecutive API requests
      Sys.sleep(1)

      # Make API request to retrieve taxonomic information
      taxon_info <- tax_name(species_name, get = c("genus", "family"), db = "ncbi")

      # Store taxonomic information for the species in the list
      chunk_taxonomic_info[[species_name]] <- data.frame(Species = species_name, 
                                                         Genus = taxon_info$genus, 
                                                         Family = taxon_info$family)
    }, error = function(e) {
      # Print error message
      cat("Error retrieving taxonomic information for", species_name, ":", conditionMessage(e), "\n")
    })
  }

  # Return the list of taxonomic information for the chunk
  return(chunk_taxonomic_info)
}

# Define the chunk size
chunk_size <- 300 # Adjust the chunk size as needed

# Split the dataframe into chunks based on the chunk size
chunks <- split(species_no_family, ceiling(seq_along(species_no_family) / chunk_size))

# Initialize an empty list to store taxonomic information for all chunks
all_taxonomic_info <- list()

# Iterate over each chunk and retrieve taxonomic information
for (chunk_index in seq_along(chunks)) {
  # Get the species names in the current chunk
  chunk_species_names <- chunks[[chunk_index]]

  # Retrieve taxonomic information for the chunk
  chunk_taxonomic_info <- get_taxonomic_info_chunk(chunk_species_names)

  # Append taxonomic information for the chunk to the list
  all_taxonomic_info <- c(all_taxonomic_info, list(chunk_taxonomic_info))

  # Add a delay 
  Sys.sleep(5)
}

# Combine taxonomic information from all chunks into a single list
all_taxonomic_info <- do.call(c, all_taxonomic_info)

# Combine the list of dataframes into a single dataframe
taxonomic_df <- do.call(rbind, all_taxonomic_info)
```

```{r}
# remove duplicates
taxonomic_df <- distinct(taxonomic_df)
# rename column to family, genus, species
taxonomic_df <- taxonomic_df %>%
  rename(family = Family,
         genus = Genus,
         species = Species)
```

```{r}
# add to species dataframe
na_species_df$family <- ifelse(is.na(na_species_df$family), taxonomic_df$family[match(na_species_df$species, taxonomic_df$species)], na_species_df$family)
na_species_df$genus <- ifelse(is.na(na_species_df$genus), taxonomic_df$genus[match(na_species_df$species, taxonomic_df$species)], na_species_df$genus)

# retrieve list of species without family & genus information
species_no_family_2 <- na_species_df %>%
  filter(is.na(family) | family == "") %>%
  select(species) %>%
  pull()
```

```{r}
# get powo id
species_no_family_2_powo <- get_pow(species_no_family_2, messages = TRUE)
```

```{r}
# Function to retrieve taxonomic information for a chunk of taxon IDs
get_taxonomic_info_chunk <- function(chunk_taxon_id) {
  # Initialize an empty list to store taxonomic information for each chunk
  chunk_taxonomic_info <- list()

  # Loop through each taxon ID in the chunk and retrieve taxonomic information
  for (taxon_id in chunk_taxon_id) {
    tryCatch({
      # Print taxon ID for debugging
      print(paste("Processing taxon ID:", taxon_id))

      # Add a delay between consecutive API requests
      Sys.sleep(1)

      # Query POWO database for taxonomic information
      taxon_info <- pow_lookup(id = taxon_id)

      # Check if taxon_info is NULL
      if (is.null(taxon_info)) {
        # If taxon_info is NULL, return NA values for species, family, and genus
        chunk_taxonomic_info[[taxon_id]] <- data.frame(taxon_id = taxon_id, species = NA, family = NA, genus = NA)
      } else {
        # Store taxonomic information for the taxon ID in the list
        chunk_taxonomic_info[[taxon_id]] <- data.frame(taxon_id = taxon_id,
                                                        species = taxon_info$meta$name,
                                                        genus = taxon_info$meta$genus, 
                                                        family = taxon_info$meta$family)
      }
    }, error = function(e) {
      # Print error message for debugging
      print(paste("Error processing taxon ID:", taxon_id))
      print(e)
    })
  }

  # Return the list of taxonomic information for the chunk
  return(chunk_taxonomic_info)
}

# Define the chunk size
chunk_size <- 300 # Adjust the chunk size as needed

# Split the vector of taxon IDs into chunks based on the chunk size
chunks <- split(species_no_family_2_powo, ceiling(seq_along(species_no_family_2_powo) / chunk_size))

# Initialize an empty list to store taxonomic information for all chunks
all_taxonomic_info <- list()

# Iterate over each chunk and retrieve taxonomic information
for (chunk_index in seq_along(chunks)) {
  # Get the taxon IDs in the current chunk
  chunk_taxon_id <- chunks[[chunk_index]]

  # Retrieve taxonomic information for the chunk
  chunk_taxonomic_info <- get_taxonomic_info_chunk(chunk_taxon_id)

  # Append taxonomic information for the chunk to the list
  all_taxonomic_info <- c(all_taxonomic_info, list(chunk_taxonomic_info))

  # Add a delay 
  Sys.sleep(5)
}

# Combine taxonomic information from all chunks into a single list
all_taxonomic_info <- do.call(c, all_taxonomic_info)

# Combine the list of dataframes into a single dataframe
taxonomic_df <- do.call(rbind, all_taxonomic_info)

```

```{r}
# add to species dataframe
na_species_df$family <- ifelse(is.na(na_species_df$family), taxonomic_df$family[match(na_species_df$species, taxonomic_df$species)], na_species_df$family)
na_species_df$genus <- ifelse(is.na(na_species_df$genus), taxonomic_df$genus[match(na_species_df$species, taxonomic_df$species)], na_species_df$genus)

# retrieve list of species without family & genus information
species_no_family_3 <- na_species_df %>%
  filter(is.na(family) | family == "") %>%
  select(species) %>%
  pull()
```

## Add genus and family info to dataframe with species and traits
```{r}
na_traits_family.genus <- merge(na_traits, na_species_df, by = "species", all.x = TRUE)
```

## Get a list for each trait of species, genus, and family to fill in gaps
```{r}
# Get unique traits
traits <- unique(na_traits_family.genus$TraitName)

# Create an empty list to store dataframes for each trait
trait_dfs <- list()

# Split the dataframe by traitName and create a dataframe for each trait
for (trait in traits) {
  # Subset the dataframe for the current trait
  trait_df <- subset(na_traits_family.genus, TraitName == trait)

  # Remove the traitName column from the subsetted dataframe
  trait_df <- trait_df[, !(names(trait_df) %in% "traitName")]

  # Add the dataframe to the list
  trait_dfs[[trait]] <- trait_df
}

# Seperate dataframes from list
Dispersal.Syndrome <- trait_dfs$Dispersal.Syndrome
Fruit.Length <- trait_dfs$Fruit.Length
Plant.Height <- trait_dfs$Plant.Height
Fruit.Type <- trait_dfs$Fruit.Type
Growth.Form <- trait_dfs$Growth.Form
```

```{r}
# Create a list of unique genus and families for each trait
Dispersal.Syndrome_families <- unique(Dispersal.Syndrome$family)
Dispersal.Syndrome_genus <- unique(Dispersal.Syndrome$genus)

Fruit.Length_families <- unique(Fruit.Length$family)
Fruit.Length_genus <- unique(Fruit.Length$genus)

Plant.Height_families <- unique(Plant.Height$family)
Plant.Height_genus <- unique(Plant.Height$genus)

Fruit.Type_families <- unique(Fruit.Type$family)
Fruit.Type_genus <- unique(Fruit.Type$genus)

Growth.Form_families <- unique(Growth.Form$family)
Growth.Form_genus <- unique(Growth.Form$genus)
```

##Retrive trait records for species using BIEN & GIFT
```{r}

```

# Set the seed for reproducibility
```{r}
set.seed(123)
```

```{r}
#replace " " with _ in column names
names(plant_traits) <- sub(" ", "_", names(plant_traits))
```

# show the missing data pattern
```{r}
md.pattern(plant_traits)
```
# Trying to prevent error
Error in edit.setup(data, setup, ...) : 
  `mice` detected constant and/or collinear variables. No predictors were left after their removal.
```{r}
# Subset to numeric plant traits
num_plant_traits <- plant_traits %>%
  select_if(is.numeric)
# Check for constant variables
constant_vars <- sapply(num_plant_traits, function(x) length(unique(x))) == 1
constant_vars <- names(constant_vars[constant_vars])

# Remove constant variables
num_plant_traits <- num_plant_traits[, !names(num_plant_traits) %in% constant_vars]

# Check for collinearity
cor_matrix <- cor(num_plant_traits)

# Find highly correlated variables
highly_correlated_pairs <- which(abs(cor_matrix) > 0.8 & upper.tri(cor_matrix, diag = TRUE), arr.ind = TRUE)

# Print highly correlated pairs
print(highly_correlated_pairs)
```

# Perform the imputation
```{r}
imp_model <- mice(plant_traits, method = "cart", maxit = 20)
```

```{r}
imputed_data <- complete(imp_model)
```