### 2024-05-10 code/Manuscript_figures.rmd

Summary plots
# Load required packages
```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(tidyr)
# devtools::install_github("pedropark99/ggfunnel")
library(ggfunnel)
library(ggpubr)
```

```{r}
Step <- c("Raw", "After cleaning", "Full trait coverage")
frug_records <- c(8089513, 1721825, 1667892)
frug_species <- c(1222, 1219, 1074)
frug_genera <- c(400, 398, 370)
frug_families <- c(66, 66, 64)

frug_records_df <- data.frame(Step, frug_records)
frug_species_df <- data.frame(Step, frug_species)
frug_genera_df <- data.frame(Step, frug_genera)
frug_families_df <- data.frame(Step, frug_families)
```

```{r}
ggplot(frug_records_df, aes(x = Step, y = frug_records)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  geom_text(aes(label = frug_records), color = "white") + 
  theme(axis.title.x = element_blank(), axis.text.x = element_blank(),axis.ticks.x = element_blank(), axis.title.y = element_blank())
```
```{r}
frug_records_fig <- ggfunnel::funnel(frug_records_df, levels = Step, values = frug_records, text_specs = list(colour = "white", size = 3)) +
  labs(title = "Frugivore records", y = NULL)
frug_records_fig
```

```{r}
frug_species_fig <- ggfunnel::funnel(frug_species_df, levels = Step, values = frug_species, text_specs = list(colour = "white", size = 3)) +
  labs(title = "Frugivore species", y = NULL)
frug_species_fig
```

```{r}
frug_genera_fig <- ggfunnel::funnel(frug_genera_df, levels = Step, values = frug_genera, text_specs = list(colour = "white", size = 3)) +
  labs(title = "Frugivore genera", y = NULL)
frug_genera_fig
```

```{r}
frug_families_fig <- ggfunnel::funnel(frug_families_df, levels = Step, values = frug_families, text_specs = list(colour = "white", size = 3)) +
  labs(title = "Frugivore families", y = NULL)
frug_families_fig
```

```{r}
figure <- ggarrange(frug_records_fig,frug_species_fig, frug_genera_fig, frug_families_fig,
                    labels = c("A", "B", "C", "D"),
                    ncol = 2, nrow = 2)
figure
```

```{r}
devtools::session_info()
```