### 2024-05-14 code/L1/L1_LastStep_DiversityInputObjects.Rmd
inital commit

---
title: ""
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "Projecting occurrence data and creating presence-absence matrix for plants and frugivores. "
data input: 
data output: 
date: "2024-05-13"
output: html_document
---

# Load required packages
```{r}
library(letsR)
library(mFD)
library(vegan)
library(rnaturalearth)
library(sf)
library(raster)
library(fasterize)
library(dplyr)
library(funbiogeo)
```

# Set file paths
```{r}
data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path_L1 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Read in Data
```{r}
TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_plant_occ_cleaned_subset.csv"))
TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_frugivore_occ_cleaned_subset.csv"))
TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path_L0, "Forest_sf.shp"), layer = "Forest_sf")
frugivore_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_Frugivoria_traits_Forest.csv"))
plant_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_imputed_plant_traits.csv"))
```

# Convert data to spatial data
```{r}
plants.sf <- st_as_sf(TropicalAndes_plant_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)

frugivores.sf <- st_as_sf(TropicalAndes_frugivore_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
```

# Polygons of countries
```{r}
worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")
Americas <- ne_countries(continent = c("North America", "South America"), returnclass = "sf")
#polygon of Tropical Andes
TApoly <- worldMap %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
```

# Transform to projected coordinate reference system (units from degrees to meters)
```{r}
Americas <- st_transform(Americas, 5389)
TApoly <- st_transform(TApoly, 5389)
TropicalAndes_IUCNHabitat_Forest <- st_transform(TropicalAndes_IUCNHabitat_Forest, 5389)
plants.sf <- st_transform(plants.sf, 5389)
frugivores.sf <- st_transform(frugivores.sf, 5389)
```

```{r}
#check units
st_crs(TApoly, parameters = TRUE)$units_gdal
st_crs(TApoly)
st_crs(TropicalAndes_IUCNHabitat_Forest)
```

# Group by species
```{r}
plants_sf_species <- plants.sf %>%
  group_by(species) %>%
  summarise()

frugivores_sf_species <- frugivores.sf %>%
  group_by(species) %>%
  summarise()
```

# Plot Base Map
```{r}
basePlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly) +
  geom_sf(data = TropicalAndes_IUCNHabitat_Forest, fill = "forestgreen")+
  labs(title = "Tropical Andes", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

basePlot
```

# Plot points
```{r}
plantsPointsPlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly) +
  geom_sf(data = plants_sf_species, pch = 21) +
  labs(title = "Tropical Andes Plant Occurrences", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

plantsPointsPlot
```

```{r}
frugivoresPointsPlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly) +
  geom_sf(data = frugivores_sf_species, pch = 21) +
  labs(title = "Tropical Andes Frugivore Occurrences", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

frugivoresPointsPlot
```

# Create Presence-Absense matrix

Use project data sf object & convert to dataframe
```{r}
plants_occ_projected <- plants.sf %>%
  mutate(lon = sf::st_coordinates(.)[,1],
         lat = sf::st_coordinates(.)[,2]) %>%
  st_drop_geometry()
frugivores_occ_projected <- frugivores.sf %>%
  mutate(lon = sf::st_coordinates(.)[,1],
         lat = sf::st_coordinates(.)[,2]) %>%
  st_drop_geometry()
```

ADD CRS argument to lets.presab.points line (in PROJ4 type description)
```{r}
x <- plants_occ_projected$lon
y <- plants_occ_projected$lat
plant_xy <- cbind(x, y)
plant_species <- plants_occ_projected$species
plants_PAM <- letsR::lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, resol = 10000, crs = "+proj=utm +zone=19 +south +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs +type=crs", count = TRUE)
summary(plants_PAM)
plot(plants_PAM, xlab = "Longitude", ylab = "Latitude", main = "Plant richness map")
```

```{r}
x <- frugivores_occ_projected$lon
y <- frugivores_occ_projected$lat
frugivore_xy <- cbind(x, y)
frugivore_species <- frugivores_occ_projected$species
frugivore_PAM <- letsR::lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, crs = "+proj=utm +zone=19 +south +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs +type=crs", count = TRUE)
summary(frugivore_PAM)
plot(frugivore_PAM, xlab = "Longitude", ylab = "Latitude", main = "Frugivore richness map")
```

Check CRS of occurrence and IUCN layer
```{r}
st_crs(TropicalAndes_frugivore_occ_forest)
st_crs(TropicalAndes_plant_occ_forest)
st_crs(TropicalAndes_IUCNHabitat_Forest)
```

## Perform the spatial intersection
```{r}
intersections_plant_TA <- st_intersection(plants.sf, TropicalAndes_IUCNHabitat_Forest)
intersections_frugivore_TA <- st_intersection(plants.sf, TropicalAndes_IUCNHabitat_Forest)

# Convert the intersected sf object back to a dataframe
coordinates_df_subset_plant <- as.data.frame(intersections_plant_TA)
coordinates_df_subset_frugivore <- as.data.frame(intersections_frugivore_TA)
```

```{r}
#Check CRS
st_crs(intersections_plant_TA)
st_crs(coordinates_df_subset_plant)
```

```{r}
# Print the resulting subset of coordinates
print(coordinates_df_subset_plant)
print(coordinates_df_subset_frugivore)

# Convert the point geometry to a dataframe of longitude and latitude
subset_plant <- as.data.frame(st_coordinates(intersections_plant_TA))
subset_frugivore <- as.data.frame(st_coordinates(intersections_frugivore_TA))
```

## Match the coordinates to the original dataframe
```{r}
#merge the subset and full dataframe together to get final TA dataset
plant_PAM_filter <- merge(plants_PAM_matrix, subset_plant, by.x = c("Longitude(x)", "Latitude(y)"), by.y = c("X", "Y"))
frugivore_PAM_filter <- merge(frugivore_PAM_matrix, subset_frugivore, by.x = c("Longitude(x)", "Latitude(y)"), by.y = c("X", "Y"))
```

## Turn PAM into matrix
```{r}
plant_PAM <- as.matrix(plant_PAM_filter)
frugivore_PAM <- as.matrix(frugivore_PAM_filter)
```

## Add rownames
```{r}
# Get the number of rows in the matrix
num_rows_plant <- nrow(plant_PAM)

# Generate unique row names
row_names_plant <- paste0("cell", 1:num_rows_plant)

# Assign row names to the matrix
rownames(plant_PAM) <- row_names_plant

column_names_plant <- colnames(plant_PAM)
clean_column_names_plant <- gsub("_", " ", column_names_plant)

#Insert clean column names
colnames(plant_PAM) <- clean_column_names_plant
```

```{r}
# Get the number of rows in the matrix
num_rows_frugivore <- nrow(frugivore_PAM)

# Generate unique row names
row_names_frugivore <- paste0("cell", 1:num_rows_frugivore)

# Assign row names to the matrix
rownames(frugivore_PAM) <- row_names_frugivore

column_names_frugivore <- colnames(frugivore_PAM)
clean_column_names_frugivore <- gsub("_", " ", column_names_frugivore)

#Insert clean column names
colnames(frugivore_PAM) <- clean_column_names_frugivore
```

## Turn data into correct data types for inputs into the trait categories dataframe
```{r}
plant_traits$Growth.Form <- as.factor(plant_traits$Growth.Form)
plant_traits$Plant.Height_m <- as.numeric(plant_traits$Plant.Height_m)
plant_traits$Fruit.Length_mm <- as.numeric(plant_traits$Fruit.Length_mm)
plant_traits$Fruit.Type <- as.factor(plant_traits$Fruit.Type)
```

```{r}
frugivore_traits$diet_cat <- as.factor(frugivore_traits$diet_cat)
frugivore_traits$body_mass_e <- as.numeric(frugivore_traits$body_mass_e)
frugivore_traits$body_size_mm <- as.numeric(frugivore_traits$body_size_mm)
frugivore_traits$generation_time <- as.numeric(frugivore_traits$generation_time)
```

## Remove the species from PAM that have no occurrences
```{r}
# Remove columns with sum equal to zero
PAM_plant_site_final <- plant_PAM[, colSums(plant_PAM) != 0]
PAM_frugivore_site_final <- frugivore_PAM[, colSums(frugivore_PAM) != 0]

#Save coordinates for later
site_loc_key_plant <- PAM_plant_site_final[,1:2]
site_loc_key_frugivore <- PAM_frugivore_site_final[,1:2]

columns_to_remove <- c(1,2)
PAM_plant_site_final <- PAM_plant_site_final[,-columns_to_remove]
PAM_frugivore_site_final <- PAM_frugivore_site_final[,-columns_to_remove]

colnames_plant <- colnames(PAM_plant_site_final)
colnames_frugivore <- colnames(PAM_frugivore_site_final)
```

## Remove species names from trait matrix not in the PAM
```{r}
plant_traits_df_subset <- plant_traits %>% filter(species %in% colnames_plant)
frugivore_traits_df_subset <- frugivore_traits %>% filter(IUCN_species_name %in% colnames_frugivore)

# Turn trait dataframe into a matrix
class(plant_traits)
class(frugivore_traits)

plant_traits_matrix <- as.matrix(plant_traits_df_subset)
frugivore_traits_matrix <- as.matrix(frugivore_traits_df_subset)

# Define row names as species names
row_names_plant <- plant_traits_matrix[,1]
row_names_frugivore <- frugivore_traits_matrix[,1]

# Assign row names to the matrix
rownames(plant_traits_matrix) <- row_names_plant
rownames(frugivore_traits_matrix) <- row_names_frugivore

# Turn back into dataframe
plant_traits_df_final <-as.data.frame(plant_traits_matrix)
frugivore_traits_df_final <-as.data.frame(frugivore_traits_matrix)

plant_traits_df_final$X <-NULL
frugivore_traits_df_final$X <-NULL
```

```{r}
#fix types
plant_traits_df_final$Growth.Form <- as.factor(plant_traits_df_final$Growth.Form)
plant_traits_df_final$Plant.Height_m <- as.numeric(plant_traits_df_final$Plant.Height_m)
plant_traits_df_final$Fruit.Length_mm <- as.numeric(plant_traits_df_final$Fruit.Length_mm)
plant_traits_df_final$Fruit.Type <- as.factor(plant_traits_df_final$Fruit.Type)
```

```{r}
frugivore_traits_df_final$diet_cat <- as.factor(frugivore_traits_df_final$diet_cat)
frugivore_traits_df_final$body_mass_e <- as.numeric(frugivore_traits_df_final$body_mass_e)
frugivore_traits_df_final$body_size_mm<- as.numeric(frugivore_traits_df_final$body_size_mm)
frugivore_traits_df_final$generation_time <- as.numeric(frugivore_traits_df_final$generation_time)
```

```{r}
# remove duplicate species name column
plant_traits_df_final <- plant_traits_df_final[, c("Growth.Form", "Plant.Height_m", "Fruit.Length_mm", "Fruit.Type")]
frugivore_traits_df_final <- frugivore_traits_df_final[, c("diet_cat", "body_mass_e", "body_size_mm", "generation_time")]
```

# Export data
```{r}
# objects to save
## sf objects
save(plants_sf_species, file = file.path(data_path_L1,"plants_sf_species.rdata"))
save(frugivores_sf_species, file = file.path(data_path_L1,"frugivores_sf_species.rdata"))
save(Americas, file = file.path(data_path_L1, "Americas.rdata"))
save(TApoly, file = file.path(data_path_L1,"TApoly.rdata"))
save(TropicalAndes_IUCNHabitat_Forest, file = file.path(data_path_L1,"TropicalAndes_IUCNHabitat_Forest.rdata"))
```

```{r}
# PAM objects
lets.save(plants_PAM, file = file.path(data_path_L1,"plants_PAM.RData"))
lets.save(frugivore_PAM, file = file.path(data_path_L1,"frugivore_PAM.RData"))
```

```{r}
# objects for functional diversity
PAM_plant_site_final
PAM_frugivore_site_final
plant_traits_df_final
frugivore_traits_df_final
```