### 2024-05-14 code/L1/L1_LastStep_DiversityInputObjects.Rmd
remove projection code for PAM and rename following variables using un-projected objects, export files as RDS files

---
title: ""
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "Projecting occurrence data and creating presence-absence matrix for plants and frugivores. "
data input: 
data output: 
date: "2024-05-13"
output: html_document
---

# Load required packages
```{r}
library(letsR)
library(mFD)
library(vegan)
library(rnaturalearth)
library(sf)
library(raster)
library(fasterize)
library(dplyr)
library(funbiogeo)
```

# Set file paths
```{r}
data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path_L1 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Read in Data
```{r}
TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_plant_occ_cleaned_subset.csv"))
TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_frugivore_occ_cleaned_subset.csv"))
TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path_L0, "Forest_sf.shp"), layer = "Forest_sf")
frugivore_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_Frugivoria_traits_Forest.csv"))
plant_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_imputed_plant_traits.csv"))
```

# Convert data to spatial data
```{r}
plants.sf <- st_as_sf(TropicalAndes_plant_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)

frugivores.sf <- st_as_sf(TropicalAndes_frugivore_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
```

# Polygons of countries
```{r}
worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")
Americas <- ne_countries(continent = c("North America", "South America"), returnclass = "sf")
#polygon of Tropical Andes
TApoly <- worldMap %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
```

# Transform to projected coordinate reference system (units from degrees to meters)
```{r}
Americas <- st_transform(Americas, 5389)
TApoly <- st_transform(TApoly, 5389)
TropicalAndes_IUCNHabitat_Forest <- st_transform(TropicalAndes_IUCNHabitat_Forest, 5389)
plants.sf <- st_transform(plants.sf, 5389)
frugivores.sf <- st_transform(frugivores.sf, 5389)
```

```{r}
#check units
st_crs(TApoly, parameters = TRUE)$units_gdal
st_crs(TApoly)
st_crs(TropicalAndes_IUCNHabitat_Forest)
```

# Group by species
```{r}
plants_sf_species <- plants.sf %>%
  group_by(species) %>%
  summarise()

frugivores_sf_species <- frugivores.sf %>%
  group_by(species) %>%
  summarise()
```

# Plot Base Map
```{r}
basePlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly) +
  geom_sf(data = TropicalAndes_IUCNHabitat_Forest, fill = "forestgreen")+
  labs(title = "Tropical Andes", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

basePlot
```

# Plot points
```{r}
plantsPointsPlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly) +
  geom_sf(data = plants_sf_species, pch = 21) +
  labs(title = "Tropical Andes Plant Occurrences", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

plantsPointsPlot
```

```{r}
frugivoresPointsPlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly) +
  geom_sf(data = frugivores_sf_species, pch = 21) +
  labs(title = "Tropical Andes Frugivore Occurrences", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

frugivoresPointsPlot
```

# Create Presence-Absense matrix

Use unprojected data because projected data is causing issues
```{r}
x <- TropicalAndes_plant_occ_forest$decimalLongitude
y <- TropicalAndes_plant_occ_forest$decimalLatitude
plant_xy <- cbind(x, y)
plant_species <- TropicalAndes_plant_occ_forest$species
plants_PAM <- letsR::lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, count = TRUE)
summary(plants_PAM)
plot(plants_PAM, xlab = "Longitude", ylab = "Latitude", main = "Plant richness map")
```

```{r}
x <- TropicalAndes_frugivore_occ_forest$decimalLongitude
y <- TropicalAndes_frugivore_occ_forest$decimalLatitude
frugivore_xy <- cbind(x, y)
frugivore_species <- TropicalAndes_frugivore_occ_forest$species
frugivore_PAM <- letsR::lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, count = TRUE)
summary(frugivore_PAM)
plot(frugivore_PAM, xlab = "Longitude", ylab = "Latitude", main = "Frugivore richness map")
```

## Turn PAM into matrix
```{r}
plants_PAM_matrix <- lets.presab.points(plant_xy, plant_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE, count = TRUE)
frugivore_PAM_matrix <- lets.presab.points(frugivore_xy, frugivore_species, xmn = -85, xmx = -60, ymn = -24, ymx = 14, show.matrix = TRUE, remove.cells = TRUE, count = TRUE)
```

## Add rownames
```{r}
# Get the number of rows in the matrix
num_rows_plant <- nrow(plants_PAM_matrix)

# Generate unique row names
row_names_plant <- paste0("cell", 1:num_rows_plant)

# Assign row names to the matrix
rownames(plants_PAM_matrix) <- row_names_plant

column_names_plant <- colnames(plants_PAM_matrix)
```

```{r}
# Get the number of rows in the matrix
num_rows_frugivore <- nrow(frugivore_PAM_matrix)

# Generate unique row names
row_names_frugivore <- paste0("cell", 1:num_rows_frugivore)

# Assign row names to the matrix
rownames(frugivore_PAM_matrix) <- row_names_frugivore

column_names_frugivore <- colnames(frugivore_PAM_matrix)
```

## Turn data into correct data types for inputs into the trait categories dataframe
```{r}
plant_traits$Growth.Form <- as.factor(plant_traits$Growth.Form)
plant_traits$Plant.Height_m <- as.numeric(plant_traits$Plant.Height_m)
plant_traits$Fruit.Length_mm <- as.numeric(plant_traits$Fruit.Length_mm)
plant_traits$Fruit.Type <- as.factor(plant_traits$Fruit.Type)
```

```{r}
frugivore_traits$diet_cat <- as.factor(frugivore_traits$diet_cat)
frugivore_traits$body_mass_e <- as.numeric(frugivore_traits$body_mass_e)
frugivore_traits$body_size_mm <- as.numeric(frugivore_traits$body_size_mm)
frugivore_traits$generation_time <- as.numeric(frugivore_traits$generation_time)
```

## Remove the species from PAM that have no occurrences
```{r}
# Remove columns with sum equal to zero
PAM_plant_site_final <- plants_PAM_matrix[, colSums(plants_PAM_matrix) != 0]
PAM_frugivore_site_final <- frugivore_PAM_matrix[, colSums(frugivore_PAM_matrix) != 0]

#Save coordinates for later
site_loc_key_plant <- PAM_plant_site_final[,1:2]
site_loc_key_frugivore <- PAM_frugivore_site_final[,1:2]

PAM_plant_site_final <- PAM_plant_site_final[,-c(1:2)]
PAM_frugivore_site_final <- PAM_frugivore_site_final[,-(1:2)]

colnames_plant <- colnames(PAM_plant_site_final)
colnames_frugivore <- colnames(PAM_frugivore_site_final)
```

```{r}
str(PAM_plant_site_final)
str(PAM_frugivore_site_final)
```

## Remove species names from trait matrix not in the PAM
```{r}
plant_traits_df_subset <- plant_traits %>%
                            filter(species %in% colnames_plant) %>%
                            distinct(species, .keep_all = TRUE)
frugivore_traits_df_subset <- frugivore_traits %>%
                                filter(IUCN_species_name %in% colnames_frugivore) %>%
                                distinct(IUCN_species_name, .keep_all = TRUE)

dim(plant_traits)
dim(plant_traits_df_subset)

dim(frugivore_traits)
dim(frugivore_traits_df_subset)
```

```{r}
# Define row names as species names
row_names_plant <- plant_traits_df_subset$species
row_names_frugivore <- frugivore_traits_df_subset$IUCN_species_name

# Assign row names to the matrix
rownames(plant_traits_df_subset) <- row_names_plant
rownames(frugivore_traits_df_subset) <- row_names_frugivore

plant_traits_df_subset$X <-NULL
frugivore_traits_df_subset$X <-NULL
# remove duplicate species name column
plant_traits_df_subset <- plant_traits_df_subset[, c("Growth.Form", "Plant.Height_m", "Fruit.Length_mm", "Fruit.Type")]
frugivore_traits_df_subset <- frugivore_traits_df_subset[, c("diet_cat", "body_mass_e", "body_size_mm", "generation_time")]
```

```{r}
str(plant_traits_df_subset)
str(frugivore_traits_df_subset)
```

# Export data
```{r}
# objects to save
## sf objects
saveRDS(plants_sf_species, file = file.path(data_path_L1,"plants_sf_species.rds"))
saveRDS(frugivores_sf_species, file = file.path(data_path_L1,"frugivores_sf_species.rds"))
saveRDS(Americas, file = file.path(data_path_L1, "Americas.rds"))
saveRDS(TApoly, file = file.path(data_path_L1,"TApoly.rds"))
saveRDS(TropicalAndes_IUCNHabitat_Forest, file = file.path(data_path_L1,"TropicalAndes_IUCNHabitat_Forest.rds"))
```

```{r}
# PAM objects
saveRDS(plants_PAM, file = file.path(data_path_L1,"plants_PAM.rds"))
saveRDS(frugivore_PAM, file = file.path(data_path_L1,"frugivore_PAM.rds"))
```

```{r}
# objects for functional diversity
saveRDS(PAM_plant_site_final, file = file.path(data_path_L1,"PAM_plant_site_final.rds"))
saveRDS(PAM_frugivore_site_final, file = file.path(data_path_L1,"PAM_frugivore_site_final.rds"))
saveRDS(plant_traits_df_subset, file = file.path(data_path_L1,"plant_traits_df_final.rds"))
saveRDS(frugivore_traits_df_subset, file = file.path(data_path_L1,"frugivore_traits_df_final.rds"))
```