### 2024-05-14 code/Manuscript_figures.rmd
add plant cleaning visualizations

---
title: "Manuscript Figures"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script generates misc figures for manuscript that corresponds to this project."
data input: ""
data output: ""
date: "2024-05-14"
output: 
    html_document:
      code_link: TRUE
      code_download: TRUE
---

# Load required packages

```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(tidyr)
# devtools::install_github("pedropark99/ggfunnel")
library(ggfunnel)
library(ggpubr)
```

# Creating cleaning occurrence data figures

These numbers are recorded from the summary function at the end of most L0 and L1 scripts. A table is available in the manuscript supplemential materials.

## Frugivores

```{r}
Step <- c("Raw", "After cleaning", "Full trait coverage")
frug_records <- c(8089513, 1721825, 1667892)
frug_species <- c(1222, 1219, 1074)
frug_genera <- c(400, 398, 370)
frug_families <- c(66, 66, 64)

frug_records_df <- data.frame(Step, frug_records)
frug_species_df <- data.frame(Step, frug_species)
frug_genera_df <- data.frame(Step, frug_genera)
frug_families_df <- data.frame(Step, frug_families)
```

```{r}
ggplot(frug_records_df, aes(x = Step, y = frug_records)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  geom_text(aes(label = frug_records), color = "white") + 
  theme(axis.title.x = element_blank(), axis.text.x = element_blank(),axis.ticks.x = element_blank(), axis.title.y = element_blank())
```

```{r}
frug_records_fig <- ggfunnel::funnel(frug_records_df, levels = Step, values = frug_records, text_specs = list(colour = "white", size = 3)) +
  labs(title = "Frugivore records", y = NULL)
frug_records_fig
```

```{r}
frug_species_fig <- ggfunnel::funnel(frug_species_df, levels = Step, values = frug_species, text_specs = list(colour = "white", size = 3)) +
  labs(title = "Frugivore species", y = NULL)
frug_species_fig
```

```{r}
frug_genera_fig <- ggfunnel::funnel(frug_genera_df, levels = Step, values = frug_genera, text_specs = list(colour = "white", size = 3)) +
  labs(title = "Frugivore genera", y = NULL)
frug_genera_fig
```

```{r}
frug_families_fig <- ggfunnel::funnel(frug_families_df, levels = Step, values = frug_families, text_specs = list(colour = "white", size = 3)) +
  labs(title = "Frugivore families", y = NULL)
frug_families_fig
```

```{r}
frug_cleaning <- ggarrange(frug_records_fig,frug_species_fig, frug_genera_fig, frug_families_fig,
                    labels = c("A", "B", "C", "D"),
                    ncol = 2, nrow = 2)
frug_cleaning
```

## Plants

```{r}
Step <- c("Raw", "After cleaning", "Full trait coverage")
plant_records <- c(4519792, 1886120, 1700044)
plant_species <- c(40288, 39972, 33746)
plant_genera <- c(4083, 3793, 3400)
plant_families <- c(307, 293, 275)

plant_records_df <- data.frame(Step, plant_records)
plant_species_df <- data.frame(Step, plant_species)
plant_genera_df <- data.frame(Step, plant_genera)
plant_families_df <- data.frame(Step, plant_families)
```

```{r}
ggplot(plant_records_df, aes(x = Step, y = plant_records)) +
  geom_bar(stat = "identity", fill = "darkgreen") +
  coord_flip() +
  geom_text(aes(label = plant_records), vjust = -0.3, size = 4) + 
  labs(title = "Plant Records", y = "Count") +
  theme_minimal()
```

```{r}
plant_records_fig <- ggfunnel::funnel(plant_records_df, levels = Step, values = plant_records, text_specs = list(colour = "white", size = 3)) +
  labs(title = "Plant records", y = NULL)
plant_records_fig
```

```{r}
plant_species_fig <- ggfunnel::funnel(plant_species_df, levels = Step, values = plant_species, text_specs = list(colour = "white", size = 3)) +
  labs(title = "Plant species", y = NULL)
plant_species_fig
```

```{r}
plant_genera_fig <- ggfunnel::funnel(plant_genera_df, levels = Step, values = plant_genera, text_specs = list(colour = "white", size = 3)) +
  labs(title = "Plant genera", y = NULL)
plant_genera_fig
```

```{r}
plant_families_fig <- ggfunnel::funnel(plant_families_df, levels = Step, values = plant_families, text_specs = list(colour = "white", size = 3)) +
  labs(title = "Plant families", y = NULL)
plant_families_fig
```

```{r}
plant_cleaning <- ggarrange(plant_records_fig,plant_species_fig, plant_genera_fig, plant_families_fig,
                    labels = c("A", "B", "C", "D"),
                    ncol = 2, nrow = 2)
plant_cleaning
```

# References
```{r}
library(report)
cite_packages()
```

# Session Info
```{r, class.source= 'fold-hide'}
devtools::session_info()
```