### 2024-05-16 code/L2/L2_1_TropicalAndes_funcdiv.Rmd
remove raster code, try using sf grid instead -- not yet working

---
title: "Tropical Andes Functional diversity for plants and Frugivores"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: ""
data input: ""
data output: ""
date: "2023-08-03"
output: html_document
---

# Load required packages
```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(smoothr)
library(purrr)
library(raster)
library(ggspatial)
library(vegan)
```

# Set file paths
```{r}
data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
```

# Read in Data

```{r}
## projected sf objects
plants_sf_species <- readRDS(file = file.path(data_path_L1,"plants_sf_species.rds"))
frugivores_sf_species <- readRDS(file = file.path(data_path_L1,"frugivores_sf_species.rds"))
Americas <- readRDS(file = file.path(data_path_L1, "Americas.rds"))
TApoly <- readRDS(file = file.path(data_path_L1,"TApoly.rds"))
TropicalAndes_IUCNHabitat_Forest <- readRDS(file = file.path(data_path_L1,"TropicalAndes_IUCNHabitat_Forest.rds"))
```

```{r}
# objects for functional diversity
PAM_plant_site_final <- readRDS(file = file.path(data_path_L1,"PAM_plant_site_final.rds"))
PAM_frugivore_site_final <- readRDS(file = file.path(data_path_L1,"PAM_frugivore_site_final.rds"))
plant_traits_df_final <- readRDS(file = file.path(data_path_L1,"plant_traits_df_final.rds"))
frugivore_traits_df_final <- readRDS(file = file.path(data_path_L1,"frugivore_traits_df_final.rds"))
site_loc_key_plant <- readRDS(file = file.path(data_path_L1,"site_loc_key_plant.rds"))
site_loc_key_frugivore <- readRDS(file = file.path(data_path_L1,"site_loc_key_frugivore.rds"))
```

# Frugivores
## Create trait type table
```{r}
trait_name <- c("diet_cat", "body_mass_e", "body_size_mm", "generation_time")
trait_type <- c("N", "Q", "Q", "Q")
frug_trait_cat <- as.data.frame(cbind(trait_name, trait_type))
```

## Summary of the assemblages * species dataframe
```{r}
asb_sp_frugivore_summ <- mFD::asb.sp.summary(asb_sp_w = PAM_frugivore_site_final)
```

## Species traits summary
```{r}
frugivore_traits_summ <- mFD::sp.tr.summary(
  tr_cat     = frug_trait_cat,   
  sp_tr      = frugivore_traits_df_final, 
  stop_if_NA = TRUE)
```

## Estimate functional trait-based distances between species
```{r}
sp_dist_frugivore <- mFD::funct.dist(
  sp_tr         = frugivore_traits_df_final,
  tr_cat        = frug_trait_cat,
  metric        = "gower",
  scale_euclid  = "scale_center",
  ordinal_var   = "classic",
  weight_type   = "equal",
  stop_if_NA    = TRUE)
```

## Generate a multidimensional space
```{r}
fspaces_quality_frugivore <- mFD::quality.fspaces(
  sp_dist             = sp_dist_frugivore,
  maxdim_pcoa         = 10,
  deviation_weighting = "absolute",
  fdist_scaling       = FALSE,
  fdendro             = "average")
```

```{r}
# Look at the quality spaces only (MAD index looks at mean absolute deviation from the dissimilary matrix; want the deviation to be low, meaning that the true distances have been retained in the PCA)
round(fspaces_quality_frugivore$"quality_fspaces", 3)     
# Plot the quality spaces (chose to look at 3D, 4D, and 5D since they had the lowest MAD). Will go with 4 dimensions.
mFD::quality.fspaces.plot(
  fspaces_quality            = fspaces_quality_frugivore,
  quality_metric             = "mad",
  fspaces_plot               = c("pcoa_3d", "pcoa_4d", 
                                 "pcoa_5d"),
  name_file                  = NULL,
  range_dist                 = NULL,
  range_dev                  = NULL,
  range_qdev                 = NULL,
  gradient_deviation         = c(neg = "darkblue", nul = "grey80", pos = "darkred"),
  gradient_deviation_quality = c(low = "yellow", high = "red"),
  x_lab                      = "Trait-based distance")
```

```{r}
#testing correlation between functional axes and traits
sp_faxes_coord_frugivore <- fspaces_quality_frugivore$"details_fspaces"$"sp_pc_coord"
```

```{r}
# View the components of the PCA axes *Remember the first few components explain the most variation in dissimilarity. Clusters into groups

# Computes linear model for continuous traits and Kruskall-Wallis tests for other types. 
frugivore_tr_faxes <- mFD::traits.faxes.cor(
  sp_tr          = frugivore_traits_df_final, 
  sp_faxes_coord = sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")], 
  plot           = TRUE)

# Print traits with significant effect:
frugivore_tr_faxes$"tr_faxes_stat"[which(frugivore_tr_faxes$"tr_faxes_stat"$"p.value" < 0.05), ]
# Return plots:
frugivore_tr_faxes$"tr_faxes_plot"
```

```{r}
#plotting functional space
sp_faxes_coord_frugivore <- fspaces_quality_frugivore$"details_fspaces"$"sp_pc_coord"

big_plot_frugivore <- mFD::funct.space.plot(
  sp_faxes_coord  = sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")],
  faxes           = c("PC1", "PC2", "PC3", "PC4"),
  name_file       = NULL,
  faxes_nm        = NULL,
  range_faxes     = c(NA, NA),
  color_bg        = "grey95",
  color_pool      = "darkgreen",
  fill_pool       = "white",
  shape_pool      = 21,
  size_pool       = 1,
  plot_ch         = TRUE,
  color_ch        = "black",
  fill_ch         = "white",
  alpha_ch        = 0.5,
  plot_vertices   = TRUE,
  color_vert      = "blueviolet",
  fill_vert       = "blueviolet",
  shape_vert      = 6,
  size_vert       = 1,
  plot_sp_nm      = NULL,
  nm_size         = 3,
  nm_color        = "black",
  nm_fontface     = "plain",
  check_input     = TRUE)
```

```{r}
big_plot_frugivore
```

```{r}
#Need to remove parts of the PAM that have values less than or equal to the number of dimensions (4)
# Calculate row sums
row_sums_frugivore <- rowSums(PAM_frugivore_site_final)
subset_matrix_frugivore <- PAM_frugivore_site_final[row_sums_frugivore >= 4, ]
```

```{r}
# match frugivore names
sp_faxes_coord_frugivore_sub <- sp_faxes_coord_frugivore[ , c("PC1", "PC2", "PC3", "PC4")]
summary(sp_faxes_coord_frugivore_sub)
```

```{r}
## check number of species names
nrow(sp_faxes_coord_frugivore_sub)
ncol(subset_matrix_frugivore) 
```

```{r}
sp_faxes_coord_frugivore_sub_names <- row.names(sp_faxes_coord_frugivore_sub)
subset_matrix_frugivore_names <- colnames(subset_matrix_frugivore)
```

```{r}
sp_faxes_coord_frugivore_sub <- as.data.frame(sp_faxes_coord_frugivore_sub)
row.names(sp_faxes_coord_frugivore_sub) <- sp_faxes_coord_frugivore_sub_names

sp_faxes_coord_frugivore_sub_names <- row.names(sp_faxes_coord_frugivore_sub)
subset_matrix_frugivore_names <- colnames(subset_matrix_frugivore)

frugivore_names <- intersect(sp_faxes_coord_frugivore_sub_names, subset_matrix_frugivore_names)
frugivore_names <- na.omit(frugivore_names)

sp_faxes_coord_frugivore_sub <- sp_faxes_coord_frugivore_sub[ which((row.names(sp_faxes_coord_frugivore_sub) %in% frugivore_names)==TRUE), ]

subset_matrix_frugivore  <- as.data.frame(subset_matrix_frugivore)
subset_matrix_frugivore <- subset_matrix_frugivore[ ,which((colnames(subset_matrix_frugivore) %in% frugivore_names)==TRUE)]

#remove NAs
sp_faxes_coord_frugivore_sub <- na.omit(sp_faxes_coord_frugivore_sub)
subset_matrix_frugivore <- na.omit(subset_matrix_frugivore)
```

```{r}
nrow(sp_faxes_coord_frugivore_sub)
ncol(subset_matrix_frugivore)
```

```{r}
sp_faxes_coord_frugivore_sub <- as.matrix(sp_faxes_coord_frugivore_sub)
subset_matrix_frugivore <- as.matrix(subset_matrix_frugivore)
```

```{r}
#computing FD
# The number of species per assemblage has to be higher or equal to the number of traits

alpha_fd_indices_frugivore <- mFD::alpha.fd.multidim(
  sp_faxes_coord   = sp_faxes_coord_frugivore_sub,
  asb_sp_w         = subset_matrix_frugivore,
  ind_vect         = c("fide", "fdis", "fmpd", "fnnd", "feve", "fori", "fspe"), #did not run fdiv or fric (slows it down)
  scaling          = TRUE,
  check_input      = TRUE,
  details_returned = TRUE)

details_list_frugivore <- alpha_fd_indices_frugivore$"details"

plots_alpha_frugivore <- mFD::alpha.multidim.plot(
  output_alpha_fd_multidim = alpha_fd_indices_frugivore,
  plot_asb_nm              = c("cell4", "cell20"),
  ind_nm                   = c("fdis", "fide", "fnnd"
                        , "fori", "fspe"),
  faxes                    = NULL,
  faxes_nm                 = NULL,
  range_faxes              = c(NA, NA),
  color_bg                 = "grey95",
  shape_sp                 = c(pool = 3, asb1 = 21, asb2 = 21),
  size_sp                  = c(pool = 0.7, asb1 = 1, asb2 = 1),
  color_sp                 = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  color_vert               = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_sp                  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_vert                = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  color_ch          
  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_ch                  = c(pool = "white", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  alpha_ch                 = c(pool = 1, asb1 = 0.3, asb2 = 0.3),
  shape_centroid_fdis      = c(asb1 = 22,  asb2 = 24),
  shape_centroid_fdiv      = c(asb1 = 22,  asb2 = 24),
  shape_centroid_fspe      = 23,
  color_centroid_fspe      = "black",
  size_sp_nm               = 3, 
  color_sp_nm              = "black",
  plot_sp_nm               = NULL,
  fontface_sp_nm           = "plain",
  save_file                = FALSE,
  check_input              = TRUE)

 #comparing convex hulls of different hulls
plots_alpha_frugivore$"fric"$"patchwork"

plots_alpha_frugivore$"fdiv"$"patchwork"

plots_alpha_frugivore$"fspe"$"patchwork"

plots_alpha_frugivore$"fdis"$"patchwork"

plots_alpha_frugivore$"fori"$"patchwork"
```

```{r}
# Get functional dispersion
fdis_frugivore <- alpha_fd_indices_frugivore$functional_diversity_indices$fdis
```

## Project FD objects
```{r}
# Generate coordinates
subset_coords_frugivore <- site_loc_key_frugivore[rowSums(PAM_frugivore_site_final) >= 5, ]
subset_coords_frugivore_sp <-subset_coords_frugivore[,1:2 ]

frugivore_fd_sp <- data.frame(subset_coords_frugivore_sp, fdis_frugivore)
```

```{r}
# Convert the dataframe to sf format
spatial_fdis_frugivore <- st_as_sf(frugivore_fd_sp, coords = c("Longitude.x.", "Latitude.y."))
```

```{r}
# set crs of sf objects
spatial_fdis_frugivore <- spatial_fdis_frugivore %>% st_set_crs(4326)
```

```{r}
spatial_fdis_frugivore_projected <- st_transform(spatial_fdis_frugivore, 5389)
```

```{r}
create_fdis_plots <- function(resolution_meters, data, variable_name) {
# Make Grid
TAGrid <- TApoly %>%
  st_make_grid(cellsize = c(resolution_meters)) %>%
  st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
  st_cast("MULTIPOLYGON") %>%
  st_sf() %>%
  mutate(cellid = row_number())

  # Calculate functional dispersion for grid
  fdis_grid <- TAGrid %>%
    st_join(data) %>%
    group_by(cellid) %>%
    summarize(sum_variable = sum(variable_name, na.rm = TRUE))

  # Make plot
  FDplot <-
    ggplot() +
    geom_sf(data = Americas, fill = "white")+
    geom_sf(data = TApoly, fill = "lightgrey", size = 0.1) +
    geom_sf(data = fdis_grid, aes(fill = fdis_frugivore)) +
    scale_fill_gradient(low = "blue", high = "red") +
    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
    theme(panel.background = element_rect(fill = "lightblue")) +
    annotation_scale(location = "bl", width_hint = 0.5) +
    labs(x = "Longitude", y = "Latitude", fill = "FDis")

  # Return a list of data
  list(FDplot = FDplot)
}
```

```{r}
create_fdis_plots(100000, spatial_fdis_frugivore_projected, fdis_frugivore)
```

# Plants
Test with small random sample as a subset to make it work before using all data
```{r}
all_plant_species <- row.names(plant_traits_df_final)
plant_random_subset <- sample(all_plant_species, 1000)
# objects that need to be subset
PAM_plant_site_final #matrix
plant_traits_df_final
```

## Create trait type table
```{r}
trait_name <- c("Growth.Form", "Plant.Height_m", "Fruit.Length_mm", "Fruit.Type")
trait_type <- c("N", "Q", "Q", "N") 
plant_trait_cat <- as.data.frame(cbind(trait_name, trait_type))
```

## Summary of the assemblages * species dataframe
```{r}
asb_sp_plant_summ <- mFD::asb.sp.summary(asb_sp_w = PAM_plant_site_final)
```

## Species traits summary
```{r}
plant_traits_summ <- mFD::sp.tr.summary(
  tr_cat     = plant_trait_cat,   
  sp_tr      = plant_traits_df_final, 
  stop_if_NA = TRUE)
```

```{r}
save(plant_traits_df_final, plant_trait_cat, file = "../../plant_funct.dist_files.rdata")
```

## Estimate functional trait-based distances between species
```{r}
sp_dist_plants <- mFD::funct.dist(
  sp_tr         = plant_traits_df_final,
  tr_cat        = plant_trait_cat,
  metric        = "gower",
  scale_euclid  = "scale_center",
  ordinal_var   = "classic",
  weight_type   = "equal",
  stop_if_NA    = TRUE)
```

## Generate a multidimensional space
```{r}
fspaces_quality_plants <- mFD::quality.fspaces(
  sp_dist             = sp_dist_plants,
  maxdim_pcoa         = 10,
  deviation_weighting = "absolute",
  fdist_scaling       = FALSE,
  fdendro             = "average")
```

```{r}
# Look at the quality spaces only (MAD index looks at mean absolute deviation from the dissimilary matrix; want the deviation to be low, meaning that the true distances have been retained in the PCA)
round(fspaces_quality_plants$"quality_fspaces", 3)            
# Plot the quality spaces (chose to look at 3D, 4D, and 5D since they had the lowest MAD). Will go with 4 dimensions.
mFD::quality.fspaces.plot(
  fspaces_quality            = fspaces_quality_plants,
  quality_metric             = "mad",
  fspaces_plot               = c("pcoa_2d", "pcoa_3d", 
                                 "pcoa_4d"),
  name_file                  = NULL,
  range_dist                 = NULL,
  range_dev                  = NULL,
  range_qdev                 = NULL,
  gradient_deviation         = c(neg = "darkblue", nul = "grey80", pos = "darkred"),
  gradient_deviation_quality = c(low = "yellow", high = "red"),
  x_lab                      = "Trait-based distance")
```

```{r}
#testing correlation between functional axes and traits
sp_faxes_coord_plants <- fspaces_quality_plants$"details_fspaces"$"sp_pc_coord"
```

```{r}
# View the components of the PCA axes *Remember the first few components explain the most variation in dissimilarity. Clusters into groups

# Computes linear model for continuous traits and Kruskall-Wallis tests for other types. 
plant_tr_faxes <- mFD::traits.faxes.cor(
  sp_tr          = plant_traits_df_final, 
  sp_faxes_coord = sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")], 
  plot           = TRUE)

# Print traits with significant effect:
plant_tr_faxes$"tr_faxes_stat"[which(plant_tr_faxes$"tr_faxes_stat"$"p.value" < 0.05), ]
# Return plots:
plant_tr_faxes$"tr_faxes_plot"
```

```{r}
#plotting functional space
sp_faxes_coord_plants <- fspaces_quality_plants$"details_fspaces"$"sp_pc_coord"

big_plot_plants <- mFD::funct.space.plot(
  sp_faxes_coord  = sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")],
  faxes           = c("PC1", "PC2", "PC3", "PC4"),
  name_file       = NULL,
  faxes_nm        = NULL,
  range_faxes     = c(NA, NA),
  color_bg        = "grey95",
  color_pool      = "darkgreen",
  fill_pool       = "white",
  shape_pool      = 21,
  size_pool       = 1,
  plot_ch         = TRUE,
  color_ch        = "black",
  fill_ch         = "white",
  alpha_ch        = 0.5,
  plot_vertices   = TRUE,
  color_vert      = "blueviolet",
  fill_vert       = "blueviolet",
  shape_vert      = 6,
  size_vert       = 1,
  plot_sp_nm      = NULL,
  nm_size         = 3,
  nm_color        = "black",
  nm_fontface     = "plain",
  check_input     = TRUE)

#Need to remove parts of the PAM that have values less than or equal to the number of dimensions (4)
# Calculate row sums
row_sums_plant <- rowSums(PAM_plant_site_final)
subset_matrix_plant <- PAM_plant_site_final[row_sums_plant >= 4, ]
```

```{r}
big_plot_plants
```

```{r}
# getting sp_faxes_coord_plants and subset_matrix_plants names to match
sp_faxes_coord_plants_sub <- sp_faxes_coord_plants[ , c("PC1", "PC2", "PC3", "PC4")]
## check number of species names
nrow(sp_faxes_coord_plants_sub) # 54
ncol(subset_matrix_plant) #82
```

```{r}
sp_faxes_coord_plants_sub_names <- row.names(sp_faxes_coord_plants_sub)
subset_matrix_plant_names <- colnames(subset_matrix_plant)

sp_faxes_coord_plants_sub_df <- as.data.frame(sp_faxes_coord_plants_sub)
subset_matrix_plant_df <- as.data.frame(subset_matrix_plant)

subset_matrix_plant <- subset_matrix_plant_df[, which((names(subset_matrix_plant_df) %in% sp_faxes_coord_plants_sub_names)==TRUE)]

ncol(subset_matrix_plant)
subset_matrix_plant <- as.matrix(subset_matrix_plant)
```

```{r}
#computing FD
# The number of species per assemblage has to be higher or equal to the number of traits
alpha_fd_indices_plant <- mFD::alpha.fd.multidim(
  sp_faxes_coord   = sp_faxes_coord_plants_sub,
  asb_sp_w         = PAM_plant_site_final,
  ind_vect         = c("fide", "fdis", "fmpd", "fnnd", "feve", "fori", "fspe"), #did not run fdiv or fric (slows it down)
  scaling          = TRUE,
  check_input      = TRUE,
  details_returned = TRUE)

details_list_plant <- alpha_fd_indices_plant$"details"

#plot
plots_alpha_plant <- mFD::alpha.multidim.plot(
  output_alpha_fd_multidim = alpha_fd_indices_plant,
  plot_asb_nm              = c("cell4", "cell20"),
  ind_nm                   = c("fdis", "fide", "fnnd"
                        , "fori", "fspe"),
  faxes                    = NULL,
  faxes_nm                 = NULL,
  range_faxes              = c(NA, NA),
  color_bg                 = "grey95",
  shape_sp                 = c(pool = 3, asb1 = 21, asb2 = 21),
  size_sp                  = c(pool = 0.7, asb1 = 1, asb2 = 1),
  color_sp                 = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  color_vert               = c(pool = "grey50", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_sp                  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_vert                = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  color_ch          
  = c(pool = NA, asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  fill_ch                  = c(pool = "white", asb1 = "#1F968BFF", asb2 = "#DCE319FF"),
  alpha_ch                 = c(pool = 1, asb1 = 0.3, asb2 = 0.3),
  shape_centroid_fdis      = c(asb1 = 22,  asb2 = 24),
  shape_centroid_fdiv      = c(asb1 = 22,  asb2 = 24),
  shape_centroid_fspe      = 23,
  color_centroid_fspe      = "black",
  size_sp_nm               = 3, 
  color_sp_nm              = "black",
  plot_sp_nm               = NULL,
  fontface_sp_nm           = "plain",
  save_file                = FALSE,
  check_input              = TRUE)

# comparing convex hulls of different hulls
plots_alpha_plant$"fric"$"patchwork"

plots_alpha_plant$"fdiv"$"patchwork"

plots_alpha_plant$"fspe"$"patchwork"

plots_alpha_plant$"fdis"$"patchwork"

plots_alpha_plant$"fori"$"patchwork"

```

```{r}
#Get functional dispersion
fdis_plant <- alpha_fd_indices_plant$functional_diversity_indices$fdis
```

## Turning the FD values into a raster
Need to subset the coordinates made at the top to the same set of the assemblages (removed values less than 5 so these coords need to be removed to do this correctly)

```{r}
# Create an empty raster with desired resolution and extent
resolution <- c(0.75, 0.75)  # Set the desired resolution
extent <- c(-85, -54, -24, 14)  # Set the desired extent
empty_raster <- raster(resolution = resolution, xmn = extent[1], xmx = extent[2], ymn = extent[3], ymx = extent[4])
```

```{r}
# Generate coordinates
subset_coords_plant <- site_loc_key_plant[rowSums(PAM_plant_site_final) >= 5, ]
subset_coords_plant_sp <-subset_coords_plant[,1:2 ]

plant_fd_sp <- data.frame(subset_coords_plant_sp, fdis_plant)

# Convert the dataframe to sf format
spatial_fdis_plant <- st_as_sf(plant_fd_sp, coords = c("Longitude.x.", "Latitude.y."))
```

```{r}
# Rasterize the sf data to create the FD raster
fd_raster_plant <- rasterize(spatial_fdis_plant, empty_raster)
writeRaster(fd_raster_plant$fdis_plant, filename = file.path(output_path,"FD_plant_TA.tif"), format="GTiff",overwrite=T)
writeRaster(fd_raster_plant$ID, filename = file.path(output_path,"FD_plant_TA_ID.tif"), format="GTiff")
```

## Project FD objects
```{r}
# set crs of sf objects
spatial_fdis_plant <- spatial_fdis_plant %>% st_set_crs(4326)
```

```{r}
spatial_fdis_plant_projected <- st_transform(spatial_fdis_plant, 5389)
```

```{r}
# Raster to raster point
spec_FD_raster_plant_point <- rasterToPoints(fd_raster_plant, spatial = TRUE)
# Convert to a 'conventional' dataframe
fd_raster_plant_point <- data.frame(spec_FD_raster_plant_point)
```

```{r}
plantgridFDTA <-
  ggplot() +
  geom_sf(data = Americas, fill = "white")+
  geom_sf(data = TApoly, fill = "grey", size = 0.1) +
  geom_tile(data= fd_raster_plant_point, aes(x=x, y=y, fill=fdis_plant)) +
  scale_fill_scico(palette = "romaO", direction = 1) +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE) +
  theme(
    plot.background = element_rect(fill = "#f1f2f3"),
    panel.background = element_rect(fill = "lightblue"),
    panel.grid = element_blank(),
    line = element_blank(),
    rect = element_blank()
  ) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
  pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  labs(title = "Tropical Andes Plant Functional Dispersion", x = "Longitude", y = "Latitude", fill = "FD")
plantgridFDTA
```

# Calculating Rao's Q
Calculate Rao's Q using occurrence data and trait dissimilarity

```{r}
## Define different spatial scales (e.g., grid cell sizes)
spatial_scales <- c(1000, 5000, 10000, 25000, 50000, 75000, 100000)  # In meters

## Create an empty dataframe to store results
plant_raos_q_results <- data.frame(spatial_scale = numeric(),
                             raos_q = numeric())

## Iterate over different spatial scales
for (scale in spatial_scales) {
  ### Aggregate occurrence data to the current spatial scale
  plant_occurrence_aggregated <- aggregate_occurrence(plant_abundance_per_species, scale)

  ### Calculate dissimilarities using trait data
  plant_trait_dissimilarities <- vegdist(plant_traits, method = "bray")

  ### Calculate dissimilarities using occurrence data
  plant_occurrence_dissimilarities <- vegdist(plant_occurrence_aggregated, method = "bray")

  ### Calculate Rao's Q index
  raos_q <- raosq(plant_trait_dissimilarities, plant_occurrence_dissimilarities)

  #### Add results to the dataframe
  plant_raos_q_results <- rbind(raos_q_results, data.frame(spatial_scale = scale, raos_q = raos_q))
}
```

```{r}
## Create an empty dataframe to store results
frugivore_raos_q_results <- data.frame(spatial_scale = numeric(),
                             raos_q = numeric())

## Iterate over different spatial scales
for (scale in spatial_scales) {
  ### Aggregate occurrence data to the current spatial scale
  frugivore_occurrence_aggregated <- aggregate_occurrence(frugivore_abundance_per_species, scale)

  ### Calculate dissimilarities using trait data
  frugivore_trait_dissimilarities <- vegdist(frugivore_traits, method = "bray")

  ### Calculate dissimilarities using occurrence data
  frugivore_occurrence_dissimilarities <- vegdist(frugivore_occurrence_aggregated, method = "bray")

  ### Calculate Rao's Q index
  raos_q <- raosq(frugivore_trait_dissimilarities, frugivore_occurrence_dissimilarities)

  #### Add results to the dataframe
  frugivore_raos_q_results <- rbind(raos_q_results, data.frame(spatial_scale = scale, raos_q = raos_q))
}
```

# Extract cell values
```{r}
plantgridFDTA <- plantgridFDTA[["data"]]
frugivoregridFDTA <- frugivoregridFDTA[["data"]]
```

```{r}
plot(spatial_fdis_plant$fdis_plant, spatial_fdis_frugivore$fdis_frugivore,
     main = "Plant FD vs Frugivore FD",
     xlab = "Plant Functional Dispersion by cell", ylab = "Frugivore Functional Dispersion by cell")
abline(lm(spatial_fdis_frugivore$fdis_frugivore ~ spatial_fdis_plant$fdis_plant))
summary(lm(spatial_fdis_frugivore$fdis_frugivore ~ spatial_fdis_plant$fdis_plant))
```