### 2024-05-16 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
move some code to L2_3_TropicalAndes_divRelationships

---
title: "Tropical Andes Taxonomic Diversity of Plants and Frugivores"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "Calculating and mapping taxonomic diversity as species richness using occurrence data for plants and frugivores in the Tropical Andes Moist Lowland and Montane forests. Note: some code is adapted from https://luisdva.github.io/rstats/richness/."
data input: "TropicalAndes_plant_occ.csv, TropicalAndes_frugivore_occ.csv"
data output: "TropicalAndes_frugivore_richness.csv, TropicalAndes_GBIF_plant_richness.csv"
date: "2023-08-01"
output: html_document
---

# Load required packages
```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(smoothr)
library(purrr)
library(raster)
library(ggspatial)
library(ggpubr)
```

# Set file paths
```{r}
data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
```

# Read in Data
```{r}
## projected sf objects
plants_sf_species <- readRDS(file = file.path(data_path_L1,"plants_sf_species.rds"))
frugivores_sf_species <- readRDS(file = file.path(data_path_L1,"frugivores_sf_species.rds"))
Americas <- readRDS(file = file.path(data_path_L1, "Americas.rds"))
TApoly <- readRDS(file = file.path(data_path_L1,"TApoly.rds"))
TropicalAndes_IUCNHabitat_Forest <- readRDS(file = file.path(data_path_L1,"TropicalAndes_IUCNHabitat_Forest.rds"))
```

# Plot TD/Richness
## Function for richness plots
```{r}
create_rich_plots <- function(resolution_meters) {
 # Make Grid
  TAGrid <- TApoly %>%
    st_make_grid(cellsize = c(resolution_meters)) %>%
    st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
    st_cast("MULTIPOLYGON") %>%
    st_sf() %>%
    mutate(cellid = row_number())

  # Calculate cell richness for plants
  plant_richness_grid <- TAGrid %>%
    st_join(plants_sf_species) %>%
    mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
    group_by(cellid) %>%
    summarize(num_species = sum(overlap))

  # Calculate cell richness for frugivores
  frugivore_richness_grid <- TAGrid %>%
    st_join(frugivores_sf_species) %>%
    mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
    group_by(cellid) %>%
    summarize(num_species = sum(overlap))

  # Helper functions to generate plots
  generate_plant_plot <- function(data) {
    ggplot(data) +
      geom_sf(data = Americas, fill = "white") +
      geom_sf(data = TApoly, fill = "grey", size = 0.1) +
      geom_sf(aes(fill = num_species), color = NA) +
      scale_fill_distiller(palette = "Greens", direction = 1, limits = c(0,6000)) +
      labs( x = "Longitude", y = "Latitude", fill = "Number of species") +
      coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
      annotation_scale(location = "bl", width_hint = 0.5)  +
      theme(panel.background = element_rect(fill = "lightblue"))
  }
  generate_frugivore_plot <- function(data) {
    ggplot(data) +
      geom_sf(data = Americas, fill = "white") +
      geom_sf(data = TApoly, fill = "grey", size = 0.1) +
      geom_sf(aes(fill = num_species), color = NA) +
      scale_fill_distiller(palette = "YlOrBr", direction = 1, limits = c(0,500)) + 
      labs(x = "Longitude", y = "Latitude", fill = "Number of species") +
      coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
      annotation_scale(location = "bl", width_hint = 0.5)  +
      theme(panel.background = element_rect(fill = "lightblue"))
  }

  # Generate plots
  plant_plot <- generate_plant_plot(plant_richness_grid)
  frugivore_plot <- generate_frugivore_plot(frugivore_richness_grid)

  # Create histograms
  plant_richness_hist <- hist(comparison_data$Plant_Richness, 
    main = paste("Histogram of Plant Richness [", resolution_meters/1000, " km] by cell", sep = ""),
                              xlab = "Plant Richness by cell")
  frugivore_richness_hist <- hist(comparison_data$Frugivore_Richness,
    main = paste("Histogram of Frugivore Richness [", resolution_meters/1000, " km] by cell", sep = ""),
                                  xlab = "Frugivore Richness by cell")
  # Return a list of data
  list(plantgridRichTA = plant_plot,
       frugivoregridRichTA = frugivore_plot,
       plant_cellRichness = plant_richness_grid,
       frugivore_cellRichness = frugivore_richness_grid, 
       plant_richness_hist = plant_richness_hist,
       frugivore_richness_hist = frugivore_richness_hist)
}

```

## 1 km resolution
```{r}
richness_1km <- create_rich_plots(1000)
print(richness_1km)

# save plots
plantgridRichTA_1km <- richness_1km$plant_plot
frugivoregridRichTA_1km <- richness_1km$frugivoregridRichTA
# save richness data
plant_cellRichness_1km <- richness_1km$plant_cellRichness
frugivore_cellRichness_1km <- richness_1km$frugivore_cellRichness
```

## 5 km resolution
```{r}
richness_5km <- create_rich_plots(5000)
print(richness_5km)

# save plots
plantgridRichTA_5km <- richness_5km$plantgridRichTA
frugivoregridRichTA_5km <- richness_5km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_5km <- richness_5km$plant_cellRichness
frugivore_cellRichness_5km <- richness_5km$frugivore_cellRichness
```

## 10 km
```{r}
richness_10km <- create_rich_plots(10000)
print(richness_10km)

# save plots
plantgridRichTA_10km <- richness_10km$plantgridRichTA
frugivoregridRichTA_10km <- richness_10km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_10km <- richness_10km$plant_cellRichness
frugivore_cellRichness_10km <- richness_10km$frugivore_cellRichness
```

## 25 km
```{r}
richness_25km <- create_rich_plots(25000)
print(richness_25km)

# save plots
plantgridRichTA_25km <- richness_25km$plantgridRichTA
frugivoregridRichTA_25km <- richness_25km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_25km <- richness_25km$plant_cellRichness
frugivore_cellRichness_25km <- richness_25km$frugivore_cellRichness
```

```{r}

```

## 50 km
```{r}
richness_50km <- create_rich_plots(50000)
print(richness_50km)
# save plots
plantgridRichTA_50km <- richness_50km$plantgridRichTA
frugivoregridRichTA_50km <- richness_50km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_50km <- richness_50km$plant_cellRichness
frugivore_cellRichness_50km <- richness_50km$frugivore_cellRichness
```

## 75 km 
```{r}
richness_75km <- create_rich_plots(75000)
print(richness_75km)
# save plots
plantgridRichTA_75km <- richness_75km$plantgridRichTA
frugivoregridRichTA_75km <- richness_75km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_75km <- richness_75km$plant_cellRichness
frugivore_cellRichness_75km <- richness_75km$frugivore_cellRichness
```

## 100 km 
```{r}
richness_100km <- create_rich_plots(100000)
print(richness_100km)
# save plots
plantgridRichTA_100km <- richness_100km$plantgridRichTA
frugivoregridRichTA_100km <- richness_100km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_100km <- richness_100km$plant_cellRichness
frugivore_cellRichness_100km <- richness_100km$frugivore_cellRichness
```

# Multi panel plots
Richness maps
```{r}
all_richness_plots <- ggpubr::ggarrange(#plantgridRichTA_1km, plantgridRichTA_5km, 
                                              plantgridRichTA_10km, plantgridRichTA_25km,
                                              plantgridRichTA_50km, plantgridRichTA_75km,
                                              plantgridRichTA_100km,
                                              #frugivoregridRichTA_1km, frugivoregridRichTA_5km, 
                                              frugivoregridRichTA_10km, frugivoregridRichTA_25km,
                                              frugivoregridRichTA_50km, frugivoregridRichTA_75km,
                                              frugivoregridRichTA_100km, 
                                              ncol = 5, nrow = 2,
                                              labels = c("10km", "25km", ))
all_richness_plots
ggsave("all_richness_plots.pdf", all_richness_plots, path = output_path, width = 20, height = 8.5, units = "in")
```

# Write data to csv
```{r}
write.csv(plant_cellRichness_1km, file.path(output_path,"TropicalAndes_plantRichness_1km.csv"))
write.csv(frugivore_cellRichness_1km, file.path(output_path,"TropicalAndes_frugivoreRichness_1km.csv"))
```

```{r}
write.csv(plant_cellRichness_5km, file.path(output_path,"TropicalAndes_plantRichness_5km.csv"))
write.csv(frugivore_cellRichness_5km, file.path(output_path,"TropicalAndes_frugivoreRichness_5km.csv"))
```

```{r}
write.csv(plant_cellRichness_10km, file.path(output_path,"TropicalAndes_plantRichness_10km.csv"))
write.csv(frugivore_cellRichness_10km, file.path(output_path,"TropicalAndes_frugivoreRichness_10km.csv"))
```

```{r}
write.csv(plant_cellRichness_25km, file.path(output_path,"TropicalAndes_plantRichness_25km.csv"))
write.csv(frugivore_cellRichness_25km, file.path(output_path,"TropicalAndes_frugivoreRichness_25km.csv"))
```

```{r}
write.csv(plant_cellRichness_50km, file.path(output_path,"TropicalAndes_plantRichness_50km.csv"))
write.csv(frugivore_cellRichness_50km, file.path(output_path,"TropicalAndes_frugivoreRichness_50km.csv"))
```

```{r}
write.csv(plant_cellRichness_75km, file.path(output_path,"TropicalAndes_plantRichness_75km.csv"))
write.csv(frugivore_cellRichness_75km, file.path(output_path,"TropicalAndes_frugivoreRichness_75km.csv"))
```

```{r}
write.csv(plant_cellRichness_100km, file.path(output_path,"TropicalAndes_plantRichness_100km.csv"))
write.csv(frugivore_cellRichness_100km, file.path(output_path,"TropicalAndes_frugivoreRichness_100km.csv"))
```