### 2024-05-22 code/L0/L0_TropicalAndes_plants_FRUBASE.Rmd
inital commit

---
title: "Tropical Andes plant trait data FRUBASE"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script retrives plant trait data from the FRUBASE database for the Tropical Andes."
data input: ""
data output: ""
date: "2024-05-14"
output: 
    html_document:
      code_link: TRUE
---

# Set file paths
```{r}
data_path_L0 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path_L0 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
#devtools::install_github("pedroj/FRUBASE")
library(frubase)
library(dplyr)
```

# Retrive data from FRUBASE
Data downloaded on 2024-05-22
For more information about FRUBASE, see https://github.com/pedroj/FRUBASE 
```{r}
data(frubase)
```