### 2024-05-29 code/L1/L1_P3_TropicalAndes_plant_fruiting_subset.Rmd
inital commit

---
title: "Subsetting plant species list by fruiting species"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script subsets plant occurrence and trait data by fruiting plant species."
data input: "ADD"
data output: "ADD"
date: "2024-05-29"
output: html_document
---

# Load required packages
```{r}
library(dplyr)
library(tidyr)
```

# Set file paths
```{r}
data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Read in harmonized data
```{r}
plant_occ <- read.csv(file.path(data_path_L1, "TropicalAndes_GBIF_plant_occ_harmonized.csv"))
TRY_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_TRY_plant_traits_harmonized.csv"))
BIEN_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_BIEN_plant_traits_harmonized.csv"))
GIFT_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_GIFT_plant_traits_harmonized.csv"))
```

# Get species list from plant_occ data
```{r}
plant_species <- unique(plant_occ$Accepted_species)
length(plant_species)
```

# Subset TRY, BIEN, and GIFT by species list
## TRY
```{r}
nrow(TRY_traits)
TRY_filtered <- TRY_traits %>%
  filter(Accepted_species %in% plant_species)
nrow(TRY_filtered)
```

```{r}
TRY_traits %>%
  distinct(Accepted_species, TraitName) %>%
  count(TraitName)
```

## BIEN
```{r}
nrow(BIEN_traits)
BIEN_filtered <- BIEN_traits %>%
  filter(Accepted_species %in% plant_species)
nrow(BIEN_filtered)
```

```{r}
BIEN_traits %>%
  count(trait_name)
```

```{r}
BIEN_traits %>%
  distinct(Accepted_species, trait_name) %>%
  count(trait_name)
```

## GIFT
```{r}
nrow(GIFT_traits)
GIFT_filtered <- GIFT_traits %>%
  filter(Accepted_species %in% plant_species)
nrow(GIFT_filtered)
```

```{r}
GIFT_traits %>%
  count(trait_name)
```

```{r}
GIFT_traits %>%
  distinct(Accepted_species, trait_name) %>%
  count(trait_name)
```

# Combine trait information from TRY, BIEN, GIFT

```{r}
TRYTraitTypes <- TRY_filtered %>%
  group_by(TraitName) %>%
  distinct(DataName)
```

```{r}
unique(TRYTraitTypes$TraitName)
```

```{r}
colnames(TRY_filtered)
```
columns to keep : Accepted_species, TraitName, OrigValueStr

```{r}
TRY_filtered_subset <- TRY_filtered[ , c("Accepted_species", "TraitName", "OrigValueStr", "UnitName")]
colnames(TRY_filtered_subset) <- c("Accepted_species", "TraitName", "TraitValue", "Unit")
```

```{r}
# save sources/references
TRY_references <- unique(TRY_filtered[ , c("Dataset", "Reference")])
write.csv(TRY_references, file.path(output_path,"TRY_references.csv"))
```

```{r}
unique(BIEN_filtered$trait_name)
```

```{r}
colnames(BIEN_filtered)
```
columns to keep : Accepted_species, trait_name, trait_value, unit

```{r}
BIEN_filtered_subset <- BIEN_filtered[ , c("Accepted_species", "trait_name", "trait_value", "unit")]
colnames(BIEN_filtered_subset) <- c("Accepted_species", "TraitName", "TraitValue", "Unit")
```

```{r}
# save sources/references
BIEN_references <- unique(BIEN_filtered[ , c("url_source", "source_citation")])
write.csv(BIEN_references, file.path(output_path,"BIEN_references.csv"))
```

```{r}
unique(GIFT_filtered$trait_name)
```

```{r}
colnames(GIFT_filtered)
```
columns to keep : Accepted_species, trait_value, trait_name

```{r}
GIFT_filtered_subset <- GIFT_filtered[ , c("Accepted_species", "trait_name", "trait_value")]
colnames(GIFT_filtered_subset) <- c("Accepted_species", "TraitName", "TraitValue")
```

Add Unit column to GIFT dataframe
```{r}
GIFT_filtered_subset <- GIFT_filtered_subset %>%
  mutate(Unit = case_when(
  TraitName == "plant_height" ~ "m",
  TraitName == "plant_height_max" ~ "m",
  TraitName == "plant_height_min" ~ "m",
  TraitName == "seed_mass_mean" ~ "g",
  TraitName == "seed_mass_min" ~ "g",
  TraitName == "seed_mass_max" ~ "g",
  TraitName == "fruit_length_min" ~ "cm",
  TraitName == "fruit_length_max" ~ "cm",
  TraitName == "fruit_length_mean" ~ "cm",
  TraitName == "seed_length_max" ~ "mm",
  TraitName == "seed_length_min" ~ "mm",
  TraitName == "seed_length_mean" ~ "mm",
  TraitName == "seed_width_max" ~ "mm",
  TraitName == "seed_width_min" ~ "mm",
  TraitName == "seed_width_mean" ~ "mm",
  TraitName == "plant_lifespan" ~ "years",
  .default = ""
  ))
```

```{r}
traits <- rbind(TRY_filtered_subset, BIEN_filtered_subset, GIFT_filtered_subset)
nrow(traits)
```

```{r}
# remove rows without accepted species names
traits_clean <- traits[!(is.na(traits$Accepted_species) | traits$Accepted_species==""), ]
nrow(traits_clean)
```

```{r}
traits %>%
  count(TraitName)
```

```{r}
traits %>%
  distinct(Accepted_species, TraitName) %>%
  count(TraitName)
```

```{r}
# Group by TraitName and summarize the unique TraitValues
unique_trait_values <- traits %>%
  group_by(TraitName) %>%
  summarize(UniqueValues = list(unique(TraitValue)))%>%
  unnest(UniqueValues)
```

# Function to get one value per species
```{r}
# categorical data
factor_data_merge <- function(data, traitname) {
# remove duplicate rows
data <- data %>%
  distinct(.keep_all = TRUE)
# remove rows with missing or empty species names
data <- data[!is.na(data$Accepted_species) & data$Accepted_species != "", ]
# Subset data to just species and TraitValue
data <- data[,c("Accepted_species", "TraitValue")]
# Find the most common trait value for each species
  most_common_values <- aggregate(. ~ Accepted_species, data, function(x) {
    # Find the mode (most common value) of the trait values
    mode_value <- names(sort(table(x), decreasing = TRUE))[1]
    return(mode_value)
  })
# Merge most common trait values back to the original data
merged_data <- merge(data, most_common_values, by = "Accepted_species", all.x = TRUE)

# Rename the merged column
colnames(merged_data)[ncol(merged_data)] <- "TraitValue"

# Remove column TraitValue.x
merged_data <- merged_data[, !colnames(merged_data) %in% "TraitValue.x", drop = FALSE]

# remove duplicate rows
merged_data <- merged_data %>%
  distinct(.keep_all = TRUE)

# Add unit & trait column
merged_data$TraitName <- traitname

  return(merged_data)
}
```

# Fruit type
```{r}
traits_fruitType <- traits %>%
  filter(grepl('Fruit type|fruit type|fruit_type', TraitName))
traits_fruitType$OrigTraitValue <- traits_fruitType$TraitValue

nrow(traits_fruitType)
length(unique(traits_fruitType$Accepted_species))
```

```{r}
traits_fruitType <- traits_fruitType %>%
  filter(is.na(as.numeric(TraitValue)))

nrow(traits_fruitType)
length(unique(traits_fruitType$Accepted_species))
```

```{r}
traits_fruitType %>%
  count(TraitValue)
```

```{r}
# standardize trait values
traits_fruitType <- traits_fruitType %>%
  filter() %>%
  mutate(TraitValue= case_when(
    grepl('Berry|berry|Berry ', OrigTraitValue, ignore.case = TRUE) ~ "berry",
    grepl('capsule|Capsule|Capsule ', OrigTraitValue, ignore.case = TRUE) ~ "capsule",
    grepl('follicle|Follicle', OrigTraitValue, ignore.case = TRUE) ~ "follicle",
  )) %>%
  filter(!is.na(TraitValue))
```

```{r}
# standardize trait name
traits_fruitType <- traits_fruitType %>%
  mutate(TraitName = "FruitType")
```

```{r}
averagetraits_fruitType <- factor_data_merge(traits_fruitType, "Fruit Type")
```

```{r}
nrow(averagetraits_fruitType)
length(unique(averagetraits_fruitType$Accepted_species))
```

```{r}
summary(averagetraits_fruitType)
```