### 2024-05-29 code/L1/L1_P6_TropicalAndes_plant_occ.Rmd

---
title: "Plant occurrence subset by species with complete trait records"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script subsets plant occurrence data to species with complete trait coverage."
data input: "ADD"
data output: "ADD"
date: "2023-10-04"
output: html_document
---

# Set file paths
```{r}
data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path_L1 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Load required packages
```{r}
library(dplyr)
#run if not installed remotes::install_github("FRBCesab/funbiogeo")
library(funbiogeo)
```

# Read in data
```{r}
plant_traits <- read.csv(file.path(data_path_L1, file = "TropicalAndes_imputed_plant_traits.csv"))
plant_occ <- read.csv(file.path(data_path_L1, file = "TropicalAndes_GBIF_plant_occ_cleaned.csv"))
```

# Check trait coverage
```{r}
summary(plant_traits)
```

```{r}
fb_plot_number_species_by_trait(plant_traits)
```

```{r}
Plant_complete <- fb_filter_species_by_trait_coverage(plant_traits, threshold_traits_proportion =  1)
```

# Create list of species with complete trait records
```{r}
plant_species <- unique(Plant_complete$species)
```

# Subset frugivore occurrence data by species list
```{r}
plant_occ_subset <- plant_occ %>%
  filter(species %in% plant_species)
```

# Summary
```{r}
glimpse(plant_occ_subset)
```

```{r}
source("~/GitHub/neotropical_plants/functions/data_summary.R")
```

```{r}
data_summary(plant_occ_subset, plant_occ_subset$species, plant_occ_subset$genus, plant_occ_subset$family)
```

# Write data to csv
```{r}
write.csv(plant_occ_subset, file.path(output_path_L1,"TropicalAndes_GBIF_plant_occ_cleaned_subset.csv"))
```