### 2024-06-03 code/L2/L2_3_TropicalAndes_divRelationships.Rmd
update input files

---
title: "Diversity relationships for plants and frugivores"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script compares taxonomic and functional diversity relationships for plants and frugivores for various spatial grains and combined diversity for each grain."
data input: "ADD"
data output: "ADD"
date: "2024-05-16"
output: html_document
---

# Set file paths
```{r}
data_path_L2 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
figure_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/figures')
```

# Load required packages
```{r}
library(dplyr)
library(ggplot2)
library(smoothr)
library(purrr)
```

# Read in Data
```{r}
plant_cellRichness_1km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantRichness_1km.csv"))
frugivore_cellRichness_1km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreRichness_1km.csv"))

plant_cellRichness_5km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantRichness_5km.csv"))
frugivore_cellRichness_5km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreRichness_5km.csv"))

plant_cellRichness_10km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantRichness_10km.csv"))
frugivore_cellRichness_10km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreRichness_10km.csv"))

plant_cellRichness_25km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantRichness_25km.csv"))
frugivore_cellRichness_25km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreRichness_25km.csv"))

plant_cellRichness_50km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantRichness_50km.csv"))
frugivore_cellRichness_50km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreRichness_50km.csv"))

plant_cellRichness_75km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantRichness_75km.csv"))
frugivore_cellRichness_75km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreRichness_75km.csv"))

plant_cellRichness_100km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantRichness_100km.csv"))
frugivore_cellRichness_100km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreRichness_100km.csv"))
```

# 1 km
# 5 km
# 10 km
# 25 km
# 50 km
# 75 km
# 100 km
```{r}
  # Extract cell values
  comparison_data_100km <- data.frame(Plant_Richness = plant_richness_grid_100km$num_species,
                               Frugivore_Richness = frugivore_richness_grid_100km$num_species)

  # Filter out rows where either plant or frugivore richness is zero
  comparison_data_filtered_100km <- comparison_data %>%
                                filter(Plant_Richness > 0 & Frugivore_Richness > 0)
```

```{r}
# Compare plant and frugivore richness
  ## Fit a linear model
  compare_richness_linear_summary_100km <- summary(lm(comparison_data_filtered_100km$Frugivore_Richness ~
                                                  comparison_data_filtered_100km$Plant_Richness))
  linear_r_squared_100km <- compare_richness_linear_summary_100km$r.squared

  ## Fit an exponential model
  compare_richness_exponential_summary_100km <- summary(lm(comparison_data_filtered_100km$Frugivore_Richness ~
                                                       log(comparison_data_filtered_100km$Plant_Richness)))
  exponential_r_squared_100km <- compare_richness_exponential_summary_100km$r.squared

  ## Create scatter plot with linear and exponential lines
  compare_richness_plot_100km <- ggplot(comparison_data_filtered_100km,
                                        aes(x = Plant_Richness, y = Frugivore_Richness)) +
  geom_point() +
  labs(title = paste("Plant Richness vs Frugivore Richness [", resolution_meters/1000, " km]", sep = ""),
       subtitle =  paste("Linear (Red) R-squared = ", round(linear_r_squared, 2), "\n",
                  "Exponential (Blue) R-squared = ", round(exponential_r_squared, 2)),
       x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
  geom_smooth(method = "lm", se = FALSE, color = "red")+
  geom_smooth(method = "lm", se = FALSE, color = "blue", formula = y ~ log(x)) +
  coord_cartesian(xlim = c(0, NA), ylim = c(0, NA))
```

```{r}
# Fit the nonlinear model using nls
nls_model <- nls(Frugivore_Richness ~ b * exp(c * Plant_Richness), 
                 data = comparison_data_filtered_100km, 
                 start = list(b = 200, c = -0.00001),
                 algorithm = "port",
                 control = nls.control(maxiter = 500, minFactor = 1e-10, warnOnly = TRUE))

# Summary of the model
summary(nls_model)
```

# Compare r-squared by resolution
## Taxonomic diversity
```{r}
# create dataframe of resolutions and r-squared values 
#[ADD 1, 5,]
resolution <- c( 10, 25, 50, 75, 100)
#ADD compare_richness_r2_1km, compare_richness_r2_5km,
r_squared <- c(compare_richness_r2_10km, compare_richness_r2_25km, compare_richness_r2_50km, compare_richness_r2_75km, compare_richness_r2_100km)
r_squared_compare <- data.frame(resolution, r_squared)

compare_r2_by_res <- ggplot(r_squared_compare, aes(x=resolution, y=r_squared)) +
    geom_point()+
    labs(title = "R-squared values of plant-frugivore richness relationships by resolution",
         x = "Resolution (km)", y = "R-squared", fill = "Number of species")
print(compare_r2_by_res)
```
## Functional Diversity
```{r}

```