### 2024-06-03 code/L1/L1_LastStep_DiversityInputObjects.Rmd
remove 1 km

---
title: ""
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "Projecting occurrence data and creating presence-absence matrix for plants and frugivores. "
data input: 
data output: 
date: "2024-05-13"
output: html_document
---

# Load required packages
```{r}
library(letsR)
library(mFD)
library(vegan)
library(rnaturalearth)
library(sf)
library(raster)
library(fasterize)
library(funbiogeo)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpubr)
```

# Set file paths
```{r}
data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path_L1 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
```

# Read in Data
```{r}
TropicalAndes_plant_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_plant_occ_cleaned_subset.csv"))
TropicalAndes_frugivore_occ_forest <- read.csv(file.path(data_path_L1,"TropicalAndes_GBIF_frugivore_occ_cleaned_subset.csv"))
TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path_L0, "Forest_sf.shp"), layer = "Forest_sf")
frugivore_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_Frugivoria_traits_Forest.csv"))
plant_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_imputed_plant_traits.csv"))
```

# Convert data to spatial data
```{r}
plants.sf <- st_as_sf(TropicalAndes_plant_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)

frugivores.sf <- st_as_sf(TropicalAndes_frugivore_occ_forest, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
```

# Polygons of countries
```{r}
worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")
Americas <- ne_countries(continent = c("North America", "South America"), returnclass = "sf")
#polygon of Tropical Andes
TApoly <- worldMap %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
```

# Transform to projected coordinate reference system (units from degrees to meters)
```{r}
Americas <- st_transform(Americas, 5389)
TApoly <- st_transform(TApoly, 5389)
TropicalAndes_IUCNHabitat_Forest <- st_transform(TropicalAndes_IUCNHabitat_Forest, 5389)
plants.sf <- st_transform(plants.sf, 5389)
frugivores.sf <- st_transform(frugivores.sf, 5389)
```

```{r}
#check units
st_crs(TApoly, parameters = TRUE)$units_gdal
st_crs(TApoly)
st_crs(TropicalAndes_IUCNHabitat_Forest)
```

# Group by species
```{r}
plants_sf_species <- plants.sf %>%
  group_by(species) %>%
  summarise()

frugivores_sf_species <- frugivores.sf %>%
  group_by(species) %>%
  summarise()
```

# Plot Base Map
```{r}
basePlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly) +
  geom_sf(data = TropicalAndes_IUCNHabitat_Forest, fill = "forestgreen")+
  labs(title = "Tropical Andes", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

basePlot
```

# Plot points
```{r}
plantsPointsPlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly) +
  geom_sf(data = plants_sf_species, pch = 21) +
  labs(title = "Tropical Andes Plant Occurrences", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

plantsPointsPlot
```

```{r}
frugivoresPointsPlot <-
ggplot() +
  geom_sf(data = Americas, fill = "white") +
  geom_sf(data = TApoly) +
  geom_sf(data = frugivores_sf_species, pch = 21) +
  labs(title = "Tropical Andes Frugivore Occurrences", x = "Latitude", y = "Longitude") +
  coord_sf(xlim = c(-85, -54), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(location = "bl", which_north = "true", 
    pad_x = unit(0.3, "in"), pad_y = unit(0.3, "in"), style = north_arrow_fancy_orienteering) +
  theme(panel.background = element_rect(fill = "lightblue"))

frugivoresPointsPlot
```

# Create Presence-Absense matrix

```{r}
create_presence_absence_matrix <- function(resolution_meters, species_sf) {
  # Make Grid
  TAGrid <- TApoly %>%
    st_make_grid(cellsize = c(resolution_meters)) %>%
    st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
    st_cast("MULTIPOLYGON") %>%
    st_sf() %>%
    mutate(cellid = row_number())

  # Join with species data and filter out NA species
  species_grid <- TAGrid %>%
    st_join(species_sf) %>%
    filter(!is.na(species)) %>%
    dplyr::select(cellid, species, geometry)

  # Extract the coordinates for each grid cell
  species_grid_coords <- species_grid %>%
    st_centroid() %>%
    st_coordinates() %>%
    as.data.frame() %>%
    rename(Longitude = X, Latitude = Y)

  # Combine coordinates with the species grid
  species_grid <- bind_cols(species_grid, species_grid_coords)

  # Create Presence-Absence Matrix
  presence_absence_matrix <- species_grid %>%
    st_set_geometry(NULL) %>%
    mutate(presence = 1) %>%
    pivot_wider(names_from = species, values_from = presence, values_fill = list(presence = 0))

  # Convert to matrix
  presence_absence_matrix_matrix <- as.matrix(presence_absence_matrix)

  # Set row names as "cell_rowNumber"
  rownames(presence_absence_matrix_matrix) <- paste0("cell_", seq_len(nrow(presence_absence_matrix_matrix)))

  # Ensure latitude and longitude are the first columns
  presence_absence_matrix_matrix <- presence_absence_matrix_matrix[, c("Latitude", "Longitude", setdiff(colnames(presence_absence_matrix_matrix), c("Latitude", "Longitude", "cellid")))]

  return(presence_absence_matrix_matrix)
}
```

## 100 km 
```{r}
plant_PAM_100km <- create_presence_absence_matrix(100000, plants_sf_species)
frugivore_PAM_100km <- create_presence_absence_matrix(100000, frugivores_sf_species)
```

```{r}
#check str 
plant_PAM_100km[1:4, 1:4]
frugivore_PAM_100km[1:4, 1:4]
```

### Remove the species from PAM that have no occurrences
```{r}
dim(plant_PAM_100km)
dim(frugivore_PAM_100km)
# Remove columns with sum equal to zero
PAM_plant_site_final_100km <- plant_PAM_100km[, colSums(plant_PAM_100km) != 0]
PAM_frugivore_site_final_100km <- frugivore_PAM_100km[, colSums(frugivore_PAM_100km) != 0]

dim(PAM_plant_site_final_100km)
dim(PAM_frugivore_site_final_100km)

# Save coordinates for later
site_loc_key_plant_100km <- PAM_plant_site_final_100km[,1:2]
site_loc_key_frugivore_100km <- PAM_frugivore_site_final_100km[,1:2]

PAM_plant_site_final_100km <- PAM_plant_site_final_100km[,-c(1:2)]
PAM_frugivore_site_final_100km <- PAM_frugivore_site_final_100km[,-(1:2)]

colnames_plant_100km <- colnames(PAM_plant_site_final_100km)
colnames_frugivore_100km <- colnames(PAM_frugivore_site_final_100km)

```

```{r}
str(PAM_plant_site_final_100km)
str(PAM_frugivore_site_final_100km)
```
## 75 km 
```{r}
plant_PAM_75km <- create_presence_absence_matrix(75000, plants_sf_species)
frugivore_PAM_75km <- create_presence_absence_matrix(75000, frugivores_sf_species)
```

```{r}
#check str 
plant_PAM_75km[1:4, 1:4]
frugivore_PAM_75km[1:4, 1:4]
```

### Remove the species from PAM that have no occurrences
```{r}
# Remove columns with sum equal to zero
PAM_plant_site_final_75km <- plant_PAM_75km[, colSums(plant_PAM_75km) != 0]
PAM_frugivore_site_final_75km <- frugivore_PAM_75km[, colSums(frugivore_PAM_75km) != 0]

#Save coordinates for later
site_loc_key_plant_75km <- PAM_plant_site_final_75km[,1:2]
site_loc_key_frugivore_75km <- PAM_frugivore_site_final_75km[,1:2]

PAM_plant_site_final_75km <- PAM_plant_site_final_75km[,-c(1:2)]
PAM_frugivore_site_final_75km <- PAM_frugivore_site_final_75km[,-(1:2)]

colnames_plant_75km <- colnames(PAM_plant_site_final_75km)
colnames_frugivore_75km <- colnames(PAM_frugivore_site_final_75km)
```

```{r}
str(PAM_plant_site_final_75km)
str(PAM_frugivore_site_final_75km)
```

## 50 km 
```{r}
plant_PAM_50km <- create_presence_absence_matrix(50000, plants_sf_species)
frugivore_PAM_50km <- create_presence_absence_matrix(50000, frugivores_sf_species)
```

```{r}
#check str 
plant_PAM_50km[1:4, 1:4]
frugivore_PAM_50km[1:4, 1:4]
```

### Remove the species from PAM that have no occurrences
```{r}
# Remove columns with sum equal to zero
PAM_plant_site_final_50km <- plant_PAM_50km[, colSums(plant_PAM_50km) != 0]
PAM_frugivore_site_final_50km <- frugivore_PAM_50km[, colSums(frugivore_PAM_50km) != 0]

#Save coordinates for later
site_loc_key_plant_50km <- PAM_plant_site_final_50km[,1:2]
site_loc_key_frugivore_50km <- PAM_frugivore_site_final_50km[,1:2]

PAM_plant_site_final_50km <- PAM_plant_site_final_50km[,-c(1:2)]
PAM_frugivore_site_final_50km <- PAM_frugivore_site_final_50km[,-(1:2)]

colnames_plant_50km <- colnames(PAM_plant_site_final_50km)
colnames_frugivore_50km <- colnames(PAM_frugivore_site_final_50km)
```

```{r}
str(PAM_plant_site_final_50km)
str(PAM_frugivore_site_final_50km)
```

## 25 km 
```{r}
plant_PAM_25km <- create_presence_absence_matrix(25000, plants_sf_species)
frugivore_PAM_25km <- create_presence_absence_matrix(25000, frugivores_sf_species)
```

```{r}
#check str 
plant_PAM_25km[1:4, 1:4]
frugivore_PAM_25km[1:4, 1:4]
```

### Remove the species from PAM that have no occurrences
```{r}
# Remove columns with sum equal to zero
PAM_plant_site_final_25km <- plant_PAM_25km[, colSums(plant_PAM_25km) != 0]
PAM_frugivore_site_final_25km <- frugivore_PAM_25km[, colSums(frugivore_PAM_25km) != 0]

#Save coordinates for later
site_loc_key_plant_25km <- PAM_plant_site_final_25km[,1:2]
site_loc_key_frugivore_25km <- PAM_frugivore_site_final_25km[,1:2]

PAM_plant_site_final_25km <- PAM_plant_site_final_25km[,-c(1:2)]
PAM_frugivore_site_final_25km <- PAM_frugivore_site_final_25km[,-(1:2)]

colnames_plant_25km <- colnames(PAM_plant_site_final_25km)
colnames_frugivore_25km <- colnames(PAM_frugivore_site_final_25km)
```

```{r}
str(PAM_plant_site_final_25km)
str(PAM_frugivore_site_final_25km)
```

## 10 km 
```{r}
plant_PAM_10km <- create_presence_absence_matrix(10000, plants_sf_species)
frugivore_PAM_10km <- create_presence_absence_matrix(10000, frugivores_sf_species)
```

```{r}
#check str 
plant_PAM_10km[1:4, 1:4]
frugivore_PAM_10km[1:4, 1:4]
```

### Remove the species from PAM that have no occurrences
```{r}
# Remove columns with sum equal to zero
PAM_plant_site_final_10km <- plant_PAM_10km[, colSums(plant_PAM_10km) != 0]
PAM_frugivore_site_final_10km <- frugivore_PAM_10km[, colSums(frugivore_PAM_10km) != 0]

#Save coordinates for later
site_loc_key_plant_10km <- PAM_plant_site_final_10km[,1:2]
site_loc_key_frugivore_10km <- PAM_frugivore_site_final_10km[,1:2]

PAM_plant_site_final_10km <- PAM_plant_site_final_10km[,-c(1:2)]
PAM_frugivore_site_final_10km <- PAM_frugivore_site_final_10km[,-(1:2)]

colnames_plant_10km <- colnames(PAM_plant_site_final_10km)
colnames_frugivore_10km <- colnames(PAM_frugivore_site_final_10km)
```

```{r}
str(PAM_plant_site_final_10km)
str(PAM_frugivore_site_final_10km)
```

## 5 km 
```{r}
plant_PAM_5km <- create_presence_absence_matrix(5000, plants_sf_species)
frugivore_PAM_5km <- create_presence_absence_matrix(5000, frugivores_sf_species)
```

```{r}
#check str 
plant_PAM_5km[1:4, 1:4]
frugivore_PAM_5km[1:4, 1:4]
```

### Remove the species from PAM that have no occurrences
```{r}
# Remove columns with sum equal to zero
PAM_plant_site_final_5km <- plant_PAM_5km[, colSums(plant_PAM_5km) != 0]
PAM_frugivore_site_final_5km <- frugivore_PAM_5km[, colSums(frugivore_PAM_5km) != 0]

#Save coordinates for later
site_loc_key_plant_5km <- PAM_plant_site_final_5km[,1:2]
site_loc_key_frugivore_5km <- PAM_frugivore_site_final_5km[,1:2]

PAM_plant_site_final_5km <- PAM_plant_site_final_5km[,-c(1:2)]
PAM_frugivore_site_final_5km <- PAM_frugivore_site_final_5km[,-(1:2)]

colnames_plant_5km <- colnames(PAM_plant_site_final_5km)
colnames_frugivore_5km <- colnames(PAM_frugivore_site_final_5km)
```

```{r}
str(PAM_plant_site_final_5km)
str(PAM_frugivore_site_final_5km)
```

## Remove species names from trait matrix not in the PAM
```{r}
plant_traits_df_subset <- plant_traits %>%
                            filter(species %in% colnames_plant_100km) %>%
                            distinct(species, .keep_all = TRUE)
frugivore_traits_df_subset <- frugivore_traits %>%
                                filter(IUCN_species_name %in% colnames_frugivore_100km) %>%
                                distinct(IUCN_species_name, .keep_all = TRUE)

dim(plant_traits)
dim(plant_traits_df_subset)

dim(frugivore_traits)
dim(frugivore_traits_df_subset)
```

```{r}
# Define row names as species names
row_names_plant <- plant_traits_df_subset$species
row_names_frugivore <- frugivore_traits_df_subset$IUCN_species_name

# Assign row names to the matrix
rownames(plant_traits_df_subset) <- row_names_plant
rownames(frugivore_traits_df_subset) <- row_names_frugivore

plant_traits_df_subset$X <-NULL
frugivore_traits_df_subset$X <-NULL
# remove duplicate species name column
plant_traits_df_subset <- plant_traits_df_subset[, c("Growth.Form", "Plant.Height_m", "Fruit.Length_mm", "Fruit.Type")]
frugivore_traits_df_subset <- frugivore_traits_df_subset[, c("diet_cat", "body_mass_e", "body_size_mm", "generation_time")]
```

```{r}
str(plant_traits_df_subset)
str(frugivore_traits_df_subset)
```

# Export data
## sf objects
```{r}
saveRDS(plants_sf_species, file = file.path(data_path_L1,"plants_sf_species.rds"))
saveRDS(frugivores_sf_species, file = file.path(data_path_L1,"frugivores_sf_species.rds"))
saveRDS(Americas, file = file.path(data_path_L1, "Americas.rds"))
saveRDS(TApoly, file = file.path(data_path_L1,"TApoly.rds"))
saveRDS(TropicalAndes_IUCNHabitat_Forest, file = file.path(data_path_L1,"TropicalAndes_IUCNHabitat_Forest.rds"))
```

# PAM objects
```{r}
saveRDS(plant_PAM_100km, file = file.path(data_path_L1,"plant_PAM_100km.rds"))
saveRDS(frugivore_PAM_100km, file = file.path(data_path_L1,"frugivore_PAM_100km.rds"))
```

```{r}
saveRDS(plant_PAM_75km, file = file.path(data_path_L1,"plant_PAM_75km.rds"))
saveRDS(frugivore_PAM_75km, file = file.path(data_path_L1,"frugivore_PAM_75km.rds"))
```

```{r}
saveRDS(plant_PAM_50km, file = file.path(data_path_L1,"plant_PAM_50km.rds"))
saveRDS(frugivore_PAM_50km, file = file.path(data_path_L1,"frugivore_PAM_50km.rds"))
```

```{r}
saveRDS(plant_PAM_25km, file = file.path(data_path_L1,"plant_PAM_25km.rds"))
saveRDS(frugivore_PAM_25km, file = file.path(data_path_L1,"frugivore_PAM_25km.rds"))
```

```{r}
saveRDS(plant_PAM_10km, file = file.path(data_path_L1,"plant_PAM_10km.rds"))
saveRDS(frugivore_PAM_10km, file = file.path(data_path_L1,"frugivore_PAM_10km.rds"))
```

```{r}
saveRDS(plant_PAM_5km, file = file.path(data_path_L1,"plant_PAM_5km.rds"))
saveRDS(frugivore_PAM_5km, file = file.path(data_path_L1,"frugivore_PAM_5km.rds"))
```

## objects for functional diversity
```{r}
saveRDS(plant_traits_df_subset, file = file.path(data_path_L1,"plant_traits_df_final.rds"))
saveRDS(frugivore_traits_df_subset, file = file.path(data_path_L1,"frugivore_traits_df_final.rds"))
```

```{r}
saveRDS(site_loc_key_plant_100km, file = file.path(data_path_L1,"site_loc_key_plant_100km.rds"))
saveRDS(site_loc_key_frugivore_100km, file = file.path(data_path_L1,"site_loc_key_frugivore_100km.rds"))
saveRDS(PAM_plant_site_final_100km, file = file.path(data_path_L1,"PAM_plant_site_final_100km.rds"))
saveRDS(PAM_frugivore_site_final_100km, file = file.path(data_path_L1,"PAM_frugivore_site_final_100km.rds"))
```

```{r}
saveRDS(site_loc_key_plant_75km, file = file.path(data_path_L1,"site_loc_key_plant_75km.rds"))
saveRDS(site_loc_key_frugivore_75km, file = file.path(data_path_L1,"site_loc_key_frugivore_75km.rds"))
saveRDS(PAM_plant_site_final_75km, file = file.path(data_path_L1,"PAM_plant_site_final_75km.rds"))
saveRDS(PAM_frugivore_site_final_75km, file = file.path(data_path_L1,"PAM_frugivore_site_final_75km.rds"))
```

```{r}
saveRDS(site_loc_key_plant_50km, file = file.path(data_path_L1,"site_loc_key_plant_50km.rds"))
saveRDS(site_loc_key_frugivore_50km, file = file.path(data_path_L1,"site_loc_key_frugivore_50km.rds"))
saveRDS(PAM_plant_site_final_50km, file = file.path(data_path_L1,"PAM_plant_site_final_50km.rds"))
saveRDS(PAM_frugivore_site_final_50km, file = file.path(data_path_L1,"PAM_frugivore_site_final_50km.rds"))
```

```{r}
saveRDS(site_loc_key_plant_25km, file = file.path(data_path_L1,"site_loc_key_plant_25km.rds"))
saveRDS(site_loc_key_frugivore_25km, file = file.path(data_path_L1,"site_loc_key_frugivore_25km.rds"))
saveRDS(PAM_plant_site_final_25km, file = file.path(data_path_L1,"PAM_plant_site_final_25km.rds"))
saveRDS(PAM_frugivore_site_final_25km, file = file.path(data_path_L1,"PAM_frugivore_site_final_25km.rds"))
```

```{r}
saveRDS(site_loc_key_plant_10km, file = file.path(data_path_L1,"site_loc_key_plant_10km.rds"))
saveRDS(site_loc_key_frugivore_10km, file = file.path(data_path_L1,"site_loc_key_frugivore_10km.rds"))
saveRDS(PAM_plant_site_final_10km, file = file.path(data_path_L1,"PAM_plant_site_final_10km.rds"))
saveRDS(PAM_frugivore_site_final_10km, file = file.path(data_path_L1,"PAM_frugivore_site_final_10km.rds"))
```

```{r}
saveRDS(site_loc_key_plant_5km, file = file.path(data_path_L1,"site_loc_key_plant_5km.rds"))
saveRDS(site_loc_key_frugivore_5km, file = file.path(data_path_L1,"site_loc_key_frugivore_5km.rds"))
saveRDS(PAM_plant_site_final_5km, file = file.path(data_path_L1,"PAM_plant_site_final_5km.rds"))
saveRDS(PAM_frugivore_site_final_5km, file = file.path(data_path_L1,"PAM_frugivore_site_final_5km.rds"))
```