### 2024-06-10 code/L1/L1_P3_TropicalAndes_plant_fruiting_subset.Rmd
update to create species list with any species that has information related to a fruiting trait

---
title: "Subsetting plant species list by fruiting species"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script subsets plant occurrence and trait data by fruiting plant species."
data input: "TropicalAndes_GBIF_plant_occ_harmonized.csv, TropicalAndes_TRY_plant_traits_harmonized.csv,
             TropicalAndes_BIEN_plant_traits_harmonized.csv, TropicalAndes_GIFT_plant_traits_harmonized.csv"
data output: "TropicalAndes_GBIF_plant_occ_harmonized_subset.csv, TropicalAndes_all_plant_traits_harmonized_subset.csv"
date: "2024-05-29"
output: html_document
---

# Load required packages
```{r}
library(dplyr)
library(tidyr)
```

# Set file paths
```{r}
data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
data_path_L1 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
figure_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/figures')
```

# Read in harmonized data
```{r}
plant_occ <- read.csv(file.path(data_path_L1, "TropicalAndes_GBIF_plant_occ_harmonized.csv"))
TRY_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_TRY_plant_traits_harmonized.csv"))
BIEN_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_BIEN_plant_traits_harmonized.csv"))
GIFT_traits <- read.csv(file.path(data_path_L1,"TropicalAndes_GIFT_plant_traits_harmonized.csv"))
```

# Get species list from plant_occ data
```{r}
plant_species <- unique(plant_occ$Accepted_species)
length(plant_species)
```

# Subset TRY, BIEN, and GIFT by species list
## TRY
```{r}
nrow(TRY_traits)
TRY_filtered <- TRY_traits %>%
  filter(Accepted_species %in% plant_species)
nrow(TRY_filtered)
```

```{r}
TRY_traits %>%
  distinct(Accepted_species, TraitName) %>%
  count(TraitName)
```

## BIEN
```{r}
nrow(BIEN_traits)
BIEN_filtered <- BIEN_traits %>%
  filter(Accepted_species %in% plant_species)
nrow(BIEN_filtered)
```

```{r}
BIEN_traits %>%
  count(trait_name)
```

```{r}
BIEN_traits %>%
  distinct(Accepted_species, trait_name) %>%
  count(trait_name)
```

## GIFT
```{r}
nrow(GIFT_traits)
GIFT_filtered <- GIFT_traits %>%
  filter(Accepted_species %in% plant_species)
nrow(GIFT_filtered)
```

```{r}
GIFT_traits %>%
  count(trait_name)
```

```{r}
GIFT_traits %>%
  distinct(Accepted_species, trait_name) %>%
  count(trait_name)
```

# Combine trait information from TRY, BIEN, GIFT

```{r}
TRYTraitTypes <- TRY_filtered %>%
  group_by(TraitName) %>%
  distinct(DataName)
```

```{r}
unique(TRYTraitTypes$TraitName)
```

```{r}
colnames(TRY_filtered)
```
columns to keep : Accepted_species, TraitName, OrigValueStr

```{r}
TRY_filtered_subset <- TRY_filtered[ , c("Accepted_species", "TraitName", "OrigValueStr", "UnitName")]
colnames(TRY_filtered_subset) <- c("Accepted_species", "TraitName", "TraitValue", "Unit")
```

```{r}
# save sources/references
TRY_references <- unique(TRY_filtered[ , c("Dataset", "Reference")])
write.csv(TRY_references, file.path(output_path,"TRY_references.csv"))
```

```{r}
unique(BIEN_filtered$trait_name)
```

```{r}
colnames(BIEN_filtered)
```
columns to keep : Accepted_species, trait_name, trait_value, unit

```{r}
BIEN_filtered_subset <- BIEN_filtered[ , c("Accepted_species", "trait_name", "trait_value", "unit")]
colnames(BIEN_filtered_subset) <- c("Accepted_species", "TraitName", "TraitValue", "Unit")
```

```{r}
# save sources/references
BIEN_references <- unique(BIEN_filtered[ , c("url_source", "source_citation")])
write.csv(BIEN_references, file.path(output_path,"BIEN_references.csv"))
```

```{r}
unique(GIFT_filtered$trait_name)
```

```{r}
colnames(GIFT_filtered)
```
columns to keep : Accepted_species, trait_value, trait_name

```{r}
GIFT_filtered_subset <- GIFT_filtered[ , c("Accepted_species", "trait_name", "trait_value")]
colnames(GIFT_filtered_subset) <- c("Accepted_species", "TraitName", "TraitValue")
```

Add Unit column to GIFT dataframe
```{r}
GIFT_filtered_subset <- GIFT_filtered_subset %>%
  mutate(Unit = case_when(
  TraitName == "plant_height" ~ "m",
  TraitName == "plant_height_max" ~ "m",
  TraitName == "plant_height_min" ~ "m",
  TraitName == "seed_mass_mean" ~ "g",
  TraitName == "seed_mass_min" ~ "g",
  TraitName == "seed_mass_max" ~ "g",
  TraitName == "fruit_length_min" ~ "cm",
  TraitName == "fruit_length_max" ~ "cm",
  TraitName == "fruit_length_mean" ~ "cm",
  TraitName == "seed_length_max" ~ "mm",
  TraitName == "seed_length_min" ~ "mm",
  TraitName == "seed_length_mean" ~ "mm",
  TraitName == "seed_width_max" ~ "mm",
  TraitName == "seed_width_min" ~ "mm",
  TraitName == "seed_width_mean" ~ "mm",
  TraitName == "plant_lifespan" ~ "years",
  .default = ""
  ))
```

```{r}
# add trait database source column to each subset
TRY_filtered_subset$DatabaseSource <- "TRY"
BIEN_filtered_subset$DatabaseSource <- "BIEN"
GIFT_filtered_subset$DatabaseSource <- "GIFT"

# combine all subsets into one dataframe
traits <- rbind(TRY_filtered_subset, BIEN_filtered_subset, GIFT_filtered_subset)
dim(traits)
```

```{r}
# remove rows without accepted species names
traits_clean <- traits[!(is.na(traits$Accepted_species) | traits$Accepted_species==""), ]
nrow(traits_clean)
```

```{r}
traits_clean %>%
  count(TraitName)
```

```{r}
traits_clean %>%
  distinct(Accepted_species, TraitName) %>%
  count(TraitName)
```

# Create a species list of species with fruiting traits
```{r}
fruit_traits <- c("Fruit dehiscence type", "Fruit dry mass", "Fruit length", "Fruit surface type",
                  "Fruit texture inside (fleshy, leathrey, …)", "Fruit type", "fruit type", "fruit_dryness", "fruit_length_max",
                  "fruit_length_mean", "fruit_length_min", "fruit_type", "maximum fruit length", "minimum fruit length")

# Filter the dataframe for fruiting traits
fruiting_df <- traits_clean %>%
  filter(TraitName %in% fruit_traits)
dim(fruiting_df)

# Extract unique species names
fruiting_species <- unique(fruiting_df$Accepted_species)
length(fruiting_species)
```

# Subset trait data and occurrence data by fruiting species list

```{r}
dim(plant_occ)
plant_occ_subset <- plant_occ %>%
  filter(Accepted_species %in% fruiting_species)
dim(plant_occ_subset)
```

```{r}
dim(traits_clean)
traits_subset <- traits_clean%>%
  filter(Accepted_species %in% fruiting_species)
dim(traits_subset)
```

```{r}
traits_subset %>%
  distinct(Accepted_species, TraitName) %>%
  count(TraitName)
```

# Write data to csv
```{r}
write.csv(plant_occ_subset, file.path(output_path, "TropicalAndes_GBIF_plant_occ_harmonized_subset.csv"))
write.csv(traits_subset, file.path(output_path, "TropicalAndes_all_plant_traits_harmonized_subset.csv"))
```