### 2024-06-24 code/L2/L2_1_TropicalAndes_taxdiv.Rmd
adjust plots before combining into multi-panel plots, remove geometry from richness dataframes before saving as csv

---
title: "Tropical Andes Taxonomic Diversity of Plants and Frugivores"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "Calculating and mapping taxonomic diversity as species richness using occurrence data for plants and frugivores in the Tropical Andes Moist Lowland and Montane forests. Note: some code is adapted from https://luisdva.github.io/rstats/richness/."
date: "2023-08-01"
output: 
  html_document:
    highlight_downlit = TRUE
    df_print: tibble
---

# Load required packages
```{r}
library(sf)
library(dplyr)
library(ggplot2)
library(parallel)
library(foreach)
library(doParallel)
library(ggspatial)
library(ggpubr)
```

# Set file paths
```{r}
data_path_L0 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
data_path_L1 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L1')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
figure_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/figures')
```

# Read in Data
```{r}
## projected sf objects
plants_sf_species <- readRDS(file = file.path(data_path_L1,"plants_sf_species.rds"))
frugivores_sf_species <- readRDS(file = file.path(data_path_L1,"frugivores_sf_species.rds"))
Americas <- readRDS(file = file.path(data_path_L1, "Americas.rds"))
TApoly <- readRDS(file = file.path(data_path_L1,"TApoly.rds"))
TropicalAndes_IUCNHabitat_Forest <- readRDS(file = file.path(data_path_L1,"TropicalAndes_IUCNHabitat_Forest.rds"))
```

# Plot TD/Richness
## Function for richness plots using parallel processing
```{r}
# Define the calculate_richness function outside
calculate_richness <- function(data_sf, species_sf) {
  result <- data_sf %>%
    st_join(species_sf) %>%
    mutate(overlap = ifelse(!is.na(species), 1, 0)) %>%
    group_by(cellid) %>%
    summarize(num_species = sum(overlap)) %>%
    ungroup()
  return(result)
}

create_rich_plots <- function(resolution_meters) {
  # Make Grid
  TAGrid <- TApoly %>%
    st_make_grid(cellsize = c(resolution_meters)) %>%
    st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
    st_cast("MULTIPOLYGON") %>%
    st_sf() %>%
    mutate(cellid = row_number())

  # Determine the chunk size
  chunk_size <- 1000

  # Split the TAGrid into chunks
  TAGrid_chunks <- split(TAGrid, ceiling(seq_along(TAGrid$cellid) / chunk_size))

  # Set up parallel processing
  no_cores <- detectCores() - 1
  cl <- makeCluster(no_cores)
  registerDoParallel(cl)

  # Export necessary objects to the cluster
  clusterExport(cl, varlist = c("plants_sf_species", "frugivores_sf_species", "calculate_richness"))

  # Process chunks in parallel
  plant_richness_grid <- foreach(chunk = TAGrid_chunks, .combine = bind_rows, .packages = c("dplyr", "sf")) %dopar% {
    calculate_richness(chunk, plants_sf_species)
  }

  frugivore_richness_grid <- foreach(chunk = TAGrid_chunks, .combine = bind_rows, .packages = c("dplyr", "sf")) %dopar% {
    calculate_richness(chunk, frugivores_sf_species)
  }

  # Stop the parallel cluster
  stopCluster(cl)

  # Helper functions to generate plots
  generate_plant_plot <- function(data) {
    ggplot(data) +
      geom_sf(data = Americas, fill = "white") +
      geom_sf(data = TApoly, fill = "grey", size = 0.1) +
      geom_sf(aes(fill = num_species), color = NA) +
      scale_fill_distiller(palette = "Greens", direction = 1, limits = c(0,1200)) +
      labs(x = "Longitude", y = "Latitude", fill = "S") +
      coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
      annotation_scale(location = "bl", width_hint = 0.5) +
      theme(panel.background = element_rect(fill = "lightblue"))
  }

  generate_frugivore_plot <- function(data) {
    ggplot(data) +
      geom_sf(data = Americas, fill = "white") +
      geom_sf(data = TApoly, fill = "grey", size = 0.1) +
      geom_sf(aes(fill = num_species), color = NA) +
      scale_fill_distiller(palette = "YlOrBr", direction = 1, limits = c(0,500)) +
      labs(x = "Longitude", y = "Latitude", fill = "S") +
      coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
      annotation_scale(location = "bl", width_hint = 0.5) +
      theme(panel.background = element_rect(fill = "lightblue"))
  }

  # Generate plots
  plant_plot <- generate_plant_plot(plant_richness_grid)
  frugivore_plot <- generate_frugivore_plot(frugivore_richness_grid)

  # Create histograms
  plant_richness_hist <- hist(plant_richness_grid$num_species,
                              main = paste("Histogram of Plant Richness [", resolution_meters / 1000, " km] by cell", sep = ""),
                              xlab = "Plant Richness by cell")
  frugivore_richness_hist <- hist(frugivore_richness_grid$num_species,
                                  main = paste("Histogram of Frugivore Richness [", resolution_meters / 1000, " km] by cell", sep = ""),
                                  xlab = "Frugivore Richness by cell")

  # Return a list of data
  list(plantgridRichTA = plant_plot,
       frugivoregridRichTA = frugivore_plot,
       plant_cellRichness = plant_richness_grid,
       frugivore_cellRichness = frugivore_richness_grid,
       plant_richness_hist = plant_richness_hist,
       frugivore_richness_hist = frugivore_richness_hist)
}
```

## 5 km resolution
```{r}
richness_5km <- create_rich_plots(5000)
```

```{r}
print(richness_5km)

# save plots
plantgridRichTA_5km <- richness_5km$plantgridRichTA
frugivoregridRichTA_5km <- richness_5km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_5km <- richness_5km$plant_cellRichness
frugivore_cellRichness_5km <- richness_5km$frugivore_cellRichness
```

## 10 km
```{r}
richness_10km <- create_rich_plots(10000)
```

```{r}
print(richness_10km)

# save plots
plantgridRichTA_10km <- richness_10km$plantgridRichTA
frugivoregridRichTA_10km <- richness_10km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_10km <- richness_10km$plant_cellRichness
frugivore_cellRichness_10km <- richness_10km$frugivore_cellRichness
```

## 25 km
```{r}
richness_25km <- create_rich_plots(25000)
```

```{r}
print(richness_25km)

# save plots
plantgridRichTA_25km <- richness_25km$plantgridRichTA
frugivoregridRichTA_25km <- richness_25km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_25km <- richness_25km$plant_cellRichness
frugivore_cellRichness_25km <- richness_25km$frugivore_cellRichness
```

## 50 km
```{r}
richness_50km <- create_rich_plots(50000)
```

```{r}
print(richness_50km)
# save plots
plantgridRichTA_50km <- richness_50km$plantgridRichTA
frugivoregridRichTA_50km <- richness_50km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_50km <- richness_50km$plant_cellRichness
frugivore_cellRichness_50km <- richness_50km$frugivore_cellRichness
```

## 75 km 
```{r}
richness_75km <- create_rich_plots(75000)
```

```{r}
print(richness_75km)
# save plots
plantgridRichTA_75km <- richness_75km$plantgridRichTA
frugivoregridRichTA_75km <- richness_75km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_75km <- richness_75km$plant_cellRichness
frugivore_cellRichness_75km <- richness_75km$frugivore_cellRichness
```

## 100 km 
```{r}
richness_100km <- create_rich_plots(100000)
```

```{r}
print(richness_100km)
# save plots
plantgridRichTA_100km <- richness_100km$plantgridRichTA
frugivoregridRichTA_100km <- richness_100km$frugivoregridRichTA
## Extract cell values
plant_cellRichness_100km <- richness_100km$plant_cellRichness
frugivore_cellRichness_100km <- richness_100km$frugivore_cellRichness
```

# Save richness objects to file
```{r}
saveRDS(richness_5km, file = file.path(output_path,"richness_5km.rds"))
saveRDS(richness_10km, file = file.path(output_path,"richness_10km.rds"))
saveRDS(richness_25km, file = file.path(output_path,"richness_25km.rds"))
saveRDS(richness_50km, file = file.path(output_path,"richness_50km.rds"))
saveRDS(richness_75km, file = file.path(output_path,"richness_75km.rds"))
saveRDS(richness_100km, file = file.path(output_path,"richness_100km.rds"))
```

# Save plots to file
```{r}
plot_list <- list(plantgridRichTA_5km = plantgridRichTA_5km, plantgridRichTA_10km = plantgridRichTA_10km,
                  plantgridRichTA_25km = plantgridRichTA_25km, plantgridRichTA_50km = plantgridRichTA_50km,
                  plantgridRichTA_75km = plantgridRichTA_75km, plantgridRichTA_100km = plantgridRichTA_100km,
                  frugivoregridRichTA_5km = frugivoregridRichTA_5km, frugivoregridRichTA_10km = frugivoregridRichTA_10km,
                  frugivoregridRichTA_25km = frugivoregridRichTA_25km, frugivoregridRichTA_50km = frugivoregridRichTA_50km,
                  frugivoregridRichTA_75km = frugivoregridRichTA_75km, frugivoregridRichTA_100km = frugivoregridRichTA_100km)
#histogram_list <- list(plant_richness_hist_5km = richness_5km$plant_richness_hist, 
#                       frugivore_richness_hist_5km = richness_5km$frugivore_richness_hist,
#                       plant_richness_hist_10km = richness_10km$plant_richness_hist,
#                       frugivore_richness_hist_10km = richness_10km$frugivore_richness_hist,
#                       plant_richness_hist_25km = richness_25km$plant_richness_hist,
#                       frugivore_richness_hist_55km = richness_25km$frugivore_richness_hist,
#                       plant_richness_hist_50km = richness_50km$plant_richness_hist,
#                       frugivore_richness_hist_50km = richness_50km$frugivore_richness_hist,
#                       plant_richness_hist_75km = richness_75km$plant_richness_hist,
#                       frugivore_richness_hist_75km = richness_75km$frugivore_richness_hist,
#                       plant_richness_hist_100km = richness_100km$plant_richness_hist,
#                       frugivore_richness_hist_100km = richness_100km$frugivore_richness_hist)
#histograms would have to be converted to ggplot to save
```

```{r}
save_plots <- function(plot_list, figure_path) {
  # Iterate over the list and save each plot
  for (plot_name in names(plot_list)) {
    plot_path <- file.path(figure_path, paste0(plot_name, ".png"))

    # Try to save the plot and handle any potential errors
    tryCatch({
      ggsave(filename = plot_path, 
             plot = plot_list[[plot_name]])
      cat("Saved plot:", plot_path, "\n")
    }, error = function(e) {
      cat("Failed to save plot:", plot_name, "\nError message:", e$message, "\n")
    })
  }
}
```

```{r}
# Save the plots
save_plots(plot_list, figure_path)
```

# Multi panel plots of richness maps
## Plants
```{r}
# Add labels and remove axis labels from all plots
plantgridRichTA_5km_labeled <- plantgridRichTA_5km + ggtitle("5 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10))
plantgridRichTA_10km_labeled <- plantgridRichTA_10km + ggtitle("10 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10))
plantgridRichTA_25km_labeled <- plantgridRichTA_25km + ggtitle("25 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10))
plantgridRichTA_50km_labeled <- plantgridRichTA_50km + ggtitle("50 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10))
plantgridRichTA_75km_labeled <- plantgridRichTA_75km + ggtitle("75 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10))
plantgridRichTA_100km_labeled <- plantgridRichTA_100km + ggtitle("100 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10))
plantgridRichTA_100km_labeled
```

```{r}
# Arrange the plots with common legend
all_plant_richness_plots <- ggarrange(plantgridRichTA_5km_labeled, plantgridRichTA_10km_labeled,
                                      plantgridRichTA_25km_labeled, plantgridRichTA_50km_labeled,
                                      plantgridRichTA_75km_labeled, plantgridRichTA_100km_labeled,
                                      ncol = 6, nrow = 1,
                                      common.legend = TRUE, legend = "left")

all_plant_richness_plots
```

```{r}
all_plant_richness_plots_labeled <- ggpubr::annotate_figure(all_plant_richness_plots,
                                                            left = ggpubr::text_grob("Plants", face = "bold", size = 20, 
                                                                                     rot = 90))
all_plant_richness_plots_labeled
ggsave("plant_richness_plots.png", all_plant_richness_plots_labeled, path = figure_path, width = 16, height = 5, units = "in")
```

## Frugivores
```{r}
# Add labels and remove axis labels from all plots
frugivoregridRichTA_5km_labeled <- frugivoregridRichTA_5km + ggtitle("5 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10))
frugivoregridRichTA_10km_labeled <- frugivoregridRichTA_10km + ggtitle("10 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10))
frugivoregridRichTA_25km_labeled <- frugivoregridRichTA_25km + ggtitle("25 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10))
frugivoregridRichTA_50km_labeled <- frugivoregridRichTA_50km + ggtitle("50 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10))
frugivoregridRichTA_75km_labeled <- frugivoregridRichTA_75km + ggtitle("75 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10))
frugivoregridRichTA_100km_labeled <- frugivoregridRichTA_100km + ggtitle("100 km") + theme(plot.title = element_text(hjust = 0.5)) + theme(axis.title = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), legend.key.height = unit(0.5, "in"), legend.text = element_text(size = 10))
frugivoregridRichTA_100km_labeled
```

```{r}
all_frugivore_richness_plots <- ggpubr::ggarrange(frugivoregridRichTA_5km_labeled,frugivoregridRichTA_10km_labeled,
                                                  frugivoregridRichTA_25km_labeled, frugivoregridRichTA_50km_labeled,
                                                  frugivoregridRichTA_75km_labeled, frugivoregridRichTA_100km_labeled,
                                                  ncol = 6, nrow = 1,
                                                  common.legend = TRUE, legend = "left")
all_frugivore_richness_plots
```

```{r}
all_frugivore_richness_plots_labeled <- ggpubr::annotate_figure(all_frugivore_richness_plots,
                                                            left = ggpubr::text_grob("Frugivores", face = "bold", size = 20,
                                                                                     rot = 90))
all_frugivore_richness_plots_labeled
ggsave("frugivore_richness_plots.png", all_frugivore_richness_plots_labeled, path = figure_path, width = 16, height = 5, units = "in")
```

## Combine plants & frugivores
```{r}
all_richness_plots <- ggpubr::ggarrange(all_plant_richness_plots_labeled,
                                        all_frugivore_richness_plots_labeled,
                                        ncol = 1, nrow = 2)
# Add a white background to the plot
all_richness_plots <- all_richness_plots + theme(plot.background = element_rect(fill = "white", color = NA))
all_richness_plots

ggsave("all_richness_plots.png", all_richness_plots, path = figure_path, width = 16, height = 10, units = "in")
```

# Write data to csv
Have to remove geometry before saving to csv
```{r}
plant_cellRichness_5km_df <- data.frame(cellid = plant_cellRichness_5km$cellid,
                                        num_species = plant_cellRichness_5km$num_species)
frugivore_cellRichness_5km_df <- data.frame(cellid = frugivore_cellRichness_5km$cellid,
                                            num_species = frugivore_cellRichness_5km$num_species)

plant_cellRichness_10km_df <- data.frame(cellid = plant_cellRichness_10km$cellid,
                                        num_species = plant_cellRichness_10km$num_species)
frugivore_cellRichness_10km_df <- data.frame(cellid = frugivore_cellRichness_10km$cellid,
                                            num_species = frugivore_cellRichness_10km$num_species)

plant_cellRichness_25km_df <- data.frame(cellid = plant_cellRichness_25km$cellid,
                                        num_species = plant_cellRichness_25km$num_species)
frugivore_cellRichness_25km_df <- data.frame(cellid = frugivore_cellRichness_25km$cellid,
                                            num_species = frugivore_cellRichness_25km$num_species)

plant_cellRichness_50km_df <- data.frame(cellid = plant_cellRichness_50km$cellid,
                                        num_species = plant_cellRichness_50km$num_species)
frugivore_cellRichness_50km_df <- data.frame(cellid = frugivore_cellRichness_50km$cellid,
                                            num_species = frugivore_cellRichness_50km$num_species)

plant_cellRichness_75km_df <- data.frame(cellid = plant_cellRichness_75km$cellid,
                                        num_species = plant_cellRichness_75km$num_species)
frugivore_cellRichness_75km_df <- data.frame(cellid = frugivore_cellRichness_75km$cellid,
                                            num_species = frugivore_cellRichness_75km$num_species)

plant_cellRichness_100km_df <- data.frame(cellid = plant_cellRichness_100km$cellid,
                                        num_species = plant_cellRichness_100km$num_species)
frugivore_cellRichness_100km_df <- data.frame(cellid = frugivore_cellRichness_100km$cellid,
                                            num_species = frugivore_cellRichness_100km$num_species)
```

```{r}
write.csv(plant_cellRichness_5km_df, file.path(output_path,"TropicalAndes_plantRichness_5km.csv"), row.names = FALSE)
write.csv(frugivore_cellRichness_5km_df, file.path(output_path,"TropicalAndes_frugivoreRichness_5km.csv"), row.names = FALSE)
```

```{r}
write.csv(plant_cellRichness_10km_df, file.path(output_path,"TropicalAndes_plantRichness_10km.csv"), row.names = FALSE)
write.csv(frugivore_cellRichness_10km_df, file.path(output_path,"TropicalAndes_frugivoreRichness_10km.csv"), row.names = FALSE)
```

```{r}
write.csv(plant_cellRichness_25km_df, file.path(output_path,"TropicalAndes_plantRichness_25km.csv"), row.names = FALSE)
write.csv(frugivore_cellRichness_25km_df, file.path(output_path,"TropicalAndes_frugivoreRichness_25km.csv"), row.names = FALSE)
```

```{r}
write.csv(plant_cellRichness_50km_df, file.path(output_path,"TropicalAndes_plantRichness_50km.csv"), row.names = FALSE)
write.csv(frugivore_cellRichness_50km_df, file.path(output_path,"TropicalAndes_frugivoreRichness_50km.csv"), row.names = FALSE)
```

```{r}
write.csv(plant_cellRichness_75km_df, file.path(output_path,"TropicalAndes_plantRichness_75km.csv"), row.names = FALSE)
write.csv(frugivore_cellRichness_75km_df, file.path(output_path,"TropicalAndes_frugivoreRichness_75km.csv"), row.names = FALSE)
```

```{r}
write.csv(plant_cellRichness_100km_df, file.path(output_path,"TropicalAndes_plantRichness_100km.csv"), row.names = FALSE)
write.csv(frugivore_cellRichness_100km_df, file.path(output_path,"TropicalAndes_frugivoreRichness_100km.csv"), row.names = FALSE)
```