### 2024-06-24 code/L2/L2_3_TropicalAndes_divRelationships.Rmd
create plots to compare tax and fun div between plants and frugivores using multiple models

---
title: "Diversity relationships for plants and frugivores"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script compares taxonomic and functional diversity relationships for plants and frugivores for various spatial grains."
date: "2024-05-16"
output: html_document
---

# Set file paths
```{r}
data_path_L2 <-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L2')
figure_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/figures')
```

# Load required packages
```{r}
library(dplyr)
library(ggplot2)
library(smoothr)
library(purrr)
library(ggtrendline)
library(ggpubr)
library(tidyr)
```

# Read in Data
```{r}
plant_cellRichness_5km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantRichness_5km.csv"))
frugivore_cellRichness_5km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreRichness_5km.csv"))

plant_cellRichness_10km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantRichness_10km.csv"))
frugivore_cellRichness_10km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreRichness_10km.csv"))

plant_cellRichness_25km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantRichness_25km.csv"))
frugivore_cellRichness_25km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreRichness_25km.csv"))

plant_cellRichness_50km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantRichness_50km.csv"))
frugivore_cellRichness_50km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreRichness_50km.csv"))

plant_cellRichness_75km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantRichness_75km.csv"))
frugivore_cellRichness_75km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreRichness_75km.csv"))

plant_cellRichness_100km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantRichness_100km.csv"))
frugivore_cellRichness_100km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreRichness_100km.csv"))
```

```{r}
plant_cellFDis_5km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantFDis_5km.csv"))
frugivore_cellFDis_5km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreFDis_5km.csv"))

plant_cellFDis_10km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantFDis_10km.csv"))
frugivore_cellFDis_10km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreFDis_10km.csv"))

plant_cellFDis_25km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantFDis_25km.csv"))
frugivore_cellFDis_25km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreFDis_25km.csv"))

plant_cellFDis_50km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantFDis_50km.csv"))
frugivore_cellFDis_50km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreFDis_50km.csv"))

plant_cellFDis_75km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantFDis_75km.csv"))
frugivore_cellFDis_75km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreFDis_75km.csv"))

plant_cellFDis_100km <- read.csv(file.path(data_path_L2,"TropicalAndes_plantFDis_100km.csv"))
frugivore_cellFDis_100km <- read.csv(file.path(data_path_L2,"TropicalAndes_frugivoreFDis_100km.csv"))
```

# Preliminary plots
Check the data and inspect possible relationships

```{r}
plot(plant_cellRichness_100km$num_species, frugivore_cellRichness_100km$num_species)
plot(plant_cellRichness_75km$num_species, frugivore_cellRichness_75km$num_species)
plot(plant_cellRichness_50km$num_species, frugivore_cellRichness_50km$num_species)
plot(plant_cellRichness_25km$num_species, frugivore_cellRichness_25km$num_species)
plot(plant_cellRichness_10km$num_species, frugivore_cellRichness_10km$num_species)
plot(plant_cellRichness_5km$num_species, frugivore_cellRichness_5km$num_species)
```

```{r}
plot(plant_cellFDis_100km$fdis_value, frugivore_cellFDis_100km$fdis_value)
plot(plant_cellFDis_75km$fdis_value, frugivore_cellFDis_75km$fdis_value)
plot(plant_cellFDis_50km$fdis_value, frugivore_cellFDis_50km$fdis_value)
plot(plant_cellFDis_25km$fdis_value, frugivore_cellFDis_25km$fdis_value)
plot(plant_cellFDis_10km$fdis_value, frugivore_cellFDis_10km$fdis_value)
plot(plant_cellFDis_5km$fdis_value, frugivore_cellFDis_5km$fdis_value)
```

```{r}
plot(plant_cellRichness_5km$num_species, plant_cellFDis_5km$fdis_value)
plot(plant_cellRichness_10km$num_species, plant_cellFDis_10km$fdis_value)
plot(plant_cellRichness_25km$num_species, plant_cellFDis_25km$fdis_value)
plot(plant_cellRichness_50km$num_species, plant_cellFDis_50km$fdis_value)
plot(plant_cellRichness_75km$num_species, plant_cellFDis_75km$fdis_value)
plot(plant_cellRichness_100km$num_species, plant_cellFDis_100km$fdis_value)
```

```{r}
plot(frugivore_cellRichness_5km$num_species, frugivore_cellFDis_5km$fdis_value)
plot(frugivore_cellRichness_10km$num_species, frugivore_cellFDis_10km$fdis_value)
plot(frugivore_cellRichness_25km$num_species, frugivore_cellFDis_25km$fdis_value)
plot(frugivore_cellRichness_50km$num_species, frugivore_cellFDis_50km$fdis_value)
plot(frugivore_cellRichness_75km$num_species, frugivore_cellFDis_75km$fdis_value)
plot(frugivore_cellRichness_100km$num_species, frugivore_cellFDis_100km$fdis_value)
```

# Compare diversity relationships between plants and frugivores
## Function for comparing diversity relationships
For comparing richness, use exp3P and line2P for line types
For comparing FDis, use line3P and line2P for line types

line colors: line2P is "red", line3P is "blue", exp3P is "purple"
```{r}
compare_Richness <- function(plant_Richness_grid, frugivore_Richness_grid, resolution_meters){
  # Extract cell values
  comparison_Richness <- data.frame(Plant_Richness = plant_Richness_grid$num_species,
                               Frugivore_Richness = frugivore_Richness_grid$num_species)

  # Filter out rows where either plant or frugivore richness is zero
  comparison_Richness_filtered <- comparison_Richness %>%
                                filter(Plant_Richness > 0 & Frugivore_Richness > 0)

  # Calculate Annotation position
  mean_x <- max(comparison_Richness_filtered$Plant_Richness) / 2
  annot_y <- max(comparison_Richness_filtered$Frugivore_Richness) + 25

  # Create scatter plots
  compare_Richness_plot_line2P <- ggtrendline(comparison_Richness_filtered$Plant_Richness,
               comparison_Richness_filtered$Frugivore_Richness, model = "line2P", linecolor = "red", linewidth = 1,
               CI.level = NA, summary = FALSE, show.eq = FALSE, show.pvalue = FALSE, rrp.x = mean_x, rrp.y = annot_y, 
               text.col = "red", eSize = 4) +
      geom_point(aes(x = comparison_Richness_filtered$Plant_Richness,
                                           y = comparison_Richness_filtered$Frugivore_Richness), alpha = 0.25, size = 1) +
      labs(title = paste("Comparing Richness [", resolution_meters/1000, " km]", sep = ""),
           subtitle = "Fitted with a linear model",
       x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
      scale_x_continuous(expand = c(0, 0)) + 
      scale_y_continuous(expand = c(0, 0)) + 
      theme_minimal() +
      theme(panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(),
            axis.line = element_line(color = "black"),
            axis.ticks = element_line(color = "black"),
            axis.text = element_text(color = "black"))

  compare_Richness_plot_exp3P <- ggtrendline(comparison_Richness_filtered$Plant_Richness,
               comparison_Richness_filtered$Frugivore_Richness, model = "exp3P", linecolor = "purple", linewidth = 1,
               CI.level = NA, summary = FALSE, show.eq = FALSE, show.pvalue = FALSE, rrp.x = mean_x, rrp.y = annot_y, 
               text.col = "purple", eSize = 4) +
      geom_point(aes(x = comparison_Richness_filtered$Plant_Richness,
                                           y = comparison_Richness_filtered$Frugivore_Richness), alpha = 0.25, size = 1) +
      labs(title = paste("Comparing Richness [", resolution_meters/1000, " km]", sep = ""),
           subtitle = "Fitted with an exponential model",
       x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
      scale_x_continuous(expand = c(0, 0)) + 
      scale_y_continuous(expand = c(0, 0)) + 
      theme_minimal() +
      theme(panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(),
            axis.line = element_line(color = "black"),
            axis.ticks = element_line(color = "black"),
            axis.text = element_text(color = "black"))

    # Summaries of models used for trendlines
  line2P_summary <- trendline_sum(comparison_Richness_filtered$Plant_Richness,
               comparison_Richness_filtered$Frugivore_Richness, model = "line2P")
  exp3P_summary <- trendline_sum(comparison_Richness_filtered$Plant_Richness,
               comparison_Richness_filtered$Frugivore_Richness, model = "exp3P")

  # Combine trendlines into one plot
  compare_Richness_plot_both <- ggtrendline(comparison_Richness_filtered$Plant_Richness,
               comparison_Richness_filtered$Frugivore_Richness, model = "exp3P", linecolor = "purple", linewidth = 1,
               CI.level = NA, summary = FALSE, show.eq = FALSE, show.pvalue = FALSE, show.Rsquare = FALSE) +
      geom_smooth(aes(x = comparison_Richness_filtered$Plant_Richness, y = comparison_Richness_filtered$Frugivore_Richness), 
              method = "lm", se = FALSE, color = "red", linetype = "solid") +
      geom_point(aes(x = comparison_Richness_filtered$Plant_Richness,
                                           y = comparison_Richness_filtered$Frugivore_Richness), alpha = 0.25, size = 1) +
      labs(title = paste("Comparing Richness [", resolution_meters/1000, " km]", sep = ""),
       x = "Plant Richness by cell", y = "Frugivore Richness by cell") +
      scale_x_continuous(expand = c(0, 0)) + 
      scale_y_continuous(expand = c(0, 0)) + 
      theme_minimal() +
      theme(panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(),
            axis.line = element_line(color = "black"),
            axis.ticks = element_line(color = "black"),
            axis.text = element_text(color = "black")) +
      annotate("text", x = mean_x, y = annot_y-10, label = paste("Linear R² =", round(line2P_summary$R.squared, 3)),
               color = "red", size = 4) +
      annotate("text", x = mean_x, y = annot_y-30, label = paste("Exponential R² =", round(exp3P_summary$R.squared, 3)),
               color = "purple", size = 4)

  list(compare_Richness_plot_line2P = compare_Richness_plot_line2P,
       compare_Richness_plot_exp3P = compare_Richness_plot_exp3P,
       line2P_summary = line2P_summary,
       exp3P_summary = exp3P_summary,
       compare_Richness_plot_both = compare_Richness_plot_both)
}
```

```{r}
compare_FDis <- function(plant_FDis_grid, frugivore_FDis_grid, resolution_meters){
  # Extract cell values
  comparison_FDis <- data.frame(Plant_FDis = plant_FDis_grid$fdis_value,
                               Frugivore_FDis = frugivore_FDis_grid$fdis_value)

  # Filter out rows where either plant or frugivore FDis is zero
  comparison_FDis_filtered <- comparison_FDis %>%
                                filter(Plant_FDis > 0 & Frugivore_FDis > 0)
  # Calculate Annotation position
  mean_x <- max(comparison_FDis_filtered$Plant_FDis) / 2
  annot_y <- max(comparison_FDis_filtered$Frugivore_FDis) + 0.25

  ## Create scatter plots
  compare_FDis_plot_line2P <- ggtrendline(comparison_FDis_filtered$Plant_FDis,
               comparison_FDis_filtered$Frugivore_FDis, model = "line2P", linecolor = "red", linewidth = 1,
               CI.level = NA, summary = FALSE, show.eq = FALSE, show.pvalue = FALSE, rrp.x = mean_x, rrp.y = annot_y, 
               text.col = "red", eSize = 4) +
      geom_point(aes(x = comparison_FDis_filtered$Plant_FDis,
                                           y = comparison_FDis_filtered$Frugivore_FDis), alpha = 0.25, size = 1) +
      labs(title = paste("Compare FDis [", resolution_meters/1000, " km]", sep = ""),
           subtitle = "Fitted with a linear model",
       x = "Plant FDis by cell", y = "Frugivore FDis by cell") +
      scale_x_continuous(expand = c(0, 0)) + 
      scale_y_continuous(expand = c(0, 0)) + 
      theme_minimal() +
      theme(panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(),
            axis.line = element_line(color = "black"),
            axis.ticks = element_line(color = "black"),
            axis.text = element_text(color = "black"))

  compare_FDis_plot_line3P <- ggtrendline(comparison_FDis_filtered$Plant_FDis,
               comparison_FDis_filtered$Frugivore_FDis, model = "line3P", linecolor = "blue", linewidth = 1,
               CI.level = NA, summary = FALSE, show.eq = FALSE, show.pvalue = FALSE, rrp.x = mean_x, rrp.y = annot_y, 
               text.col = "blue", eSize = 4) +
      geom_point(aes(x = comparison_FDis_filtered$Plant_FDis,
                                           y = comparison_FDis_filtered$Frugivore_FDis), alpha = 0.25, size = 1) +
      labs(title = paste("Compare FDis [", resolution_meters/1000, " km]", sep = ""),
           subtitle = "Fitted with a quadratic model",
       x = "Plant FDis by cell", y = "Frugivore FDis by cell") +
      scale_x_continuous(expand = c(0, 0)) + 
      scale_y_continuous(expand = c(0, 0)) + 
      theme_minimal() +
      theme(panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(),
            axis.line = element_line(color = "black"),
            axis.ticks = element_line(color = "black"),
            axis.text = element_text(color = "black"))

  # Summaries of models used for trendlines
  line2P_summary <- trendline_sum(comparison_FDis_filtered$Plant_FDis,
               comparison_FDis_filtered$Frugivore_FDis, model = "line2P")
  line3P_summary <- trendline_sum(comparison_FDis_filtered$Plant_FDis,
               comparison_FDis_filtered$Frugivore_FDis, model = "line3P")

  # Combine trendlines into one plot
  compare_FDis_plot_both <- ggtrendline(comparison_FDis_filtered$Plant_FDis,
               comparison_FDis_filtered$Frugivore_FDis, model = "line3P", linecolor = "blue", linewidth = 1,
               CI.level = NA, summary = FALSE, show.eq = FALSE, show.pvalue = FALSE, show.Rsquare = FALSE) +
      geom_smooth(aes(x = comparison_FDis_filtered$Plant_FDis, y = comparison_FDis_filtered$Frugivore_FDis), 
              method = "lm", se = FALSE, color = "red", linetype = "solid") +
      geom_point(aes(x = comparison_FDis_filtered$Plant_FDis,
                                           y = comparison_FDis_filtered$Frugivore_FDis), alpha = 0.25, size = 1) +
      labs(title = paste("Comparing FDis [", resolution_meters/1000, " km]", sep = ""),
       x = "Plant FDis by cell", y = "Frugivore FDis by cell") +
      scale_x_continuous(expand = c(0, 0)) + 
      scale_y_continuous(expand = c(0, 0)) + 
      theme_minimal() +
      theme(panel.grid.major = element_blank(), 
            panel.grid.minor = element_blank(),
            axis.line = element_line(color = "black"),
            axis.ticks = element_line(color = "black"),
            axis.text = element_text(color = "black")) +
      annotate("text", x = mean_x, y = annot_y, label = paste("Linear R² =", round(line2P_summary$R.squared, 3)),
               color = "red", size = 4) +
      annotate("text", x = mean_x, y = annot_y-0.5, label = paste("Quadratic R² =", round(line3P_summary$R.squared, 3)),
               color = "blue", size = 4)

  list(compare_FDis_plot_line2P = compare_FDis_plot_line2P,
       compare_FDis_plot_line3P = compare_FDis_plot_line3P,
       line2P_summary = line2P_summary,
       line3P_summary = line3P_summary,
       compare_FDis_plot_both = compare_FDis_plot_both)
}
```

## 5 km
```{r}
compare_Richness_5km <- compare_Richness(plant_cellRichness_5km, frugivore_cellRichness_5km, 5000)
(compare_Richness_5km_linear <- compare_Richness_5km$compare_Richness_plot_line2P)
(compare_Richness_5km_exp <- compare_Richness_5km$compare_Richness_plot_exp3P)
(compare_Richness_5km_plots <- compare_Richness_5km$compare_Richness_plot_both)
compare_Richness_r2_5km_linear <- compare_Richness_5km$line2P_summary$R.squared
compare_Richness_r2_5km_exp <- compare_Richness_5km$exp3P_summary$R.squared
compare_Richness_slopecoefficent_5km_linear <- compare_Richness_5km$line2P_summary$parameter["a"]
compare_Richness_slopecoefficent_5km_exp <- compare_Richness_5km$exp3P_summary$parameter["b"]
```

```{r}
compare_FDis_5km <- compare_FDis(plant_cellFDis_5km, frugivore_cellFDis_5km, 5000)
(compare_FDis_5km_linear <- compare_FDis_5km$compare_FDis_plot_line2P)
(compare_FDis_5km_quad <- compare_FDis_5km$compare_FDis_plot_line3P)
(compare_FDis_5km_plots <- compare_FDis_5km$compare_FDis_plot_both)
compare_FDis_r2_5km_linear <- compare_FDis_5km$line2P_summary$R.squared
compare_FDis_r2_5km_quad <- compare_FDis_5km$line3P_summary$R.squared
compare_FDis_slopecoefficent_5km_linear <- compare_FDis_5km$line2P_summary$parameter["a"]
compare_FDis_slopecoefficent_5km_quad <- compare_FDis_5km$line3P_summary$parameter["b"]
```

## 10 km
```{r}
compare_Richness_10km <- compare_Richness(plant_cellRichness_10km, frugivore_cellRichness_10km, 10000)
(compare_Richness_10km_linear <- compare_Richness_10km$compare_Richness_plot_line2P)
(compare_Richness_10km_exp <- compare_Richness_10km$compare_Richness_plot_exp3P)
(compare_Richness_10km_plots <- compare_Richness_10km$compare_Richness_plot_both)
compare_Richness_r2_10km_linear <- compare_Richness_10km$line2P_summary$R.squared
compare_Richness_r2_10km_exp <- compare_Richness_10km$exp3P_summary$R.squared
compare_Richness_slopecoefficent_10km_linear <- compare_Richness_10km$line2P_summary$parameter["a"]
compare_Richness_slopecoefficent_10km_exp <- compare_Richness_10km$exp3P_summary$parameter["b"]
```

```{r}
compare_FDis_10km <- compare_FDis(plant_cellFDis_10km, frugivore_cellFDis_10km, 10000)
(compare_FDis_10km_linear <- compare_FDis_10km$compare_FDis_plot_line2P)
(compare_FDis_10km_quad <- compare_FDis_10km$compare_FDis_plot_line3P)
(compare_FDis_10km_plots <- compare_FDis_10km$compare_FDis_plot_both)
compare_FDis_r2_10km_linear <- compare_FDis_10km$line2P_summary$R.squared
compare_FDis_r2_10km_quad <- compare_FDis_10km$line3P_summary$R.squared
compare_FDis_slopecoefficent_10km_linear <- compare_FDis_10km$line2P_summary$parameter["a"]
compare_FDis_slopecoefficent_10km_quad <- compare_FDis_10km$line3P_summary$parameter["b"]
```

## 25 km
```{r}
compare_Richness_25km <- compare_Richness(plant_cellRichness_25km, frugivore_cellRichness_25km, 25000)
(compare_Richness_25km_linear <- compare_Richness_25km$compare_Richness_plot_line2P)
(compare_Richness_25km_exp <- compare_Richness_25km$compare_Richness_plot_exp3P)
(compare_Richness_25km_plots <- compare_Richness_25km$compare_Richness_plot_both)
compare_Richness_r2_25km_linear <- compare_Richness_25km$line2P_summary$R.squared
compare_Richness_r2_25km_exp <- compare_Richness_25km$exp3P_summary$R.squared
compare_Richness_slopecoefficent_25km_linear <- compare_Richness_25km$line2P_summary$parameter["a"]
compare_Richness_slopecoefficent_25km_exp <- compare_Richness_25km$exp3P_summary$parameter["b"]
```

```{r}
compare_FDis_25km <- compare_FDis(plant_cellFDis_25km, frugivore_cellFDis_25km, 25000)
(compare_FDis_25km_linear <- compare_FDis_25km$compare_FDis_plot_line2P)
(compare_FDis_25km_quad <- compare_FDis_25km$compare_FDis_plot_line3P)
(compare_FDis_25km_plots <- compare_FDis_25km$compare_FDis_plot_both)
compare_FDis_r2_25km_linear <- compare_FDis_25km$line2P_summary$R.squared
compare_FDis_r2_25km_quad <- compare_FDis_25km$line3P_summary$R.squared
compare_FDis_slopecoefficent_25km_linear <- compare_FDis_25km$line2P_summary$parameter["a"]
compare_FDis_slopecoefficent_25km_quad <- compare_FDis_25km$line3P_summary$parameter["b"]
```

## 50 km
```{r}
compare_Richness_50km <- compare_Richness(plant_cellRichness_50km, frugivore_cellRichness_50km, 50000)
(compare_Richness_50km_linear <- compare_Richness_50km$compare_Richness_plot_line2P)
(compare_Richness_50km_exp <- compare_Richness_50km$compare_Richness_plot_exp3P)
(compare_Richness_50km_plots <- compare_Richness_50km$compare_Richness_plot_both)
compare_Richness_r2_50km_linear <- compare_Richness_50km$line2P_summary$R.squared
compare_Richness_r2_50km_exp <- compare_Richness_50km$exp3P_summary$R.squared
compare_Richness_slopecoefficent_50km_linear <- compare_Richness_50km$line2P_summary$parameter["a"]
compare_Richness_slopecoefficent_50km_exp <- compare_Richness_50km$exp3P_summary$parameter["b"]
```

```{r}
compare_FDis_50km <- compare_FDis(plant_cellFDis_50km, frugivore_cellFDis_50km, 50000)
(compare_FDis_50km_linear <- compare_FDis_50km$compare_FDis_plot_line2P)
(compare_FDis_50km_quad <- compare_FDis_50km$compare_FDis_plot_line3P)
(compare_FDis_50km_plots <- compare_FDis_50km$compare_FDis_plot_both)
compare_FDis_r2_50km_linear <- compare_FDis_50km$line2P_summary$R.squared
compare_FDis_r2_50km_quad <- compare_FDis_50km$line3P_summary$R.squared
compare_FDis_slopecoefficent_50km_linear <- compare_FDis_50km$line2P_summary$parameter["a"]
compare_FDis_slopecoefficent_50km_quad <- compare_FDis_50km$line3P_summary$parameter["b"]
```

## 75 km
```{r}
compare_Richness_75km <- compare_Richness(plant_cellRichness_75km, frugivore_cellRichness_75km, 75000)
(compare_Richness_75km_linear <- compare_Richness_75km$compare_Richness_plot_line2P)
(compare_Richness_75km_exp <- compare_Richness_75km$compare_Richness_plot_exp3P)
(compare_Richness_75km_plots <- compare_Richness_75km$compare_Richness_plot_both)
compare_Richness_r2_75km_linear <- compare_Richness_75km$line2P_summary$R.squared
compare_Richness_r2_75km_exp <- compare_Richness_75km$exp3P_summary$R.squared
compare_Richness_slopecoefficent_75km_linear <- compare_Richness_75km$line2P_summary$parameter["a"]
compare_Richness_slopecoefficent_75km_exp <- compare_Richness_75km$exp3P_summary$parameter["b"]
```

```{r}
compare_FDis_75km <- compare_FDis(plant_cellFDis_75km, frugivore_cellFDis_75km, 75000)
(compare_FDis_75km_linear <- compare_FDis_75km$compare_FDis_plot_line2P)
(compare_FDis_75km_quad <- compare_FDis_75km$compare_FDis_plot_line3P)
(compare_FDis_75km_plots <- compare_FDis_75km$compare_FDis_plot_both)
compare_FDis_r2_75km_linear <- compare_FDis_75km$line2P_summary$R.squared
compare_FDis_r2_75km_quad <- compare_FDis_75km$line3P_summary$R.squared
compare_FDis_slopecoefficent_75km_linear <- compare_FDis_75km$line2P_summary$parameter["a"]
compare_FDis_slopecoefficent_75km_quad <- compare_FDis_75km$line3P_summary$parameter["b"]
```

## 100 km
```{r}
compare_Richness_100km <- compare_Richness(plant_cellRichness_100km, frugivore_cellRichness_100km, 100000)
(compare_Richness_100km_linear <- compare_Richness_100km$compare_Richness_plot_line2P)
(compare_Richness_100km_exp <- compare_Richness_100km$compare_Richness_plot_exp3P)
(compare_Richness_100km_plots <- compare_Richness_100km$compare_Richness_plot_both)
compare_Richness_r2_100km_linear <- compare_Richness_100km$line2P_summary$R.squared
compare_Richness_r2_100km_exp <- compare_Richness_100km$exp3P_summary$R.squared
compare_Richness_slopecoefficent_100km_linear <- compare_Richness_100km$line2P_summary$parameter["a"]
compare_Richness_slopecoefficent_100km_exp <- compare_Richness_100km$exp3P_summary$parameter["b"]
```

```{r}
compare_FDis_100km <- compare_FDis(plant_cellFDis_100km, frugivore_cellFDis_100km, 100000)
(compare_FDis_100km_linear <- compare_FDis_100km$compare_FDis_plot_line2P)
(compare_FDis_100km_quad <- compare_FDis_100km$compare_FDis_plot_line3P)
(compare_FDis_100km_plots <- compare_FDis_100km$compare_FDis_plot_both)
compare_FDis_r2_100km_linear <- compare_FDis_100km$line2P_summary$R.squared
compare_FDis_r2_100km_quad <- compare_FDis_100km$line3P_summary$R.squared
compare_FDis_slopecoefficent_100km_linear <- compare_FDis_100km$line2P_summary$parameter["a"]
compare_FDis_slopecoefficent_100km_quad <- compare_FDis_100km$line3P_summary$parameter["b"]
```

## Combine plots using for richness and fdis into one figure

```{r}
(compare_Richness_all_plots <- ggarrange(compare_Richness_5km_plots, compare_Richness_10km_plots,
                                     compare_Richness_25km_plots, compare_Richness_50km_plots,
                                     compare_Richness_75km_plots, compare_Richness_100km_plots,
                                        ncol = 3, nrow = 2))
```

```{r}
(compare_FDis_all_plots <- ggarrange(compare_FDis_5km_plots, compare_FDis_10km_plots,
                                     compare_FDis_25km_plots, compare_FDis_50km_plots,
                                     compare_FDis_75km_plots, compare_FDis_100km_plots,
                                        ncol = 3, nrow = 2))
```

# Compare r-squared by resolution
## Taxonomic diversity
```{r}
# create dataframe of resolutions and r-squared values 
resolution <- c(5, 10, 25, 50, 75, 100)
taxdiv_r_squared_linear <- c(compare_Richness_r2_5km_linear, compare_Richness_r2_10km_linear, compare_Richness_r2_25km_linear, compare_Richness_r2_50km_linear, compare_Richness_r2_75km_linear, compare_Richness_r2_100km_linear)
taxdiv_r_squared_exp <- c(compare_Richness_r2_5km_exp, compare_Richness_r2_10km_exp, compare_Richness_r2_25km_exp, compare_Richness_r2_50km_exp, compare_Richness_r2_75km_exp, compare_Richness_r2_100km_exp)
taxdiv_r_squared_compare <- data.frame(resolution, taxdiv_r_squared_linear, taxdiv_r_squared_exp)

# reshape to long format for ease of plotting
taxdiv_r_squared_long <- taxdiv_r_squared_compare %>%
  pivot_longer(cols = c(taxdiv_r_squared_linear, taxdiv_r_squared_exp), 
               names_to = "model", 
               values_to = "taxdiv_r_squared")

```

```{r}
# Define custom colors for the points
custom_colors <- c(taxdiv_r_squared_linear = "red", taxdiv_r_squared_exp = "purple")

# Define custom legend labels
custom_labels <- c(taxdiv_r_squared_linear = "Linear Model", taxdiv_r_squared_exp = "Exponential Model")

(taxdiv_r_squared_plot <- ggplot(taxdiv_r_squared_long,aes(x = resolution, y = taxdiv_r_squared, color = model)) +
  geom_line() +
  geom_point(size = 3) +
  scale_color_manual(values = custom_colors, labels = custom_labels) +
  scale_x_continuous(breaks = taxdiv_r_squared_compare$resolution) + 
  labs(title = "Comparison of Richness R-squared Values by Resolution",
       x = "Resolution (km)",
       y = "R-squared Value",
       color = "Model Type") +
  theme_minimal() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black"),
        axis.ticks = element_line(color = "black"),
        axis.text = element_text(color = "black"),
        legend.position = "inside",
        legend.position.inside = c(.85, .3),
        legend.box.margin = margin(10, 10, 10, 10)))

# Calculate R-squared values for linear model
taxdiv_lm_linear <- lm(taxdiv_r_squared_linear ~ resolution, data = taxdiv_r_squared_compare)
taxdiv_r_squared_linear <- summary(taxdiv_lm_linear)$r.squared
taxdiv_slopecoefficent_linear <- coef(taxdiv_lm_linear)["resolution"]

# Calculate R-squared values for exponential model
taxdiv_lm_exp <- lm(taxdiv_r_squared_exp ~ resolution, data = taxdiv_r_squared_compare)
taxdiv_r_squared_exp <- summary(taxdiv_lm_exp)$r.squared
taxdiv_slopecoefficent_exp <- coef(taxdiv_lm_exp)["resolution"]

(taxdiv_r_squared_plot_trendlines <- ggplot() +
  geom_smooth(aes(x = taxdiv_r_squared_compare$resolution, y = taxdiv_r_squared_compare$taxdiv_r_squared_linear), 
              method = "lm", se = FALSE, color = "red", linetype = "solid") +
  geom_smooth(aes(x = taxdiv_r_squared_compare$resolution, y = taxdiv_r_squared_compare$taxdiv_r_squared_exp), 
              method = "lm", se = FALSE, color = "purple", linetype = "solid") +
  geom_point(aes(x = taxdiv_r_squared_long$resolution, y = taxdiv_r_squared_long$taxdiv_r_squared, color = taxdiv_r_squared_long$model),
             size = 3) +
  scale_color_manual(values = custom_colors, labels = custom_labels) +
  scale_x_continuous(breaks = taxdiv_r_squared_compare$resolution) + 
  labs(title = "Comparison of Richness R-squared Values by Resolution",
       x = "Resolution (km)",
       y = "R-squared Value",
       color = "Model Type") +
  theme_minimal() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black"),
        axis.ticks = element_line(color = "black"),
        axis.text = element_text(color = "black"),
        legend.position = "inside",
        legend.position.inside = c(.85, .3),
        legend.box.margin = margin(10, 10, 10, 10)) +
  annotate("text", x = 55, y = 0.4, label = paste("R² =", round(taxdiv_r_squared_linear, 2)), color = "red", size = 4) +
  annotate("text", x = 60, y = 0.7, label = paste("R² =", round(taxdiv_r_squared_exp, 2)), color = "purple", size = 4))
```

```{r}
ggtrendline(taxdiv_r_squared_compare$resolution, taxdiv_r_squared_compare$taxdiv_r_squared_linear, model = "line2P", summary = FALSE, linecolor = "red", linewidth = 1, CI.level = NA, show.eq = FALSE, show.pvalue = FALSE, text.col = "red", eSize = 4)
ggtrendline(taxdiv_r_squared_compare$resolution, taxdiv_r_squared_compare$taxdiv_r_squared_exp, model = "line2P", summary = FALSE, linecolor = "purple", linewidth = 1, CI.level = NA, show.eq = FALSE, show.pvalue = FALSE, text.col = "purple", eSize = 4)

ggtrendline(taxdiv_r_squared_compare$resolution, taxdiv_r_squared_compare$taxdiv_r_squared_linear, model = "line3P", summary = FALSE, linecolor = "red", linewidth = 1, CI.level = NA, show.eq = FALSE, show.pvalue = FALSE, text.col = "red", eSize = 4)
ggtrendline(taxdiv_r_squared_compare$resolution, taxdiv_r_squared_compare$taxdiv_r_squared_exp, model = "line3P", summary = FALSE, linecolor = "purple", linewidth = 1, CI.level = NA, show.eq = FALSE, show.pvalue = FALSE, text.col = "purple", eSize = 4)
```

## Functional Diversity
```{r}
# create dataframe of resolutions and r-squared values 
resolution <- c(5, 10, 25, 50, 75, 100)
fundiv_r_squared_linear <- c(compare_FDis_r2_5km_linear, compare_FDis_r2_10km_linear, compare_FDis_r2_25km_linear, compare_FDis_r2_50km_linear, compare_FDis_r2_75km_linear, compare_FDis_r2_100km_linear)
fundiv_r_squared_quad <- c(compare_FDis_r2_5km_quad, compare_FDis_r2_10km_quad, compare_FDis_r2_25km_quad, compare_FDis_r2_50km_quad, compare_FDis_r2_75km_quad, compare_FDis_r2_100km_quad)
fundiv_r_squared_compare <- data.frame(resolution, fundiv_r_squared_linear, fundiv_r_squared_quad)

# reshape to long format for ease of plotting
fundiv_r_squared_long <- fundiv_r_squared_compare %>%
  pivot_longer(cols = c(fundiv_r_squared_linear, fundiv_r_squared_quad), 
               names_to = "model", 
               values_to = "fundiv_r_squared")

```

```{r}
# Define custom colors for the points
custom_colors <- c(fundiv_r_squared_linear = "red", fundiv_r_squared_quad = "blue")

# Define custom legend labels
custom_labels <- c(fundiv_r_squared_quad = "Quadratic Model", fundiv_r_squared_linear = "Linear Model")

(fundiv_r_squared_plot <- ggplot(fundiv_r_squared_long,aes(x = resolution, y = fundiv_r_squared, color = model)) +
  geom_line() +
  geom_point(size = 3) +
  scale_color_manual(values = custom_colors, labels = custom_labels) +
  scale_x_continuous(breaks = fundiv_r_squared_compare$resolution) + 
  labs(title = "Comparison of FDis R-squared Values by Resolution",
       x = "Resolution (km)",
       y = "R-squared Value",
       color = "Model Type") +
  theme_minimal() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black"),
        axis.ticks = element_line(color = "black"),
        axis.text = element_text(color = "black"),
        legend.position = "inside",
        legend.position.inside = c(.9, .2),
        legend.box.margin = margin(10, 10, 10, 10)))

# Calculate R-squared values for linear model
fundiv_lm_linear <- lm(fundiv_r_squared_linear ~ resolution, data = fundiv_r_squared_compare)
fundiv_r_squared_linear <- summary(fundiv_lm_linear)$r.squared

# Calculate R-squared values for quadratic model
fundiv_lm_quad <- lm(fundiv_r_squared_quad ~ resolution, data = fundiv_r_squared_compare)
fundiv_r_squared_quad <- summary(fundiv_lm_quad)$r.squared

(fundiv_r_squared_plot_trendlines <- ggplot() +
  geom_smooth(aes(x = fundiv_r_squared_compare$resolution, y = fundiv_r_squared_compare$fundiv_r_squared_linear), 
              method = "lm", se = FALSE, color = "red", linetype = "solid") +
  geom_smooth(aes(x = fundiv_r_squared_compare$resolution, y = fundiv_r_squared_compare$fundiv_r_squared_quad), 
              method = "lm", se = FALSE, color = "blue", linetype = "solid") +
  geom_point(aes(x = fundiv_r_squared_long$resolution, y = fundiv_r_squared_long$fundiv_r_squared, color = fundiv_r_squared_long$model),
             size = 3) +
  scale_color_manual(values = custom_colors, labels = custom_labels) +
  scale_x_continuous(breaks = fundiv_r_squared_compare$resolution) + 
  labs(title = "Comparison of FDis R-squared Values by Resolution",
       x = "Resolution (km)",
       y = "R-squared Value",
       color = "Model Type") +
  theme_minimal() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black"),
        axis.ticks = element_line(color = "black"),
        axis.text = element_text(color = "black"),
        legend.position = "inside",
        legend.position.inside = c(.9, .2),
        legend.box.margin = margin(10, 10, 10, 10)) +
  annotate("text", x = 70, y = 0.2, label = paste("R² =", round(fundiv_r_squared_linear, 2)), color = "red", size = 4) +
  annotate("text", x = 70, y = 0.6, label = paste("R² =", round(fundiv_r_squared_quad, 2)), color = "blue", size = 4))
```

# Compare slope coefficents by resolution
## Taxonomic diversity
```{r}
# Create a tibble
taxdiv_slopecoefficent_compare <- tibble(
  resolution = c(5, 10, 25, 50, 75, 100),
  taxdiv_slopecoefficent_linear = c(
    as.numeric(compare_Richness_slopecoefficent_5km_linear$a), 
    as.numeric(compare_Richness_slopecoefficent_10km_linear$a), 
    as.numeric(compare_Richness_slopecoefficent_25km_linear$a), 
    as.numeric(compare_Richness_slopecoefficent_50km_linear$a), 
    as.numeric(compare_Richness_slopecoefficent_75km_linear$a), 
    as.numeric(compare_Richness_slopecoefficent_100km_linear$a)
  ),
  taxdiv_slopecoefficent_exp = c(
    as.numeric(compare_Richness_slopecoefficent_5km_exp$b), 
    as.numeric(compare_Richness_slopecoefficent_10km_exp$b), 
    as.numeric(compare_Richness_slopecoefficent_25km_exp$b), 
    as.numeric(compare_Richness_slopecoefficent_50km_exp$b), 
    as.numeric(compare_Richness_slopecoefficent_75km_exp$b), 
    as.numeric(compare_Richness_slopecoefficent_100km_exp$b)
  )
)

# reshape to long format for ease of plotting
taxdiv_slopecoefficent_long <- taxdiv_slopecoefficent_compare %>%
  pivot_longer(cols = c(taxdiv_slopecoefficent_linear, taxdiv_slopecoefficent_exp), 
               names_to = "model", 
               values_to = "slope_coefficent")
```

```{r}
# Define custom colors for the points
custom_colors <- c(taxdiv_slopecoefficent_linear = "red", taxdiv_slopecoefficent_exp = "purple")

# Define custom legend labels
custom_labels <- c(taxdiv_slopecoefficent_linear = "Linear Model", taxdiv_slopecoefficent_exp = "Exponential Model")

(taxdiv_slopecoefficent_plot <- ggplot(taxdiv_slopecoefficent_long,aes(x = resolution, y = slope_coefficent, color = model)) +
  geom_line() +
  geom_point(size = 3) +
  scale_color_manual(values = custom_colors, labels = custom_labels) +
  scale_x_continuous(breaks = taxdiv_slopecoefficent_compare$resolution) + 
  labs(title = "Comparison of Richness Slope Coefficents by Resolution",
       x = "Resolution (km)",
       y = "Slope Coefficent",
       color = "Model Type") +
  theme_minimal() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black"),
        axis.ticks = element_line(color = "black"),
        axis.text = element_text(color = "black"),
        legend.position = "inside",
        legend.position.inside = c(.85, .3),
        legend.box.margin = margin(10, 10, 10, 10)))
```

## Functional diversity
```{r}
# Create a tibble
fundiv_slopecoefficent_compare <- tibble(
  resolution = c(5, 10, 25, 50, 75, 100),
  fundiv_slopecoefficent_linear = c(
    as.numeric(compare_FDis_slopecoefficent_5km_linear$a), 
    as.numeric(compare_FDis_slopecoefficent_10km_linear$a), 
    as.numeric(compare_FDis_slopecoefficent_25km_linear$a), 
    as.numeric(compare_FDis_slopecoefficent_50km_linear$a), 
    as.numeric(compare_FDis_slopecoefficent_75km_linear$a), 
    as.numeric(compare_FDis_slopecoefficent_100km_linear$a)
  ),
  fundiv_slopecoefficent_quad = c(
    as.numeric(compare_FDis_slopecoefficent_5km_quad$b), 
    as.numeric(compare_FDis_slopecoefficent_10km_quad$b), 
    as.numeric(compare_FDis_slopecoefficent_25km_quad$b), 
    as.numeric(compare_FDis_slopecoefficent_50km_quad$b), 
    as.numeric(compare_FDis_slopecoefficent_75km_quad$b), 
    as.numeric(compare_FDis_slopecoefficent_100km_quad$b)
  )
)

# reshape to long format for ease of plotting
fundiv_slopecoefficent_long <- fundiv_slopecoefficent_compare %>%
  pivot_longer(cols = c(fundiv_slopecoefficent_linear, fundiv_slopecoefficent_quad), 
               names_to = "model", 
               values_to = "slope_coefficent")
```

```{r}
# Define custom colors for the points
custom_colors <- c(fundiv_slopecoefficent_linear = "red", fundiv_slopecoefficent_quad = "blue")

# Define custom legend labels
custom_labels <- c(fundiv_slopecoefficent_linear = "Linear Model", fundiv_slopecoefficent_quad = "Quadratic Model")

(fundiv_slopecoefficent_plot <- ggplot(fundiv_slopecoefficent_long,aes(x = resolution, y = slope_coefficent, color = model)) +
  geom_line() +
  geom_point(size = 3) +
  scale_color_manual(values = custom_colors, labels = custom_labels) +
  scale_x_continuous(breaks = fundiv_slopecoefficent_compare$resolution) + 
  labs(title = "Comparison of FDis Slope Coefficents by Resolution",
       x = "Resolution (km)",
       y = "Slope Coefficent",
       color = "Model Type") +
  theme_minimal() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black"),
        axis.ticks = element_line(color = "black"),
        axis.text = element_text(color = "black"),
        legend.position = "inside",
        legend.position.inside = c(.85, .2),
        legend.box.margin = margin(10, 10, 10, 10)))
```

## Compare the linear model slope coefficents for taxonomic and functional diversity
```{r}
# Create a tibble
bothdiv_slopecoefficent_compare <- tibble(
  resolution = c(5, 10, 25, 50, 75, 100),
  taxdiv_slopecoefficent_linear = c(
    as.numeric(compare_Richness_slopecoefficent_5km_linear$a), 
    as.numeric(compare_Richness_slopecoefficent_10km_linear$a), 
    as.numeric(compare_Richness_slopecoefficent_25km_linear$a), 
    as.numeric(compare_Richness_slopecoefficent_50km_linear$a), 
    as.numeric(compare_Richness_slopecoefficent_75km_linear$a), 
    as.numeric(compare_Richness_slopecoefficent_100km_linear$a)
  ),
  fundiv_slopecoefficent_linear = c(
    as.numeric(compare_FDis_slopecoefficent_5km_linear$a), 
    as.numeric(compare_FDis_slopecoefficent_10km_linear$a), 
    as.numeric(compare_FDis_slopecoefficent_25km_linear$a), 
    as.numeric(compare_FDis_slopecoefficent_50km_linear$a), 
    as.numeric(compare_FDis_slopecoefficent_75km_linear$a), 
    as.numeric(compare_FDis_slopecoefficent_100km_linear$a)
  )
)

# reshape to long format for ease of plotting
bothdiv_slopecoefficent_long <- bothdiv_slopecoefficent_compare %>%
  pivot_longer(cols = c(taxdiv_slopecoefficent_linear, fundiv_slopecoefficent_linear), 
               names_to = "type", 
               values_to = "slope_coefficent")
```

```{r}
# Define custom colors for the points
custom_colors <- c(taxdiv_slopecoefficent_linear = "goldenrod", fundiv_slopecoefficent_linear = "forestgreen")

# Define custom legend labels
custom_labels <- c(taxdiv_slopecoefficent_linear = "Taxonomic Diversity", fundiv_slopecoefficent_linear = "Functional Diversity")

(bothdiv_slopecoefficent_plot <- ggplot(bothdiv_slopecoefficent_long,aes(x = resolution, y = slope_coefficent, color = type)) +
  geom_line() +
  geom_point(size = 3) +
  scale_color_manual(values = custom_colors, labels = custom_labels) +
  scale_x_continuous(breaks = bothdiv_slopecoefficent_compare$resolution) + 
  labs(x = "Resolution (km)",
       y = "Slope Coefficent",
       color = "Diversity Measure") +
  theme_minimal() +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black"),
        axis.ticks = element_line(color = "black"),
        axis.text = element_text(color = "black"),
        legend.position = "inside",
        legend.position.inside = c(.85, .2),
        legend.box.margin = margin(10, 10, 10, 10)))
```

# Save plots to file
```{r}
ggsave("compare_Richness_all_plots.png", compare_Richness_all_plots, path = figure_path, width = 10, height = 8, units = "in")
ggsave("compare_FDis_all_plots.png", compare_FDis_all_plots, path = figure_path, width = 10, height = 8, units = "in")
```

```{r}
ggsave("taxdiv_r_squared_plot.png", taxdiv_r_squared_plot, path = figure_path)
ggsave("taxdiv_r_squared_plot_trendlines.png", taxdiv_r_squared_plot_trendlines, path = figure_path)
ggsave("fundiv_r_squared_plot.png", fundiv_r_squared_plot, path = figure_path)
ggsave("fundiv_r_squared_plot_trendlines.png", fundiv_r_squared_plot_trendlines, path = figure_path)
```

```{r}
ggsave("taxdiv_slopecoefficent_plot.png", taxdiv_slopecoefficent_plot, path = figure_path)
ggsave("fundiv_slopecoefficent_plot.png", fundiv_slopecoefficent_plot, path = figure_path)
ggsave("bothdiv_slopecoefficent_plot.png", bothdiv_slopecoefficent_plot, path = figure_path)
```

# Compare same tax taxanomic and functional diversity
```{r}
same_taxa_diversity <- function(richness_grid, fdis_grid, resolution_meters, z){
  # Extract cell values
  comparison_diversity <- data.frame(Richness = richness_grid$num_species,
                                    FDis = fdis_grid$fdis_value)

  # Filter out rows where either richness or FDis is zero
  comparison_diversity_filtered <- comparison_diversity %>%
                                filter(Richness > 0 & FDis > 0)

  ## Create scatter plots
  (compare_diversity_plot_line2P <- ggtrendline(comparison_diversity_filtered$Richness,
               comparison_diversity_filtered$FDis, model = z, linewidth = 1, CI.level = NA, eSize = 4) +
      geom_point(aes(x = comparison_diversity_filtered$Richness,
                                           y = comparison_diversity_filtered$FDis)) +
      labs(title = paste("Richness vs FDis [", resolution_meters/1000, " km]", sep = ""),
       x = "Richness by cell", y = "FDis by cell"))

}
```

```{r}
same_taxa_diversity(plant_cellRichness_100km, plant_cellFDis_100km, 100000, "line2P")
same_taxa_diversity(plant_cellRichness_100km, plant_cellFDis_100km, 100000, "line3P")
same_taxa_diversity(plant_cellRichness_100km, plant_cellFDis_100km, 100000, "log2P")
same_taxa_diversity(plant_cellRichness_100km, plant_cellFDis_100km, 100000, "exp2P")
same_taxa_diversity(plant_cellRichness_100km, plant_cellFDis_100km, 100000, "exp3P")
same_taxa_diversity(plant_cellRichness_100km, plant_cellFDis_100km, 100000, "power2P")
same_taxa_diversity(plant_cellRichness_100km, plant_cellFDis_100km, 100000, "power3P")
```

```{r}
same_taxa_diversity(plant_cellRichness_75km, plant_cellFDis_75km, 75000, "line2P")
same_taxa_diversity(plant_cellRichness_75km, plant_cellFDis_75km, 75000, "line3P")
same_taxa_diversity(plant_cellRichness_75km, plant_cellFDis_75km, 75000, "log2P")
same_taxa_diversity(plant_cellRichness_75km, plant_cellFDis_75km, 75000, "exp2P")
same_taxa_diversity(plant_cellRichness_75km, plant_cellFDis_75km, 75000, "exp3P")
same_taxa_diversity(plant_cellRichness_75km, plant_cellFDis_75km, 75000, "power2P")
same_taxa_diversity(plant_cellRichness_75km, plant_cellFDis_75km, 75000, "power3P")
```