### 2024-09-03 code/Manuscript_figures.rmd
change figure colors, combine figures into images for manuscripts

---
title: "Manuscript Figures"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script generates misc figures for manuscript that corresponds to this project."
date: "2024-05-14"
output: 
    html_document:
      code_link: TRUE
---

# Set file paths
```{r}
figure_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/figures')
```

# Load required packages
```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggpubr)
library(networkD3)
#devtools::install_github("davidsjoberg/ggsankey")
library(ggsankey)
```

# Creating cleaning occurrence data figures

These numbers are recorded from the summary function at the end of most L0 and L1 scripts. A table is available in the manuscript supplemental materials.

## Frugivores
```{r}
Step <- c("Raw", "After cleaning", "Full trait coverage")
frug_records <- c(8885330, 1721825, 1667892)
frug_species <- c(1360, 1219, 1074)
frug_genera <- c(400, 398, 370)
frug_families <- c(66, 66, 64)

frug_summary_df <- data.frame(Step, frug_records, frug_species, frug_genera, frug_families)

# Reshape data to long format suitable for plotting
frug_summary_long <- frug_summary_df %>%
  pivot_longer(cols = starts_with("frug"), names_to = "Category", values_to = "Value")
```

```{r}
# Define the order of steps with "Raw" at the top
step_order <- c("Full trait coverage", "After cleaning", "Raw")

# Convert Step to factor with defined order
frug_summary_long$Step <- factor(frug_summary_long$Step, levels = step_order)

# Replace Category labels with more descriptive names
frug_summary_long <- frug_summary_long %>%
  mutate(Category = case_when(
    Category == "frug_records"   ~ "Records",
    Category == "frug_species"   ~ "Species",
    Category == "frug_genera"    ~ "Genera",
    Category == "frug_families"  ~ "Families",
    TRUE                         ~ as.character(Category)  # Handle any other cases
  ))

# Define the order of categories
category_order <- c("Records", "Species", "Genera", "Families")

# Convert Category to factor with defined order
frug_summary_long$Category <- factor(frug_summary_long$Category, levels = category_order)

# Define custom colors
frug_colors <- c("Records" = "darkgoldenrod", "Species" = "darkgoldenrod3", "Genera" = "darkgoldenrod2", "Families" = "darkgoldenrod1")

```

```{r}
(frugivore_summary_plot <- ggplot(frug_summary_long, aes(y = Step, x = Value, fill = Category)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(. ~ Category, scales = "free_x", nrow = 1) +
  scale_fill_manual(values = frug_colors) +
  labs(title = "Frugivore",
       x = "Value",
       y = NULL) +
  theme_minimal() +
  theme(strip.background = element_blank(),  # Remove background from facet labels
        strip.text = element_text(size = 12, face = "bold"),
        legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)))
ggsave("frugivore_summary_plot.png", frugivore_summary_plot, path = figure_path)
```

## Plants

```{r}
Step <- c("Raw", "After cleaning", "Full trait coverage")
plant_records <- c(4519792, 1886120, 658609)
plant_species <- c(40288, 39972, 3128)
plant_genera <- c(4083, 3793, 1359)
plant_families <- c(307, 293, 213)

plant_summary_df <- data.frame(Step, plant_records, plant_species, plant_genera, plant_families)

# Reshape data to long format suitable for plotting
plant_summary_long <- plant_summary_df %>%
  pivot_longer(cols = starts_with("plant"), names_to = "Category", values_to = "Value")
```

```{r}
# Define the order of steps with "Raw" at the top
step_order <- c("Full trait coverage", "After cleaning", "Raw")

# Convert Step to factor with defined order
plant_summary_long$Step <- factor(plant_summary_long$Step, levels = step_order)

# Replace Category labels with more descriptive names
plant_summary_long <- plant_summary_long %>%
  mutate(Category = case_when(
    Category == "plant_records"   ~ "Records",
    Category == "plant_species"   ~ "Species",
    Category == "plant_genera"    ~ "Genera",
    Category == "plant_families"  ~ "Families",
    TRUE                         ~ as.character(Category)  # Handle any other cases
  ))

# Define the order of categories
category_order <- c("Records", "Species", "Genera", "Families")

# Convert Category to factor with defined order
plant_summary_long$Category <- factor(plant_summary_long$Category, levels = category_order)

# Define custom colors
plant_colors <- c("Records" = "palegreen4", "Species" = "palegreen3", "Genera" = "palegreen2", "Families" = "palegreen1")

```

```{r}
(plant_summary_plot <- ggplot(plant_summary_long, aes(y = Step, x = Value, fill = Category)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(. ~ Category, scales = "free_x", nrow = 1) +
  scale_fill_manual(values = plant_colors) +
  labs(title = "Plant",
       x = "Value",
       y = NULL) +
  theme_minimal() +
  theme(strip.background = element_blank(),  # Remove background from facet labels
        strip.text = element_text(size = 12, face = "bold"),
        legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)))
ggsave("plant_summary_plot.png", plant_summary_plot, path = figure_path)
```

## Combine data summary figures
```{r}
(both_summary_plots <- ggarrange(plant_summary_plot, frugivore_summary_plot,
                                        ncol = 1, nrow = 2))
ggsave("both_summary_plots.png", plot = both_summary_plots, path = figure_path, height = 4, width = 10, units = "in")
```

# Create spatial grain maps
```{r}
library(rnaturalearth)
library(sf)
```

```{r}
data_path_L0<-file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path_L0, "Forest_sf.shp"), layer = "Forest_sf")
```

```{r}
worldMap <- ne_countries(scale = "medium", type = "countries", returnclass = "sf")
Americas <- ne_countries(continent = c("North America", "South America"), returnclass = "sf")
#polygon of Tropical Andes
TApoly <- worldMap %>% filter(sovereignt == "Bolivia" |sovereignt == "Ecuador" | sovereignt == "Venezuela" | sovereignt == "Colombia" | sovereignt == "Peru")
Ecuador <- worldMap %>% filter(sovereignt == "Ecuador")
```

```{r}
# Check and set CRS
# Assuming your data is initially in WGS 84 (EPSG:4326), reproject to a suitable projected CRS
TApoly <- st_transform(TApoly, crs = 32718) # UTM Zone 18S for the Andes region
TropicalAndes_IUCNHabitat_Forest <- st_transform(TropicalAndes_IUCNHabitat_Forest, crs = 32718)
Ecuador <- st_transform(Ecuador, crs = 32718)
Americas <- st_transform(Americas, crs = 32718)
```

Make & map grid for TA
```{r}
TAGrid_5km <- TApoly %>%
    st_make_grid(cellsize = c(5000, 5000)) %>%
    st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
    st_cast("MULTIPOLYGON") %>%
    st_sf() %>%
    mutate(cellid = row_number())
```

```{r}
(TAGrid_5km_plot <- ggplot(data = TAGrid_5km) +
    geom_sf(data = Americas) +
    geom_sf(data = TAGrid_5km) +
    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
    theme_minimal() +
    labs(title = "5 km") +
    theme(plot.title = element_text(hjust = 0.5)))
```

```{r}
TAGrid_10km <- TApoly %>%
    st_make_grid(cellsize = c(10000, 10000)) %>%
    st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
    st_cast("MULTIPOLYGON") %>%
    st_sf() %>%
    mutate(cellid = row_number())
```

```{r}
(TAGrid_10km_plot <- ggplot(data = TAGrid_10km) +
    geom_sf(data = Americas) +
    geom_sf(data = TAGrid_10km) +
    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
    theme_minimal() +
    labs(title = "10 km") +
    theme(plot.title = element_text(hjust = 0.5)))
```

```{r}
TAGrid_25km <- TApoly %>%
    st_make_grid(cellsize = c(25000, 25000)) %>%
    st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
    st_cast("MULTIPOLYGON") %>%
    st_sf() %>%
    mutate(cellid = row_number())
```

```{r}
(TAGrid_25km_plot <- ggplot(data = TAGrid_25km) +
    geom_sf(data = Americas) +
    geom_sf(data = TAGrid_25km) +
    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
    theme_minimal() +
    labs(title = "25 km") +
    theme(plot.title = element_text(hjust = 0.5)))
```

```{r}
TAGrid_50km <- TApoly %>%
    st_make_grid(cellsize = c(50000, 50000)) %>%
    st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
    st_cast("MULTIPOLYGON") %>%
    st_sf() %>%
    mutate(cellid = row_number())
```

```{r}
(TAGrid_50km_plot <- ggplot(data = TAGrid_50km) +
    geom_sf(data = Americas) +
    geom_sf(data = TAGrid_50km) +
    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
    theme_minimal() +
    labs(title = "50 km") +
    theme(plot.title = element_text(hjust = 0.5)))
```

```{r}
TAGrid_75km <- TApoly %>%
    st_make_grid(cellsize = c(75000, 75000)) %>%
    st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
    st_cast("MULTIPOLYGON") %>%
    st_sf() %>%
    mutate(cellid = row_number())
```

```{r}
(TAGrid_75km_plot <- ggplot(data = TAGrid_75km) +
    geom_sf(data = Americas) +
    geom_sf(data = TAGrid_75km) +
    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
    theme_minimal() +
    labs(title = "75 km") +
    theme(plot.title = element_text(hjust = 0.5)))
```

```{r}
TAGrid_100km <- TApoly %>%
    st_make_grid(cellsize = c(100000, 100000)) %>%
    st_intersection(TropicalAndes_IUCNHabitat_Forest) %>%
    st_cast("MULTIPOLYGON") %>%
    st_sf() %>%
    mutate(cellid = row_number())
```

```{r}
(TAGrid_100km_plot <- ggplot(data = TAGrid_100km) +
    geom_sf(data = Americas) +
    geom_sf(data = TAGrid_100km) +
    coord_sf(xlim = c(-82, -60), ylim = c(-24, 14), expand = FALSE, crs = 4326) +
    theme_minimal() +
    labs(title = "100 km") +
    theme(plot.title = element_text(hjust = 0.5)))
```

Combine grid plots into one figure
```{r}
library(ggpubr)
(TAGrid_all_plots <- ggpubr::ggarrange(TAGrid_5km_plot, TAGrid_10km_plot, TAGrid_25km_plot,
                  TAGrid_50km_plot, TAGrid_75km_plot, TAGrid_100km_plot,
                  ncol = 6, nrow = 1))
ggsave("TAGrid_all_plots.png", TAGrid_all_plots, path = figure_path, width = 16, height = 10, units = "in")
```

```{r}
resolution <- c(5, 10, 25, 50, 75, 100)
num_cells <- c(length(TAGrid_5km$cellid), length(TAGrid_10km$cellid),length(TAGrid_25km$cellid),
               length(TAGrid_50km$cellid), length(TAGrid_75km$cellid),length(TAGrid_100km$cellid))
plot(resolution, num_cells)

```

Make & map grid for just Ecuador
```{r}
EcGrid_5km <- TApoly %>%
    st_make_grid(cellsize = c(5000, 5000)) %>%
    st_intersection(Ecuador) %>%
    st_cast("MULTIPOLYGON") %>%
    st_sf() %>%
    mutate(cellid = row_number())
```

```{r}
(EcGrid_5km_plot <- ggplot(data = EcGrid_5km) +
    geom_sf() +
    theme_minimal() +
    labs(title = "5 km") +
    coord_sf(xlim = c(-81.5, -75), ylim = c(-5.5, 2), expand = FALSE, crs = 4326))
(Grid_5km_plot <- ggplot(data = EcGrid_5km) +
    geom_sf() +
    theme_minimal() +
    labs(title = "5 km") +
    theme(axis.text = element_blank())+
    coord_sf(xlim = c(-79.5, -77.3), ylim = c(-2.5, -0.5), expand = FALSE, crs = 4326))
```

```{r}
EcGrid_10km <- TApoly %>%
    st_make_grid(cellsize = c(10000, 10000)) %>%
    st_intersection(Ecuador) %>%
    st_cast("MULTIPOLYGON") %>%
    st_sf() %>%
    mutate(cellid = row_number())
```

```{r}
(EcGrid_10km_plot <- ggplot(data = EcGrid_10km) +
    geom_sf() +
    theme_minimal() +
    labs(title = "10 km") +
    coord_sf(xlim = c(-81.5, -75), ylim = c(-5.5, 2), expand = FALSE, crs = 4326))
(Grid_10km_plot <- ggplot(data = EcGrid_10km) +
    geom_sf() +
    theme_minimal() +
    labs(title = "10 km") +
    theme(axis.text = element_blank())+
    coord_sf(xlim = c(-79.5, -77.3), ylim = c(-2.5, -0.5), expand = FALSE, crs = 4326))
```

```{r}
EcGrid_25km <- TApoly %>%
    st_make_grid(cellsize = c(25000, 25000)) %>%
    st_intersection(Ecuador) %>%
    st_cast("MULTIPOLYGON") %>%
    st_sf() %>%
    mutate(cellid = row_number())
```

```{r}
(EcGrid_25km_plot <- ggplot(data = EcGrid_25km) +
    geom_sf() +
    theme_minimal() +
    labs(title = "25 km") +
    coord_sf(xlim = c(-81.5, -75), ylim = c(-5.5, 2), expand = FALSE, crs = 4326))
(Grid_25km_plot <- ggplot(data = EcGrid_25km) +
    geom_sf() +
    theme_minimal() +
    labs(title = "25 km") +
    theme(axis.text = element_blank())+
    coord_sf(xlim = c(-79.5, -77.3), ylim = c(-2.5, -0.5), expand = FALSE, crs = 4326))
```

```{r}
EcGrid_50km <- TApoly %>%
    st_make_grid(cellsize = c(50000, 50000)) %>%
    st_intersection(Ecuador) %>%
    st_cast("MULTIPOLYGON") %>%
    st_sf() %>%
    mutate(cellid = row_number())
```

```{r}
(EcGrid_50km_plot <- ggplot(data = EcGrid_50km) +
    geom_sf() +
    theme_minimal() +
    labs(title = "50 km") +
    coord_sf(xlim = c(-81.5, -75), ylim = c(-5.5, 2), expand = FALSE, crs = 4326))
(Grid_50km_plot <- ggplot(data = EcGrid_50km) +
    geom_sf() +
    theme_minimal() +
    labs(title = "50 km") +
    theme(axis.text = element_blank())+ 
    coord_sf(xlim = c(-79.5, -77.3), ylim = c(-2.5, -0.5), expand = FALSE, crs = 4326))
```

```{r}
EcGrid_75km <- TApoly %>%
    st_make_grid(cellsize = c(75000, 75000)) %>%
    st_intersection(Ecuador) %>%
    st_cast("MULTIPOLYGON") %>%
    st_sf() %>%
    mutate(cellid = row_number())
```

```{r}
(EcGrid_75km_plot <- ggplot(data = EcGrid_75km) +
    geom_sf() +
    theme_minimal() +
    labs(title = "75 km") +
    coord_sf(xlim = c(-81.5, -75), ylim = c(-5.5, 2), expand = FALSE, crs = 4326))
(Grid_75km_plot <- ggplot(data = EcGrid_75km) +
    geom_sf() +
    theme_minimal() +
    labs(title = "75 km") +
    theme(axis.text = element_blank())+
    coord_sf(xlim = c(-79.5, -77.3), ylim = c(-2.5, -0.5), expand = FALSE, crs = 4326))
```

```{r}
EcGrid_100km <- TApoly %>%
    st_make_grid(cellsize = c(100000, 100000)) %>%
    st_intersection(Ecuador) %>%
    st_cast("MULTIPOLYGON") %>%
    st_sf() %>%
    mutate(cellid = row_number())
```

```{r}
(EcGrid_100km_plot <- ggplot(data = EcGrid_100km) +
    geom_sf() +
    theme_minimal() +
    labs(title = "100 km") +
    coord_sf(xlim = c(-81.5, -75), ylim = c(-5.5, 2), expand = FALSE, crs = 4326))
(Grid_100km_plot <- ggplot(data = EcGrid_100km) +
    geom_sf() +
    theme_minimal() +
    labs(title = "100 km") +
    theme(axis.text = element_blank())+
    coord_sf(xlim = c(-79.5, -77.3), ylim = c(-2.5, -0.5), expand = FALSE, crs = 4326))
```

Combine Ecuador grid plots into one figure
```{r}
library(ggpubr)
(Grid_all_plots <- ggpubr::ggarrange(Grid_5km_plot, Grid_10km_plot, Grid_25km_plot,
                  Grid_50km_plot, Grid_75km_plot, Grid_100km_plot,
                  ncol = 6, nrow = 1))
ggsave("Grid_all_plots.png", Grid_all_plots, path = figure_path, width = 16, height = 10, units = "in")
```

# Create Sankey diagram for taxonomic harmonization
Inspiration figures: https://conbio-onlinelibrary-wiley-com.proxy2.cl.msu.edu/doi/10.1111/csp2.311#csp2311-fig-0003 and https://bsapubs-onlinelibrary-wiley-com.proxy2.cl.msu.edu/doi/full/10.1002/aps3.11575#aps311575-fig-0003

Package used (databases)    Plant species inputted    Matches   Disparities
TNRS                        487,201                   472,680   14,548
bdc (COL)                   14,548                    5,177     9,371

*The following code doesn't generate the figure like I want.*
```{r}
# Create connection dataframe
links <- data.frame(
  source = c("Plant species inputted", "TNRS", "TNRS", "Disparities", "bdc (COL)", "bdc (COL)", "Matches"),
  target = c("TNRS", "Matches", "Disparities", "bdc (COL)", "Matches", "Species dropped", "Species harmonized"),
  value = c(487201, 472680, 14548, 14548, 5177, 9371, 476010)
)

# Create nodes
nodes <- data.frame(
  name = c(as.character(links$source),
           as.character(links$target) %>% unique())
)

# Reformat connections
links$IDsource <- match(links$source, nodes$name)-1
links$IDtarget <- match(links$target, nodes$name)-1

# Create the Sankey diagram
sankeyNetwork(Links = links, Nodes = nodes, Source = "IDsource", Target = "IDtarget",
              Value = "value", NodeID = "name")
```

```{r}
# Create the data frames for the nodes and links
sankey_data <- data.frame(
  x = c("source", "target", "target", "source", "target", "target", "source", "target", "source"),
  node = c("Plant species inputted", "TNRS", "TNRS", "Disparities", "bdc (COL)", "bdc (COL)", "Species dropped", "Matches", "Species harmonized"),
  next_x = c("target", "source", "source", "target", "source", "source", NA, "source", NA),
  next_node = c("TNRS", "Matches", "Disparities", "bdc (COL)", "Matches", "Species dropped", NA, "Species harmonized", NA),
  value = c(487201, 472680, 14548, 14548, 5177, 9371, 9371, 476010, 476010)
)

```

```{r}
# Create the Sankey diagram
ggplot(sankey_data, aes(x = x, next_x = next_x, node = node, next_node = next_node, value = value)) +
  geom_sankey(flow.fill = "darkgrey", node.fill = "darkblue", node.color = "darkblue", alpha = 0.5) +
  geom_sankey_label(aes(label = node), size = 3, fill = "white") +
  theme_sankey(base_size = 12) +
  theme_void() +
  labs(title = "Plant Species Matching Workflow",
       subtitle = "Visualization of species input, TNRS and BDC matching process",
       x = NULL,
       y = NULL) +
  theme(legend.position = "none")
```

# References

```{r}
library(report)
cite_packages()
```

```{r, class.source= 'fold-hide'}
devtools::session_info()
```