### 2025-08-14 code/L0/L0_TropicalAndes_frugivores_GBIF.Rmd
Added name and few minor changes;
computer did not like running line 78- return to this

---
title: "Tropical Andes montane and lowland forest bird and mammal species GBIF"
author: "Hazel J. Anderson, Jenna B. Baljunas"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske"
overview: "This script retrives plant occurrence data from GBIF using the species list from Frugivoria."
data input: "TropicalAndes_Frugivoria_frugivore_traits_species.csv"
data output: "TropicalAndes_GBIF_frugivore_occ.csv and TropicalAndes_GBIF_frugivore_occ_species.csv"
date: "2023-07-25"
output: html_document
---

# Set file paths
```{r}
data_path_L0 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path_L0 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(dplyr)
library(rgbif) 
library(sf)
```

# Read in Frugivoria species list
```{r}
frugivore_species <- read.csv(file.path(data_path_L0,"TropicalAndes_Frugivoria_frugivore_traits_species.csv"))
```

# Download data
Data downloaded on 2023-08-02
```{r}
file_name <- "0009771-231002084531237.zip"

if (!file.exists(file.path(output_path_L0, file_name))) {
    # Match species names to GBIF  
    gbif_taxon_keys <- frugivore_species$x %>% 
      name_backbone_checklist() %>% # match to backbone 
      filter(!matchType == "NONE") %>% # get matched names
      pull(usageKey) 
    # Make GBIF request
    occ_download(
      pred_in("taxonKey", gbif_taxon_keys),
      pred("hasCoordinate", TRUE),
      pred("hasGeospatialIssue", FALSE),
      pred_in("country",c("EC","CO","VE", "PE", "BO")), 
      format = "SIMPLE_CSV"
    )
    # Download GBIF data
  d <- occ_download_get("0009771-231002084531237",
                        path = output_path_L0, overwrite = TRUE) %>%
    occ_download_import()
  } else {
    cat("File already exists:", file_name, "\n")
    d <- occ_download_get("0009771-231002084531237",
                        path = output_path_L0, overwrite = TRUE) %>%
                        occ_download_import()
  }
```

# Subset occurrence to Tropical Andes IUCN Montane and Lowland forest
## Read in Forest shape
```{r}
TropicalAndes_IUCNHabitat_Forest <- read_sf(file.path(data_path_L0, "Forest_sf.shp"), layer = "Forest_sf")
```

## Convert data frame to sf object
```{r}
TropicalAndes_frugivore_occ <- st_as_sf(x = d, 
                        coords = c("decimalLongitude", "decimalLatitude"),
                        crs = "+proj=longlat +datum=WGS84 +no_defs")
```

# Crop occurrence data to forest extent
```{r}
TropicalAndes_frugivore_occ_forest <- TropicalAndes_frugivore_occ[TropicalAndes_IUCNHabitat_Forest,]
```

## Extract lat and long to seperate columns
```{r}
TropicalAndes_frugivore_occ_forest <- TropicalAndes_frugivore_occ_forest %>%
    mutate(decimalLongitude = sf::st_coordinates(.)[,1],
           decimalLatitude = sf::st_coordinates(.)[,2])
```

```{r}
TropicalAndes_frugivore_occ_forest <- as.data.frame(TropicalAndes_frugivore_occ_forest)
```

# Fix column type 
```{r}
TropicalAndes_frugivore_occ_forest$gbifID <- as.character(TropicalAndes_frugivore_occ_forest$gbifID)
TropicalAndes_frugivore_occ_forest$geometry <- NULL
```

# Summary 
```{r}
glimpse(TropicalAndes_frugivore_occ_forest)
```

```{r}
source("~/GitHub/neotropical_plants/code/data_summary.R")

data_summary(TropicalAndes_frugivore_occ_forest, TropicalAndes_frugivore_occ_forest$species, TropicalAndes_frugivore_occ_forest$genus, TropicalAndes_frugivore_occ_forest$family)
```

# Extract species list
```{r}
GBIF_occ_SpeciesList <- unique(TropicalAndes_frugivore_occ_forest$species)
```

# Write data to csv
```{r}
write.csv(TropicalAndes_frugivore_occ_forest, file = file.path(output_path_L0,"TropicalAndes_GBIF_frugivore_occ.csv"))

write.csv(GBIF_occ_SpeciesList, file = file.path(output_path_L0,"TropicalAndes_GBIF_frugivore_occ_species.csv"))
```

# Package Citations and Session Info
```{r}
library(report)
report::cite_packages()

devtools::session_info()
```