### 2025-08-27 code/L0/L0_TropicalAndes_plants_BIEN_traits.Rmd
Small error in output_path name; could not download AllDesired_BIEN_plant_traits, only TropicalAndes

---
title: "Tropical Andes plant trait data BIEN"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske, and Jenna B. Baljunas"
overview: "This script retrieves plant trait data from the BIEN database for Tropical Andean countries."
data input: "none"
data output: "TropicalAndes_BIEN_traits.csv, AllDesired_BIEN_plant_traits.csv"
date: "2023-10-17"
output: html_document
---

# Set file paths
```{r}
data_path_L0 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path_L0 <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(BIEN)
library(tidyr)
library(dplyr)
```

# Retrived desired trait data from BIEN
Data downloaded on 2023-10-25
```{r}
BIEN_trait_list()
```

## List of desired traits
```{r}
trait_list <- c("whole plant dispersal syndrome", "whole plant height", "maximum whole plant longevity", "maximum whole plant height", "minimum whole plant height", "longest whole plant longevity", "fruit type", "maximum fruit length", "minimum fruit length", "seed mass", "whole plant growth form", "whole plant growth form diversity")
```

## List of desired countries
```{r}
TA_countries <- c("Bolivia", "Ecuador", "Venezuela", "Colombia", "Peru")
```

## Retrive all BIEN records for desired traits
```{r}
AllDesired_BIEN_plant_traits <- BIEN_trait_trait(trait_list, all.taxonomy = TRUE, political.boundaries = TRUE, source.citation = TRUE)
```

## Retrive all BIEN records for desired traits in Tropical Andes countries
```{r}
TropicalAndes_BIEN_plant_traits <- BIEN_trait_country(TA_countries, trait_list, all.taxonomy = TRUE, political.boundaries = TRUE, source.citation = TRUE)
```

# Data summary

## Desired traits
```{r}
glimpse(AllDesired_BIEN_plant_traits)
```

```{r}
source("C:/GitHub_projects/plant-frugivore diversity/neotropical_plants/code/data_summary.R")

data_summary(AllDesired_BIEN_plant_traits, AllDesired_BIEN_plant_traits$scrubbed_species_binomial, AllDesired_BIEN_plant_traits$scrubbed_genus, AllDesired_BIEN_plant_traits$scrubbed_family)
```

```{r}
AllDesired_BIEN_plant_traits %>% count(trait_name)
```

## Desired traits in Tropical Andes countries
```{r}
glimpse(TropicalAndes_BIEN_plant_traits)
```

```{r}
source("C:/GitHub_projects/plant-frugivore diversity/neotropical_plants/code/data_summary.R")

data_summary(AllDesired_BIEN_plant_traits, TropicalAndes_BIEN_plant_traits$scrubbed_species_binomial, TropicalAndes_BIEN_plant_traits$scrubbed_genus, TropicalAndes_BIEN_plant_traits$scrubbed_family)
```

```{r}
TropicalAndes_BIEN_plant_traits %>% count(trait_name)
```

# Write data to csv
```{r}
write.csv(AllDesired_BIEN_plant_traits, file.path(output_path_L0,"AllDesired_BIEN_plant_traits.csv"))
write.csv(TropicalAndes_BIEN_plant_traits, file.path(output_path_L0,"TropicalAndes_BIEN_plant_traits.csv"))
```