### 2025-08-27 code/L0/L0_TropicalAndes_plants_TRY.Rmd
Still cannot import TRY data

---
title: "Tropical Andes plant trait data TRY"
author: "Hazel J. Anderson"
project: "Plant-Frugivore Diversity"
collaborators: "Beth E. Gerstner, Phoebe L. Zarnetske, and Jenna B. Baljunas"
overview: "This script retrieves plant trait data from the TRY database for plant species list."
data input: "none"
data output: "TropicalAndes_TRY_traits.csv"
date: "2023-07-18"
output: html_document
---

# Set file paths
```{r}
data_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
output_path <- file.path('G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0')
```

# Load required packages
```{r}
library(rtry)
library(dplyr)
library(tidyr)
```

# Load TRY data
Data was obtained from https://www.try-db.org/TryWeb/Prop0.php on 2023-10-24
```{r}
try_data <- rtry_import("G:/Shared drives/SpaCE_Lab_FRUGIVORIA/data/plants/L0/TRY_2023/29548_24102023041242/29548.txt")
```

# Summary
```{r}
glimpse(try_data)
```

# Keep only record with non-NA trait IDs
```{r}
try_data <- try_data %>%
  drop_na(TraitID)
nrow(try_data)
```

```{r}
print("The number of records is")
nrow(try_data)
print("The number of species is" )
length(unique(try_data$AccSpeciesName))
```

```{r}
try_data %>% count(TraitName)
```

# Get units for each trait
```{r}
try_data %>% distinct(TraitName, UnitName)
```

# Write data to csv
```{r}
write.csv(try_data, file = file.path(output_path,"TropicalAndes_TRY_plant_traits.csv") )
```